---
title: "Dry_Tortugas_RVC"
author: "Megan Hepner"
date: "6/9/2017"
output: html_document
---

## Inputs 
1. species-trait matrix
1. domain abundance dataframe
1. strata abundance dataframe 
1. psu abundance dataframe
1. domain biomass dataframe
1. strata biomass dataframe 
1. psu biomass dataframe

## Outputs 
1.  Functional dendogram
1.  Trait-matrix visualization 
1.  By *domain* for each sampling year and region  
    + Species weighted abundance and plot
    + Species weighted biomass and plot
    + Species biodiversity 
      + species richness and plot
      + Simpson diversity as effective number of species and plot
      + Shannon diversity as effective number of species and plot
      + Functional diversity as effective number of species and plot
1. By *upper keys and lower keys domain* for each sampling year and region 
    + Species weighted abundance and plot
    + Species weighted biomass and plot
    + Species biodiversity 
      + species richness and plot
      + Simpson diversity as effective number of species and plot
      + Shannon diversity as effective number of species and plot
      + Functional diversity as effective number of species and plot
1. By *strata* for each sampling year and region 
    + Species weighted abundance and plot
    + Species weighted biomass and plot
    + Species biodiversity 
      + species richness and plot
      + Simpson diversity as effective number of species and plot
      + Shannon diversity as effective number of species and plot
      + Functional diversity as effective number of species and plot
1. By *primary sampling unit* for each sampling year and region and plot
    + Species weighted abundance and plot
    + Species weighted biomass and plot
    + Species biodiversity 
      + species richness and plot
      + Simpson diversity as effective number of species and plot
      + Shannon diversity as effective number of species and plot
      + Functional diversity as effective number of species and plot
    + statistical tests by strata and year: Shapiro, homocasdicity, 2-way ANOVA
1. Environmental factors from SERC 
1. Heat map of the Florida Keys and Dry Tortugas for the diversity indices 

### Dry Tortugas domain
Protected status:
- 0 = Unprotected 
- 1 = Dry Tortugas National Park (DRTO)
- 2 = Tortugas Northeast Reserve (TNER)
- All = All

```{r domain abundance}
domain_dt_abun <- read_csv('big_csv/abundance_domain/domain_dt_abun.csv',
    col_types = cols(
      YEAR = col_integer(),
      REGION = col_character(),
      SPECIES_CD = col_character(),
      abundance = col_double(),
      var = col_double(),
      n = col_integer(),
      nm = col_integer(),
      N = col_double(),
      NM = col_double(),
      protected_status = col_character()
      ))

domain_dt_abun_no_spp = domain_dt_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #370 species
```

```{r domain biomass}

domain_dt_bio <- read_csv('big_csv/biomass_domain/domain_dt_biomass.csv',
    col_types = cols(
      protected_status = col_character()
      ))

domain_dt_bio_no_spp = domain_dt_bio %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #370 species

```

```{r domain biodiversity}
domain_dt_abun <-
  read_csv('big_csv/abundance_domain/domain_dt_abun.csv',
    col_types = cols(
      YEAR = col_integer(),
      REGION = col_character(),
      SPECIES_CD = col_character(),
      abundance = col_double(),
      var = col_double(),
      n = col_integer(),
      nm = col_integer(),
      N = col_double(),
      NM = col_double(),
      protected_status = col_character()
      ))

domain_dt_abun_no_spp = domain_dt_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #370 species

#protected status: 0 - not protected; 1 - protected; all - both 

dt_domain_abun_diversity_no_spp = domain_dt_abun_no_spp %>% #12,140 * 10
  select(YEAR, protected_status, SPECIES_CD, abundance) %>% #12,140 * 4 
  group_by(YEAR, protected_status) %>% ##36 groups (fk 9*4)
  nest(-YEAR) %>%  
  mutate(
    data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill =0)), 
    dom_richness = map( #species richness 
      data_wide, #why are there 116 rows??? 
      function(x) specnumber(x)),
    dom_simpson = map( #simpson diversity as effective number of species 
      data_wide,
      function(x) 1/(1 - diversity(x, index = 'simpson'))),
    dom_shannon = map( #shannon diversity as effective number of species 
      data_wide,
      function(x) exp(diversity(x, index = 'shannon')))) %>%
  unnest(dom_richness, dom_simpson, dom_shannon)

domain_dt_abun_diversity_no_spp = dt_domain_abun_diversity_no_spp %>%
  select(YEAR, protected_status, dom_richness, dom_simpson, dom_shannon) 
write_csv(domain_dt_abun_diversity_no_spp, 'big_csv/abundance_domain/domain_dt_abun_diversity_no_spp.csv')

```

```{r domain biodiversity plots}

domain_dt_abun_diversity = read_csv("big_csv/abundance_domain/domain_dt_abun_diversity.csv")

dt_protected_status_names <- c(
  "0" = "Unprotected",
  "1" = "DRTO",
  "2" = "TNER",
  "all" = "All"
  )

dt = domain_dt_abun_diversity %>% 
  filter(protected_status != 0) %>% 
  filter(protected_status != 1) %>% 
  filter(protected_status != 2) 

## DT richness, shannon, and simpson for all 
dt_rich_simp_shannon = dt %>%
  select(YEAR, dom_simpson, dom_shannon, dom_richness)

dygraph(dt_rich_simp_shannon, main = "Dry Tortugas Reef Fish Biodiversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 350)) %>% 
  dyAxis("x", label = "Year") %>%
  dySeries("dom_richness", label = "Richness", color = "purple") %>%
  dySeries("dom_shannon", label = "Shannon", color = "blue") %>% 
  dySeries("dom_simpson", label = "Simpson", color = "red") %>% 
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 1.5) %>% #drawGrid = F
  dyLegend(width = 400)

## Dry Tortugas Shannon and Simpson for all 
dt_simp_shannon = dt %>%
  select(YEAR, dom_simpson, dom_shannon)

dygraph(dt_simp_shannon, main = "Dry Tortugas Reef Fish Biodiversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 70)) %>%
  dyAxis("x", label = "Year") %>%
  dySeries("dom_shannon", label = "Shannon Diversity", color = "blue") %>% 
  dySeries("dom_simpson", label = "Simpson Diversity", color = "red") %>% 
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 1.5) %>% #drawGrid = F
  dyLegend(width = 400)

#richness, simpson, shannon for domain in Dry Tortugas   
ggplot(data = dt_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_richness, color = "dom_richness")) +
  geom_point(aes(y = dom_simpson, color = "dom_simpson")) +
  geom_point(aes(y = dom_shannon, color = "dom_shannon")) +
  geom_line(aes(y = dom_richness, color = "dom_richness")) +
  geom_line(aes(y = dom_simpson, color = "dom_simpson")) +
  geom_line(aes(y = dom_shannon, color = "dom_shannon")) +
  facet_grid(protected_status ~ ., labeller = as_labeller(dt_protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Diversity indices") +
  scale_colour_manual(values = c("darkturquoise", "magenta", "blue"), labels=c("Species Richness", "Shannon Diversity", "Simpson Diversity")) +
  ggtitle("Reef Fish Biodiversity in Dry Tortugas") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(breaks = c(1999:2016))+
  scale_y_continuous(limits= c(0, 300), breaks = c(0,25,50,75,100,125,150,175,200,225,250,275,300))


#simpson, shannon by protected status for domain FKNMS
ggplot(data = dt_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_simpson, color = "dom_simpson")) +
  geom_point(aes(y = dom_shannon, color = "dom_shannon")) +
  geom_line(aes(y = dom_simpson, color = "dom_simpson")) +
  geom_line(aes(y = dom_shannon, color = "dom_shannon")) +
  facet_grid(protected_status ~ ., labeller = as_labeller(dt_protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Diversity index") +
  scale_colour_manual(values = c("magenta", "blue"), labels=c("Shannon diversity", "Simpson Diversity")) +
  ggtitle("Reef Fish Biodiversity in Dry Tortugas") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 35), breaks = c(5,10,15,20,25,30,35))

#species richness by protected status for domain FKNMS
ggplot(data = dt_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_richness, color = "dom_richness"),color = "darkturquoise") +
  geom_line(aes(y = dom_richness, color = "dom_richness"), color = "darkturquoise") +
  facet_grid(protected_status ~ ., labeller = as_labeller(dt_protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Level of protection") +
  ggtitle("Species Richness of Reef Fish in Dry Tortugas") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 250), breaks = c(0,50, 100, 150, 200, 250, 300))
## wonky 

#shannon by protected status for domain FKNMS
ggplot(data = dt_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_shannon, color = "dom_shannon"),color = "magenta") +
  geom_line(aes(y = dom_shannon, color = "dom_shannon"), color = "magenta") +
  facet_grid(protected_status ~ ., labeller = as_labeller(dt_protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Level of protection") +
  #scale_colour_manual(labels=c("Unprotected", "Protected", "All")) +
  ggtitle("Shannon Diveristy of Reef Fish in Dry Tortugas") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 50), breaks = c(5,10,15,20,25,30,35,40,45,50))

#simpson by protected status for domain FKNMS
ggplot(data = dt_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_simpson, color = "dom_simpson"),color = "blue") +
  geom_line(aes(y = dom_simpson, color = "dom_simpson"), color = "blue") +
  facet_grid(protected_status ~ ., labeller = as_labeller(dt_protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Level of protection") +
  #scale_colour_manual(labels=c("Unprotected", "Protected", "All")) +
  ggtitle("Simpson Diveristy of Reef Fish in Dry Tortugas") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 25), breaks = c(5,10,15,20,25))

```
### Dry Tortugas strata's:
1. CONT_LR (continuous low relief)
2. CONT_MR (continuous mid relief)
3. CONT_HR (continuous high relief)
4. ISOL_LR (isolated low relief)
5. ISOL_MR (isolated mid relief)
6. ISOL_HR (isolated high relief)
7. SPGR_LR (spur and groove low relief)
8. SPGR_HR (spur and groove high relief)

```{r strata abundance}
## DT's abundance by strata 
strata_dt_abun <- read_csv('big_csv/abundance_strata/strata_dt_abun.csv')
# Parsed with column specification:
# cols(
#   YEAR = col_integer(),
#   REGION = col_character(),
#   STRAT = col_character(),
#   PROT = col_integer(),
#   SPECIES_CD = col_character(),
#   abundance = col_double(),
#   var = col_double(),
#   n = col_integer(),
#   nm = col_integer(),
#   N = col_double(),
#   NM = col_double()
# )

strata_dt_abun_no_spp = strata_dt_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #370 species
```

```{r strata biomass}

strata_dt_bio <- read_csv('big_csv/biomass_strata/strata_dt_biomass.csv')
# Parsed with column specification:
# cols(
#   YEAR = col_integer(),
#   REGION = col_character(),
#   STRAT = col_character(),
#   PROT = col_integer(),
#   SPECIES_CD = col_character(),
#   biomass = col_double(),
#   var = col_double(),
#   n = col_integer(),
#   nm = col_integer(),
#   N = col_integer(),
#   NM = col_double()
# )
strata_dt_bio_no_spp = strata_dt_bio %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #370 species
```

```{r strata biodiversity, eval = F}

## DT's abundance by strata 
strata_dt_abun <- read_csv('big_csv/abundance_strata/strata_dt_abun.csv')
# Parsed with column specification:
# cols(
#   YEAR = col_integer(),
#   REGION = col_character(),
#   STRAT = col_character(),
#   PROT = col_integer(),
#   SPECIES_CD = col_character(),
#   abundance = col_double(),
#   var = col_double(),
#   n = col_integer(),
#   nm = col_integer(),
#   N = col_double(),
#   NM = col_double()
# )

strata_dt_abun_no_spp = strata_dt_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #370 species

dt_strata_abun_diversity_no_spp = strata_dt_abun_no_spp %>% #133,762 * 12
  group_by(YEAR, STRAT, PROT) %>% #150 groups 
  nest(-YEAR) %>%
  mutate(
    data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill =0)), 
    richness = map( #species richness
      data_wide, 
      function(x) specnumber(x %>% select(-REGION, -var, -n, -nm, -N, -NM))),
    simpson = map( #simpson diversity as effective number of species 
      data_wide,
      function(x) 1/(1 - diversity(x %>% select(-REGION, -var, -n, -nm, -N, -NM), index = 'simpson'))),
    shannon = map( #shannon diversity as effective number of species 
      data_wide,
      function(x) exp(diversity(x %>% select(-REGION, -var, -n, -nm, -N, -NM), index = 'shannon')))) %>%
  unnest(richness, simpson, shannon)

strata_dt_abun_diversity_no_spp = dt_strata_abun_diversity_no_spp %>%
  select(YEAR, STRAT, PROT, richness, simpson, shannon)
write_csv(strata_dt_abun_diversity_no_spp, "big_csv/abundance_strata/strata_dt_abun_diversity_no_spp.csv")

```

```{r strata biodiversity plots}

strata_dt_abun_diversity = read_csv("big_csv/abundance_strata/strata_dt_abun_diversity_no_spp.csv")

dt_strata_names = c(
"CONT_LR" = "CONT_LR",#continuous low relief,
"CONT_MR" = "CONT_MR", #continuous mid relief,
"CONT_HR" = "CONT_HR", #continuous high relief,
"ISOL_LR" = "ISOL_LR", #isolated low relief,
"ISOL_MR" = "ISOL_MR", #isolated mid relief,
"ISOL_HR" = "ISOL_HR", #isolated high relief,
"SPGR_LR" = "SPGR_LR", #spur and groove low relief,
"SPGR_HR" = "SPGR_HR") #spur and groove high relief)

dt_protected_status <- c(
  "0" = "Unprotected",
  "1" = "DRTO",
  "2" = "TNER",
  "all" = "All"
  )

dt_strata_ab_shannon <- ggplot(data = strata_dt_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = shannon, color = "shannon"), color = "magenta") +
  geom_line(aes(y = shannon, color = "shannon"), color = "magenta") +
  facet_grid(STRAT ~ PROT, 
             labeller = labeller(.rows = dt_strata_names, .cols = dt_protected_status)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Level of protection") +
  ggtitle("Shannon Diveristy of Reef Fish in the Florida Keys by Strata") + 
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 40), breaks = c(5,10,15,20,25,30,35,40))

```

### Dry Tortugas PSU 

```{r psu abundance}
dt_psu_abun <- read_csv("big_csv/abundance_psu/psu_dt_abun.csv")
# Parsed with column specification:
# cols(
#   YEAR = col_integer(),
#   REGION = col_character(),
#   STRAT = col_character(),
#   PROT = col_integer(),
#   PRIMARY_SAMPLE_UNIT = col_character(),
#   SPECIES_CD = col_character(),
#   m = col_integer(),
#   var = col_double(),
#   abundance = col_double()
# )

dt_psu_abun_no_spp = domain_dt_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #370 species 

```

```{r psu biomass}

psu_dt_bio <- read_csv('big_csv/biomass_psu/psu_dt_biomass.csv')

psu_dt_bio_no_spp = psu_dt_bio %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #370 species
```

```{r psu biodiversity, eval = F}

dt_psu_abun <- read_csv("big_csv/abundance_psu/psu_dt_abun.csv")
# Parsed with column specification:
# cols(
#   YEAR = col_integer(),
#   REGION = col_character(),
#   STRAT = col_character(),
#   PROT = col_integer(),
#   PRIMARY_SAMPLE_UNIT = col_character(),
#   SPECIES_CD = col_character(),
#   m = col_integer(),
#   var = col_double(),
#   abundance = col_double()
# )

dt_psu_abun_no_spp = domain_dt_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) 

dt_psu_abun_diversity_no_spp = dt_psu_abun_no_spp %>% #934,402 * 9 
  select(YEAR, STRAT, PRIMARY_SAMPLE_UNIT, PROT, SPECIES_CD, abundance) %>% 
  group_by(YEAR, PRIMARY_SAMPLE_UNIT)%>% #2,702 groups 
  nest(-YEAR) %>% #
  mutate(
    data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill =0)), 
    psu_richness = map( #species richness
      data_wide, 
      function(x) specnumber(x %>% select(-STRAT, -PROT))),
    psu_simpson = map( #simpson diversity as effective number of species 
      data_wide,
      function(x) 1/(1 - diversity(x %>% select(-STRAT, -PROT), index = 'simpson'))),
    psu_shannon = map( #shannon diversity as effective number of species 
      data_wide,
      function(x) exp(diversity(x %>% select(-STRAT, - PROT), index = 'shannon')))) %>% 
  unnest(psu_richness,psu_simpson, psu_shannon)
 
psu_dt_abun_diversity_no_spp = dt_psu_abun_diversity_no_spp %>%
  select(YEAR, PRIMARY_SAMPLE_UNIT, psu_richness, psu_simpson, psu_shannon)

write_csv(psu_dt_abun_diversity_no_spp, "big_csv/abundance_psu/psu_dt_abun_diversity_no_spp.csv") 

```

```{r psu biodiversity plots and maps}

psu_dt_abun_diversity = read_csv("big_csv/abundance_psu/psu_dt_abun_diversity.csv")

```


## Biodiversity metrics by computed with weighted species density 

### Florida Keys domain
```{r FK domain biodiversity with density}

domain_fk_density <- read_csv("big_csv/density_domain/domain_fk_density.csv", 
col_types = cols(
  YEAR = col_integer(),
  REGION = col_character(),
  SPECIES_CD = col_character(),
  density = col_double(),
  var = col_double(),
  n = col_integer(),
  nm = col_integer(),
  N = col_double(),
  NM = col_double(),
  protected_status = col_character()
))

## FK's domain density 
domain_fk_density = domain_fk_density %>% #15,771 * 10
  select(YEAR, protected_status, SPECIES_CD, density) %>% #27,911 * 4 
  group_by(YEAR, protected_status) %>% ##45 groups (fk 15*3)
  nest(-YEAR) %>%  
  mutate(
    data_wide = map(data, ~ spread(data=.x, SPECIES_CD, density, fill =0)), 
    dom_richness = map( #species richness 
      data_wide, #why are there 116 rows??? 
      function(x) specnumber(x)),
    dom_simpson = map( #simpson diversity as effective number of species 
      data_wide,
      function(x) 1/(1 - diversity(x, index = 'simpson'))),
    dom_shannon = map( #shannon diversity as effective number of species 
      data_wide,
      function(x) exp(diversity(x, index = 'shannon')))) %>%
  unnest(dom_richness, dom_simpson, dom_shannon)

domain_fk_density_diversity = domain_fk_density %>%
  select(YEAR, protected_status, dom_richness, dom_simpson, dom_shannon) 

write_csv(domain_fk_density_diversity, 'big_csv/density_domain/domain_fk_density_diversity.csv')
  
```

```{r FK domain biodiversity plots - density}

#simpson, shannon by protected status for domain FKNMS
ggplot(data = domain_fk_density, aes(x = YEAR)) +
  geom_point(aes(y = dom_simpson, color = "dom_simpson")) +
  geom_point(aes(y = dom_shannon, color = "dom_shannon")) +
  geom_line(aes(y = dom_simpson, color = "dom_simpson")) +
  geom_line(aes(y = dom_shannon, color = "dom_shannon")) +
  facet_grid(protected_status ~ ., labeller = as_labeller(protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Diversity index") +
  scale_colour_manual(values = c("magenta", "blue"), labels=c("Shannon diversity", "Simpson Diversity")) +
  ggtitle("Reef Fish Biodiversity in the Florida Keys") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 50), breaks = c(5,10,15,20,25,30,35,40,45,50))

```

### Dry Tortugas domain
```{r DT domain biodiversity with density, eval = F}

domain_dt_density <- read_csv("big_csv/density_domain/domain_dt_density.csv", 
col_types = cols(
  YEAR = col_integer(),
  REGION = col_character(),
  SPECIES_CD = col_character(),
  density = col_double(),
  var = col_double(),
  n = col_integer(),
  nm = col_integer(),
  N = col_double(),
  NM = col_double(),
  protected_status = col_character()
))

#protected status: 0 - not protected; 1 - protected; all - both 

domain_dt_density = domain_dt_density %>% #12,140 * 10
  select(YEAR, protected_status, SPECIES_CD, density()) %>% #12,140 * 4 
  group_by(YEAR, protected_status) %>% ##36 groups (fk 9*4)
  nest(-YEAR) %>%  
  mutate(
    data_wide = map(data, ~ spread(data=.x, SPECIES_CD, density, fill =0)), 
    dom_richness = map( #species richness 
      data_wide, #why are there 116 rows??? 
      function(x) specnumber(x)),
    dom_simpson = map( #simpson diversity as effective number of species 
      data_wide,
      function(x) 1/(1 - diversity(x, index = 'simpson'))),
    dom_shannon = map( #shannon diversity as effective number of species 
      data_wide,
      function(x) exp(diversity(x, index = 'shannon')))) %>%
  unnest(dom_richness, dom_simpson, dom_shannon)

domain_dt_density_diversity = domain_dt_density %>%
  select(YEAR, protected_status, dom_richness, dom_simpson, dom_shannon) 

write_csv(domain_dt_density_diversity, 'big_csv/density_domain/domain_dt_density_diversity.csv')

#TODO: why does DT have 4 options underprotected status?? 0,1,2,all 
```

```{r DT domain biodiversity plots - density}

```

### Florida Keys strata 

```{r FK strata biodiversity - density}

## FK's density by strata 
strata_fkdensity_1999_2016 <- read_csv("big_csv/density_strata/strata_fk_density.csv",
col_types = cols(
  YEAR = col_integer(),
  REGION = col_character(),
  SPECIES_CD = col_character(),
  density = col_double(),
  var = col_double(),
  n = col_integer(),
  nm = col_integer(),
  N = col_double(),
  NM = col_double(),
  protected_status = col_character()
))

fk_strata_density_diversity = strata_fkdensity_1999_2016 %>% #133,762 * 12
  filter(protected_status!='all') %>% #66,881 * 12
  select(YEAR, STRAT, protected_status, SPECIES_CD, density) %>% #66,881 * 5
  group_by(YEAR, STRAT, protected_status) %>% #294 groups (15 years *7 strata *3prot = 315)??
  nest() %>%
   mutate(
    data_wide = map(data, ~ spread(data=.x, SPECIES_CD, density, fill =0)), 
    strata_richness = map( #species richness
      data_wide, 
      function(x) specnumber(x)),
    strata_simpson = map( #simpson diversity as effective number of species 
      data_wide,
      function(x) 1/(1 - diversity(x, index = 'simpson'))),
    strata_shannon = map( #shannon diversity as effective number of species 
      data_wide,
      function(x) exp(diversity(x, index = 'shannon')))) %>%
  unnest(strata_richness, strata_simpson, strata_shannon)

```

```{r FK strata biodiversity plots - density, eval = F}

ggplot(data = fk_strata_density_diversity, aes(x = YEAR)) +
  geom_point(aes(y = strata_simpson, color = "strata_simpson"),color = "blue") +
  geom_line(aes(y = strata_simpson, color = "strata_simpson"), color = "blue") +
  facet_grid(STRAT ~ protected_status, 
             labeller = labeller(.rows = strata_names, .cols = protected_status_names)) +
             #labeller(strata_names = label_wrap_gen(3))) + 
  labs(x = "Year", y = "Effective Number of Species") +
  ggtitle("Simpson Diveristy of Reef Fish in the Florida Keys") + 
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12)) +
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 25), breaks = c(5,10,15,20,25))

```

### Dry Tortugas strata 
```{r DT strata biodiversity - density, eval = F}

## DT's density by strata 
strata_dtdensity_1999_2016 <- read_csv("big_csv/density_strata/strata_dtdensity_1999_2016.csv",
col_types = cols(
  YEAR = col_integer(),
  REGION = col_character(),
  SPECIES_CD = col_character(),
  density = col_double(),
  var = col_double(),
  n = col_integer(),
  nm = col_integer(),
  N = col_double(),
  NM = col_double(),
  protected_status = col_character()
))

dt_strata_diversity = strata_dtdensity_1999_2016 %>% #133,762 * 12
  group_by(YEAR, STRAT, protected_status) %>% #294 groups 
  nest(-YEAR) %>%
  mutate(
    data_wide = map(data, ~ spread(data=.x, SPECIES_CD, density, fill =0)), 
    strata_richness = map( #species richness
      data_wide, 
      function(x) specnumber(x %>% select(-PROT, -var, -n, -nm, -N, -NM))),
    strata_simpson = map( #simpson diversity as effective number of species 
      data_wide,
      function(x) 1/(1 - diversity(x %>% select(-PROT, -var, -n, -nm, -N, -NM), index = 'simpson'))),
    strata_shannon = map( #shannon diversity as effective number of species 
      data_wide,
      function(x) exp(diversity(x %>% select(-PROT, -var, -n, -nm, -N, -NM), index = 'shannon')))) %>%
  unnest(strata_richness, strata_simpson, strata_shannon)

strata_dt_density_diversity = dt_strata_diversity %>%
  select(YEAR, STRAT, protected_status, strata_richness, strata_simpson, strata_shannon)

write_csv(strata_dt_density_diversity, "density_strata/strata_dt_density_diversity")
```

```{r DT  strata biodiversity plots - density}

```

### Florida Keys PSU 
```{r FK psu biodiversity - density, eval = F}

psu_fkdensity_1999_2016 <- read_csv("big_csv/density_psu/psu_fkdensity_1999_2016.csv",
col_types = cols(
  YEAR = col_integer(),
  REGION = col_character(),
  SPECIES_CD = col_character(),
  density = col_double(),
  var = col_double(),
  n = col_integer(),
  nm = col_integer(),
  N = col_double(),
  NM = col_double(),
  protected_status = col_character()
))

fk_strata_diversity = strata_fkdensity_1999_2016 %>%
  group_by(YEAR, Primary_Sampling_Unit, protected_status)%>% 
  nest(-YEAR) %>%
  mutate(
    data_wide = map(data, ~ spread(data=.x, SPECIES_CD, density, fill =0)), 
    psu_richness = map( #species richness
      data_wide, 
      function(x) specnumber(x %>% select(-REGION, protected_status))),
    psu_simpson = map( #simpson diversity as effective number of species 
      data_wide,
      function(x) 1/(1 - diversity(x %>% select(-REGION, -protected_status), index = 'simpson'))),
    psu_shannon = map( #shannon diversity as effective number of species 
      data_wide,
      function(x) exp(diversity(x %>% select(-REGION, -protected_status), index = 'shannon'))))

```

```{r FK psu biodiversity plots and maps - density}

```

### Dry Tortugas PSU 
```{r DT psu biodiversity - density, eval = F}

psu_dtdensity_1999_2016 <- read_csv("big_csv/density_psu/psu_dtdensity_1999_2016.csv",
col_types = cols(
  YEAR = col_integer(),
  REGION = col_character(),
  SPECIES_CD = col_character(),
  density = col_double(),
  var = col_double(),
  n = col_integer(),
  nm = col_integer(),
  N = col_double(),
  NM = col_double(),
  protected_status = col_character()
))

dt_strata_diversity = strata_dtdensity_1999_2016 %>%
  group_by(YEAR, SPECIES_CD, density, protected_status)%>% 
  nest(-YEAR) %>%
  mutate(
      data_wide = map(data, ~ spread(data=.x, SPECIES_CD, density, fill =0)), 
    psu_richness = map( #species richness
      data_wide, 
      function(x) specnumber(x %>% select(-REGION, protected_status))),
    psu_simpson = map( #simpson diversity as effective number of species 
      data_wide,
      function(x) 1/(1 - diversity(x %>% select(-REGION, -protected_status), index = 'simpson'))),
    psu_shannon = map( #shannon diversity as effective number of species 
      data_wide,
      function(x) exp(diversity(x %>% select(-REGION, -protected_status), index = 'shannon'))))

```

```{r DT psu biodiversity plots and maps - density}

```

```{r}
## FK's domain abundance including specis to genus level (spp.)  - 377 species 
# fk_domain_abun_diversity = domain_fk_abun %>% #15,771 * 10
#   select(YEAR, protected_status, SPECIES_CD, abundance) %>% #27,911 * 4 
#   group_by(YEAR, protected_status) %>% ##45 groups (fk 15*3)
#   nest(-YEAR) %>%  
#   mutate(
#     data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill =0)), 
#     dom_richness = map( #species richness 
#       data_wide, #why are there 116 rows??? 
#       function(x) specnumber(x)),
#     dom_simpson = map( #simpson diversity as effective number of species 
#       data_wide,
#       function(x) 1/(1 - diversity(x, index = 'simpson'))),
#     dom_shannon = map( #shannon diversity as effective number of species 
#       data_wide,
#       function(x) exp(diversity(x, index = 'shannon')))) %>%
#   unnest(dom_richness, dom_simpson, dom_shannon)
# 
# fk_domain_abun_diversity = fk_domain_abun_diversity %>%
#   select(dom_richness,dom_simpson,dom_shannon)

#write_csv(fk_domain_abun_diversity, 'big_csv/abundance_domain/domain_fk_abun_diversity')
```

```{r practice codes, eval = F}
# 
# spp_list <- read_csv('spp_list.csv') #402 species 
# 
# fk1999 <- getRvcData(1999, "FLA KEYS", server = 'https://www.sefsc.noaa.gov/rvc_analysis20/') 
# 
# spp_list <- fk1999$taxonomic_data$COMNAME
# sp_list <- as.data.frame(spp_list, row.names = F)
# write_csv(sp_list, "spp_list.csv") 
# 
# trial <- getDomainAbundance(fk1999, "ocy chry", merge_protected = F ) # returns: YEAR, REGION, SPECIES_CD, density, var, n, nm, N, NM, protected_status
# 
# trial2 <- getDomainDensity(fk1999, "ocy chry", merge_protected = F ) # returns: YEAR, REGION, SPECIES_CD, density, var, n, nm, N, NM, protected_status
# 
# #getDomainDensity(fk1999, "ocy chry", is_protected = T) # returns: YEAR, REGION, SPECIES_CD, density, var, n, nm, N, NM for only protected areas 
# 
# #getDomainDensity(fk1999, "ocy chry", when_present = T)
# 
# abun1999 <- getDomainAbundance(fk1999, spp_list, merge_protected = F)
# 
# write_csv(abun1999, 'abunfk1999.csv') #save domain abundance as csv 
# 
# div_1999_tbl = abun1999 %>%
#   select(YEAR, REGION, SPECIES_CD, abundance, protected_status) %>%
#   nest(-YEAR) %>% 
#   mutate(
#     data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill =0)), 
#     dom_richness = map(
#       data_wide, 
#       function(x) specnumber(x %>% select(-REGION, protected_status))),
#     dom_simpson = map(
#       data_wide,
#       function(x) 1/(1 - diversity(x %>% select(-REGION, -protected_status), index = 'simpson'))),
#     dom_shannon = map( 
#       data_wide,
#       function(x) exp(diversity(x %>% select(-REGION, -protected_status), index = 'shannon'))))
``` 

8.2% of the abundance data is from species identified to the genus and/or family level. 
```{r percent spp}
##Percent of the abundance data is from species identified to the genus? 
# 
# domain_fk_abun <- read_csv('big_csv/abundance_domain/domain_fk_abun.csv', 
# col_types = cols(
#   YEAR = col_integer(),
#   REGION = col_character(),
#   SPECIES_CD = col_character(),
#   abundance = col_double(),
#   var = col_double(),
#   n = col_integer(),
#   nm = col_integer(),
#   N = col_double(),
#   NM = col_double(),
#   protected_status = col_character()
# ))
# 
# spp_abun = domain_fk_abun %>%
#   filter(stringr::str_detect(SPECIES_CD, 'SPE\\.$')) %>%
#   #sum(.$abundance, na.rm=T)
#   .$abundance %>% 
#   sum() #1923525258 
# 
# total_abun = sum(domain_fk_abun$abundance) #23,434,984,377 counts over 15 years in FK
# 
# percent_spp = (spp_abun/total_abun)*100 #8.207% 

```

Matched species codes in trait matrix with the species in the RVC data. Total of 348 species identified to species level. 
```{r matching species, eval = F}
## Matching species in RVC data with species in trait matrix

# domain_fk_abun_no_spp = domain_fk_abun %>%
#   filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$'))
# 
# length(unique(domain_fk_abun$SPECIES_CD)) #377 species in RVC with .SPE (29 end in SPE.)
# length(unique(domain_fk_abun_no_spp$SPECIES_CD)) #348 species in RVC without .SPE
# 
# spp_list_without_spp = domain_fk_abun_no_spp %>%
#   select(SPECIES_CD) %>%
#   arrange(SPECIES_CD) %>%
#   mutate(
#     SPECIES_CD = toupper(as.character(SPECIES_CD)))
# 
# spp_list_without_spp = unique(spp_list_without_spp) #348 species
# 
# trait_matrix = read_csv('C:/Users/Megan Hepner/Dropbox/Thesis/Functional_Diversity/species_trait_matrix.csv')
# 
# length(unique(trait_matrix$SPECIES_CD)) #371 species
# 
# trait_matrix_spp = trait_matrix %>%
#   as.tibble() %>%
#   mutate(
#     SPECIES_CD = toupper(as.character(SPECIES_CD))) %>%
#   select(SPECIES_CD) #371 species
# 
# missing_species = anti_join(trait_matrix_spp, spp_list_without_spp, by = "SPECIES_CD") #species in trait_matrix not in RVC species list #23
# 
# trait_matrix_348_spp = read_csv('C:/Users/Megan Hepner/Dropbox/Thesis/Functional_Diversity/species_trait_matrix_348_spp.csv')
# 
# length(unique(trait_matrix_348_spp$SPECIES_CD)) #348 species
# 
# trait_matrix_348_spp = trait_matrix_348_spp %>%
#   as.tibble() %>%
#   mutate(
#     SPECIES_CD = toupper(as.character(SPECIES_CD))) #%>%
# select(SPECIES_CD) #348 species
# 
# missing_species_2 = semi_join(trait_matrix_spp_348_spp, spp_list_without_spp, by = "SPECIES_CD") #perfect match :)

```
