---
title: "iNEXT"
author: "Megan Hepner"
date: "6/11/2017"
output: html_document
---
## example for abundance based data (list of vector)
data(spider)
out1 <- iNEXT(spider, q=0, datatype="abundance")
out1$DataInfo # showing basic data information.
out1$iNextEst # showing diversity estimates with rarefied and extrapolated and related statistics.
out1$AsyEst # showing asymptotic diversity estimates with related statistics.


$DataInfo: 
n - returns basic data information including the reference sample size 
S.obs - observed species richness 
SC - a sample coverage estimate 
f1-f10 - the first ten frequency counts

$iNextEst: 
m - the sample size 
method - the method (interpolated, observed, or extrapolated, depending on whether the size m is less than, equal to, or greater than the reference sample size
order - the diversity order
qD - the diversity estimate of order q
qD.LCL, qD.UCL- the 95% lower and upper confidence limits of diversity 
SC - the sample coverage estimate
SC.LCL, SC.UC - the 95% lower and upper confidence limits of sample coverage.

$AsyEst: asymptotic diversity estimates along with related statistics.
Observed - observed diversity, 
Estimator - asymptotic estimates, 
EST_s.e. - estimated bootstrap s.e. 
Lower_CI, Upper_CI - 95% confidence intervals 

```{r psu inext}
library(iNEXT)
library(tidyverse)
library(data.table)

psu_fk_abun <- read_csv("big_csv/abundance_psu/psu_fk_abun.csv",
                        col_types = cols(
                          YEAR = col_integer(),
                          REGION = col_character(),
                          STRAT = col_character(),
                          PROT = col_integer(),
                          PRIMARY_SAMPLE_UNIT = col_character(),
                          SPECIES_CD = col_character(),
                          m = col_integer(),
                          var = col_double(),
                          abundance = col_double(),
                          protected_status = col_character()))

psu_fk_abun_no_spp = psu_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species 

trait_matrix = read_csv('functional_diversity/species_trait_matrix_348_spp.csv')

trait_matrix = trait_matrix %>%
   as.tibble() %>%
   mutate(
     SPECIES_CD = toupper(as.character(SPECIES_CD))) %>%
   arrange(SPECIES_CD)

 fk_psu_abun_diversity_no_spp_merged = psu_fk_abun_no_spp %>% #3,107,408 x 10
   select(YEAR, PRIMARY_SAMPLE_UNIT, STRAT, protected_status, SPECIES_CD, abundance) %>% #3,107,408*5
   filter(protected_status == 'all') %>% #1,553,704*10
   group_by(YEAR, STRAT, PRIMARY_SAMPLE_UNIT, SPECIES_CD) %>% #1,553,704 groups
   summarise(
     abun_merged = sum(abundance))

idk = psu_fk_abun_no_spp %>% #3,107,408 x 10
  group_by(STRAT, SPECIES_CD) %>% #2,436*3 groups 
  mutate(abundance = as.integer(abundance)) %>% 
  summarise(abun = sum(abundance)) %>% 
  #filter(abun != 0) %>% #1,417*3
  spread(SPECIES_CD, abun) %>% #7*293
  t() 
  
colnames(idk) = idk[1,] #make the first row "STRAT" as the header
idk = idk[-1,] #remove the first row "year"
spec_by_strata = as.data.frame(idk)


iNEXT(spec_by_strata, q = 0, datatype = "abundance")
# Error in FUN(newX[, i], ...) : invalid data structure
# In addition: Warning messages:
# 1: In EstiBootComm.Ind(Spec) :
#   This site has only one species. Estimation is not robust.

#species_by_strata = iNEXT(idk, q=c(0,1,2), datatype = "abundance")

```

```{r iNEXT example}
data(spider)
out1 <- iNEXT(spider, q=0, datatype="abundance")

data(bird)
str(bird)
out2 <- iNEXT(bird, q=0, datatype="abundance")
ggiNEXT(out2)

```


