---
title: "Florida Keys Reef Fish Biodiversity"
author: "Megan Hepner"
date: "April 28, 2017"
output: html_document
---

## Inputs 
1. species-trait matrix
1. domain abundance dataframe
1. strata abundance dataframe 
1. psu abundance dataframe
1. domain biomass dataframe
1. strata biomass dataframe 
1. psu biomass dataframe

## Outputs 
1.  Functional dendogram
+ Trait-matrix visualization 

1.  By *domain* for each sampling year and region  
+ Species weighted abundance and plot
+ Species weighted biomass and plot
+ Species biodiversity 
+ species richness and plot
+ Simpson diversity as effective number of species and plot
+ Shannon diversity as effective number of species and plot
+ Functional diversity as effective number of species and plot

1. By *upper keys and lower keys domain* for each sampling year and region 
+ Species weighted abundance and plot
+ Species weighted biomass and plot
+ Species biodiversity 
+ species richness and plot
+ Simpson diversity as effective number of species and plot
+ Shannon diversity as effective number of species and plot
+ Functional diversity as effective number of species and plot

1. By *strata* for each sampling year and region 
+ Species weighted abundance and plot
+ Species weighted biomass and plot
+ Species biodiversity 
+ species richness and plot
+ Simpson diversity as effective number of species and plot
+ Shannon diversity as effective number of species and plot
+ Functional diversity as effective number of species and plot

1. By *primary sampling unit* for each sampling year and region and plot
+ Species weighted abundance and plot
+ Species weighted biomass and plot
+ Species biodiversity 
+ species richness and plot
+ Simpson diversity as effective number of species and plot
+ Shannon diversity as effective number of species and plot
+ Functional diversity as effective number of species and plot
1. Bivariate Relationships 
+ correlation with spearman rank 

1. Parametric tets
+ statistical tests by strata and year: Shapiro, homoscedacity, 2-way ANOVA

1. Generalized Additive Model
+ temporal variable: YEAR
+ spatial variable: Latitude, longitude
+ environmental variable: STRAT, habitat_cd, depth, bottom temperature 

1.Diversity heat Map
+ Kriging  

## Reef Visual Census Background 

Reef fish community data used in this study were collected as part of the multi-agency Reef Visual Census (RVC) [Brandt et al., 2009](https://www.coris.noaa.gov/activities/fish_monitoring_protocol/). Fish communities were visually surveyed underwater annually in the mainland FKNMS (Lower Keys, Middle Keys and Upper Keys) from 1999 to 2012, and biennially from 2012 to 2016. The Dry Tortugas region was surveyed from 1999 to 2000 and biennially from 2004 to 2016.     

The RVC uses a stationary point count method within a randomly selected 7.5 m radius circular plot with a 200m map grid from 1999 to 2012 and 100m map grid from 2014 to 2016 to optimize the observation of conspicuous and diurnally active reef fish, specifically economically and ecologically important reef fish species [Bohnsack and Bannerot, 1986](https://docs.lib.noaa.gov/noaa_documents/NMFS/TR_NMFS/TR_NMFS_41.pdf). 

## Objective 

To analyze and compare changes in species richness, Simpson diversity, Shannon diversity and functional diversity in the Florida Keys National Marine Sanctuary (FKNMS) to reveal changes in temporal and spatial variability from 1999 – 2016. The indices were assessed throughout the Florida Keys, between the upper and lower keys, between habitat types, and for different levels of management (Ecological Reserves (ER), Sanctuary Preservations Areas (SPA), and Special Use/Research Only Areas (SU/RO)). 

## Diversity indicse as effective number of species 

All indices were computed as effective number of species. Effective number of species is the number of equally abundance species needed to produce the observed value of a diversity index (Jost, 2006; Jost et al., 2010; MacArthur, 1965). Jost, 2006 refers to these as true measures of diversity, where prior to the conversion the diversity indices are simply measures of entropy. 

- Species Richness is the number of species in a habitat sampled  

$Richness = \sum_{i=1}^S p_i^0$  

- Simpson diversity measures the probability that two individuals randomly selected from a sample will belong to different species.   

$Simpson = 1 - \sum_{i=1}^S p_i^2$  

- Shannon diveristy is a measure of entropy, it is the uncertainty in the species identity of a sample  

$Shannon = - \sum_{i=1}^S p_i \log_b p_i$  

- Functional diversity (Rao's Q) is a measure of pairwise functional differences between species weighted by their relative abundances   

$Rao Q = \sum_{i=1}^S-1 \sum_{j=i+1}^S d_i p_i p_j$

```{r setup, include= F}
knitr::opts_chunk$set(echo = T)
#rm(list = ls())

#install.packages('devtools')
#devtools::install_github('jeremiaheb/rvc')
library(rvc) #calls RVC data from SEFSC server 
#install.packages("tidyverse")
library(tidyverse)
#install.packages("vegan")
library(vegan) # biodiveristy functions for richness, simpson, and shannon 
#install.packages("dygraphs")
library(dygraphs)
#install.packages("plotly")
#library(plotly)
#install.packages("webshot")
library(webshot)
#install.packages("htmlwidgets")
library(htmlwidgets)
#install.packages("FD")
library(FD)
#install.packages("iNEXT")
library(iNEXT) 
RVCdata_FK = read_rds('big_csv/RVCdata_FK.rds')

map = purrr::map # override maps::map
select = dplyr::select #override MASS::select 
group_by =  dplyr::group_by #override plotly::group_by
summarise = dplyr::summarise #override plotly::summarise
```

## Functional distance trait based matrix (from Lefcheck et al.,)

Gowdis computes the Gower (1971) similarity coefficient exactly as described by Podani (1999), then converts it to a dissimilarity coefficient by using D = 1 - S

```{r Functional distance matrix, eval = F}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("clue")
library(clue) #calls cl_ultrametric 
#install.packages("fpc")
library(fpc) #calls pamk
#install.packages("ape")
library(ape)
#install.packages("phytools")
library(phytools)
#install.packages("FD")
library(FD)
#biocLite("BiocUpgrade") ## you may need this
#biocLite("ggtree", type = "source")
library(tidyverse)

trait_matrix = read_csv('functional_diversity/species_trait_matrix_348_spp.csv')

# Parsed with column specification:
# cols(
#   Species_code = col_integer(),
#   SPECIES_CD = col_character(),
#   Latin_name = col_character(),
#   Common_name = col_character(),
#   Maxlength = col_double(),
#   Trophic_level = col_double(),
#   Trophic_group = col_character(),
#   Water_column = col_character(),
#   Diel_activity = col_character(),
#   Substrate_type = col_character(),
#   Complexity = col_character(),
#   Gregariousness = col_integer()

trait_matrix = trait_matrix %>%
  as.tibble() %>%
  mutate(
    SPECIES_CD = toupper(as.character(SPECIES_CD))) %>% 
  arrange(SPECIES_CD)%>% 
  select(SPECIES_CD, Maxlength, Trophic_level, Trophic_group, Water_column, Diel_activity, Substrate_type, Complexity, Gregariousness) %>% #348*9
  group_by(
    SPECIES_CD, Maxlength, Trophic_level, Trophic_group, Water_column, Diel_activity, Substrate_type, Complexity, Gregariousness) %>%
  #summarize(n = n()) %>%
  #select(-n) %>%
  ungroup() %>%
  mutate(
    # ordinal traits
    Complexity = ordered(Complexity, levels=c("Low","Medium","High")),
    Gregariousness = ordered(Gregariousness, levels=c("1","2","3"))) %>%
  as.data.frame()

#Set species names as row.names and remove extra columns
rownames(trait_matrix)=trait_matrix$SPECIES_CD 

#Calculate Gower distances. Variables have equal weight. 
traits.dist = gowdis(trait_matrix, ord="podani")

#Use clustering method to produce ultrametric dendrogram because values of Rao's Q can be maximized when fewer than the max number of functional types are present unless distances are ultramtetric 

#to account for sensitivity in clustering use multiple algorithms  (Mouchet et al., 2008) 
tree_methods = c("single","complete","average","mcquitty","ward.D") #average is best clustering method 
trees=lapply(tree_methods,  function(i) hclust(traits.dist, method=i))
par(mfrow=c(3,2))
for(i in 1:length(trees)) {plot(trees[[i]])} #plots the 5 different kinds of trees 

#convert trees to ultrametric
trees.ultra= lapply(trees,function(i) cl_ultrametric(as.hclust(i)))

#Plot each tree
par(mfrow=c(3,2))
for (i in 1:length(trees.ultra)) {plot(trees.ultra[[i]])} #plots the 5 different kinds of trees 

#Build the consensus tree (Mouchet et al 2008 Oikos)  
ensemble.trees=cl_ensemble(list=trees) #list of clusterings 
class(ensemble.trees)
consensus.tree=cl_consensus(ensemble.trees) #synthesizes the information in the elements of a cluster ensemble into a single clustering 
par(mar=c(1,1,1,1))
plot(consensus.tree, horiz=T)

#Calculate dissimilarity values for each tree using 2-norm (Merigot et al 2010 Ecology) to determine which tree best preserves orignial distances 
#spectral norm (2-norm) of the differences of the ultrametrics
all.trees=c(trees.ultra,consensus.tree[1])
names(all.trees)=c(tree_methods,"consensus")
(trees.dissim=lapply(all.trees,function(i) cl_dissimilarity(i,traits.dist,method="spectral"))) 
#Cross-dissimilarities using spectral ultrametric distance:
#Single = 63.17399
#complete = 116.8516
#average = 14.99288
#mcquitty = 36.06599
#ward.D = 3161.581
#consensus =  1065.058 (1067.782) (1091.628)

#Identify best tree and isolate
trees.dissim2=do.call(rbind,trees.dissim)
min.tree=which.min(trees.dissim2) #average 
names(all.trees)[min.tree]
func.dist=all.trees[names(all.trees)==names(all.trees)[min.tree]][[1]]

#Confirm lowest 2-norm value,  spectral norm (2-norm) of the differences if the ultrametrics 
cl_dissimilarity(func.dist,traits.dist,method="spectral") #Cross-dissimilarities using spectral ultrametric distance:  14.99288

#Scale by the max value so that all values are between 0-1
func.dist=func.dist/max(func.dist)

func.dist_common=func.dist/max(func.dist)

#Plot the best tree - average 
pdf("functional_dendrogram.pdf", width = 30, height = 70)
par(mfrow=c(1,1))         #nc-by-nr array 
par(mar=c(2,1,0,2))       #number of lines of margins for bottom, left, top, right 
par(pin=c(25,58))         #plot dimensions 
plot(func.dist, horiz=TRUE, main = "Functional Dendrogram", cex = 1, cex.main = 2, cex.lab = 0.5)
dev.off()

#Phylogenetic tree with 348 tips and 347 internal nodes. 
func.dist.phy = as.phylo(as.hclust(func.dist))

#Write newick tree
write.tree(func.dist.phy, "functional_diversity/functional_dendrogram.nwk")
#write.nexus(func.dist.phy, "functional_diversity/functional_dendrogram.nex") 

tree <- read.tree("functional_diversity/functional_dendrogram.nwk")
# List of 4
#  $ edge       : int [1:694, 1:2] 
#  $ Nnode      : int 347
#  $ tip.label  : chr [1:348] 
#  $ edge.length: num
# - attr(*, "class")= chr "phylo"
# - attr(*, "order")= chr "cladewise"

# Phylogenetic tree with 348 tips and 347 internal nodes.
# Tip labels:
# 	KYP_SECT, ALE_CILI, ELO_SAUR, HAR_JAGU, CEN_UNDE, TYL_CROC, ...
# Rooted; includes branch lengths.

plot(tree, label.offset=0.2, edge.width = 2, type = "cladogram")  
nodelabels() #add node numbers
tiplabels() #add tip numbers

ggtree(tree, branch.length = T) +
  ggtitle("Functional Dendogram") +
  geom_tiplab()+
  geom_text(aes(x = branch, label=branch.length), vjust = -.5) +
  theme_tree2()

ggtree(tree, layout= "circular") +
  ggtitle("Functional Dendogram") +
  geom_tiplab(size = 1, aes(angle=angle))+
  theme_tree2()

# transformed to a tidy data.frame by fortify method
tree_data <- fortify(tree) #node; parent; branch.length; x; y; label; isTip; branch; angle 
head(tree_data)

#Visualize species' differences in multivariate trait space
#Perform k-means clustering with no a priori specification for k
traits.kclus= pamk(traits.dist, krange=2:10) 

#krange: integer vector. Numbers of clusters which are to be compared by the average silhouette width criterion.
#traits.kclus$nc = 10 
#traits.kclus$crit 0.0000000 0.2374706 0.2362513 0.2355555 0.2681756 0.2581352 0.2759547 0.2852216 0.3006077 0.3062123

# #Perform multidimensional scaling on functional dendrogram
traits_nmds = metaMDS(traits.dist,k=traits.kclus$nc,trymax=20)
#k = Number of dimensions

#*** No convergence -- monoMDS stopping criteria:
#499: no. of iterations >= maxit
#1: stress ratio > sratmax

# Dimensions: 10 
# Stress:     0.03545178 
# Stress type 1, weak ties
# No convergent solutions - best solution after 500 tries
# Scaling: centring, PC rotation 
# Species: scores missing


# #Plot in two dimensions - doesn't work 
par(mar=c(4,4,1,1))
ordiplot(traits_nmds,type="n")

#Assign colors to different groups
groups=levels(factor(traits.kclus$pamobject$clustering))
points.symbols=15:16
points.colors=c("firebrick3","cornflowerblue")
for(i in seq_along(groups)) {
  points(traits_nmds$points[traits.kclus$pamobject$clustering==groups[i],],
         pch=points.symbols[i],col=points.colors[i],cex=1.4) }
ordispider(traits_nmds,factor(traits.kclus$pamobject$clustering),label=F)
ordihull(traits_nmds,factor(traits.kclus$pamobject$clustering),lty="dotted")
orditorp(traits_nmds,dis="sites",pcex=0,air=0.5,col="grey10",cex=0.8)

```

## Trait matrix visualization 

[http://jkunst.com/r/pokemon-visualize-em-all/]
[https://d3js.org/]
[http://www.buildingwidgets.com/blog/2015/7/22/week-29-d3treer-v2]

```{r trait visualization}

library(tidyverse)
library(treemap)
devtools::install_github("gluc/data.tree")
devtools::install_github("timelyportfolio/d3treeR")
library(d3treeR)
devtools::install_github("GuangchuangYu/ggtree") #requires R>= 3.3.2
library(ggtree)

trait_matrix = read_csv('functional_diversity/species_trait_matrix_348_spp.csv')

trait_matrix_treemap = trait_matrix %>%
  select(Common_name, Maxlength, Trophic_level, Trophic_group, Water_column,Diel_activity,Substrate_type,Complexity, Gregariousness) %>% 
  group_by(Maxlength,
           Trophic_level,
           Trophic_group,
           Water_column,
           Diel_activity,
           Substrate_type,
           Complexity,
           Gregariousness) %>%
  summarise(n = n())%>% 
  ungroup() %>%
  as.data.frame()

treemap2 = 
  d3treeR::d3tree2(
    treemap(
      dtf = trait_matrix_treemap,
      index = c("Trophic_group", "Trophic_level"), #, "Water_column", "Complexity", "Substrate_type", "Diel_activity"),
      vSize = "n", 
      vColor = "Trophic_group",
      type = "index"),
    #title = "Reef Fish Traits",
    #fontsize.title = 14, 
    #algorithm = "squarified"
    rootname = "Species_Traits")

#treemap example 
data(GNI2014)
d3tree2(
  treemap(
    GNI2014
    ,index=c("continent", "iso3")
    ,vSize="population"
    ,vColor="GNI"
    ,type="value"
  )
  , rootname = "World"
)

```

```{r practice FD code}

trait_matrix = read_csv('functional_diversity/species_trait_matrix_348_spp.csv')
trait_matrix = trait_matrix %>%
  as.tibble() %>%
  mutate(
    SPECIES_CD = toupper(as.character(SPECIES_CD))) %>% 
  arrange(SPECIES_CD)

domain_fk_abun <- read_csv('big_csv/abundance_domain/domain_fk_abun.csv', 
                           col_types = cols(
                             YEAR = col_integer(),
                             REGION = col_character(),
                             SPECIES_CD = col_character(),
                             abundance = col_double(),
                             var = col_double(),
                             n = col_integer(),
                             nm = col_integer(),
                             N = col_double(),
                             NM = col_double(),
                             protected_status = col_character()
                           ))

domain_fk_abun_no_spp = domain_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species 

# DEBUG with own dbFD function for use of browser() to inspect variables before stop() or other errors
source('dbFD_debug.R')
i_dbg <<- 0
# rm(dbFD) # remove function when ready to try FD:dbFD

## FK's domain abundance only identified to species level (without spp.) 
fk_domain_abun_diversity_no_spp = domain_fk_abun_no_spp %>% #15,651 x 10
  filter(protected_status != "all") %>% 
  select(YEAR, protected_status, SPECIES_CD, abundance) %>% #10,434 * 4
  group_by(YEAR, protected_status) %>% ##32 groups (fk 16*2)
  nest(-YEAR) %>%  
  mutate(
    data_wide = map(data, function(x) 
      full_join(x, trait_matrix %>% select(SPECIES_CD), by='SPECIES_CD') %>%
        mutate(abundance = ifelse(is.na(abundance), 0, abundance)) %>%
        spread(SPECIES_CD, abundance, fill =0)), 
    # functional = map(data_wide,
    #                  function(x) 1/(1 - dbFD(as.matrix(func.dist), x))))
    spp_idx = map(data_wide, function(x) apply(x, 2, sum) > 0),
    #fk_domain_abun_diversity_no_spp$spp_idx[[1]]
    dbFD_abundance = map2(data_wide, spp_idx, function(x,y) x[,y]), 
    dbFD_traits    = map(spp_idx, function(y) as.matrix(func.dist)[y,y]),
    #test = 1/(1 - dbFD(dbFD_traits, dbFD_abundance))
    functional = map2(dbFD_traits, dbFD_abundance, 
                      function(x, a) 1/(1 - dbFD(x, a, stand.FRic=F)))) #x = distance matrix, a
                      #function(x, a) 1/(1 - divc(x, a))))

#FRic: No dimensionality reduction was required. All 176 PCoA axes were kept as 'traits'.  Called from: dbFD(x, a, stand.FRic = F)

### PACKAGE ade4 Rao's Q with divc()
fk_domain_abun_diversity_no_spp = domain_fk_abun_no_spp %>% #15,651 x 10
  filter(protected_status != "all") %>% 
  select(YEAR, protected_status, SPECIES_CD, abundance) %>% #10,434 * 4
  group_by(YEAR, protected_status) %>% ##32 groups (fk 16*2)
  nest(-YEAR) %>%  
  mutate(
    data_wide = map(data, function(x) 
      full_join(x, trait_matrix %>% select(SPECIES_CD), by='SPECIES_CD') %>%
        mutate(abundance = ifelse(is.na(abundance), 0, abundance)) %>%
        spread(SPECIES_CD, abundance, fill =0)), 
    raoQ = map(data_wide, 
                function(x) 1/(1 - divc(x, as.dist(func.dist))))) 
  #Error in mutate_impl(.data, dots) : Non convenient df

#divc(df, dis, scale) 
#df: a data frame with elements as rows, samples as columns, and abundance, presence-absence or frequencies as entries
#dis: an object of class dist containing distances or dissimilarities among elements. If dis is NULL, Gini-Simpson index is performed

```
## Fetch the RVC data from 
[NOAA's Southeast Fisheries Science Center server](https://grunt.sefsc.noaa.gov/rvc_analysis20/) using the [RVC package](https://github.com/jeremiaheb/rvc)
- Fetched weighted abundance and biomass data by domain, strata, and primary sampling units 

### Sample Data Variables
- PRIMARY_SAMPLE_UNIT: A code indicating the primary sample unit in which a sample was collected.
- YEAR: A number indicating the calendar year.
- MONTH: A number indicating the month of the year.
- DAY: A number indicating the day of the month (EST).
- STATION_NR: A number indicating the secondary sampling unit within a given primary sample unit.
- LAT_DEGREES: Latitude of secondary sampling unit in decimal degrees).
- LON_DEGREES: Longitude of secondary sampling unit in decimal degrees).
- DEPTH: Average depth, in meters, of secondary sampling unit).
- UNDERWATER_VISIBILITY: visibility, in meters, at secondary sampling unit.
- MAPGRID_NR: A number indicating the primary sample unit.  
- HABITAT_CD: A code indicating the habitat type. 
- ZONE_NR: A code indicating the distance offshore: 
+ 1 - Inshore
+ 2 - Midchannel
+ 3 - Offshore
+ 4 - Fore-reef
- SUBREGION_NR: A number indicating the subregion.
- MPA_NR: A number identifying the marine protected area in which the sample was collected. Zero indicates unprotected status)
- SPECIES_NR: A number indicating the species for a sample 
- SPECIES_CD: A code indicating the species for a sample. Consists of the first three letters of the generic name and the first four of the specific name   
- LEN: The length, in cm, of a sample.      
- NUM: The number of individuals of a given species and length observed in a secondary sampling unit
- TIME_SEEN: A number indicating when, during sampling, an individual was observed. 1: In the first five minutes, 2: From 5-10 minutes, 3: After 10 minutes. 
- PROT: A boolean value indicating whether a sample was in a protected area or not 
+ 1 - protected area
+ 0 - not protected 
- STRAT: A code indicating the stratum in which a sample was taken. Differs by region.   
+ FMLR
+ FSLR
+ HRRF
+ INPR
+ MCPR 
+ OFPR
- REGION: A code indicating the region in which a sample was taken.
+ FLA KEYS (florida keys)
+ DRY TORT (dry tortugas)

### Stratum Data Variables 
- REGION 
- YEAR 
- PROT 
- STRAT 
- NTOT: The number of possible primary sample units for a given year, region, stratum, and protected status 
- GRID_SIZE: The length (in meters) to a side of a primary sample unit for a given year, region, stratum, and protected status.

### Taxonomic Data Variables 
- SPECIES_CD
- FAMILY
- SCINAME 
- COMNAME
- LC: Minimum length at capture, in centimeters
- LM: Median length at maturity, in centimeters.
- WLEN_A: The linear coefficient of the allometric growth equation in grams per millimeter (g/mm)
- WLEN_B: The exponential coefficient of the allometric growth equation

### Benthic Data Variables 
- REGION: A code indicating the region. 
- YEAR: A number indicating the calendar year.
- PRIMARY_SAMPLE_UNIT: A code indicating the primary sample unit in which a sample was collected.
- STATION_NR: A number indicating the secondary sampling unit within a given primary sample unit.
- DEPTH: Average depth, in meters, of secondary sampling unit.
- MAX_HARD_RELIEF: The maximum height, in meters, of hard relief (e.g. hard corals, rock).
- MAX_SOFT_RELIEF: The maximum height, in meters, of soft relief (e.g. soft corals, sponges).
- AVG_HARD_RELIEF: The average height, in meters, of hard relief.
- HARD_REL_PCT_0: Percentage of hard relief less than 0.2 meters.
- HARD_REL_PCT_1: Percentage of hard relief between 0.2 and 0.5 meters.
- HARD_REL_PCT_2: Percentage of hard relief between 0.5 and 1.0 meters.
- HARD_REL_PCT_3: Percentage of hard relief between 1.0 and 1.5 meters.
- HARD_REL_PCT_4: Percentage of hard relief greater than 1.5 meters.
- PCT_SAND: Percentage of abiotic cover that is sand.
- PCT_HARD_BOTTOM: Percentage of abiotic cover that is hard bottom.
- PCT_RUBBLE: Percentage of abiotic cover that is rubble.
- PCT_CORAL: Percentage of biotic hardbottom that is coral.
- PCT_OCTO: Percentage of biotic hardbottom that is octocoral.
- PCT_SPONGE: Percentage of biotic hardbottom that is sponge.

### Function: getDomainDensity 
- YEAR
- REGION
- SPECIES_CD
- density: Domain-wide mean density for a stratified random survey  
- var: Variance in average density per secondary sampling unit
- n: Number of primary sampling units sampled
- nm: Number of secondary sampling units sampled
- N: Number of total possible primary sample units
- NM: Number of possible secondary sampling units
- length_class: The length class or bin. Only present if length_bins is not NULL. The notation, [lower, upper), is inclusive of the lower bound, but exclusive of the upper bound
- protected_status: The protected status. Only present if merge_protected is FALSE

### Florida Keys domain

```{r domain abundance}

domain_fk_abun <- read_csv('big_csv/abundance_domain/domain_fk_abun.csv', 
                           col_types = cols(
                             YEAR = col_integer(),
                             REGION = col_character(),
                             SPECIES_CD = col_character(),
                             abundance = col_double(),
                             var = col_double(),
                             n = col_integer(),
                             nm = col_integer(),
                             N = col_double(),
                             NM = col_double(),
                             protected_status = col_character()
                           ))

domain_fk_abun_no_spp = domain_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species

```

```{r domain biomass}

domain_fk_bio <- read_csv('big_csv/biomass_domain/domain_fk_biomass.csv', 
                          col_types = cols(
                            protected_status = col_character()
                          ))

domain_fk_bio_no_spp = domain_fk_bio %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species
```

```{r domain biodiversity}

trait_matrix = read_csv('functional_diversity/species_trait_matrix_348_spp.csv')
trait_matrix = trait_matrix %>%
  as.tibble() %>%
  mutate(
    SPECIES_CD = toupper(as.character(SPECIES_CD))) %>% 
  arrange(SPECIES_CD)

domain_fk_abun <- read_csv('big_csv/abundance_domain/domain_fk_abun.csv', 
                           col_types = cols(
                             YEAR = col_integer(),
                             REGION = col_character(),
                             SPECIES_CD = col_character(),
                             abundance = col_double(),
                             var = col_double(),
                             n = col_integer(),
                             nm = col_integer(),
                             N = col_double(),
                             NM = col_double(),
                             protected_status = col_character()
                           ))

domain_fk_abun_no_spp = domain_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species 

#length(unique(domain_fk_abun_no_spp$SPECIES_CD)) # 348
#length(unique(trait_matrix$SPECIES_CD))          # 348



## FK's domain abundance only identified to species level (without spp.) 
fk_domain_abun_diversity_no_spp = domain_fk_abun_no_spp %>% #14,634 * 10
  #filter(protected_status != "all") %>% 
  select(YEAR, protected_status, SPECIES_CD, abundance) %>% #14,634 * 4
  group_by(YEAR, protected_status) %>% ##45 groups (fk 15*3)
  nest(-YEAR) %>%  
  mutate(
    data_wide = map(data, function(x) 
      full_join(x, trait_matrix %>% select(SPECIES_CD), by='SPECIES_CD') %>%
        mutate(abundance = ifelse(is.na(abundance), 0, abundance)) %>%
        spread(SPECIES_CD, abundance, fill =0)), 
    #species richness
    richness=map(data_wide,   
                 function(x) specnumber(x)),
    #simpson diversity as effective number of species 
    simpson=map(data_wide, 
                function(x) 1/(1 - diversity(x, index = 'simpson'))),
    #shannon diversity as effective number of species 
    shannon=map(data_wide, 
                function(x) exp(diversity(x, index = 'shannon')))) %>%
  unnest(richness,shannon,simpson)  
  
#fk_domain_abun_diversity_no_spp = fk_domain_abun_diversity_no_spp %>%
  #select(YEAR, protected_status,richness,shannon,simpson)
  
#write_csv(fk_domain_abun_diversity_no_spp, 'big_csv/diversity_without_spe/domain_fk_abun_diversity_no_spp.csv')

```


```{r dygraphs}

domain_fk_abun_diversity = read_csv('big_csv/abundance_domain/domain_fk_abun_diversity.csv')

## Simpson for protected and unprotected stacked 
fk_simpson = domain_fk_abun_diversity %>% 
  filter(protected_status != 'all') %>%
  select(YEAR, protected_status, dom_simpson) %>%
  mutate(
    protected_status = recode(
      protected_status, '0'='Unprotected', '1'='Protected')) %>%
  spread(protected_status, dom_simpson)

dygraph(fk_simpson, main = "Simpson Reef Fish Diversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 50)) %>%
  dyAxis("x", label = "Year") %>%
  dyOptions(stackedGraph = T) 
#dyRangeSelector(height = 20) 

## Shannon for protected and unprotected stacked 
fk_shannon = domain_fk_abun_diversity %>% 
  filter(protected_status != 'all') %>%
  select(YEAR, protected_status, dom_shannon) %>%
  mutate(
    protected_status = recode(
      protected_status, '0'='Unprotected', '1'='Protected')) %>%
  spread(protected_status, dom_shannon)

dygraph(fk_shannon, main = "Shannon Reef Fish Diversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 75)) %>%
  dyAxis("x", label = "Year") %>%
  dyOptions(stackedGraph = T)
#dyRangeSelector(height = 20) 


d = domain_fk_abun_diversity %>% 
  filter(protected_status != 0) %>% 
  filter(protected_status != 1)

#Simpson for "all" filled 
dygraph_simp = d %>%
  select(YEAR, dom_simpson)

dygraph(dygraph_simp, main = "Simpson Diversity of Reef Fish in FKNMS") %>%
  dyOptions(fillGraph = T, fillAlpha = 0.4, drawPoints = T, pointSize = 2) %>%
  dySeries("dom_simpson", label = "Simpson Diversity") %>% 
  dyAxis("y", label = "Effective Number of Species") %>% 
  dyAxis("x", label = "Year")

#Shannon for "all" filled 
dygraph_shannon = d %>%
  select(YEAR, dom_shannon)

dygraph(dygraph_shannon, main = "Shannon Diversity of Reef Fish in FKNMS") %>%
  dyOptions(fillGraph = T, fillAlpha = 0.4, drawPoints = T, pointSize = 2) %>%
  dySeries("dom_shannon", label = "Shannon Diversity") %>% 
  dyAxis("y", label = "Effective Number of Species") %>% 
  dyAxis("x", label = "Year")

#Richness for "all" filled 
dygraph_richness = d %>%
  select(YEAR, dom_richness)

dygraph(dygraph_richness, main = "Species Richness of Reef Fish in FKNMS") %>%
  dyOptions(fillGraph = T, fillAlpha = 0.4, drawPoints = T, pointSize = 2) %>%
  dySeries("dom_richness", label = "Species Richness") %>% 
  dyAxis("y", label = "Number of Species") %>% 
  dyAxis("x", label = "Year")

## Simpson and Shannon stacked for "all"
dygraph_simp_shannon = d %>%
  select(YEAR, dom_simpson, dom_shannon)

dygraph(dygraph_simp_shannon, main = "Florida Keys Reef Fish Biodiversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 70)) %>%
  dyAxis("x", label = "Year") %>%
  dySeries("dom_shannon", label = "Shannon Diversity", color = "blue") %>% 
  dySeries("dom_simpson", label = "Simpson Diversity", color = "red") %>% 
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 1.5) %>% #drawGrid = F
  dyLegend(width = 400)

# m <- plot_ly(x = 1:10)
# saveWidget(as.widget(m), "temp.html")
# webshot("temp.html", file = "test.png",
#         cliprect = "viewport")

#Richness, Shannon, Simpson stacked for "all"
dygraph_rich_simp_shannon = d %>%
  select(YEAR, dom_simpson, dom_shannon, dom_richness)

dygraph(dygraph_rich_simp_shannon, main = "Florida Keys Reef Fish Biodiversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 350)) %>% 
  dyAxis("x", label = "Year") %>%
  dySeries("dom_richness", label = "Richness", color = "purple") %>%
  dySeries("dom_shannon", label = "Shannon", color = "blue") %>% 
  dySeries("dom_simpson", label = "Simpson", color = "red") %>% 
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 1.5) %>% #drawGrid = F
  dyLegend(width = 400)

##without spp 
d_no_spp = fk_domain_abun_diversity_no_spp %>% 
  filter(protected_status != 0) %>% 
  filter(protected_status != 1)

dygraph_rich_simp_shannon_no_spp = d_no_spp %>%
  select(YEAR, simpson, shannon, richness)

dygraph(dygraph_rich_simp_shannon_no_spp, main = "Florida Keys Reef Fish Biodiversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 350)) %>% 
  dyAxis("x", label = "Year") %>%
  dySeries("richness", label = "Richness", color = "purple") %>%
  dySeries("shannon", label = "Shannon", color = "blue") %>% 
  dySeries("simpson", label = "Simpson", color = "red") %>% 
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 1.5) %>% #drawGrid = F
  dyLegend(width = 400)

dygraph_simp_shannon_no_spp = d_no_spp %>%
  select(YEAR, simpson, shannon)

dygraph(dygraph_simp_shannon_no_spp, main = "Florida Keys Reef Fish Biodiversity") %>% 
  dyAxis("y", label = "Effective Number of Species") %>%  #valueRange = c(0, 350)
  dyAxis("x", label = "Year") %>%
  dySeries("shannon", label = "Shannon", color = "blue") %>% 
  dySeries("simpson", label = "Simpson", color = "red") %>% 
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 1.5) %>% #drawGrid = F
  dyLegend(width = 400)

```

```{r domain biodiversity plots}

domain_fk_abun_diversity = read_csv('big_csv/abundance_domain/domain_fk_abun_diversity.csv')
# Parsed with column specification:
# cols(
#   YEAR = col_integer(),
#   protected_status = col_character(),
#   dom_richness = col_integer(),
#   dom_simpson = col_double(),
#   dom_shannon = col_double()
# )
protected_status_names <- c(
  "0" = "Unprotected",
  "1" = "Protected",
  "all" = "All"
)
#richness, simpson, shannon for domain FKNMS  
FK_abun_RShSim <- ggplot(data = fk_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_richness, color = "dom_richness")) +
  geom_point(aes(y = dom_simpson, color = "dom_simpson")) +
  geom_point(aes(y = dom_shannon, color = "dom_shannon")) +
  geom_line(aes(y = dom_richness, color = "dom_richness")) +
  geom_line(aes(y = dom_simpson, color = "dom_simpson")) +
  geom_line(aes(y = dom_shannon, color = "dom_shannon")) +
  facet_grid(protected_status ~ ., labeller = as_labeller(protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Diversity indices") +
  scale_colour_manual(values = c("darkturquoise", "magenta", "blue"), labels=c("Species Richness", "Shannon Diversity", "Simpson Diversity")) +
  ggtitle("Reef Fish Biodiversity in the FKNMS") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(breaks = c(1999:2016))+
  scale_y_continuous(limits= c(0, 300), breaks = c(0,25,50,75,100,125,150,175,200,225,250,275,300))


#simpson, shannon by protected status for domain FKNMS
FK_abun_ShSim <-ggplot(data = fk_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_simpson, color = "dom_simpson")) +
  geom_point(aes(y = dom_shannon, color = "dom_shannon")) +
  geom_line(aes(y = dom_simpson, color = "dom_simpson")) +
  geom_line(aes(y = dom_shannon, color = "dom_shannon")) +
  facet_grid(protected_status ~ ., labeller = as_labeller(protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Diversity index") +
  scale_colour_manual(values = c("magenta", "blue"), labels=c("Shannon diversity", "Simpson Diversity")) +
  ggtitle("Reef Fish Biodiversity in the Florida Keys") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 250), breaks = c(5,10,15,20,25,30,35,40,45,50))

#species richness by protected status for domain FKNMS
FK_abun_Rich <- ggplot(data = fk_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_richness, color = "dom_richness"),color = "darkturquoise") +
  geom_line(aes(y = dom_richness, color = "dom_richness"), color = "darkturquoise") +
  facet_grid(protected_status ~ ., labeller = as_labeller(protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Level of protection") +
  ggtitle("Species Richness of Reef Fish in the Florida Keys") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 250), breaks = c(0,50, 100, 150, 200, 250, 300))
## wonky 

#shannon by protected status for domain FKNMS
FK_abun_Shan <- ggplot(data = fk_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_shannon, color = "dom_shannon"),color = "magenta") +
  geom_line(aes(y = dom_shannon, color = "dom_shannon"), color = "magenta") +
  facet_grid(protected_status ~ ., labeller = as_labeller(protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Level of protection") +
  #scale_colour_manual(labels=c("Unprotected", "Protected", "All")) +
  ggtitle("Shannon Diveristy of Reef Fish in the Florida Keys") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 50), breaks = c(5,10,15,20,25,30,35,40,45,50))

#simpson by protected status for domain FKNMS
FK_abun_Simp <- ggplot(data = fk_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_simpson, color = "dom_simpson"),color = "blue") +
  geom_line(aes(y = dom_simpson, color = "dom_simpson"), color = "blue") +
  facet_grid(protected_status ~ ., labeller = as_labeller(protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Level of protection") +
  #scale_colour_manual(labels=c("Unprotected", "Protected", "All")) +
  ggtitle("Simpson Diveristy of Reef Fish in the Florida Keys") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 25), breaks = c(5,10,15,20,25))

```

### Florida Keys Strata
1. FSLR - Forereef Shallow Linear Reef (<6m)
2. FMLR - Forereef Midchannel Linear Reef (6-18m)
3. FDLR - Forereef Deep Linear Reef (18-33m)
4. INPR - Inshore patch reef
5. MCPR - Midchannel Patch Reef 
6. OFPR - Offshore Patch Reef
7. HRRF - High Relief Reef (Spur and Groove)

```{r strata abundance}
## FK's abundance by strata 
strata_fk_abun <- read_csv("big_csv/abundance_strata/strata_fk_abun.csv",
                           col_types = cols(
                             YEAR = col_integer(),
                             REGION = col_character(),
                             STRAT = col_character(),
                             SPECIES_CD = col_character(),
                             abundance = col_double(),
                             var = col_double(),
                             n = col_integer(),
                             nm = col_integer(),
                             N = col_double(),
                             NM = col_double(),
                             protected_status = col_character()
                           ))

strata_fk_abun_no_spp = strata_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species 
```

```{r strata biomass}

strata_fk_bio <- read_csv('big_csv/biomass_strata/strata_fk_biomass.csv', 
                          col_types = cols(
                            protected_status = col_character()
                          ))

strata_fk_bio_no_spp = strata_fk_bio %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species
```

Is reef fish biodiversity and strata type related? 

```{r strata biodiversity}

## FK's abundance by strata 
strata_fk_abun <- read_csv("big_csv/abundance_strata/strata_fk_abun.csv",
                           col_types = cols(
                             YEAR = col_integer(),
                             REGION = col_character(),
                             STRAT = col_character(),
                             SPECIES_CD = col_character(),
                             abundance = col_double(),
                             var = col_double(),
                             n = col_integer(),
                             nm = col_integer(),
                             N = col_double(),
                             NM = col_double(),
                             protected_status = col_character()
                           ))

strata_fk_abun_no_spp = strata_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species 

fk_strata_abun_diversity_no_spp = strata_fk_abun_no_spp %>% #124,116 * 12
  filter(protected_status == 'all') %>% #66,058 * 12
  select(YEAR, STRAT, protected_status, SPECIES_CD, abundance) %>% #62,058 * 5
  group_by(YEAR, STRAT, SPECIES_CD) %>% #33,468 gruops    
  summarise(
    abun_merged = sum(abundance))

trait_matrix = read_csv('functional_diversity/species_trait_matrix_348_spp.csv')

trait_matrix = trait_matrix %>%
  as.tibble() %>%
  mutate(
    SPECIES_CD = toupper(as.character(SPECIES_CD))) %>% 
  arrange(SPECIES_CD)

strata_merged = fk_strata_abun_diversity_no_spp %>%
  group_by(YEAR, STRAT) %>% #103 groups 
  nest() %>%  #data 339*2
  mutate(
    data_wide = map(data, function(x) 
      full_join(x, trait_matrix %>% select(SPECIES_CD), by='SPECIES_CD') %>% 
        mutate(abun_merged = ifelse(is.na(abun_merged), 0, abun_merged)) %>%
        spread(SPECIES_CD, abun_merged, fill =0)),
    richness = map( #species richness
      data_wide, 
      function(x) specnumber(x)),
    simpson = map( #simpson diversity as effective number of species 
      data_wide,
      function(x) 1/(1 - diversity(x, index = 'simpson'))),
    shannon = map( #shannon diversity as effective number of species 
      data_wide,
      function(x) exp(diversity(x, index = 'shannon')))) %>%
  unnest(richness, simpson, shannon)

strata_diversity_no_spp_merged = strata_merged %>%
  select(YEAR, STRAT, richness, simpson, shannon) 

write_csv(strata_diversity_no_spp_merged, 'diversity_without_spe/strata_fk_abun_diversity_no_spp_merged.csv')

```

```{r strata biodiversity plots, eval = F}

strata_diversity_no_spp_merged = read_csv("diversity_without_spe/strata_fk_abun_diversity_no_spp_merged.csv")
# Parsed with column specification:
# cols(
#   YEAR = col_integer(),
#   STRAT = col_character(),
#   richness = col_integer(),
#   simpson = col_double(),
#   shannon = col_double()
# )

# strata_names <- c(
#  "FMLR" = "Forereef Midchannel Linear Reef",
#  "FSLR" = "Forereef Shallow Linear Reef",
#  "HRRF" = "High Relief Reef", #(Spur and Groove)
#  "INPR" = "Inshore patch reef",
#  "MCPR" = "Midchannel Patch Reef",
#  "OFPR" = "Offshore Patch Reef",
#  "FDLR" = "Forereef Deep Linear Reef")

##protected by strata 
fk_strata_protected = strata_fk_abun_diversity %>% 
  filter(protected_status != 1)

#richness protected by strata 
fk_rich_protected = fk_strata_protected %>% 
  select(YEAR, STRAT, richness) %>%
  spread(STRAT, richness)

dygraph(fk_rich_protected, main = "Florida Keys Richness by Strata - NTMR") %>% 
  dyAxis("y", label = "Effective Number of Species") %>%
  dyAxis("x", label = "Year") %>% 
  dyOptions(stackedGraph = T, colors = colorRamps::blue2red(7), fillAlpha = .8, axisLineWidth = 2)

##shannon protected by strata 
fk_shan_protected = fk_strata_protected %>% 
  select(YEAR, STRAT, shannon) %>%
  spread(STRAT, shannon)

dygraph(fk_shan_protected, main = "Florida Keys Shannon Diversity by Strata - NTMR") %>% 
  dyAxis("y", label = "Effective Number of Species") %>%
  dyAxis("x", label = "Year") %>% 
  dyOptions(stackedGraph = T, colors = colorRamps::blue2red(7), fillAlpha = .8, axisLineWidth = 2)

##simpson protected by strata 
fk_simp_protected = fk_strata_protected %>% 
  select(YEAR, STRAT, simpson) %>%
  spread(STRAT, simpson)

dygraph(fk_simp_protected, main = "Florida Keys Simpson Diversity by Strata - NTMR") %>% 
  dyAxis("y", label = "Effective Number of Species") %>%
  dyAxis("x", label = "Year") %>% 
  dyOptions(stackedGraph = T, colors = colorRamps::blue2red(7), fillAlpha = .8, axisLineWidth = 2)#topo.colors(n = 7, alpha = 1) RColorBrewer::brewer.pal(7, "Paired") grDevices::topo.colors(7, alpha = 1) grDevices::rainbow(7) colorspace::rainbow_hcl(7) colorspace::heat_hcl(7)

#unprotected by strata 
fk_strata_unprotected = strata_fk_abun_diversity %>% 
  filter(protected_status != 0)

#richness unprotected by strata 
fk_rich_unprotected = fk_strata_unprotected %>% 
  select(YEAR, STRAT, richness) %>%
  spread(STRAT, richness)

dygraph(fk_rich_unprotected, main = "Florida Keys Richness by Strata - Unprotected") %>% 
  dyAxis("y", label = "Effective Number of Species") %>%
  dyAxis("x", label = "Year") %>% 
  dyOptions(stackedGraph = T, colors = colorRamps::blue2red(7), fillAlpha = .8, axisLineWidth = 2)

#simpson unprotected by strata
fk_simp_unprotected = fk_strata_unprotected %>% 
  select(YEAR, STRAT, simpson) %>%
  spread(STRAT, simpson)

dygraph(fk_simp_unprotected, main = "Florida Keys Simpson Diversity by Strata - Unprotected") %>% 
  dyAxis("y", label = "Effective Number of Species") %>%
  dyAxis("x", label = "Year") %>% 
  dyOptions(stackedGraph = T, colors = colorRamps::blue2red(7), fillAlpha = .8, axisLineWidth = 2)

strata_names <- c(
  "FMLR" = "FMLR",
  "FSLR" = "FSLR",
  "HRRF" = "HRRF", #(Spur and Groove)
  "INPR" = "INPR",
  "MCPR" = "MCPR",
  "OFPR" = "OFPR",
  "FDLR" = "FDLR")

protected_status_names <- c(
  "0" = "Unprotected",
  "1" = "Protected",
  "all" = "All"
)

#strata_diversity_no_spp_merged - not by level of protection  

#simpson diversity by protected status by strata in FKNMS
ggplot(data = strata_diversity_no_spp_merged, aes(x = YEAR)) +
  geom_point(aes(y = simpson, color = "simpson"),color = "blue") +
  geom_line(aes(y = simpson, color = "simpson"), color = "blue") +
  facet_grid(STRAT ~.) +
  labs(x = "Year", y = "Effective Number of Species") +
  ggtitle("Simpson Diveristy of Reef Fish in the Florida Keys") + 
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12)) +
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 25), breaks = c(5,10,15,20,25))

#shannon diversity by protected status by strata in FKNMS
FK_strata_ab_shannon <- ggplot(data = fk_strata_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = shannon, color = "shannon"), color = "magenta") +
  geom_line(aes(y = shannon, color = "shannon"), color = "magenta") +
  facet_grid(STRAT ~ protected_status, 
             labeller = labeller(.rows = strata_names, .cols = protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Level of protection") +
  ggtitle("Shannon Diveristy of Reef Fish in the Florida Keys by Strata") + 
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 50), breaks = c(5,10,15,20,25))

```

### Florida Keys Primary Sampling Unit 

```{r psu abundance}

psu_fk_abun <- read_csv("big_csv/abundance_psu/psu_fk_abun.csv",
                        col_types = cols(
                          YEAR = col_integer(),
                          REGION = col_character(),
                          STRAT = col_character(),
                          PROT = col_integer(),
                          PRIMARY_SAMPLE_UNIT = col_character(),
                          SPECIES_CD = col_character(),
                          m = col_integer(),
                          var = col_double(),
                          abundance = col_double(),
                          protected_status = col_character())) #3,643,306 x 10

psu_fk_abun_no_spp = psu_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species and 3,379,964 x 10

psu_fk_abun_no_spp_merged = psu_fk_abun_no_spp %>%
  filter(stringr::str_detect(protected_status, 'all')) # 1,689,982 x 10 (this is half the number of rows as the original as it should be)

write_csv(psu_fk_abun_no_spp_merged, "big_csv/abundance_psu/psu_fk_abundance_no_spp_merged.csv")

```

```{r psu biomass}

psu_fk_bio <- read_csv('big_csv/biomass_psu/psu_fk_biomass.csv', 
                       col_types = cols(
                          YEAR = col_integer(),
                          REGION = col_character(),
                          STRAT = col_character(),
                          PROT = col_integer(),
                          PRIMARY_SAMPLE_UNIT = col_character(),
                          SPECIES_CD = col_character(),
                          m = col_integer(),
                          var = col_double(),
                          biomass = col_double(),
                          protected_status = col_character())) #3,429,476 x 10

psu_fk_bio_no_spp = psu_fk_bio %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species and 3,219,344 x 10
psu_fk_bio_no_spp_merged = psu_fk_bio_no_spp %>%
  filter(stringr::str_detect(protected_status, 'all')) #1,609,672 x 10 (this is half the number of rows as the original as it should be)

write_csv(psu_fk_bio_no_spp_merged, "big_csv/biomass_psu/psu_fk_biomass_no_spp_merged.csv")

```

```{r psu biodiversity}
#read in psu abundance data 
psu_fk_abun <- read_csv("big_csv/abundance_psu/psu_fk_abun.csv",
                        col_types = cols(
                          YEAR = col_integer(),
                          REGION = col_character(),
                          STRAT = col_character(),
                          PROT = col_integer(),
                          PRIMARY_SAMPLE_UNIT = col_character(),
                          SPECIES_CD = col_character(),
                          m = col_integer(),
                          var = col_double(),
                          abundance = col_double(),
                          protected_status = col_character()))

#remove species not identified to species 
psu_fk_abun_no_spp = psu_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species 

#read in trait matrix 
trait_matrix = read_csv('functional_diversity/species_trait_matrix_348_spp.csv')

#trait matrix needs to be alphabetically for FD code 
trait_matrix = trait_matrix %>%
  as.tibble() %>%
  mutate(
    SPECIES_CD = toupper(as.character(SPECIES_CD))) %>% 
  arrange(SPECIES_CD)

# fk_psu_abun_diversity_no_spp_merged = psu_fk_abun_no_spp %>% #3,107,408 x 10
#  filter(protected_status != 'all') %>% #1,689,982 x 6
#  group_by(YEAR,  PRIMARY_SAMPLE_UNIT, SPECIES_CD) %>% #1,689,982 groups  
#  mutate(abun_merged = sum(abundance)) 

#compute diversity indices by PSU  
psu_diversity_no_spp_merged = psu_fk_abun_no_spp %>% #tibble: 3,379,964 x 10 %>%
  filter(protected_status != "all") %>% #tibble: 1,689,982 x 10
  select(YEAR, PRIMARY_SAMPLE_UNIT,STRAT,PROT,SPECIES_CD,abundance) %>% #tibble: 1,689,982 x 6
  group_by(YEAR, PRIMARY_SAMPLE_UNIT) %>% #4,839 groups
  nest(-YEAR) %>%
  mutate(
    data_wide = map(data, function(x) 
      full_join(x, trait_matrix %>% select(SPECIES_CD), by='SPECIES_CD') %>%
        mutate(abundance = ifelse(is.na(abundance), 0, abundance)) %>%
        spread(SPECIES_CD, abundance, fill =0)),
    data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill =0)), 
    richness = map( #species richness
      data_wide, 
      function(x) specnumber(x %>% select(-STRAT, -PROT))),
    simpson = map( #simpson diversity as effective number of species 
      data_wide,
      function(x) 1/(1 - diversity(x %>% select(-STRAT, -PROT), index = 'simpson'))),
    shannon = map( #shannon diversity as effective number of species 
      data_wide, 
      function(x) exp(diversity(x %>% select(-STRAT, -PROT), index = 'shannon')))) %>% 
  unnest(richness,simpson, shannon)

#get strata and prot information and remove duplicates
psu_diversity_no_spp_merged_strat = psu_diversity_no_spp_merged %>%
  unnest(data) #tibble: 1,689,982 x 9 
  
#save dataframe as csv
write_csv(psu_diversity_no_spp_merged_strat, "big_csv/diversity_without_spe/psu_diversity_no_spp_merged.csv") 

psu_div = psu_diversity_no_spp_merged_strat %>% 
  dplyr::group_by(YEAR, PRIMARY_SAMPLE_UNIT) %>% #A tibble: 1,689,982 x 7; groups 5,241
  dplyr::summarize(
    richness = first(richness),
    simpson = first(simpson),
    shannon = first(shannon))
# 5241 total PSU's over 15 years 

```

```{r psu abundance plots and maps}

psu_abun <- read_csv('big_csv/abundance_psu/psu_fk_abundance_no_spp_merged.csv', 
                       col_types = cols(
                          YEAR = col_integer(),
                          REGION = col_character(),
                          STRAT = col_character(),
                          PROT = col_integer(),
                          PRIMARY_SAMPLE_UNIT = col_character(),
                          SPECIES_CD = col_character(),
                          m = col_integer(),
                          var = col_double(),
                          abundance = col_double(),
                          protected_status = col_character())) # 1,689,982  x 10

# the psu_abun dataframe has an abundance for each species for each psu so sum abundance across species for each psu...skip this step for diversity indices
psu_abun_sum <- psu_abun %>%
  group_by(YEAR, PRIMARY_SAMPLE_UNIT, STRAT) %>%
  summarize(abundance_sum = sum(abundance)) # 5241 x 4 (5241 = total # of PSUs sampled from 1999 to 2016)
  
# (optional) the psu_bio_sum_YRsubset dataframe produced with the code below has years with low sample sizes (few PSUs)/missing strata removed (i.e., 1999, 2000, and 2004)
psu_abun_sum_YRsubset <- psu_abun_sum %>%
  filter(YEAR!=1999,YEAR!=2000,YEAR!=2004) # 4725 x 4

# 2. Plot data by year and stratum separately and together
# summarize data to get mean and standard error for each year/stratum combination  
psu_abun_forPlot = psu_abun_sum %>% 
  group_by(YEAR, STRAT) %>%
  summarize(
    abundance_mean = mean(abundance_sum),
    abundance_n = length(abundance_sum),
    abundance_sd = sd(abundance_sum),
    abundance_se = abundance_sd / sqrt(abundance_n),
    abundance_min = min(abundance_sum),
    abundance_max = max(abundance_sum))
# plot mean and se across strat and year using bar graphs
# by STRAT
ggplot(psu_abun_forPlot, aes(x=STRAT, y=abundance_mean)) + geom_bar(position=position_dodge(), stat="identity") + geom_errorbar(aes(ymin=abundance_mean-abundance_se, ymax=abundance_mean+abundance_se), width=.2, position=position_dodge(.9)) + theme_classic() + theme(axis.text.x = element_text(angle=90)) + facet_wrap(~YEAR, scales='free_y')
ggsave(file="fk_MeanAbundance_ByStratumForEachYear.pdf", path="big_csv/plots")
# by YEAR
ggplot(psu_abun_forPlot, aes(x=YEAR, y=abundance_mean, group=YEAR)) + geom_bar(position=position_dodge(), stat="identity") + geom_errorbar(aes(ymin=abundance_mean-abundance_se, ymax=abundance_mean+abundance_se), width=.2, position=position_dodge(.9)) + theme_classic() + theme(axis.text.x = element_text(angle=90)) + facet_wrap(~STRAT, scales='free_y')
ggsave(file="fk_MeanAbundance_ByYearForEachStratum.pdf", path="big_csv/plots")
# by STRAT and YEAR
ggplot(psu_abun_forPlot, aes(x=YEAR, y=abundance_mean, colour=STRAT, group=STRAT)) + geom_point() + geom_line() + geom_errorbar(aes(ymin=abundance_mean-abundance_se, ymax=abundance_mean+abundance_se), width=.2)
ggsave(file="fk_MeanAbundance_ByYearANDStratum.pdf", path="big_csv/plots")

# do this for dataframe with subset of years as well
# summarize data to get mean and standard error for each year/stratum combination  
psu_abun_forPlot_YRsubset = psu_abun_sum_YRsubset %>% 
  group_by(YEAR, STRAT) %>%
  summarize(
    abundance_mean = mean(abundance_sum),
    abundance_n = length(abundance_sum),
    abundance_sd = sd(abundance_sum),
    abundance_se = abundance_sd / sqrt(abundance_n),
    abundance_min = min(abundance_sum),
    abundance_max = max(abundance_sum))
# plot mean and se across strat and year using bar graphs
# by STRAT
ggplot(psu_abun_forPlot_YRsubset, aes(x=STRAT, y=abundance_mean)) + geom_bar(position=position_dodge(), stat="identity") + geom_errorbar(aes(ymin=abundance_mean-abundance_se, ymax=abundance_mean+abundance_se), width=.2, position=position_dodge(.9)) + theme_classic() + theme(axis.text.x = element_text(angle=90)) + facet_wrap(~YEAR, scales='free_y')
ggsave(file="fk_MeanAbundance_ByStratumForEachYear_YRsubset.pdf", path="big_csv/plots")
# by YEAR
ggplot(psu_abun_forPlot_YRsubset, aes(x=YEAR, y=abundance_mean, group=YEAR)) + geom_bar(position=position_dodge(), stat="identity") + geom_errorbar(aes(ymin=abundance_mean-abundance_se, ymax=abundance_mean+abundance_se), width=.2, position=position_dodge(.9)) + theme_classic() + theme(axis.text.x = element_text(angle=90)) + facet_wrap(~STRAT, scales='free_y')
ggsave(file="fk_MeanAbundance_ByYearForEachStratum_YRsubset.pdf", path="big_csv/plots")
# by STRAT and YEAR
ggplot(psu_abun_forPlot_YRsubset, aes(x=YEAR, y=abundance_mean, colour=STRAT, group=STRAT)) + geom_point() + geom_line() + geom_errorbar(aes(ymin=abundance_mean-abundance_se, ymax=abundance_mean+abundance_se), width=.2)
ggsave(file="fk_MeanAbundance_ByYearANDStratum_YRsubset.pdf", path="big_csv/plots")

### tried these for abundance but didn't look good (too many very low values mixed with a small number of very high values)
#ggplot(data = psu_abun_sum) + geom_boxplot(mapping = aes(x = STRAT, y = abundance_sum)) + facet_wrap(~YEAR, scales="free_y")
#ggplot(data = psu_abun_sum) + geom_boxplot(mapping = aes(x = YEAR, y = abundance_sum, group = YEAR)) + facet_wrap(~STRAT, scales="free_y")
```

```{r psu biomass plots and maps}

psu_bio <- read_csv('big_csv/biomass_psu/psu_fk_biomass_no_spp_merged.csv', 
                       col_types = cols(
                          YEAR = col_integer(),
                          REGION = col_character(),
                          STRAT = col_character(),
                          PROT = col_integer(),
                          PRIMARY_SAMPLE_UNIT = col_character(),
                          SPECIES_CD = col_character(),
                          m = col_integer(),
                          var = col_double(),
                          biomass = col_double(),
                          protected_status = col_character())) # 1,609,672  x 10

# the psu_bio dataframe has a biomass for each species for each psu so sum biomass across species for each psu...skip this step for diversity indices
psu_bio_sum <- psu_bio %>%
  group_by(YEAR, PRIMARY_SAMPLE_UNIT, STRAT) %>%
  summarize(biomass_sum = sum(biomass)) # 5241 x 4 (5241 = total # of PSUs sampled from 1999 to 2016)
  
# (optional) the psu_bio_sum_YRsubset dataframe produced with the code below has years with low sample sizes (few PSUs)/missing strata removed (i.e., 1999, 2000, and 2004)
psu_bio_sum_YRsubset <- psu_bio_sum %>%
  filter(YEAR!=1999,YEAR!=2000,YEAR!=2004) # 4725 x 4

# plot data by year and stratum separately and together
# summarize data to get mean and standard error for each year/stratum combination  
psu_bio_forPlot = psu_bio_sum %>% 
  group_by(YEAR, STRAT) %>%
  summarize(
    biomass_mean = mean(biomass_sum),
    biomass_n = length(biomass_sum),
    biomass_sd = sd(biomass_sum),
    biomass_se = biomass_sd / sqrt(biomass_n),
    biomass_min = min(biomass_sum),
    biomass_max = max(biomass_sum))
# plot mean and se across strat and year using bar graphs
# by STRAT
ggplot(psu_bio_forPlot, aes(x=STRAT, y=biomass_mean)) + geom_bar(position=position_dodge(), stat="identity") + geom_errorbar(aes(ymin=biomass_mean-biomass_se, ymax=biomass_mean+biomass_se), width=.2, position=position_dodge(.9)) + theme_classic() + theme(axis.text.x = element_text(angle=90)) + facet_wrap(~YEAR, scales='free_y')
ggsave(file="fk_MeanBiomass_ByStratumForEachYear.pdf", path="big_csv/plots")
# by YEAR
ggplot(psu_bio_forPlot, aes(x=YEAR, y=biomass_mean, group=YEAR)) + geom_bar(position=position_dodge(), stat="identity") + geom_errorbar(aes(ymin=biomass_mean-biomass_se, ymax=biomass_mean+biomass_se), width=.2, position=position_dodge(.9)) + theme_classic() + theme(axis.text.x = element_text(angle=90)) + facet_wrap(~STRAT, scales='free_y')
ggsave(file="fk_MeanBiomass_ByYearForEachStratum.pdf", path="big_csv/plots")
# by STRAT and YEAR
ggplot(psu_bio_forPlot, aes(x=YEAR, y=biomass_mean, colour=STRAT, group=STRAT)) + geom_point() + geom_line() + geom_errorbar(aes(ymin=biomass_mean-biomass_se, ymax=biomass_mean+biomass_se), width=.2)
ggsave(file="fk_MeanBiomass_ByYearANDStratum.pdf", path="big_csv/plots")

# do this for dataframe with subset of years as well
# summarize data to get mean and standard error for each year/stratum combination  
psu_bio_forPlot_YRsubset = psu_bio_sum_YRsubset %>% 
  group_by(YEAR, STRAT) %>%
  summarize(
    biomass_mean = mean(biomass_sum),
    biomass_n = length(biomass_sum),
    biomass_sd = sd(biomass_sum),
    biomass_se = biomass_sd / sqrt(biomass_n),
    biomass_min = min(biomass_sum),
    biomass_max = max(biomass_sum))
# plot mean and se across strat and year using bar graphs
# by STRAT
ggplot(psu_bio_forPlot_YRsubset, aes(x=STRAT, y=biomass_mean)) + geom_bar(position=position_dodge(), stat="identity") + geom_errorbar(aes(ymin=biomass_mean-biomass_se, ymax=biomass_mean+biomass_se), width=.2, position=position_dodge(.9)) + theme_classic() + theme(axis.text.x = element_text(angle=90)) + facet_wrap(~YEAR, scales='free_y')
ggsave(file="fk_MeanBiomass_ByStratumForEachYear_YRsubset.pdf", path="big_csv/plots")
# by YEAR
ggplot(psu_bio_forPlot_YRsubset, aes(x=YEAR, y=biomass_mean, group=YEAR)) + geom_bar(position=position_dodge(), stat="identity") + geom_errorbar(aes(ymin=biomass_mean-biomass_se, ymax=biomass_mean+biomass_se), width=.2, position=position_dodge(.9)) + theme_classic() + theme(axis.text.x = element_text(angle=90)) + facet_wrap(~STRAT, scales='free_y')
ggsave(file="fk_MeanBiomass_ByYearForEachStratum_YRsubset.pdf", path="big_csv/plots")
# by STRAT and YEAR
ggplot(psu_bio_forPlot_YRsubset, aes(x=YEAR, y=biomass_mean, colour=STRAT, group=STRAT)) + geom_point() + geom_line() + geom_errorbar(aes(ymin=biomass_mean-biomass_se, ymax=biomass_mean+biomass_se), width=.2)
ggsave(file="fk_MeanBiomass_ByYearANDStratum_YRsubset.pdf", path="big_csv/plots")

### tried using boxplots but they don't look good
#ggplot(data = psu_bio_sum) + geom_boxplot(mapping = aes(x = STRAT, y = biomass_sum)) + facet_wrap(~YEAR, scales="free_y")
#ggplot(data = psu_bio_sum) + geom_boxplot(mapping = aes(x = YEAR, y = biomass_sum, group = YEAR)) + facet_wrap(~STRAT, scales="free_y")

```

```{r psu biodiversity plots and maps}

psu_fk_abun_diversity = read_csv("big_csv/diversity_without_spe/psu_fk_abun_diversity_no_spp.csv",
                                 col_types = cols(
                                   protected_status = col_character())) # this was already here in the chunk so i didn't want to delete it, but not using it in code below - KS


psu_diversity = read_csv("big_csv/diversity_without_spe/psu_diversity_no_spp_merged.csv",
                      col_types = cols(
                          YEAR = col_integer(),
                          PRIMARY_SAMPLE_UNIT = col_character(),
                          richness = col_integer(),
                          simpson = col_double(),
                          shannon = col_double(),
                          STRAT = col_character(),
                          SPECIES_CD = col_character(),
                          abun_merged = col_double())) # 1,689,982 x 9

# the psu_diversity dataframe has a value for each diversity metric that is repeated for each YEAR/PSU for each species record....need to remove duplicate rows and can also select columns of interest for the analysis of diversity metrics
psu_div = psu_diversity %>% 
  select(YEAR, PRIMARY_SAMPLE_UNIT, STRAT, richness, simpson, shannon) %>%
  group_by(YEAR, PRIMARY_SAMPLE_UNIT, STRAT) %>%
  summarize(
    richness = first(richness),
    simpson = first(simpson),
    shannon = first(shannon)) # 5241 x 6
  
# (optional) the psu_div_YRsubset dataframe produced with the code below has years with low sample sizes (few PSUs)/missing strata removed (i.e., 1999, 2000, and 2004)
psu_div_YRsubset <- psu_div %>%
  filter(YEAR!=1999,YEAR!=2000,YEAR!=2004) # 4725 x 6


# plot data by year and stratum separately and together
# summarize data to get mean and standard error for each year/stratum combination  
psu_div_forPlot = psu_div %>% 
  group_by(YEAR, STRAT) %>%
  summarize(
    richness_mean = mean(richness),
    richness_n = length(richness),
    richness_sd = sd(richness),
    richness_se = richness_sd / sqrt(richness_n),
    richness_min = min(richness),
    richness_max = max(richness),
    simpson_mean = mean(simpson),
    simpson_n = length(simpson),
    simpson_sd = sd(simpson),
    simpson_se = simpson_sd / sqrt(simpson_n),
    simpson_min = min(simpson),
    simpson_max = max(simpson),
    shannon_mean = mean(shannon),
    shannon_n = length(shannon),
    shannon_sd = sd(shannon),
    shannon_se = shannon_sd / sqrt(shannon_n),
    shannon_min = min(shannon),
    shannon_max = max(shannon))
# boxplots look better for diversity metrics than barplots (the opposite is true for biomass and abundance)...skip barplots
# by STRAT
#richness
#ggplot(psu_div_forPlot, aes(x=STRAT, y=richness_mean)) + geom_bar(position=position_dodge(), stat="identity") + geom_errorbar(aes(ymin=richness_mean-richness_se, ymax=richness_mean+richness_se), width=.2, position=position_dodge(.9)) + theme_classic() + theme(axis.text.x = element_text(angle=90)) + facet_wrap(~YEAR, scales='free_y')
#ggsave(file="fk_MeanRichness_ByStratumForEachYear.pdf", path="big_csv/plots")
# by YEAR
#ggplot(psu_div_forPlot, aes(x=YEAR, y=richness_mean, group=YEAR)) + geom_bar(position=position_dodge(), stat="identity") + geom_errorbar(aes(ymin=richness_mean-richness_se, ymax=richness_mean+richness_se), width=.2, position=position_dodge(.9)) + theme_classic() + theme(axis.text.x = element_text(angle=90)) + facet_wrap(~STRAT, scales='free_y')
#ggsave(file="fk_MeanRichness_ByYearForEachStratum.pdf", path="big_csv/plots")
# by STRAT and YEAR
# richness
ggplot(psu_div_forPlot, aes(x=YEAR, y=richness_mean, colour=STRAT, group=STRAT)) + geom_point() + geom_line() + geom_errorbar(aes(ymin=richness_mean-richness_se, ymax=richness_mean+richness_se), width=.2)
ggsave(file="fk_MeanRichness_ByYearANDStratum.pdf", path="big_csv/plots")
# simpson
ggplot(psu_div_forPlot, aes(x=YEAR, y=simpson_mean, colour=STRAT, group=STRAT)) + geom_point() + geom_line() + geom_errorbar(aes(ymin=simpson_mean-simpson_se, ymax=simpson_mean+simpson_se), width=.2)
ggsave(file="fk_MeanSimpson_ByYearANDStratum.pdf", path="big_csv/plots")
# shannon
ggplot(psu_div_forPlot, aes(x=YEAR, y=shannon_mean, colour=STRAT, group=STRAT)) + geom_point() + geom_line() + geom_errorbar(aes(ymin=shannon_mean-shannon_se, ymax=shannon_mean+shannon_se), width=.2)
ggsave(file="fk_MeanShannon_ByYearANDStratum.pdf", path="big_csv/plots")

# do this for dataframe with subset of years as well
# summarize data to get mean and standard error for each year/stratum combination  
psu_div_forPlot_YRsubset = psu_div_YRsubset %>% 
  group_by(YEAR, STRAT) %>%
  summarize(
    richness_mean = mean(richness),
    richness_n = length(richness),
    richness_sd = sd(richness),
    richness_se = richness_sd / sqrt(richness_n),
    richness_min = min(richness),
    richness_max = max(richness),
    simpson_mean = mean(simpson),
    simpson_n = length(simpson),
    simpson_sd = sd(simpson),
    simpson_se = simpson_sd / sqrt(simpson_n),
    simpson_min = min(simpson),
    simpson_max = max(simpson),
    shannon_mean = mean(shannon),
    shannon_n = length(shannon),
    shannon_sd = sd(shannon),
    shannon_se = shannon_sd / sqrt(shannon_n),
    shannon_min = min(shannon),
    shannon_max = max(shannon))
# boxplots look better for diversity metrics than barplots (the opposite is true for biomass and abundance)...skip barplots
# by STRAT
#ggplot(psu_div_forPlot_YRsubset, aes(x=STRAT, y=richness_mean)) + geom_bar(position=position_dodge(), stat="identity") + geom_errorbar(aes(ymin=richness_mean-richness_se, ymax=richness_mean+richness_se), width=.2, position=position_dodge(.9)) + theme_classic() + theme(axis.text.x = element_text(angle=90)) + facet_wrap(~YEAR, scales='free_y')
#ggsave(file="fk_MeanRichness_ByStratumForEachYear_YRsubset.pdf", path="big_csv/plots")
# by YEAR
#ggplot(psu_div_forPlot_YRsubset, aes(x=YEAR, y=richness_mean, group=YEAR)) + geom_bar(position=position_dodge(), stat="identity") + geom_errorbar(aes(ymin=richness_mean-richness_se, ymax=richness_mean+richness_se), width=.2, position=position_dodge(.9)) + theme_classic() + theme(axis.text.x = element_text(angle=90)) + facet_wrap(~STRAT, scales='free_y')
#ggsave(file="fk_MeanRichness_ByYearForEachStratum_YRsubset.pdf", path="big_csv/plots")
# by STRAT and YEAR
## richness
ggplot(psu_div_forPlot_YRsubset, aes(x=YEAR, y=richness_mean, colour=STRAT, group=STRAT)) + geom_point() + geom_line() + geom_errorbar(aes(ymin=richness_mean-richness_se, ymax=richness_mean+richness_se), width=.2)
ggsave(file="fk_MeanRichness_ByYearANDStratum_YRsubset.pdf", path="big_csv/plots")
## simpson
ggplot(psu_div_forPlot_YRsubset, aes(x=YEAR, y=richness_mean, colour=STRAT, group=STRAT)) + geom_point() + geom_line() + geom_errorbar(aes(ymin=richness_mean-richness_se, ymax=richness_mean+richness_se), width=.2)
ggsave(file="fk_MeanSimpson_ByYearANDStratum_YRsubset.pdf", path="big_csv/plots")
## shannon
ggplot(psu_div_forPlot_YRsubset, aes(x=YEAR, y=richness_mean, colour=STRAT, group=STRAT)) + geom_point() + geom_line() + geom_errorbar(aes(ymin=richness_mean-richness_se, ymax=richness_mean+richness_se), width=.2)
ggsave(file="fk_MeanShannon_ByYearANDStratum_YRsubset.pdf", path="big_csv/plots")

### although boxplots don't work well for biomass and abundance, they look great for diversity metrics
# richness
# by STRAT
ggplot(data = psu_div) + geom_boxplot(mapping = aes(x = STRAT, y = richness)) + facet_wrap(~YEAR, scales="free_y")
ggsave(file="fk_Richness_boxplot_ByStratumForEachYear.pdf", path="big_csv/plots")
# by YEAR
ggplot(data = psu_div) + geom_boxplot(mapping = aes(x = YEAR, y = richness, group = YEAR)) + facet_wrap(~STRAT, scales="free_y")
ggsave(file="fk_Richness_boxplot_ByYearForEachStratum.pdf", path="big_csv/plots")
# simpson
# by STRAT
ggplot(data = psu_div) + geom_boxplot(mapping = aes(x = STRAT, y = simpson)) + facet_wrap(~YEAR, scales="free_y")
ggsave(file="fk_Simpson_boxplot_ByStratumForEachYear.pdf", path="big_csv/plots")
# by YEAR
ggplot(data = psu_div) + geom_boxplot(mapping = aes(x = YEAR, y = simpson, group = YEAR)) + facet_wrap(~STRAT, scales="free_y")
ggsave(file="fk_Simpson_boxplot_ByYearForEachStratum.pdf", path="big_csv/plots")
# shannon
# by STRAT
ggplot(data = psu_div) + geom_boxplot(mapping = aes(x = STRAT, y = shannon)) + facet_wrap(~YEAR, scales="free_y")
ggsave(file="fk_Shannon_boxplot_ByStratumForEachYear.pdf", path="big_csv/plots")
# by YEAR
ggplot(data = psu_div) + geom_boxplot(mapping = aes(x = YEAR, y = shannon, group = YEAR)) + facet_wrap(~STRAT, scales="free_y")
ggsave(file="fk_Shannon_boxplot_ByYearForEachStratum.pdf", path="big_csv/plots")

# do this for dataframe with subset of years as well
# richness
# by STRAT
ggplot(data = psu_div_YRsubset) + geom_boxplot(mapping = aes(x = STRAT, y = richness)) + facet_wrap(~YEAR, scales="free_y")
ggsave(file="fk_Richness_boxplot_ByStratumForEachYear_YRsubset.pdf", path="big_csv/plots")
# by YEAR
ggplot(data = psu_div_YRsubset) + geom_boxplot(mapping = aes(x = YEAR, y = richness, group = YEAR)) + facet_wrap(~STRAT, scales="free_y")
ggsave(file="fk_Richness_boxplot_ByYearForEachStratum_YRsubset.pdf", path="big_csv/plots")
# simpson
# by STRAT
ggplot(data = psu_div_YRsubset) + geom_boxplot(mapping = aes(x = STRAT, y = simpson)) + facet_wrap(~YEAR, scales="free_y")
ggsave(file="fk_Simpson_boxplot_ByStratumForEachYear_YRsubset.pdf", path="big_csv/plots")
# by YEAR
ggplot(data = psu_div_YRsubset) + geom_boxplot(mapping = aes(x = YEAR, y = simpson, group = YEAR)) + facet_wrap(~STRAT, scales="free_y")
ggsave(file="fk_Simpson_boxplot_ByYearForEachStratum_YRsubset.pdf", path="big_csv/plots")
# shannon
# by STRAT
ggplot(data = psu_div_YRsubset) + geom_boxplot(mapping = aes(x = STRAT, y = shannon)) + facet_wrap(~YEAR, scales="free_y")
ggsave(file="fk_Shannon_boxplot_ByStratumForEachYear_YRsubset.pdf", path="big_csv/plots")
# by YEAR
ggplot(data = psu_div_YRsubset) + geom_boxplot(mapping = aes(x = YEAR, y = shannon, group = YEAR)) + facet_wrap(~STRAT, scales="free_y")
ggsave(file="fk_Shannon_boxplot_ByYearForEachStratum_YRsubset.pdf", path="big_csv/plots")

```

# Statistical analysis 

For continuous response variable (diversity - richness, shannon, simpson, functional) and categorical predictor variable (strata)

```{r libraries}

library(MASS) # contains boxcox function
library(moments) # contains skewness function
library(DescTools)
library(rpsychi) #statistical tests using summary data
library(AICcmodavg) #AIC
library(ade4) #mantel.rtest
library(ggplot2) #geom_boxplot and other plots
library(nortest) #checks for normality 
library(psych) #pairs.panels 
library(tidyverse)
library(magrittr) # needed to call this library separately from one in tidyverse in order to get functionality of github version not found in CRAN version

```

```{r stat testing with abundance data}

# 1. Read in data and remove years (optional)
psu_abun <- read_csv('big_csv/abundance_psu/psu_fk_abundance_no_spp_merged.csv', 
                       col_types = cols(
                          YEAR = col_integer(),
                          REGION = col_character(),
                          STRAT = col_character(),
                          PROT = col_integer(),
                          PRIMARY_SAMPLE_UNIT = col_character(),
                          SPECIES_CD = col_character(),
                          m = col_integer(),
                          var = col_double(),
                          abundance = col_double(),
                          protected_status = col_character())) # 1,553,704  x 10

# the psu_abun dataframe has an abundance for each species for each psu so sum abundance across species for each psu...skip this step for diversity indices
psu_abun_sum <- psu_abun %>%
  group_by(YEAR, PRIMARY_SAMPLE_UNIT, STRAT) %>%
  summarize(abundance_sum = sum(abundance)) # 4839 x 4 (5241 = total # of PSUs sampled from 1999 to 2016)
  
# (optional) the psu_bio_sum_YRsubset dataframe produced with the code below has years with low sample sizes (few PSUs)/missing strata removed (i.e., 1999, 2000, and 2004)
psu_abun_sum_YRsubset <- psu_abun_sum %>%
  filter(YEAR!=1999,YEAR!=2000,YEAR!=2004) # 4323 x 4


# 2. Run ANOVA to get residuals so we can use them to test assumptions of normality and equal variances
psu_abun_sum_YRsubset$fYEAR <- as.factor(psu_abun_sum_YRsubset$YEAR) # categorical variables are factors
psu_abun_sum_YRsubset$fSTRAT <- as.factor(psu_abun_sum_YRsubset$STRAT) # categorical variables are factors
abundance_aov <- aov(abundance_sum~fYEAR*fSTRAT, data=psu_abun_sum_YRsubset) # since there might be an interaction between YEAR and STRAT (i.e., biomass may not change through time consistently across strata or vice versa) use "*" in your model to include an interaction term
psu_abun_sum_YRsubset$resids <- residuals(abundance_aov) # get residuals
psu_abun_sum_YRsubset$pred_val <- predict(abundance_aov) # get predicted values

# 2a. Examine plots of residuals to qualitatively check for normality and equal variances
ggplot() + geom_density(aes(x=psu_abun_sum_YRsubset$resids)) # look at distribution of residuals
  ### distribution looks terrible....few very high values and many very low values
plot(abundance_aov) 
  # plot 1: Residuals vs Fitted values - data points should be distibuted evenly above and below the line and across the plot...if they are not, this indicates a violation of assumptions
  # plot 2: Q-Q plot - data points\ should fall more or less along dashed line...if they do not, this indicates a violation of assumptions
  ### plots look terribles....few very high values and many very low values

# 2b. Shapiro-Wilk test - p-value < 0.05 means the data are NOT normally distributed
psu_abun_sum_YRsubset %$% shapiro.test(resids) 
  ### W = 0.17422, p-value < 2.2e-16; non-normal distribution

# 2c. Test for skew - greater than zero is evidence for having skew; test with t-test - p-value < 0.05 means data are skewed
skewness(psu_abun_sum_YRsubset$resids)
t<- skewness(psu_abun_sum_YRsubset$resids) %>% divide_by(sqrt(6/length(psu_abun_sum_YRsubset$resids))) # t-value
1-pt(t,length(psu_abun_sum_YRsubset$resids)) # probability of the t-value by chance alone when skew is zero; pt=cummulative density function of t-dist, first number=calculated t-value; second=df
  ### skew = 44.67215 and p = 0; significantly skewed

# 2d. Bartlett test - p-value < 0.05 means data do not have equal variances across groups
bartlett.test(psu_abun_sum_YRsubset$resids~psu_abun_sum_YRsubset$fYEAR) # Bartlett's K-squared = 7315.7, df = 12, p-value < 2.2e-16
bartlett.test(psu_abun_sum_YRsubset$resids~psu_abun_sum_YRsubset$fSTRAT) # Bartlett's K-squared = 5467, df = 6, p-value < 2.2e-16
  ### significant different variances across both YEAR groups and STRAT groups


# 3. Transform data using Boxcox method 
psu_abun_sum_YRsubset %>%
  boxcox(abundance_sum~fYEAR, data = .) %$%
  x[which.max(y)] #x is lambda and y is values of likelihood
# x = 0.1010101; this output is the lambda value that you should use to transform data
psu_abun_sum_YRsubset %>%
  boxcox(abundance_sum~fSTRAT, data = .) %$%
  x[which.max(y)] #x is lambda and y is values of likelihood
# x = 0.1010101
psu_abun_sum_YRsubset <- psu_abun_sum_YRsubset %>% mutate(abundance_sum_bc = (abundance_sum %>%
  raise_to_power(.1) %>% subtract(1) %>% divide_by(.1))) # use x/lambda in raise_to_power() and divide_by()

# 4. Retest assumptions using transformed data
abundance_aov <- aov(abundance_sum_bc~fYEAR*fSTRAT, data=psu_abun_sum_YRsubset) # since there might be an interaction between YEAR and STRAT (i.e., biomass may not change through time consistently across strata or vice versa) use "*" in your model to include an interaction term
psu_abun_sum_YRsubset$resids_bc <- residuals(abundance_aov) # get residuals
psu_abun_sum_YRsubset$pred_va_bcl <- predict(abundance_aov) # get predicted values

# 4a. Examine plots of residuals to qualitatively check for normality and equal variances
ggplot() + geom_density(aes(x=psu_abun_sum_YRsubset$resids_bc)) # look at distribution of residuals
  ### better looking distribution, but kind of skinny/not much spread around 0
plot(abundance_aov) 
  # plot 1: Residuals vs Fitted values - data points should be distibuted evenly above and below the line and across the plot...if they are not, this indicates a violation of assumptions
  # plot 2: Q-Q plot - data points should fall more or less along dashed line...if they do not, this indicates a violation of assumptions
  ### still don't look great

# 4b. Shapiro-Wilk test - p-value < 0.05 means the data are NOT normally distributed
psu_abun_sum_YRsubset %$% shapiro.test(resids_bc) 
  ### W = 0.95965, p-value < 2.2e-16; still not normal

# 4c. Test for skew - greater than zero is evidence for having skew; test with t-test - p-value < 0.05 means data are skewed
skewness(psu_abun_sum_YRsubset$resids_bc)
t<- skewness(psu_abun_sum_YRsubset$resids_bc) %>% divide_by(sqrt(6/length(psu_abun_sum_YRsubset$resids_bc))) # t-value
1-pt(t,length(psu_abun_sum_YRsubset$resids_bc)) # probability of the t-value by chance alone when skew is zero; pt=cummulative density function of t-dist, first number=calculated t-value; second=df
  ### skew = 0.4274009 and p = 0; still skewed?

# 4d. Bartlett test - p-value < 0.05 means data do not have equal variances across groups
bartlett.test(psu_abun_sum_YRsubset$resids_bc~psu_abun_sum_YRsubset$fYEAR) # Bartlett's K-squared = 241.18, df = 12, p-value < 2.2e-16
bartlett.test(psu_abun_sum_YRsubset$resids_bc~psu_abun_sum_YRsubset$fSTRAT) # Bartlett's K-squared = 263.38, df = 6, p-value < 2.2e-16
  ### still have unequal variancess


### If data conform to assumptions proceed with parametric test (e.g., ANOVA, t-test); if data do not conform to assumptions proceed with non-parametric test (e.g., Kruksal-Wallis, Mann-Whitney U (aka Wilcoxon Rank Sum))


# 5. 2-way ANOVA
psu_abun_sum_YRsubset$fYEAR <- as.factor(psu_abun_sum_YRsubset$YEAR)
psu_abun_sum_YRsubset$fSTRAT <- as.factor(psu_abun_sum_YRsubset$STRAT)
abundance_aov <- aov(abundance_sum_bc~fYEAR*fSTRAT, data=psu_abun_sum_YRsubset) #since there might be an interaction between YEAR and STRAT (i.e., biomass may not change through time consistently across strata or vice versa) use "*" in your model to include an interaction term
summary(abundance_aov) #displays Type I ANOVA table
drop1(abundance_aov,~.,test="F") #displays Type III ANOVA table (default of most stats packages...use this one)
#Single term deletions

#Model:
#abundance_sum_bc ~ fYEAR * fSTRAT
#             Df Sum of Sq    RSS    AIC F value    Pr(>F)    
#<none>                    6682.7 1819.9                      
#fYEAR        12    50.055 6732.7 1831.2  2.8925 0.0005387 ***
#fSTRAT        6   186.704 6869.4 1938.1 21.5779 < 2.2e-16 ***
#fYEAR:fSTRAT 72   251.339 6934.0 1850.4  2.4207 3.283e-10 ***
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

# 5a. If 2-way ANOVA is significant, do a posteriori post-hoc pairwise comparisons
Tuk_fYEAR <- TukeyHSD(abundance_aov, "fYEAR", ordered=T)
Tuk_fSTRAT <- TukeyHSD(abundance_aov, "fSTRAT", ordered=T)
Tuk_fYEARfSTRAT <- TukeyHSD(abundance_aov, "fYEAR:fSTRAT", ordered=T)
Tuk <- as.data.frame(rbind(Tuk_fYEAR$fYEAR,Tuk_fSTRAT$fSTRAT,Tuk_fYEARfSTRAT$`fYEAR:fSTRAT`))
write.csv(Tuk, "big_csv/plots/abundance_2wayANOVA_TukeyHSDvalues.csv") 

```

```{r stat testing with biomass data}

# 1. Read in data and remove years (optional)
psu_bio <- read_csv('big_csv/biomass_psu/psu_fk_biomass_no_spp_merged.csv', 
                       col_types = cols(
                          YEAR = col_integer(),
                          REGION = col_character(),
                          STRAT = col_character(),
                          PROT = col_integer(),
                          PRIMARY_SAMPLE_UNIT = col_character(),
                          SPECIES_CD = col_character(),
                          m = col_integer(),
                          var = col_double(),
                          biomass = col_double(),
                          protected_status = col_character())) # 1,609,672  x 10

# the psu_bio dataframe has a biomass for each species for each psu so sum biomass across species for each psu...skip this step for diversity indices
psu_bio_sum <- psu_bio %>%
  group_by(YEAR, PRIMARY_SAMPLE_UNIT, STRAT) %>%
  summarize(biomass_sum = sum(biomass)) # 5241 x 4 (5241 = total # of PSUs sampled from 1999 to 2016)
  
# (optional) the psu_bio_sum_YRsubset dataframe produced with the code below has years with low sample sizes (few PSUs)/missing strata removed (i.e., 1999, 2000, and 2004)
psu_bio_sum_YRsubset <- psu_bio_sum %>%
  filter(YEAR!=1999,YEAR!=2000,YEAR!=2004) # 4725 x 4

# (optional) tried removing two points with very high biomass values (at least an order of magnitude higher than all other values)....data still violated assumptions even after transformation
#psu_bio_sum_YRsubset <- psu_bio_sum %>%
#  filter(YEAR!=1999,YEAR!=2000,YEAR!=2004) %>%
#  filter(biomass_sum<1000)


# 2. Run ANOVA to get residuals so we can use them to test assumptions of normality and equal variances
psu_bio_sum_YRsubset$fYEAR <- as.factor(psu_bio_sum_YRsubset$YEAR) # categorical variables are factors
psu_bio_sum_YRsubset$fSTRAT <- as.factor(psu_bio_sum_YRsubset$STRAT) # categorical variables are factors
biomass_aov <- aov(biomass_sum~fYEAR*fSTRAT, data=psu_bio_sum_YRsubset) # since there might be an interaction between YEAR and STRAT (i.e., biomass may not change through time consistently across strata or vice versa) use "*" in your model to include an interaction term
psu_bio_sum_YRsubset$resids <- residuals(biomass_aov) # get residuals
psu_bio_sum_YRsubset$pred_val <- predict(biomass_aov) # get predicted values

# 2a. Examine plots of residuals to qualitatively check for normality and equal variances
ggplot() + geom_density(aes(x=psu_bio_sum_YRsubset$resids)) # look at distribution of residuals
  ### distribution of our residuals does not look good...few very large values and many very small values
plot(biomass_aov) 
  # plot 1: Residuals vs Fitted values - data points should be distibuted evenly above and below the line and across the plot...if they are not, this indicates a violation of assumptions
  # plot 2: Q-Q plot - data points\ should fall more or less along dashed line...if they do not, this indicates a violation of assumptions
  ### two data points are way out there and possibly causing violation of assumptions (i.e., 524, 26, less so with 1080)

# 2b. Shapiro-Wilk test - p-value < 0.05 means the data are NOT normally distributed
psu_bio_sum_YRsubset %$% shapiro.test(resids) 
  ### W = 0.2177, p-value < 2.2e-16; our residuals are NOT normally distributed

# 2c. Test for skew - greater than zero is evidence for having skew; test with t-test - p-value < 0.05 means data are skewed
skewness(psu_bio_sum_YRsubset$resids)
t<- skewness(psu_bio_sum_YRsubset$resids) %>% divide_by(sqrt(6/length(psu_bio_sum_YRsubset$resids))) # t-value
1-pt(t,length(psu_bio_sum_YRsubset$resids)) # probability of the t-value by chance alone when skew is zero; pt=cummulative density function of t-dist, first number=calculated t-value; second=df
  ### skew = 34.50221 and p = 0; our residuals are skewed

# 2d. Bartlett test - p-value < 0.05 means data do not have equal variances across groups
bartlett.test(psu_bio_sum_YRsubset$resids~psu_bio_sum_YRsubset$fYEAR) # Bartlett's K-squared = 6471, df = 12, p-value < 2.2e-16
bartlett.test(psu_bio_sum_YRsubset$resids~psu_bio_sum_YRsubset$fSTRAT) # Bartlett's K-squared = 2625.7, df = 6, p-value < 2.2e-16
  ### our variances are unequal across both YEAR and STRAT


# 3. Transform data using Boxcox method 
psu_bio_sum_YRsubset %>%
  boxcox(biomass_sum~fYEAR, data = .) %$%
  x[which.max(y)] #x is lambda and y is values of likelihood
# x = 0.1010101; this output is the lambda value that you should use to transform data
psu_bio_sum_YRsubset %>%
  boxcox(biomass_sum~fSTRAT, data = .) %$%
  x[which.max(y)] #x is lambda and y is values of likelihood
# x = 0.1010101
psu_bio_sum_YRsubset <- psu_bio_sum_YRsubset %>% mutate(biomass_sum_bc = (biomass_sum %>%
  raise_to_power(.1) %>% subtract(1) %>% divide_by(.1))) # use x/lambda in raise_to_power() and divide_by()

# 3a. Try sqrt transformation
psu_bio_sum_YRsubset$biomass_sum_sqrt = sqrt(psu_bio_sum_YRsubset$biomass_sum)
  ### this transformation did even worse than the boxcox


# 4. Retest assumptions using transformed data
biomass_aov <- aov(biomass_sum_bc~fYEAR*fSTRAT, data=psu_bio_sum_YRsubset) # since there might be an interaction between YEAR and STRAT (i.e., biomass may not change through time consistently across strata or vice versa) use "*" in your model to include an interaction term
psu_bio_sum_YRsubset$resids_bc <- residuals(biomass_aov) # get residuals
psu_bio_sum_YRsubset$pred_va_bcl <- predict(biomass_aov) # get predicted values

# 4a. Examine plots of residuals to qualitatively check for normality and equal variances
ggplot() + geom_density(aes(x=psu_bio_sum_YRsubset$resids_bc)) # look at distribution of residuals
  ### distribution of our residuals looks better but still not much spread around 0
plot(biomass_aov) 
  # plot 1: Residuals vs Fitted values - data points should be distibuted evenly above and below the line and across the plot...if they are not, this indicates a violation of assumptions
  # plot 2: Q-Q plot - data points should fall more or less along dashed line...if they do not, this indicates a violation of assumptions
  ### better but not great; data point 524 still is pretty far out there

# 4b. Shapiro-Wilk test - p-value < 0.05 means the data are NOT normally distributed
psu_bio_sum_YRsubset %$% shapiro.test(resids_bc) 
  ### W = 0.98867, p-value < 2.2e-16; our residuals are still NOT normally distributed

# 4c. Test for skew - greater than zero is evidence for having skew; test with t-test - p-value < 0.05 means data are skewed
skewness(psu_bio_sum_YRsubset$resids_bc)
t<- skewness(psu_bio_sum_YRsubset$resids_bc) %>% divide_by(sqrt(6/length(psu_bio_sum_YRsubset$resids_bc))) # t-value
1-pt(t,length(psu_bio_sum_YRsubset$resids_bc)) # probability of the t-value by chance alone when skew is zero; pt=cummulative density function of t-dist, first number=calculated t-value; second=df
  ### skew = 0.06864448 and p = 0.02706148; our residuals are still slightly skewed but look way better

# 4d. Bartlett test - p-value < 0.05 means data do not have equal variances across groups
bartlett.test(psu_bio_sum_YRsubset$resids_bc~psu_bio_sum_YRsubset$fYEAR) # Bartlett's K-squared = 181.45, df = 12, p-value < 2.2e-16
bartlett.test(psu_bio_sum_YRsubset$resids_bc~psu_bio_sum_YRsubset$fSTRAT) # Bartlett's K-squared = 149.52, df = 6, p-value < 2.2e-16
  ### our variances are still quite unequal across both YEAR and STRAT


### If data conform to assumptions proceed with parametric test (e.g., ANOVA, t-test); if data do not conform to assumptions proceed with non-parametric test (e.g., Kruksal-Wallis, Mann-Whitney U (aka Wilcoxon Rank Sum))


# 5. 2-way ANOVA
psu_bio_sum_YRsubset$fYEAR <- as.factor(psu_bio_sum_YRsubset$YEAR)
psu_bio_sum_YRsubset$fSTRAT <- as.factor(psu_bio_sum_YRsubset$STRAT)
biomass_aov <- aov(biomass_sum_bc~fYEAR*fSTRAT, data=psu_bio_sum_YRsubset) #since there might be an interaction between YEAR and STRAT (i.e., biomass may not change through time consistently across strata or vice versa) use "*" in your model to include an interaction term
summary(biomass_aov) #displays Type I ANOVA table
drop1(biomass_aov,~.,test="F") #displays Type III ANOVA table (default of most stats packages...use this one)
#Single term deletions

#Model:
#biomass_sum_bc ~ fYEAR * fSTRAT
#             Df Sum of Sq    RSS    AIC F value    Pr(>F)    
#<none>                    8166.8 2767.6                      
#fYEAR        12    30.060 8196.8 2760.9  1.4214     0.148    
#fSTRAT        6    97.877 8264.7 2811.8  9.2562 4.164e-10 ***
#fYEAR:fSTRAT 72   246.562 8413.3 2764.1  1.9431 3.792e-06 ***
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

# 5a. If 2-way ANOVA is significant, do a posteriori post-hoc pairwise comparisons
Tuk_fYEAR <- TukeyHSD(biomass_aov, "fYEAR", ordered=T)
Tuk_fSTRAT <- TukeyHSD(biomass_aov, "fSTRAT", ordered=T)
Tuk_fYEARfSTRAT <- TukeyHSD(biomass_aov, "fYEAR:fSTRAT", ordered=T)
Tuk <- as.data.frame(rbind(Tuk_fYEAR$fYEAR,Tuk_fSTRAT$fSTRAT,Tuk_fYEARfSTRAT$`fYEAR:fSTRAT`))
write.csv(Tuk, "big_csv/plots/biomass_2wayANOVA_TukeyHSDvalues.csv") 

```


```{r stat testing with diversity data}

# 1. Read in data and remove years (optional)
psu_diversity = read_csv("big_csv/diversity_without_spe/psu_diversity_no_spp_merged.csv",
                      col_types = cols(
                          YEAR = col_integer(),
                          PRIMARY_SAMPLE_UNIT = col_character(),
                          richness = col_integer(),
                          simpson = col_double(),
                          shannon = col_double(),
                          STRAT = col_character(),
                          SPECIES_CD = col_character(),
                          abun_merged = col_double())) # 1,689,982 x 9

# the psu_diversity dataframe has a value for each diversity metric that is repeated for each YEAR/PSU for each species record....need to remove duplicate rows and can also select columns of interest for the analysis of diversity metrics
psu_div = psu_diversity %>% 
  select(YEAR, PRIMARY_SAMPLE_UNIT, STRAT, richness, simpson, shannon) %>%
  group_by(YEAR, PRIMARY_SAMPLE_UNIT, STRAT) %>%
  summarize(
    richness = first(richness),
    simpson = first(simpson),
    shannon = first(shannon)) # 5241 x 6
  
# (optional) the psu_div_YRsubset dataframe produced with the code below has years with low sample sizes (few PSUs)/missing strata removed (i.e., 1999, 2000, and 2004)
psu_div_YRsubset <- psu_div %>%
  filter(YEAR!=1999,YEAR!=2000,YEAR!=2004) # 4725 x 6

# 2. Run ANOVA to get residuals so we can use them to test assumptions of normality and equal variances
psu_div_YRsubset$fYEAR <- as.factor(psu_div_YRsubset$YEAR) # categorical variables are factors
psu_div_YRsubset$fSTRAT <- as.factor(psu_div_YRsubset$STRAT) # categorical variables are factors
# richness
richness_aov <- aov(richness~fYEAR*fSTRAT, data=psu_div_YRsubset) # since there might be an interaction between YEAR and STRAT (i.e., diversity may not change through time consistently across strata or vice versa) use "*" in your model to include an interaction term
psu_div_YRsubset$resids_rich <- residuals(richness_aov) # get residuals
psu_div_YRsubset$pred_val_rich <- predict(richness_aov) # get predicted values
# simpson
simpson_aov <- aov(simpson~fYEAR*fSTRAT, data=psu_div_YRsubset) # since there might be an interaction between YEAR and STRAT (i.e., diversity may not change through time consistently across strata or vice versa) use "*" in your model to include an interaction term
psu_div_YRsubset$resids_simp <- residuals(simpson_aov) # get residuals
psu_div_YRsubset$pred_val_simp <- predict(simpson_aov) # get predicted values
# shannon
shannon_aov <- aov(shannon~fYEAR*fSTRAT, data=psu_div_YRsubset) # since there might be an interaction between YEAR and STRAT (i.e., diversity may not change through time consistently across strata or vice versa) use "*" in your model to include an interaction term
psu_div_YRsubset$resids_shan <- residuals(shannon_aov) # get residuals
psu_div_YRsubset$pred_val_shan <- predict(shannon_aov) # get predicted values

# 2a. Examine plots of residuals to qualitatively check for normality and equal variances
#richness
ggplot() + geom_density(aes(x=psu_div_YRsubset$resids_rich)) # look at distribution of residuals
  ### doesn't look bad
plot(richness_aov) 
  # plot 1: Residuals vs Fitted values - data points should be distibuted evenly above and below the line and across the plot...if they are not, this indicates a violation of assumptions
  # plot 2: Q-Q plot - data points\ should fall more or less along dashed line...if they do not, this indicates a violation of assumptions
  ### these plots also don't look bad
#simpson
ggplot() + geom_density(aes(x=psu_div_YRsubset$resids_simp)) # look at distribution of residuals
  ### distribution of residuals looks ok
plot(simpson_aov) 
  # plot 1: Residuals vs Fitted values - data points should be distibuted evenly above and below the line and across the plot...if they are not, this indicates a violation of assumptions
  # plot 2: Q-Q plot - data points\ should fall more or less along dashed line...if they do not, this indicates a violation of assumptions
  ### residuals vs fitted values plot doesn't look great - spread across line increases to the right of plot
#shannon
ggplot() + geom_density(aes(x=psu_div_YRsubset$resids_shan)) # look at distribution of residuals
  ### distribution looks ok 
plot(shannon_aov) 
  # plot 1: Residuals vs Fitted values - data points should be distibuted evenly above and below the line and across the plot...if they are not, this indicates a violation of assumptions
  # plot 2: Q-Q plot - data points\ should fall more or less along dashed line...if they do not, this indicates a violation of assumptions
  ### residuals vs fitted values plot doesn't look great - spread across line seems to increase to the right of plot

# 2b. Shapiro-Wilk test - p-value < 0.05 means the data are NOT normally distributed
#richness
psu_div_YRsubset %$% shapiro.test(resids_rich) 
  ### W = 0.99612, p-value = 8.243e-10; this test result says distribution of residuals is significantly different from normal 
#simpson
psu_div_YRsubset %$% shapiro.test(resids_simp) 
  ### W = 0.99557, p-value = 8.865e-11; non-normal distribution
#shannon
psu_div_YRsubset %$% shapiro.test(resids_shan) 
  ### W = 0.99978, p-value = 0.928; normal distribution without transformation!


# 2c. Test for skew - greater than zero is evidence for having skew; test with t-test - p-value < 0.05 means data are skewed
# richness
skewness(psu_div_YRsubset$resids_rich)
t<- skewness(psu_div_YRsubset$resids_rich) %>% divide_by(sqrt(6/length(psu_div_YRsubset$resids_rich))) # t-value
1-pt(t,length(psu_div_YRsubset$resids_rich)) # probability of the t-value by chance alone when skew is zero; pt=cummulative density function of t-dist, first number=calculated t-value; second=df
  ### skew = -0.2038638 and p = 1; distribution of residuals is NOT significantly skewed
# simpson
skewness(psu_div_YRsubset$resids_simp)
t<- skewness(psu_div_YRsubset$resids_simp) %>% divide_by(sqrt(6/length(psu_div_YRsubset$resids_simp))) # t-value
1-pt(t,length(psu_div_YRsubset$resids_simp)) # probability of the t-value by chance alone when skew is zero; pt=cummulative density function of t-dist, first number=calculated t-value; second=df
  ### skew = 0.2653317 and p = 5.684342e-14; significant skew in distribution
# shannon
skewness(psu_div_YRsubset$resids_shan)
t<- skewness(psu_div_YRsubset$resids_shan) %>% divide_by(sqrt(6/length(psu_div_YRsubset$resids_shan))) # t-value
1-pt(t,length(psu_div_YRsubset$resids_shan)) # probability of the t-value by chance alone when skew is zero; pt=cummulative density function of t-dist, first number=calculated t-value; second=df
  ### skew = 0.004875036 and p = 0.4455952; no significant skew

# 2d. Bartlett test - p-value < 0.05 means data do not have equal variances across groups
# richness
bartlett.test(psu_div_YRsubset$resids_rich~psu_div_YRsubset$fYEAR) # Bartlett's K-squared = 81.852, df = 12, p-value = 1.828e-12
bartlett.test(psu_div_YRsubset$resids_rich~psu_div_YRsubset$fSTRAT) # Bartlett's K-squared = 70.023, df = 6, p-value = 4.044e-13
  ### variances of residuals are significantly different among groups for both YEAR and STRAT
# simpson
bartlett.test(psu_div_YRsubset$resids_simp~psu_div_YRsubset$fYEAR) # Bartlett's K-squared = 32.026, df = 12, p-value = 0.001371
bartlett.test(psu_div_YRsubset$resids_simp~psu_div_YRsubset$fSTRAT) # Bartlett's K-squared = 63.763, df = 6, p-value = 7.713e-12
  ### variances of residuals are significantly different among groups for both YEAR and STRAT, but more so for STRAT
# shannon
bartlett.test(psu_div_YRsubset$resids_shan~psu_div_YRsubset$fYEAR) # Bartlett's K-squared = 49.285, df = 12, p-value = 1.865e-06
bartlett.test(psu_div_YRsubset$resids_shan~psu_div_YRsubset$fSTRAT) # Bartlett's K-squared = 45.932, df = 6, p-value = 3.054e-08
  ### variances of residuals are significantly different among groups for both YEAR and STRAT


# 3. Transform data using Boxcox method 
# richness
psu_div_YRsubset %>%
  boxcox(richness~fYEAR, data = .) %$%
  x[which.max(y)] #x is lambda and y is values of likelihood
# x = 1.232323; this output is the lambda value that you should use to transform data
psu_div_YRsubset %>%
  boxcox(richness~fSTRAT, data = .) %$%
  x[which.max(y)] #x is lambda and y is values of likelihood
# x = 1.313131
psu_div_YRsubset <- psu_div_YRsubset %>% mutate(richness_bc = (richness %>%
  raise_to_power(1.27) %>% subtract(1) %>% divide_by(1.27))) # since x for YEAR and STRAT were so similar, I used something in between; use x/lambda in raise_to_power() and divide_by()
# simpson
psu_div_YRsubset %>%
  boxcox(simpson~fYEAR, data = .) %$%
  x[which.max(y)] #x is lambda and y is values of likelihood
# x = 0.6666667; this output is the lambda value that you should use to transform data
psu_div_YRsubset %>%
  boxcox(simpson~fSTRAT, data = .) %$%
  x[which.max(y)] #x is lambda and y is values of likelihood
# x = 0.6666667
psu_div_YRsubset <- psu_div_YRsubset %>% mutate(simpson_bc = (simpson %>%
  raise_to_power(.7) %>% subtract(1) %>% divide_by(.7))) # use x/lambda in raise_to_power() and divide_by()
# shannon
psu_div_YRsubset %>%
  boxcox(shannon~fYEAR, data = .) %$%
  x[which.max(y)] #x is lambda and y is values of likelihood
# x = 0.9090909; this output is the lambda value that you should use to transform data
psu_div_YRsubset %>%
  boxcox(shannon~fSTRAT, data = .) %$%
  x[which.max(y)] #x is lambda and y is values of likelihood
# x = 0.9494949
psu_div_YRsubset <- psu_div_YRsubset %>% mutate(shannon_bc = (shannon %>%
  raise_to_power(.9) %>% subtract(1) %>% divide_by(.9))) # use x/lambda in raise_to_power() and divide_by()


# 4. Retest assumptions using transformed data
# richness
richness_aov <- aov(richness_bc~fYEAR*fSTRAT, data=psu_div_YRsubset) # since there might be an interaction between YEAR and STRAT (i.e., diversity may not change through time consistently across strata or vice versa) use "*" in your model to include an interaction term
psu_div_YRsubset$resids_rich_bc <- residuals(richness_aov) # get residuals
psu_div_YRsubset$pred_val_rich_bc <- predict(richness_aov) # get predicted values
# simpson
simpson_aov <- aov(simpson_bc~fYEAR*fSTRAT, data=psu_div_YRsubset) # since there might be an interaction between YEAR and STRAT (i.e., diversity may not change through time consistently across strata or vice versa) use "*" in your model to include an interaction term
psu_div_YRsubset$resids_simp_bc <- residuals(simpson_aov) # get residuals
psu_div_YRsubset$pred_val_simp_bc <- predict(simpson_aov) # get predicted values
# shannon
shannon_aov <- aov(shannon_bc~fYEAR*fSTRAT, data=psu_div_YRsubset) # since there might be an interaction between YEAR and STRAT (i.e., diversity may not change through time consistently across strata or vice versa) use "*" in your model to include an interaction term
psu_div_YRsubset$resids_shan_bc <- residuals(shannon_aov) # get residuals
psu_div_YRsubset$pred_val_shan_bc <- predict(shannon_aov) # get predicted values

# 4a. Examine plots of residuals to qualitatively check for normality and equal variances
# richness
ggplot() + geom_density(aes(x=psu_div_YRsubset$resids_rich_bc)) # look at distribution of residuals
  ### plot still looks ok to me
plot(richness_aov) 
  # plot 1: Residuals vs Fitted values - data points should be distibuted evenly above and below the line and across the plot...if they are not, this indicates a violation of assumptions
  # plot 2: Q-Q plot - data points should fall more or less along dashed line...if they do not, this indicates a violation of assumptions
  ### plots still look ok to me
ggplot(data = psu_div_YRsubset) + geom_boxplot(mapping = aes(x = STRAT, y = richness_bc)) + facet_wrap(~YEAR, scales="free_y")
# simpson
ggplot() + geom_density(aes(x=psu_div_YRsubset$resids_simp_bc)) # look at distribution of residuals
  ### plot looks ok to me 
plot(simpson_aov) 
  # plot 1: Residuals vs Fitted values - data points should be distibuted evenly above and below the line and across the plot...if they are not, this indicates a violation of assumptions
  # plot 2: Q-Q plot - data points should fall more or less along dashed line...if they do not, this indicates a violation of assumptions
  ### still same pattern in plot of residuals vs fitted values
# shannon
ggplot() + geom_density(aes(x=psu_div_YRsubset$resids_shan_bc)) # look at distribution of residuals
  ### looks fine
plot(shannon_aov) 
  # plot 1: Residuals vs Fitted values - data points should be distibuted evenly above and below the line and across the plot...if they are not, this indicates a violation of assumptions
  # plot 2: Q-Q plot - data points should fall more or less along dashed line...if they do not, this indicates a violation of assumptions
  ### still same pattern in plot of residuals vs fitted values

# 4b. Shapiro-Wilk test - p-value < 0.05 means the data are NOT normally distributed
# richness
psu_div_YRsubset %$% shapiro.test(resids_rich_bc) 
  ### W = 0.999, p-value = 0.006514; distribution of residuals is a little closer to normal
# simpson
psu_div_YRsubset %$% shapiro.test(resids_simp_bc) 
  ### W = 0.99936, p-value = 0.09424; woohoo!!!!!!! normal distribution!!!!!
# shannon
psu_div_YRsubset %$% shapiro.test(resids_shan_bc) 
  ### W = 0.99921, p-value = 0.03159; slightly non-normal distribution

# 4c. Test for skew - greater than zero is evidence for having skew; test with t-test - p-value < 0.05 means data are skewed
# richness
skewness(psu_div_YRsubset$resids_rich_bc)
t<- skewness(psu_div_YRsubset$resids_rich_bc) %>% divide_by(sqrt(6/length(psu_div_YRsubset$resids_rich_bc))) # t-value
1-pt(t,length(psu_div_YRsubset$resids_rich_bc)) # probability of the t-value by chance alone when skew is zero; pt=cummulative density function of t-dist, first number=calculated t-value; second=df
  ### skew = 0.01062075 and p = 0.3828413; distribution of residuals is still NOT significantly skewed
# simpson
skewness(psu_div_YRsubset$resids_simp_bc)
t<- skewness(psu_div_YRsubset$resids_simp_bc) %>% divide_by(sqrt(6/length(psu_div_YRsubset$resids_simp_bc))) # t-value
1-pt(t,length(psu_div_YRsubset$resids_simp_bc)) # probability of the t-value by chance alone when skew is zero; pt=cummulative density function of t-dist, first number=calculated t-value; second=df
  ### skew =-0.04192732 and p = 0.8802892; yay! no significant skew!
# shannon
skewness(psu_div_YRsubset$resids_shan_bc)
t<- skewness(psu_div_YRsubset$resids_shan_bc) %>% divide_by(sqrt(6/length(psu_div_YRsubset$resids_shan_bc))) # t-value
1-pt(t,length(psu_div_YRsubset$resids_shan_bc)) # probability of the t-value by chance alone when skew is zero; pt=cummulative density function of t-dist, first number=calculated t-value; second=df
  ### skew = -0.09237946 and p = 0.9952198; no skew

# 4d. Bartlett test - p-value < 0.05 means data do not have equal variances across groups
# richness
bartlett.test(psu_div_YRsubset$resids_rich_bc~psu_div_YRsubset$fYEAR) # Bartlett's K-squared = 64.355, df = 12, p-value = 3.587e-09
bartlett.test(psu_div_YRsubset$resids_rich_bc~psu_div_YRsubset$fSTRAT) # Bartlett's K-squared = 42.823, df = 6, p-value = 1.264e-07
  ### better but variances of residuals still significantly different among groups for both YEAR and STRAT
# simpson
bartlett.test(psu_div_YRsubset$resids_simp_bc~psu_div_YRsubset$fYEAR) # Bartlett's K-squared = 33.481, df = 12, p-value = 0.0008139
bartlett.test(psu_div_YRsubset$resids_simp_bc~psu_div_YRsubset$fSTRAT) # Bartlett's K-squared = 46.817, df = 6, p-value = 2.036e-08
  ### variances still significantly different among groups especially for STRAT
# shannon
bartlett.test(psu_div_YRsubset$resids_shan_bc~psu_div_YRsubset$fYEAR) # Bartlett's K-squared = 51.741, df = 12, p-value = 6.889e-07
bartlett.test(psu_div_YRsubset$resids_shan_bc~psu_div_YRsubset$fSTRAT) # Bartlett's K-squared = 43.221, df = 6, p-value = 1.055e-07
  ### variances still unequal across groups


### If data conform to assumptions proceed with parametric test (e.g., ANOVA, t-test); if data do not conform to assumptions proceed with non-parametric test (e.g., Kruksal-Wallis, Mann-Whitney U (aka Wilcoxon Rank Sum))


# 5. 2-way ANOVA
# richness
psu_div_YRsubset$fYEAR <- as.factor(psu_div_YRsubset$YEAR)
psu_div_YRsubset$fSTRAT <- as.factor(psu_div_YRsubset$STRAT)
richness_aov <- aov(richness_bc~fYEAR*fSTRAT, data=psu_div_YRsubset) #since there might be an interaction between YEAR and STRAT (i.e., biomass may not change through time consistently across strata or vice versa) use "*" in your model to include an interaction term
summary(richness_aov) #displays Type I ANOVA table
drop1(richness_aov,~.,test="F") #displays Type III ANOVA table (default of most stats packages...use this one)
#Single term deletions

#Model:
#richness_bc ~ fYEAR * fSTRAT
#             Df Sum of Sq     RSS   AIC F value    Pr(>F)    
#<none>                    2576671 29956                      
#fYEAR        12     19995 2596666 29969  2.9967 0.0003412 ***
#fSTRAT        6     78540 2655211 30086 23.5417 < 2.2e-16 ***
#fYEAR:fSTRAT 72     86991 2663662 29969  2.1729 5.056e-08 ***
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# simpson
psu_div_YRsubset$fYEAR <- as.factor(psu_div_YRsubset$YEAR)
psu_div_YRsubset$fSTRAT <- as.factor(psu_div_YRsubset$STRAT)
simpson_aov <- aov(simpson_bc~fYEAR*fSTRAT, data=psu_div_YRsubset) #since there might be an interaction between YEAR and STRAT (i.e., biomass may not change through time consistently across strata or vice versa) use "*" in your model to include an interaction term
summary(simpson_aov) #displays Type I ANOVA table
drop1(simpson_aov,~.,test="F") #displays Type III ANOVA table (default of most stats packages...use this one)
#Single term deletions

#Model:
#simpson_bc ~ fYEAR * fSTRAT
#             Df Sum of Sq   RSS    AIC F value    Pr(>F)    
#<none>                    13990 5310.9                      
#fYEAR        12    102.76 14093 5321.5  2.8365 0.0006868 ***
#fSTRAT        6    176.59 14167 5358.2  9.7489 1.067e-10 ***
#fYEAR:fSTRAT 72    401.96 14392 5300.7  1.8492 1.967e-05 ***
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# shannon
psu_div_YRsubset$fYEAR <- as.factor(psu_div_YRsubset$YEAR)
psu_div_YRsubset$fSTRAT <- as.factor(psu_div_YRsubset$STRAT)
shannon_aov <- aov(shannon_bc~fYEAR*fSTRAT, data=psu_div_YRsubset) #since there might be an interaction between YEAR and STRAT (i.e., biomass may not change through time consistently across strata or vice versa) use "*" in your model to include an interaction term
summary(shannon_aov) #displays Type I ANOVA table
drop1(shannon_aov,~.,test="F") #displays Type III ANOVA table (default of most stats packages...use this one)
#Single term deletions

#Model:
#shannon_bc ~ fYEAR * fSTRAT
#             Df Sum of Sq   RSS   AIC F value    Pr(>F)    
#<none>                    53740 11670                      
#fYEAR        12    453.61 54193 11686  3.2596 0.0001055 ***
#fSTRAT        6    820.96 54561 11729 11.7987 3.570e-13 ***
#fYEAR:fSTRAT 72   1519.41 55259 11658  1.8197 3.248e-05 ***
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

# 5a. If 2-way ANOVA is significant, do a posteriori post-hoc pairwise comparisons
Tuk_fYEAR <- TukeyHSD(richness_aov, "fYEAR", ordered=T)
Tuk_fSTRAT <- TukeyHSD(richness_aov, "fSTRAT", ordered=T)
Tuk_fYEARfSTRAT <- TukeyHSD(richness_aov, "fYEAR:fSTRAT", ordered=T)
Tuk <- as.data.frame(rbind(Tuk_fYEAR$fYEAR,Tuk_fSTRAT$fSTRAT,Tuk_fYEARfSTRAT$`fYEAR:fSTRAT`))
write.csv(Tuk, "big_csv/plots/richness_2wayANOVA_TukeyHSDvalues.csv") 

```

Parametric test require the data to meet assumptions of having a normal distribution and homoscedasticity

- Are diversity values and strata related? 

```{r parametric test with psu - do we still need this chunk?}
##########################GET FILE TO READ IN AND THEN MAKE SURE CODE WORKS
#1. Read in data 
psu_diversity = read_csv("big_csv/diversity_without_spe/psu_diversity_no_spp_merged.csv")  
# Parsed with column specification:
# cols(
#   YEAR = col_integer(),
#   PRIMARY_SAMPLE_UNIT = col_character(),
#   richness = col_integer(),
#   simpson = col_double(),
#   shannon = col_double(),
#   STRAT = col_character(),
#   PROT = col_integer(),
#   SPECIES_CD = col_character(),
#   abundance = col_double()
# )
 
psu_div = psu_diversity %>% #A tibble: 5,241 x 7
  dplyr::group_by(YEAR, PRIMARY_SAMPLE_UNIT, STRAT, PROT) %>%
  dplyr::summarize(
    richness = first(richness),
    simpson = first(simpson),
    shannon = first(shannon))
  
#2. Plot data - look at your distribution 
ggplot(data = psu_div, aes(x = STRAT)) +
  geom_point(aes(y = simpson, color = "simpson"), color = "blue") +
  geom_line(aes(y = simpson, color = "simpson"), color = "blue")

ggplot(data = psu_div) +
 (mapping = aes(x = YEAR, y = simpson)) 
# ggplot(data = psu_div, aes(x = YEAR, color = simpson)) + 
#   geom_density()
# 
# hist(psu_div$simpson, main ="")

ggplot(data = psu_div, aes(x = YEAR)) +
  geom_point(aes(y = simpson, color = "simpson"),color = "blue") +
  geom_line(aes(y = simpson, color = "simpson"), color = "blue") +
  facet_grid(STRAT ~ .) 

summary(psu_div$richness)
#   Min.    1st Qu.  Median    Mean    3rd Qu.    Max. 
#   1.00    32.00    39.00     38.64   46.00      73.00
summary(psu_div$shannon)
#   Min.    1st Qu.  Median    Mean    3rd Qu.    Max.
#   1.00    10.27    13.48     13.45   16.55      30.15
summary(psu_div$simpson)
#   Min.    1st Qu.  Median    Mean    3rd Qu.    Max. 
#   1.000   5.804    8.103     8.247   10.465     22.017 

#3. Check for normality using quantile-quantile plot 
qqnorm(psu_div$simpson) 
qqline(psu_div$simpson, lty=2)  # note that lty is the line type (2=medium dashes; try different values betweem 1-6)

#4. Test whether data is NOT normal using Shapiro-Wilk Normaility Test
shapiro.test(psu_div$simpson)
#W = 0.99184, p-value = 4.496e-16      
#p-value is less then 0.5 so the data is NOT normally distributed. 

#5. Check skewness 
skew<-function(x){ 
  m3<-sum((x-mean(x))^3)/length(x) # note that the length expression automatically determines the sample size of your data
  s3<-sqrt(var(x))^3
  m3/s3} 
skew(psu_div$simpson) #0.3071657

#6. Check whether the calculated skew value is significantly different than zero (i.e., not normal) using a t-test. To calulate the t-value, divide the skew by its standard error:
skew(psu_div$simpson)/sqrt(6/length(psu_div$simpson)) 
#8.723182
# note that a rough measure of the standard error of the skewness is "sqrt{6/n}" where n is the sample size

length(psu_div$simpson) #4839

#7. check the probability of this t-value by chance alone when the skew is zero using the following:
1-pt(8.723182, 4838) #pt=cummulative density function of t-dist, first number=calculated t-value; second=df
#0.981466

#8. Check for homoscedasticity: can use var.test, fligner.test, or bartlett.test
bartlett.test(psu_div$simpson~psu_div$YEAR) #between simpson and year 
#Bartlett's K-squared = 38.613, df = 14, p-value = 0.0004185

bartlett.test(psu_div$simpson~psu_div$STRAT) #between simpson and strata
#Bartlett's K-squared = 71.143, df = 6, p-value = 2.382e-13

#9. Transform data using Boxcox method 
#need non-zero data for boxcox method to add a small amount to every value
psu_div$simpson_t = psu_div$simpson+0.01
b = boxcox(psu_div$simpson_t~psu_div$YEAR) 
lamda = b$x #lambda values  
lik = b$y  #y is values of liklihood %>%
bc = cbind(lamda,lik) 
bc[order(-lik),] #use the lamda with the highest likelihood to transform your data #lambda of 0.70707071 is best transformation
psu_div$simpson_boxcox = psu_div$simpson^0.70707071 #simpson_boxcox is now transformed using the boxcox method 
psu_div$simpson_boxcox = psu_div$simpson-1
psu_div$simpson_boxcox = psu_div$simpson/0.70707071

#10. Check for normality of transformed data 
ggplot() + geom_density(aes(x=psu_div$simpson_boxcox))

qqnorm(psu_div$simpson_boxcox) 
qqline(psu_div$simpson_boxcox, lty=2)
shapiro.test(psu_div$simpson_boxcox)
#W = 0.99184, p-value = 4.496e-16
skew(psu_div$simpson_boxcox)
#0.3071657
skew(psu_div$simpson_boxcox)/sqrt(6/length(psu_div$simpson_boxcox)) 
#8.723182
1-pt(8.723182, 4838)
# 0

# if data is normal after transformation, proceed with parametric test 

#10. Mann-Whitney or t-test parametric test 

#11. ANOVA Parametric test 
  ANOVA = aov(psu_div$simpson_boxcox~psu_div$YEAR) 
  #plot(ANOVA)
  summary(ANOVA)
#               Df Sum Sq Mean Sq F value Pr(>F)
#psu_div$YEAR    1     26   25.82   1.155  0.283
#Residuals    4837 108147   22.36    
#??WHY IS DF 1 AND NOT 14?
  
drop1(ANOVA,~.,test="F") # type III SS and F Tests 
#                      Df Sum of Sq    RSS    AIC F value Pr(>F)
#<none>                    108147 15038               
#psu_div$YEAR  1    25.816 108172 15037  1.1546 0.2826

ANOVA = aov(psu_div$simpson_boxcox~psu_div$STRAT) 
  #plot(ANOVA)
  summary(ANOVA)
  
ANOVA = aov(simpson~YEAR, data = psu_div) 
  #plot(ANOVA)
  summary(ANOVA)
  
ANOVA = aov(simpson~STRAT, data = psu_div) 
  #plot(ANOVA)
summary(ANOVA)


# for post-hoc pairwise tests
#TukeyHSD(ANOVA) # where fit comes from aov()


```

```{r psu removed 2004}

#1. Read in data 
psu_diversity = read_csv("big_csv/diversity_without_spe/psu_diversity_no_spp_merged.csv")  
# Parsed with column specification:
# cols(
#   YEAR = col_integer(),
#   PRIMARY_SAMPLE_UNIT = col_character(),
#   richness = col_integer(),
#   simpson = col_double(),
#   shannon = col_double(),
#   STRAT = col_character(),
#   SPECIES_CD = col_character(),
#   abun_merged = col_double()
# )
#wide format 

#2. remove 2004 
psu_div = psu_diversity %>% 
  select(YEAR, PRIMARY_SAMPLE_UNIT, STRAT, richness, simpson, shannon) %>%
  filter(YEAR != "2004") %>%
  group_by(YEAR, PRIMARY_SAMPLE_UNIT, STRAT) %>%
  summarize(
    richness = first(richness),
    simpson = first(simpson),
    shannon = first(shannon))
  
#2. Plot data - look at your distribution 
ggplot(data = psu_div, aes(x = STRAT)) +
  geom_point(aes(y = simpson, color = "simpson"), color = "blue") +
  geom_line(aes(y = simpson, color = "simpson"), color = "blue")

ggplot(data = psu_div) +
 (mapping = aes(x = YEAR, y = simpson)) 

ggplot(data = psu_div, aes(x = YEAR, color = simpson)) + 
   geom_density()
## 
hist(psu_div$simpson, main ="")

ggplot(data = psu_div, aes(x = YEAR)) +
  geom_point(aes(y = simpson, color = "simpson"),color = "blue") +
  geom_line(aes(y = simpson, color = "simpson"), color = "blue") +
  facet_grid(STRAT ~ .) 

summary(psu_div$richness)
#   Min.    1st Qu.  Median    Mean    3rd Qu.    Max. 
#   1.00    32.00    39.00     38.53   46.00      73.00
summary(psu_div$shannon)
#   Min.    1st Qu.  Median    Mean    3rd Qu.    Max.
#   1.00    10.25    13.42     13.41   16.50      30.15
summary(psu_div$simpson)
#   Min.    1st Qu.  Median    Mean    3rd Qu.    Max. 
#   1.000   5.795    8.071     8.225   10.442     22.017 

#3. Check for normality using quantile-quantile plot 
qqnorm(psu_div$simpson) 
qqline(psu_div$simpson, lty=2)  # note that lty is the line type (2=medium dashes; try different values betweem 1-6)

#4. Test whether data is NOT normal using Shapiro-Wilk Normaility Test
shapiro.test(psu_div$simpson)
#W = 0.99165, p-value = 4.895e-16     
#p-value is less then 0.5 so the data is NOT normally distributed. 

#5. Check skewness 
skew<-function(x){ 
  m3<-sum((x-mean(x))^3)/length(x) # note that the length expression automatically determines the sample size of your data
  s3<-sqrt(var(x))^3
  m3/s3} 
skew(psu_div$simpson) #0.3103343

#6. Check whether the calculated skew value is significantly different than zero (i.e., not normal) using a t-test. To calulate the t-value, divide the skew by its standard error:
skew(psu_div$simpson)/sqrt(6/length(psu_div$simpson)) 
#8.696745
# note that a rough measure of the standard error of the skewness is "sqrt{6/n}" where n is the sample size

length(psu_div$simpson) #4712

#7. check the probability of this t-value by chance alone when the skew is zero using the following:
1-pt(8.696745, 4711) #pt=cummulative density function of t-dist, first number=calculated t-value; second=df
#0

#8. Check for homoscedasticity: can use var.test, fligner.test, or bartlett.test
bartlett.test(psu_div$simpson~psu_div$YEAR) #between simpson and year 
#Bartlett's K-squared = 38.364, df = 13, p-value = 0.0002525

bartlett.test(psu_div$simpson~psu_div$STRAT) #between simpson and strata
#Bartlett's K-squared = 71.222, df = 6, p-value = 2.295e-1

#9. Transform data using Boxcox method 
#need non-zero data for boxcox method to add a small amount to every value
psu_div$simpson_t = psu_div$simpson+0.01
b = boxcox(psu_div$simpson_t~psu_div$YEAR) 
lamda = b$x #lambda values  
lik = b$y  #y is values of liklihood %>%
bc = cbind(lamda,lik) 
bc[order(-lik),] #use the lamda with the highest likelihood to transform your data #lambda of 0.70707071 is best transformation
psu_div$simpson_boxcox = psu_div$simpson^0.70707071 #simpson_boxcox is now transformed using the boxcox method 
psu_div$simpson_boxcox = psu_div$simpson-1
psu_div$simpson_boxcox = psu_div$simpson/0.70707071

#10. Check for normality of transformed data 
ggplot() + geom_density(aes(x=psu_div$simpson_boxcox))

qqnorm(psu_div$simpson_boxcox) 
qqline(psu_div$simpson_boxcox, lty=2)
shapiro.test(psu_div$simpson_boxcox)
#W = 0.99165, p-value = 4.895e-16
skew(psu_div$simpson_boxcox)
#0.3103343
skew(psu_div$simpson_boxcox)/sqrt(6/length(psu_div$simpson_boxcox)) 
#8.696745
1-pt(8.723182, 4838)
# 0

# if data is normal after transformation, proceed with parametric test 

#10. Mann-Whitney or t-test parametric test 

#11. ANOVA Parametric test 
  ANOVA = aov(psu_div$simpson_boxcox~psu_div$YEAR) 
  #plot(ANOVA)
  summary(ANOVA)
#               Df Sum Sq Mean Sq F value Pr(>F)
#psu_div$YEAR    1     26   25.82   1.155  0.283
#Residuals    4837 108147   22.36    
#??WHY IS DF 1 AND NOT 14?
  
drop1(ANOVA,~.,test="F") # type III SS and F Tests 
#                      Df Sum of Sq    RSS    AIC F value Pr(>F)
#<none>                    108147 15038               
#psu_div$YEAR  1    25.816 108172 15037  1.1546 0.2826

ANOVA = aov(psu_div$simpson_boxcox~psu_div$STRAT) 
  #plot(ANOVA)
  summary(ANOVA)
  
ANOVA = aov(simpson~YEAR, data = psu_div) 
  #plot(ANOVA)
  summary(ANOVA)
  
ANOVA = aov(simpson~STRAT, data = psu_div) 
  #plot(ANOVA)
summary(ANOVA)


# for post-hoc pairwise tests
#TukeyHSD(ANOVA) # where fit comes from aov()
```

## Sample Data Variables
- PRIMARY_SAMPLE_UNIT: A code indicating the primary sample unit in which a sample was collected.
- YEAR: A number indicating the calendar year.
- STATION_NR: A number indicating the secondary sampling unit within a given primary sample unit.
- LAT_DEGREES: Latitude of secondary sampling unit in decimal degrees).
- LON_DEGREES: Longitude of secondary sampling unit in decimal degrees).
- DEPTH: Average depth, in meters, of secondary sampling unit).
- UNDERWATER_VISIBILITY: visibility, in meters, at secondary sampling unit..  
- HABITAT_CD: A code indicating the habitat type. 
- ZONE_NR: A code indicating the distance offshore: 
+ 1 - Inshore
+ 2 - Midchannel
+ 3 - Offshore
+ 4 - Fore-reef
- MPA_NR: A number identifying the marine protected area in which the sample was collected. Zero indicates unprotected status)
- PROT: A boolean value indicating whether a sample was in a protected area or not 
+ 1 - protected area
+ 0 - not protected 
- STRAT: A code indicating the stratum in which a sample was taken. Differs by region.   
+ FMLR
+ FSLR
+ HRRF
+ INPR
+ MCPR 
+ OFPR

```{r sample data}

RVCdata_FK = read_rds("big_csv/RVCdata_FK.rds")

FK_sample_data = RVCdata_FK$sample_data

sample_data = FK_sample_data %>%
  select(YEAR, PRIMARY_SAMPLE_UNIT, LAT_DEGREES, LON_DEGREES, HABITAT_CD, ZONE_NR, MPA_NR, STRAT) %>% #4420893*8
  dplyr::distinct() %>% #8,678*8  
  group_by(YEAR, PRIMARY_SAMPLE_UNIT) #5,241 groups 

psu_diversity = psu_diversity_no_spp_merged %>%
  select(YEAR, PRIMARY_SAMPLE_UNIT, richness, simpson, shannon) %>%
  group_by(YEAR, PRIMARY_SAMPLE_UNIT)

bind_psu_variables = full_join(sample_data, psu_diversity, by="PRIMARY_SAMPLE_UNIT")

write_csv(bind_psu_variables, "big_csv/bind_psu_variables.csv")

##why is this joining funny 
```

## GAM 
Temporal variables:
- Year (categorical factor)

Spatial variables:
- latitude and longitude (continuous covariate)

Environmental variables: 
- Rugosity (categorical factor)
- bottom depth (continuous covariate)
- vertical relief (continuous covariate)
- % coral cover (continuous covariate)
- % hard bottom (continuous covariate)

Conitinuous covariates: modeled using non-parametric smoothing 
- thin plate regression splines with shrinkage terms (Ciannelli et al., 2008) 

Categorical factors: models parametrically to determine their mean effect sizes (Zurr et al., 2009)

```{r strata environmental characteristics}

# read in strata data for each year and psu 
#year, psu, latitude, longitude, rugosity, bottom_depth, vertical_relief, percent_coral_cover, percent_hardbottom

# bind to psu diversity dataframe 
```

Build GAM model with temporal, spatial and environmental variables 
```{r GAM}

#Build full model and null model based on (transformed) response variable
  #s() smooth term in gam model 
  #bs() a two letter character string indicating the smoothing basis to use (ts plate regression splines)

gam = data %>% 
  select() %>% 
  mutate(
    rich.model = gam(formula(richness,"~year+rugosity+s(Long,Lat,by=year,bs='ts')+s(depth,bs='ts')+s(vert_relief,bs='ts')+s(per_coral_cover,bs='ts')+s(per_hardbottom,bs='ts')"), family=poisson),
    rich.nullmodel = gam(formula(richness,"~1"), family=poisson),
    shan.model = gam(formula(shannon,"~year+rugosity+s(Long,Lat,by=year,bs='ts')+s(depth,bs='ts')+s(vert_relief,bs='ts')+s(per_coral_cover,bs='ts')+s(per_hardbottom,bs='ts')"), family=poisson),
    nullmodel = gam(formula("log(",shannon,")~1"))) 

#Set up empty vectors for deviances
time.dev=c(); space.dev=c(); env.dev=c()

#Sequence 1: time->space->environment

#Sequence 2: time->environment->space

#Sequence 3: space->time->environment

#Sequence 4: space->environment->time

#Sequence 5: environment->time->space

#Sequence 6: environment->space->time

```

```{r shapefiles}
library(foreign)
dbf = read.dbf('FL_files/FLGrid_100m_work8.dbf')
dbf2 = read.dbf('FL_files/Unified_Florida_Hab.dbf')
library(maptools)
shape = readShapePoly("FL_files/Unified_Florida_Hab.shp")
library(shapefiles)
shx = read.shx("FL_files/Unified_Florida_Hab.shx")
library(spatstat)
library(lattice)
library(maps)
library(rgeos)
library(rgdal)
library(ggmap)
library(tmap)
library(marmap) #marine maps 
```
