---
title: "Florida Keys Reef Fish Biodiversity"
author: "Megan Hepner"
date: "April 28, 2017"
output: html_document
---

## Inputs 
1. species-trait matrix
1. domain abundance dataframe
1. strata abundance dataframe 
1. psu abundance dataframe
1. domain biomass dataframe
1. strata biomass dataframe 
1. psu biomass dataframe

## Outputs 
1.  Functional dendogram
+ Trait-matrix visualization 

1.  By *domain* for each sampling year and region  
+ Species weighted abundance and plot
+ Species weighted biomass and plot
+ Species biodiversity 
+ species richness and plot
+ Simpson diversity as effective number of species and plot
+ Shannon diversity as effective number of species and plot
+ Functional diversity as effective number of species and plot

1. By *upper keys and lower keys domain* for each sampling year and region 
+ Species weighted abundance and plot
+ Species weighted biomass and plot
+ Species biodiversity 
+ species richness and plot
+ Simpson diversity as effective number of species and plot
+ Shannon diversity as effective number of species and plot
+ Functional diversity as effective number of species and plot

1. By *strata* for each sampling year and region 
+ Species weighted abundance and plot
+ Species weighted biomass and plot
+ Species biodiversity 
+ species richness and plot
+ Simpson diversity as effective number of species and plot
+ Shannon diversity as effective number of species and plot
+ Functional diversity as effective number of species and plot

1. By *primary sampling unit* for each sampling year and region and plot
+ Species weighted abundance and plot
+ Species weighted biomass and plot
+ Species biodiversity 
+ species richness and plot
+ Simpson diversity as effective number of species and plot
+ Shannon diversity as effective number of species and plot
+ Functional diversity as effective number of species and plot
1. Bivariate Relationships 
+ correlation with spearman rank 

1. Parametric tets
+ statistical tests by strata and year: Shapiro, homoscedacity, 2-way ANOVA

1. Generalized Additive Model
+ temporal variable: YEAR
+ spatial variable: Latitude, longitude
+ environmental variable: STRAT, habitat_cd, depth, bottom temperature 

1.Diversity heat Map
+ Kriging  

## Reef Visual Census Background 

Reef fish community data used in this study were collected as part of the multi-agency Reef Visual Census (RVC) [Brandt et al., 2009](https://www.coris.noaa.gov/activities/fish_monitoring_protocol/). Fish communities were visually surveyed underwater annually in the mainland FKNMS (Lower Keys, Middle Keys and Upper Keys) from 1999 to 2012, and biennially from 2012 to 2016. The Dry Tortugas region was surveyed from 1999 to 2000 and biennially from 2004 to 2016.     

The RVC uses a stationary point count method within a randomly selected 7.5 m radius circular plot with a 200m map grid from 1999 to 2012 and 100m map grid from 2014 to 2016 to optimize the observation of conspicuous and diurnally active reef fish, specifically economically and ecologically important reef fish species [Bohnsack and Bannerot, 1986](https://docs.lib.noaa.gov/noaa_documents/NMFS/TR_NMFS/TR_NMFS_41.pdf). 

## Objective 

To analyze and compare changes in species richness, Simpson diversity, Shannon diversity and functional diversity in the Florida Keys National Marine Sanctuary (FKNMS) to reveal changes in temporal and spatial variability from 1999 – 2016. The indices were assessed throughout the Florida Keys, between the upper and lower keys, between habitat types, and for different levels of management (Ecological Reserves (ER), Sanctuary Preservations Areas (SPA), and Special Use/Research Only Areas (SU/RO)). 

## Diversity indicse as effective number of species 

All indices were computed as effective number of species. Effective number of species is the number of equally abundance species needed to produce the observed value of a diversity index (Jost, 2006; Jost et al., 2010; MacArthur, 1965). Jost, 2006 refers to these as true measures of diversity, where prior to the conversion the diversity indices are simply measures of entropy. 

- Species Richness is the number of species in a habitat sampled  

$Richness = \sum_{i=1}^S p_i^0$  

- Simpson diversity measures the probability that two individuals randomly selected from a sample will belong to different species.   

$Simpson = 1 - \sum_{i=1}^S p_i^2$  

- Shannon diveristy is a measure of entropy, it is the uncertainty in the species identity of a sample  

$Shannon = - \sum_{i=1}^S p_i \log_b p_i$  

- Functional diversity (Rao's Q) is a measure of pairwise functional differences between species weighted by their relative abundances   

$Rao Q = \sum_{i=1}^S-1 \sum_{j=i+1}^S d_i p_i p_j$

## Functional distance trait based matrix (from Lefcheck et al.,)

Gowdis computes the Gower (1971) similarity coefficient exactly as described by Podani (1999), then converts it to a dissimilarity coefficient by using D = 1 - S

```{r Functional distance matrix, eval = F}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("clue")
library(clue) #calls cl_ultrametric 
#install.packages("fpc")
library(fpc) #calls pamk
#install.packages("ape")
library(ape)
#install.packages("phytools")
library(phytools)
#install.packages("FD")
library(FD)
#biocLite("BiocUpgrade") ## you may need this
#biocLite("ggtree", type = "source")
devtools::install_github("GuangchuangYu/ggtree") #requires R>= 3.3.2
library(ggtree)
library(tidyverse)

trait_matrix = read_csv('functional_diversity/species_trait_matrix_348_spp.csv')

# Parsed with column specification:
# cols(
#   Species_code = col_integer(),
#   SPECIES_CD = col_character(),
#   Latin_name = col_character(),
#   Common_name = col_character(),
#   Maxlength = col_double(),
#   Trophic_level = col_double(),
#   Trophic_group = col_character(),
#   Water_column = col_character(),
#   Diel_activity = col_character(),
#   Substrate_type = col_character(),
#   Complexity = col_character(),
#   Gregariousness = col_integer()

trait_matrix = trait_matrix %>%
  as.tibble() %>%
  mutate(
    SPECIES_CD = toupper(as.character(SPECIES_CD))) %>% 
  arrange(SPECIES_CD)%>% 
  select(SPECIES_CD, Maxlength, Trophic_level, Trophic_group, Water_column, Diel_activity, Substrate_type, Complexity, Gregariousness) %>% #348*9
  group_by(
    SPECIES_CD, Maxlength, Trophic_level, Trophic_group, Water_column, Diel_activity, Substrate_type, Complexity, Gregariousness) %>%
  summarize(
    n = n()) %>%
  select(-n) %>%
  ungroup() %>%
  mutate(
    # ordinal traits
    Complexity = ordered(Complexity, levels=c("Low","Medium","High")),
    Gregariousness = ordered(Gregariousness, levels=c("1","2","3"))) %>%
  as.data.frame()

#Set species names as row.names and remove extra columns
rownames(trait_matrix)=trait_matrix$SPECIES_CD 

#Calculate Gower distances. Variables have equal weight. 
traits.dist = gowdis(trait_matrix, ord="podani")

#Use clustering method to produce ultrametric dendrogram because values of Rao's Q can be maximized when fewer than the max number of functional types are present unless distances are ultramtetric 

#to account for sensitivity in clustering use multiple algorithms  (Mouchet et al., 2008) 
tree_methods = c("single","complete","average","mcquitty","ward.D") #average is best clustering method 
trees=lapply(tree_methods,  function(i) hclust(traits.dist, method=i))
par(mfrow=c(3,2))
for(i in 1:length(trees)) {plot(trees[[i]])} #plots the 5 different kinds of trees 

#convert trees to ultrametric
trees.ultra= lapply(trees,function(i) cl_ultrametric(as.hclust(i)))

#Plot each tree
par(mfrow=c(3,2))
for (i in 1:length(trees.ultra)) {plot(trees.ultra[[i]])} #plots the 5 different kinds of trees 

#Build the consensus tree (Mouchet et al 2008 Oikos)  
ensemble.trees=cl_ensemble(list=trees) #list of clusterings 
class(ensemble.trees)
consensus.tree=cl_consensus(ensemble.trees) #synthesizes the information in the elements of a cluster ensemble into a single clustering 
par(mar=c(1,1,1,1))
plot(consensus.tree, horiz=T)

#Calculate dissimilarity values for each tree using 2-norm (Merigot et al 2010 Ecology) to determine which tree best preserves orignial distances 
#spectral norm (2-norm) of the differences of the ultrametrics
all.trees=c(trees.ultra,consensus.tree[1])
names(all.trees)=c(tree_methods,"consensus")
(trees.dissim=lapply(all.trees,function(i) cl_dissimilarity(i,traits.dist,method="spectral"))) 
#Cross-dissimilarities using spectral ultrametric distance:
#Single = 63.17399
#complete = 116.8516
#average = 14.99288
#mcquitty = 36.06599
#ward.D = 3161.581
#consensus =  1065.058 (1067.782) (1091.628)

#Identify best tree and isolate
trees.dissim2=do.call(rbind,trees.dissim)
min.tree=which.min(trees.dissim2) #average 
names(all.trees)[min.tree]
func.dist=all.trees[names(all.trees)==names(all.trees)[min.tree]][[1]]

#Confirm lowest 2-norm value,  spectral norm (2-norm) of the differences if the ultrametrics 
cl_dissimilarity(func.dist,traits.dist,method="spectral") #Cross-dissimilarities using spectral ultrametric distance:  14.99288

#Scale by the max value so that all values are between 0-1
func.dist=func.dist/max(func.dist)

#Plot the best tree - average 
par(mfrow=c(1,1))
par(mar=c(2,1,0,2))
#par(pin=c(10,15))
plot(func.dist,horiz=TRUE)
#Save plot: 10" x 15"

#Phylogenetic tree with 348 tips and 347 internal nodes. 
func.dist.phy = as.phylo(as.hclust(func.dist))

#Write newick tree
write.tree(func.dist.phy, "functional_diversity/functional_dendrogram.nwk")
#write.nexus(func.dist.phy, "functional_diversity/functional_dendrogram.nex") 

tree <- read.tree("functional_diversity/functional_dendrogram.nwk")
# List of 4
#  $ edge       : int [1:694, 1:2] 
#  $ Nnode      : int 347
#  $ tip.label  : chr [1:348] 
#  $ edge.length: num
# - attr(*, "class")= chr "phylo"
# - attr(*, "order")= chr "cladewise"

# Phylogenetic tree with 348 tips and 347 internal nodes.
# Tip labels:
# 	KYP_SECT, ALE_CILI, ELO_SAUR, HAR_JAGU, CEN_UNDE, TYL_CROC, ...
# Rooted; includes branch lengths.

plot(tree, label.offset=0.2, edge.width = 2, type = "cladogram")  
nodelabels() #add node numbers
tiplabels() #add tip numbers

ggtree(tree, branch.length = T) +
  ggtitle("Functional Dendogram") +
  geom_tiplab()+
  geom_text(aes(x = branch, label=branch.length), vjust = -.5) +
  theme_tree2()

ggtree(tree, layout= "circular") +
  ggtitle("Functional Dendogram") +
  geom_tiplab(size = 1, aes(angle=angle))+
  theme_tree2()

# transformed to a tidy data.frame by fortify method
tree_data <- fortify(tree) #node; parent; branch.length; x; y; label; isTip; branch; angle 
head(tree_data)

#Visualize species' differences in multivariate trait space
#Perform k-means clustering with no a priori specification for k
traits.kclus= pamk(traits.dist, krange=2:10) 

#krange: integer vector. Numbers of clusters which are to be compared by the average silhouette width criterion.
#traits.kclus$nc = 10 
#traits.kclus$crit 0.0000000 0.2374706 0.2362513 0.2355555 0.2681756 0.2581352 0.2759547 0.2852216 0.3006077 0.3062123

# #Perform multidimensional scaling on functional dendrogram
traits_nmds = metaMDS(traits.dist,k=traits.kclus$nc,trymax=20)
#k = Number of dimensions

#*** No convergence -- monoMDS stopping criteria:
#499: no. of iterations >= maxit
#1: stress ratio > sratmax

# Dimensions: 10 
# Stress:     0.03545178 
# Stress type 1, weak ties
# No convergent solutions - best solution after 500 tries
# Scaling: centring, PC rotation 
# Species: scores missing


# #Plot in two dimensions - doesn't work 
par(mar=c(4,4,1,1))
ordiplot(traits_nmds,type="n")

#Assign colors to different groups
groups=levels(factor(traits.kclus$pamobject$clustering))
points.symbols=15:16
points.colors=c("firebrick3","cornflowerblue")
for(i in seq_along(groups)) {
  points(traits_nmds$points[traits.kclus$pamobject$clustering==groups[i],],
         pch=points.symbols[i],col=points.colors[i],cex=1.4) }
ordispider(traits_nmds,factor(traits.kclus$pamobject$clustering),label=F)
ordihull(traits_nmds,factor(traits.kclus$pamobject$clustering),lty="dotted")
orditorp(traits_nmds,dis="sites",pcex=0,air=0.5,col="grey10",cex=0.8)

```

## Trait matrix visualization 

[http://jkunst.com/r/pokemon-visualize-em-all/]
[https://d3js.org/]
[http://www.buildingwidgets.com/blog/2015/7/22/week-29-d3treer-v2]

```{r trait visualization}

library(tidyverse)
library(treemap)
devtools::install_github("gluc/data.tree")
devtools::install_github("timelyportfolio/d3treeR")
library(d3treeR)

trait_matrix = read_csv('functional_diversity/species_trait_matrix_348_spp.csv')

trait_matrix_treemap = trait_matrix %>%
  select(Common_name, Maxlength, Trophic_level, Trophic_group, Water_column,Diel_activity,Substrate_type,Complexity, Gregariousness) %>% 
  group_by(Maxlength,
           Trophic_level,
           Trophic_group,
           Water_column,
           Diel_activity,
           Substrate_type,
           Complexity,
           Gregariousness) %>%
  summarise(n = n())%>% 
  ungroup() %>%
  as.data.frame()

treemap2 = 
  d3treeR::d3tree2(
    treemap(
      dtf = trait_matrix_treemap,
      index = c("Trophic_group", "Trophic_level"), #, "Water_column", "Complexity", "Substrate_type", "Diel_activity"),
      vSize = "n", 
      vColor = "Trophic_group",
      type = "index"),
    #title = "Reef Fish Traits",
    #fontsize.title = 14, 
    #algorithm = "squarified"
    rootname = "Species_Traits")

#treemap example 
data(GNI2014)
d3tree2(
  treemap(
    GNI2014
    ,index=c("continent", "iso3")
    ,vSize="population"
    ,vColor="GNI"
    ,type="value"
  )
  , rootname = "World"
)

```

## Fetch the RVC data from 
[NOAA's Southeast Fisheries Science Center server](https://grunt.sefsc.noaa.gov/rvc_analysis20/) using the [RVC package](https://github.com/jeremiaheb/rvc)
- Fetched weighted abundance and biomass data by domain, strata, and primary sampling units 

### Sample Data Variables
- PRIMARY_SAMPLE_UNIT: A code indicating the primary sample unit in which a sample was collected.
- YEAR: A number indicating the calendar year.
- MONTH: A number indicating the month of the year.
- DAY: A number indicating the day of the month (EST).
- STATION_NR: A number indicating the secondary sampling unit within a given primary sample unit.
- LAT_DEGREES: Latitude of secondary sampling unit in decimal degrees).
- LON_DEGREES: Longitude of secondary sampling unit in decimal degrees).
- DEPTH: Average depth, in meters, of secondary sampling unit).
- UNDERWATER_VISIBILITY: visibility, in meters, at secondary sampling unit.
- MAPGRID_NR: A number indicating the primary sample unit.  
- HABITAT_CD: A code indicating the habitat type. 
- ZONE_NR: A code indicating the distance offshore: 
+ 1 - Inshore
+ 2 - Midchannel
+ 3 - Offshore
+ 4 - Fore-reef
- SUBREGION_NR: A number indicating the subregion.
- MPA_NR: A number identifying the marine protected area in which the sample was collected. Zero indicates unprotected status)
- SPECIES_NR: A number indicating the species for a sample 
- SPECIES_CD: A code indicating the species for a sample. Consists of the first three letters of the generic name and the first four of the specific name   
- LEN: The length, in cm, of a sample.      
- NUM: The number of individuals of a given species and length observed in a secondary sampling unit
- TIME_SEEN: A number indicating when, during sampling, an individual was observed. 1: In the first five minutes, 2: From 5-10 minutes, 3: After 10 minutes. 
- PROT: A boolean value indicating whether a sample was in a protected area or not 
+ 1 - protected area
+ 0 - not protected 
- STRAT: A code indicating the stratum in which a sample was taken. Differs by region.   
+ FMLR
+ FSLR
+ HRRF
+ INPR
+ MCPR 
+ OFPR
- REGION: A code indicating the region in which a sample was taken.
+ FLA KEYS (florida keys)
+ DRY TORT (dry tortugas)

### Stratum Data Variables 
- REGION 
- YEAR 
- PROT 
- STRAT 
- NTOT: The number of possible primary sample units for a given year, region, stratum, and protected status 
- GRID_SIZE: The length (in meters) to a side of a primary sample unit for a given year, region, stratum, and protected status.

### Taxonomic Data Variables 
- SPECIES_CD
- FAMILY
- SCINAME 
- COMNAME
- LC: Minimum length at capture, in centimeters
- LM: Median length at maturity, in centimeters.
- WLEN_A: The linear coefficient of the allometric growth equation in grams per millimeter (g/mm)
- WLEN_B: The exponential coefficient of the allometric growth equation

### Benthic Data Variables 
- REGION: A code indicating the region. 
- YEAR: A number indicating the calendar year.
- PRIMARY_SAMPLE_UNIT: A code indicating the primary sample unit in which a sample was collected.
- STATION_NR: A number indicating the secondary sampling unit within a given primary sample unit.
- DEPTH: Average depth, in meters, of secondary sampling unit.
- MAX_HARD_RELIEF: The maximum height, in meters, of hard relief (e.g. hard corals, rock).
- MAX_SOFT_RELIEF: The maximum height, in meters, of soft relief (e.g. soft corals, sponges).
- AVG_HARD_RELIEF: The average height, in meters, of hard relief.
- HARD_REL_PCT_0: Percentage of hard relief less than 0.2 meters.
- HARD_REL_PCT_1: Percentage of hard relief between 0.2 and 0.5 meters.
- HARD_REL_PCT_2: Percentage of hard relief between 0.5 and 1.0 meters.
- HARD_REL_PCT_3: Percentage of hard relief between 1.0 and 1.5 meters.
- HARD_REL_PCT_4: Percentage of hard relief greater than 1.5 meters.
- PCT_SAND: Percentage of abiotic cover that is sand.
- PCT_HARD_BOTTOM: Percentage of abiotic cover that is hard bottom.
- PCT_RUBBLE: Percentage of abiotic cover that is rubble.
- PCT_CORAL: Percentage of biotic hardbottom that is coral.
- PCT_OCTO: Percentage of biotic hardbottom that is octocoral.
- PCT_SPONGE: Percentage of biotic hardbottom that is sponge.

### Function: getDomainDensity 
- YEAR
- REGION
- SPECIES_CD
- density: Domain-wide mean density for a stratified random survey  
- var: Variance in average density per secondary sampling unit
- n: Number of primary sampling units sampled
- nm: Number of secondary sampling units sampled
- N: Number of total possible primary sample units
- NM: Number of possible secondary sampling units
- length_class: The length class or bin. Only present if length_bins is not NULL. The notation, [lower, upper), is inclusive of the lower bound, but exclusive of the upper bound
- protected_status: The protected status. Only present if merge_protected is FALSE

```{r setup, include= F}
knitr::opts_chunk$set(echo = T)
#rm(list = ls())

#install.packages('devtools')
#devtools::install_github('jeremiaheb/rvc')
library(rvc) #calls RVC data from SEFSC server 
#install.packages("tidyverse")
library(tidyverse)
#install.packages("vegan")
library(vegan) # biodiveristy functions for richness, simpson, and shannon 
#install.packages("dygraphs")
library(dygraphs)
#install.packages("plotly")
library(plotly)
#install.packages("webshot")
library(webshot)
#install.packages("htmlwidgets")
library(htmlwidgets)
#install.packages("FD")
library(FD)
#install.packages("iNEXT")
library(iNEXT)

map = purrr::map # override maps::map
select = dplyr::select #override MASS::select 
```

### Florida Keys domain

```{r domain abundance}

domain_fk_abun <- read_csv('big_csv/abundance_domain/domain_fk_abun.csv', 
                           col_types = cols(
                             YEAR = col_integer(),
                             REGION = col_character(),
                             SPECIES_CD = col_character(),
                             abundance = col_double(),
                             var = col_double(),
                             n = col_integer(),
                             nm = col_integer(),
                             N = col_double(),
                             NM = col_double(),
                             protected_status = col_character()
                           ))

domain_fk_abun_no_spp = domain_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species

```

```{r domain biomass}

domain_fk_bio <- read_csv('big_csv/biomass_domain/domain_fk_biomass.csv', 
                          col_types = cols(
                            protected_status = col_character()
                          ))

domain_fk_bio_no_spp = domain_fk_bio %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species
```

```{r domain biodiversity}

trait_matrix = read_csv('functional_diversity/species_trait_matrix_348_spp.csv')
trait_matrix = trait_matrix %>%
  as.tibble() %>%
  mutate(
    SPECIES_CD = toupper(as.character(SPECIES_CD))) %>% 
  arrange(SPECIES_CD)

domain_fk_abun <- read_csv('big_csv/abundance_domain/domain_fk_abun.csv', 
                           col_types = cols(
                             YEAR = col_integer(),
                             REGION = col_character(),
                             SPECIES_CD = col_character(),
                             abundance = col_double(),
                             var = col_double(),
                             n = col_integer(),
                             nm = col_integer(),
                             N = col_double(),
                             NM = col_double(),
                             protected_status = col_character()
                           ))

domain_fk_abun_no_spp = domain_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species 

#length(unique(domain_fk_abun_no_spp$SPECIES_CD)) # 348
#length(unique(trait_matrix$SPECIES_CD))          # 348



## FK's domain abundance only identified to species level (without spp.) 
fk_domain_abun_diversity_no_spp = domain_fk_abun_no_spp %>% #14,634 * 10
  #filter(protected_status != "all") %>% 
  select(YEAR, protected_status, SPECIES_CD, abundance) %>% #14,634 * 4
  group_by(YEAR, protected_status) %>% ##45 groups (fk 15*3)
  nest(-YEAR) %>%  
  mutate(
    data_wide = map(data, function(x) 
      full_join(x, trait_matrix %>% select(SPECIES_CD), by='SPECIES_CD') %>%
        mutate(abundance = ifelse(is.na(abundance), 0, abundance)) %>%
        spread(SPECIES_CD, abundance, fill =0)), 
    #species richness
    richness=map(data_wide,   
                 function(x) specnumber(x)),
    #simpson diversity as effective number of species 
    simpson=map(data_wide, 
                function(x) 1/(1 - diversity(x, index = 'simpson'))),
    #shannon diversity as effective number of species 
    shannon=map(data_wide, 
                function(x) exp(diversity(x, index = 'shannon')))) %>%
  unnest(richness,shannon,simpson)  
  
#fk_domain_abun_diversity_no_spp = fk_domain_abun_diversity_no_spp %>%
  #select(YEAR, protected_status,richness,shannon,simpson)
    

fd_attempt = fk_domain_abun_diversity_no_spp %>%
  mutate(
    #functional diversity as effective number of species 
    abun.sum2=map(data_wide, 
                   function(x) apply(x, 2, sum))) %>%
  unnest(abun.sum2)

if(any(fd_attempt$abun.sum2 == 0)) 
  stop("At least one species does not occur in any community (zero total abundance across all communities).") #,
    #functional = map(data_wide, 
                   #function(x) 1/(1 - dbFD(as.matrix(func.dist), x)))) #%>% 
  
#6/12/90 - Error in mutate_impl(.data, dots) : At least one species does not occur in any community (zero total abundance across all communities).

# check if one species has total abundance of zero (never occurs)
abun.sum2 <- apply(a, 2, sum)
if (any(abun.sum2 == 0) ) stop("At least one species does not occur in any community (zero total abundance across all communities).","\n")
  


#write_csv(fk_domain_abun_diversity_no_spp, 'big_csv/diversity_without_spe/domain_fk_abun_diversity_no_spp.csv')

```


```{r dygraphs}

domain_fk_abun_diversity = read_csv('big_csv/abundance_domain/domain_fk_abun_diversity.csv')

## Simpson for protected and unprotected stacked 
fk_simpson = domain_fk_abun_diversity %>% 
  filter(protected_status != 'all') %>%
  select(YEAR, protected_status, dom_simpson) %>%
  mutate(
    protected_status = recode(
      protected_status, '0'='Unprotected', '1'='Protected')) %>%
  spread(protected_status, dom_simpson)

dygraph(fk_simpson, main = "Simpson Reef Fish Diversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 50)) %>%
  dyAxis("x", label = "Year") %>%
  dyOptions(stackedGraph = T) 
#dyRangeSelector(height = 20) 

## Shannon for protected and unprotected stacked 
fk_shannon = domain_fk_abun_diversity %>% 
  filter(protected_status != 'all') %>%
  select(YEAR, protected_status, dom_shannon) %>%
  mutate(
    protected_status = recode(
      protected_status, '0'='Unprotected', '1'='Protected')) %>%
  spread(protected_status, dom_shannon)

dygraph(fk_shannon, main = "Shannon Reef Fish Diversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 75)) %>%
  dyAxis("x", label = "Year") %>%
  dyOptions(stackedGraph = T)
#dyRangeSelector(height = 20) 


d = domain_fk_abun_diversity %>% 
  filter(protected_status != 0) %>% 
  filter(protected_status != 1)

#Simpson for "all" filled 
dygraph_simp = d %>%
  select(YEAR, dom_simpson)

dygraph(dygraph_simp, main = "Simpson Diversity of Reef Fish in FKNMS") %>%
  dyOptions(fillGraph = T, fillAlpha = 0.4, drawPoints = T, pointSize = 2) %>%
  dySeries("dom_simpson", label = "Simpson Diversity") %>% 
  dyAxis("y", label = "Effective Number of Species") %>% 
  dyAxis("x", label = "Year")

#Shannon for "all" filled 
dygraph_shannon = d %>%
  select(YEAR, dom_shannon)

dygraph(dygraph_shannon, main = "Shannon Diversity of Reef Fish in FKNMS") %>%
  dyOptions(fillGraph = T, fillAlpha = 0.4, drawPoints = T, pointSize = 2) %>%
  dySeries("dom_shannon", label = "Shannon Diversity") %>% 
  dyAxis("y", label = "Effective Number of Species") %>% 
  dyAxis("x", label = "Year")

#Richness for "all" filled 
dygraph_richness = d %>%
  select(YEAR, dom_richness)

dygraph(dygraph_richness, main = "Species Richness of Reef Fish in FKNMS") %>%
  dyOptions(fillGraph = T, fillAlpha = 0.4, drawPoints = T, pointSize = 2) %>%
  dySeries("dom_richness", label = "Species Richness") %>% 
  dyAxis("y", label = "Number of Species") %>% 
  dyAxis("x", label = "Year")

## Simpson and Shannon stacked for "all"
dygraph_simp_shannon = d %>%
  select(YEAR, dom_simpson, dom_shannon)

dygraph(dygraph_simp_shannon, main = "Florida Keys Reef Fish Biodiversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 70)) %>%
  dyAxis("x", label = "Year") %>%
  dySeries("dom_shannon", label = "Shannon Diversity", color = "blue") %>% 
  dySeries("dom_simpson", label = "Simpson Diversity", color = "red") %>% 
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 1.5) %>% #drawGrid = F
  dyLegend(width = 400)

# m <- plot_ly(x = 1:10)
# saveWidget(as.widget(m), "temp.html")
# webshot("temp.html", file = "test.png",
#         cliprect = "viewport")

#Richness, Shannon, Simpson stacked for "all"
dygraph_rich_simp_shannon = d %>%
  select(YEAR, dom_simpson, dom_shannon, dom_richness)

dygraph(dygraph_rich_simp_shannon, main = "Florida Keys Reef Fish Biodiversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 350)) %>% 
  dyAxis("x", label = "Year") %>%
  dySeries("dom_richness", label = "Richness", color = "purple") %>%
  dySeries("dom_shannon", label = "Shannon", color = "blue") %>% 
  dySeries("dom_simpson", label = "Simpson", color = "red") %>% 
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 1.5) %>% #drawGrid = F
  dyLegend(width = 400)

##without spp 
d_no_spp = fk_domain_abun_diversity_no_spp %>% 
  filter(protected_status != 0) %>% 
  filter(protected_status != 1)

dygraph_rich_simp_shannon_no_spp = d_no_spp %>%
  select(YEAR, simpson, shannon, richness)

dygraph(dygraph_rich_simp_shannon_no_spp, main = "Florida Keys Reef Fish Biodiversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 350)) %>% 
  dyAxis("x", label = "Year") %>%
  dySeries("richness", label = "Richness", color = "purple") %>%
  dySeries("shannon", label = "Shannon", color = "blue") %>% 
  dySeries("simpson", label = "Simpson", color = "red") %>% 
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 1.5) %>% #drawGrid = F
  dyLegend(width = 400)

dygraph_simp_shannon_no_spp = d_no_spp %>%
  select(YEAR, simpson, shannon)

dygraph(dygraph_simp_shannon_no_spp, main = "Florida Keys Reef Fish Biodiversity") %>% 
  dyAxis("y", label = "Effective Number of Species") %>%  #valueRange = c(0, 350)
  dyAxis("x", label = "Year") %>%
  dySeries("shannon", label = "Shannon", color = "blue") %>% 
  dySeries("simpson", label = "Simpson", color = "red") %>% 
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 1.5) %>% #drawGrid = F
  dyLegend(width = 400)

```

```{r domain biodiversity plots}

domain_fk_abun_diversity = read_csv('big_csv/abundance_domain/domain_fk_abun_diversity.csv')
# Parsed with column specification:
# cols(
#   YEAR = col_integer(),
#   protected_status = col_character(),
#   dom_richness = col_integer(),
#   dom_simpson = col_double(),
#   dom_shannon = col_double()
# )
protected_status_names <- c(
  "0" = "Unprotected",
  "1" = "Protected",
  "all" = "All"
)
#richness, simpson, shannon for domain FKNMS  
FK_abun_RShSim <- ggplot(data = fk_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_richness, color = "dom_richness")) +
  geom_point(aes(y = dom_simpson, color = "dom_simpson")) +
  geom_point(aes(y = dom_shannon, color = "dom_shannon")) +
  geom_line(aes(y = dom_richness, color = "dom_richness")) +
  geom_line(aes(y = dom_simpson, color = "dom_simpson")) +
  geom_line(aes(y = dom_shannon, color = "dom_shannon")) +
  facet_grid(protected_status ~ ., labeller = as_labeller(protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Diversity indices") +
  scale_colour_manual(values = c("darkturquoise", "magenta", "blue"), labels=c("Species Richness", "Shannon Diversity", "Simpson Diversity")) +
  ggtitle("Reef Fish Biodiversity in the FKNMS") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(breaks = c(1999:2016))+
  scale_y_continuous(limits= c(0, 300), breaks = c(0,25,50,75,100,125,150,175,200,225,250,275,300))


#simpson, shannon by protected status for domain FKNMS
FK_abun_ShSim <-ggplot(data = fk_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_simpson, color = "dom_simpson")) +
  geom_point(aes(y = dom_shannon, color = "dom_shannon")) +
  geom_line(aes(y = dom_simpson, color = "dom_simpson")) +
  geom_line(aes(y = dom_shannon, color = "dom_shannon")) +
  facet_grid(protected_status ~ ., labeller = as_labeller(protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Diversity index") +
  scale_colour_manual(values = c("magenta", "blue"), labels=c("Shannon diversity", "Simpson Diversity")) +
  ggtitle("Reef Fish Biodiversity in the Florida Keys") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 250), breaks = c(5,10,15,20,25,30,35,40,45,50))

#species richness by protected status for domain FKNMS
FK_abun_Rich <- ggplot(data = fk_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_richness, color = "dom_richness"),color = "darkturquoise") +
  geom_line(aes(y = dom_richness, color = "dom_richness"), color = "darkturquoise") +
  facet_grid(protected_status ~ ., labeller = as_labeller(protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Level of protection") +
  ggtitle("Species Richness of Reef Fish in the Florida Keys") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 250), breaks = c(0,50, 100, 150, 200, 250, 300))
## wonky 

#shannon by protected status for domain FKNMS
FK_abun_Shan <- ggplot(data = fk_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_shannon, color = "dom_shannon"),color = "magenta") +
  geom_line(aes(y = dom_shannon, color = "dom_shannon"), color = "magenta") +
  facet_grid(protected_status ~ ., labeller = as_labeller(protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Level of protection") +
  #scale_colour_manual(labels=c("Unprotected", "Protected", "All")) +
  ggtitle("Shannon Diveristy of Reef Fish in the Florida Keys") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 50), breaks = c(5,10,15,20,25,30,35,40,45,50))

#simpson by protected status for domain FKNMS
FK_abun_Simp <- ggplot(data = fk_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_simpson, color = "dom_simpson"),color = "blue") +
  geom_line(aes(y = dom_simpson, color = "dom_simpson"), color = "blue") +
  facet_grid(protected_status ~ ., labeller = as_labeller(protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Level of protection") +
  #scale_colour_manual(labels=c("Unprotected", "Protected", "All")) +
  ggtitle("Simpson Diveristy of Reef Fish in the Florida Keys") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 25), breaks = c(5,10,15,20,25))

```

### Florida Keys Strata
1. FSLR - Forereef Shallow Linear Reef (<6m)
2. FMLR - Forereef Midchannel Linear Reef (6-18m)
3. FDLR - Forereef Deep Linear Reef (18-33m)
4. INPR - Inshore patch reef
5. MCPR - Midchannel Patch Reef 
6. OFPR - Offshore Patch Reef
7. HRRF - High Relief Reef (Spur and Groove)

```{r strata abundance}
## FK's abundance by strata 
strata_fk_abun <- read_csv("big_csv/abundance_strata/strata_fk_abun.csv",
                           col_types = cols(
                             YEAR = col_integer(),
                             REGION = col_character(),
                             STRAT = col_character(),
                             SPECIES_CD = col_character(),
                             abundance = col_double(),
                             var = col_double(),
                             n = col_integer(),
                             nm = col_integer(),
                             N = col_double(),
                             NM = col_double(),
                             protected_status = col_character()
                           ))

strata_fk_abun_no_spp = strata_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species 
```

```{r strata biomass}

strata_fk_bio <- read_csv('big_csv/biomass_strata/strata_fk_biomass.csv', 
                          col_types = cols(
                            protected_status = col_character()
                          ))

strata_fk_bio_no_spp = strata_fk_bio %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species
```

Is reef fish biodiversity and strata type related? 

```{r strata biodiversity}

## FK's abundance by strata 
strata_fk_abun <- read_csv("big_csv/abundance_strata/strata_fk_abun.csv",
                           col_types = cols(
                             YEAR = col_integer(),
                             REGION = col_character(),
                             STRAT = col_character(),
                             SPECIES_CD = col_character(),
                             abundance = col_double(),
                             var = col_double(),
                             n = col_integer(),
                             nm = col_integer(),
                             N = col_double(),
                             NM = col_double(),
                             protected_status = col_character()
                           ))

strata_fk_abun_no_spp = strata_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species 

fk_strata_abun_diversity_no_spp = strata_fk_abun_no_spp %>% #124,116 * 12
  filter(protected_status == 'all') %>% #66,058 * 12
  select(YEAR, STRAT, protected_status, SPECIES_CD, abundance) %>% #62,058 * 5
  group_by(YEAR, STRAT, SPECIES_CD) %>% #33,468 gruops    
  summarise(
    abun_merged = sum(abundance))

trait_matrix = read_csv('functional_diversity/species_trait_matrix_348_spp.csv')

trait_matrix = trait_matrix %>%
  as.tibble() %>%
  mutate(
    SPECIES_CD = toupper(as.character(SPECIES_CD))) %>% 
  arrange(SPECIES_CD)

strata_merged = fk_strata_abun_diversity_no_spp %>%
  group_by(YEAR, STRAT) %>% #103 groups 
  nest() %>%  #data 339*2
  mutate(
    data_wide = map(data, function(x) 
      full_join(x, trait_matrix %>% select(SPECIES_CD), by='SPECIES_CD') %>% 
        mutate(abun_merged = ifelse(is.na(abun_merged), 0, abun_merged)) %>%
        spread(SPECIES_CD, abun_merged, fill =0)),
    richness = map( #species richness
      data_wide, 
      function(x) specnumber(x)),
    simpson = map( #simpson diversity as effective number of species 
      data_wide,
      function(x) 1/(1 - diversity(x, index = 'simpson'))),
    shannon = map( #shannon diversity as effective number of species 
      data_wide,
      function(x) exp(diversity(x, index = 'shannon')))) %>%
  unnest(richness, simpson, shannon)

strata_diversity_no_spp_merged = strata_merged %>%
  select(YEAR, STRAT, richness, simpson, shannon) 

write_csv(strata_diversity_no_spp_merged, 'diversity_without_spe/strata_fk_abun_diversity_no_spp_merged.csv')

```

```{r strata biodiversity plots, eval = F}

strata_fk_abun_diversity = read_csv("big_csv/abundance_strata/strata_fk_abun_diversity.csv")

# strata_names <- c(
#  "FMLR" = "Forereef Midchannel Linear Reef",
#  "FSLR" = "Forereef Shallow Linear Reef",
#  "HRRF" = "High Relief Reef", #(Spur and Groove)
#  "INPR" = "Inshore patch reef",
#  "MCPR" = "Midchannel Patch Reef",
#  "OFPR" = "Offshore Patch Reef",
#  "FDLR" = "Forereef Deep Linear Reef")

##protected by strata 
fk_strata_protected = strata_fk_abun_diversity %>% 
  filter(protected_status != 1)

#richness protected by strata 
fk_rich_protected = fk_strata_protected %>% 
  select(YEAR, STRAT, richness) %>%
  spread(STRAT, richness)

dygraph(fk_rich_protected, main = "Florida Keys Richness by Strata - NTMR") %>% 
  dyAxis("y", label = "Effective Number of Species") %>%
  dyAxis("x", label = "Year") %>% 
  dyOptions(stackedGraph = T, colors = colorRamps::blue2red(7), fillAlpha = .8, axisLineWidth = 2)

##shannon protected by strata 
fk_shan_protected = fk_strata_protected %>% 
  select(YEAR, STRAT, shannon) %>%
  spread(STRAT, shannon)

dygraph(fk_shan_protected, main = "Florida Keys Shannon Diversity by Strata - NTMR") %>% 
  dyAxis("y", label = "Effective Number of Species") %>%
  dyAxis("x", label = "Year") %>% 
  dyOptions(stackedGraph = T, colors = colorRamps::blue2red(7), fillAlpha = .8, axisLineWidth = 2)

##simpson protected by strata 
fk_simp_protected = fk_strata_protected %>% 
  select(YEAR, STRAT, simpson) %>%
  spread(STRAT, simpson)

dygraph(fk_simp_protected, main = "Florida Keys Simpson Diversity by Strata - NTMR") %>% 
  dyAxis("y", label = "Effective Number of Species") %>%
  dyAxis("x", label = "Year") %>% 
  dyOptions(stackedGraph = T, colors = colorRamps::blue2red(7), fillAlpha = .8, axisLineWidth = 2)#topo.colors(n = 7, alpha = 1) RColorBrewer::brewer.pal(7, "Paired") grDevices::topo.colors(7, alpha = 1) grDevices::rainbow(7) colorspace::rainbow_hcl(7) colorspace::heat_hcl(7)

#unprotected by strata 
fk_strata_unprotected = strata_fk_abun_diversity %>% 
  filter(protected_status != 0)

#richness unprotected by strata 
fk_rich_unprotected = fk_strata_unprotected %>% 
  select(YEAR, STRAT, richness) %>%
  spread(STRAT, richness)

dygraph(fk_rich_unprotected, main = "Florida Keys Richness by Strata - Unprotected") %>% 
  dyAxis("y", label = "Effective Number of Species") %>%
  dyAxis("x", label = "Year") %>% 
  dyOptions(stackedGraph = T, colors = colorRamps::blue2red(7), fillAlpha = .8, axisLineWidth = 2)

#simpson unprotected by strata
fk_simp_unprotected = fk_strata_unprotected %>% 
  select(YEAR, STRAT, simpson) %>%
  spread(STRAT, simpson)

dygraph(fk_simp_unprotected, main = "Florida Keys Simpson Diversity by Strata - Unprotected") %>% 
  dyAxis("y", label = "Effective Number of Species") %>%
  dyAxis("x", label = "Year") %>% 
  dyOptions(stackedGraph = T, colors = colorRamps::blue2red(7), fillAlpha = .8, axisLineWidth = 2)

strata_names <- c(
  "FMLR" = "FMLR",
  "FSLR" = "FSLR",
  "HRRF" = "HRRF", #(Spur and Groove)
  "INPR" = "INPR",
  "MCPR" = "MCPR",
  "OFPR" = "OFPR",
  "FDLR" = "FDLR")

protected_status_names <- c(
  "0" = "Unprotected",
  "1" = "Protected",
  "all" = "All"
)

#simpson diversity by protected status by strata in FKNMS
fk_strata_ab_simpson <- ggplot(data = strata_fk_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = simpson, color = "simpson"),color = "blue") +
  geom_line(aes(y = simpson, color = "simpson"), color = "blue") +
  facet_grid(STRAT ~ protected_status, 
             labeller = labeller(.rows = strata_names, .cols = protected_status_names)) +
  #labeller(strata_names = label_wrap_gen(3))) + 
  labs(x = "Year", y = "Effective Number of Species") +
  ggtitle("Simpson Diveristy of Reef Fish in the Florida Keys") + 
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12)) +
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 25), breaks = c(5,10,15,20,25))

#shannon diversity by protected status by strata in FKNMS
FK_strata_ab_shannon <- ggplot(data = fk_strata_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = shannon, color = "shannon"), color = "magenta") +
  geom_line(aes(y = shannon, color = "shannon"), color = "magenta") +
  facet_grid(STRAT ~ protected_status, 
             labeller = labeller(.rows = strata_names, .cols = protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Level of protection") +
  ggtitle("Shannon Diveristy of Reef Fish in the Florida Keys by Strata") + 
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 50), breaks = c(5,10,15,20,25))

```

### Florida Keys Primary Sampling Unit 

```{r psu abundance}

psu_fk_abun <- read_csv("big_csv/abundance_psu/psu_fk_abun.csv",
                        col_types = cols(
                          YEAR = col_integer(),
                          REGION = col_character(),
                          STRAT = col_character(),
                          PROT = col_integer(),
                          PRIMARY_SAMPLE_UNIT = col_character(),
                          SPECIES_CD = col_character(),
                          m = col_integer(),
                          var = col_double(),
                          abundance = col_double(),
                          protected_status = col_character()))

psu_fk_abun_no_spp = psu_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species 
```

```{r psu biomass}
psu_fk_bio <- read_csv('big_csv/biomass_psu/psu_fk_biomass.csv', 
                       col_types = cols(
                         protected_status = col_character()
                       ))

psu_fk_bio_no_spp = psu_fk_bio %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species
```

```{r psu biodiversity}

psu_fk_abun <- read_csv("big_csv/abundance_psu/psu_fk_abun.csv",
                        col_types = cols(
                          YEAR = col_integer(),
                          REGION = col_character(),
                          STRAT = col_character(),
                          PROT = col_integer(),
                          PRIMARY_SAMPLE_UNIT = col_character(),
                          SPECIES_CD = col_character(),
                          m = col_integer(),
                          var = col_double(),
                          abundance = col_double(),
                          protected_status = col_character()))

psu_fk_abun_no_spp = psu_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species 

trait_matrix = read_csv('functional_diversity/species_trait_matrix_348_spp.csv')

trait_matrix = trait_matrix %>%
  as.tibble() %>%
  mutate(
    SPECIES_CD = toupper(as.character(SPECIES_CD))) %>% 
  arrange(SPECIES_CD)

fk_psu_abun_diversity_no_spp_merged = psu_fk_abun_no_spp %>% #3,107,408 x 10
  select(YEAR, PRIMARY_SAMPLE_UNIT, STRAT, protected_status, SPECIES_CD, abundance) %>% #3,107,408*5
  filter(protected_status == 'all') %>% #1,553,704*10
  group_by(YEAR, STRAT, PRIMARY_SAMPLE_UNIT, SPECIES_CD) %>% #1,553,704 groups  
  summarise(
    abun_merged = sum(abundance)) 
  
psu_diversity_no_spp_merged = fk_psu_abun_diversity_no_spp_merged %>% #1,553,704 x 5
  group_by(YEAR, PRIMARY_SAMPLE_UNIT) %>% #4,839 groups
  nest(-YEAR) %>%
  mutate(
    data_wide = map(data, function(x) 
      full_join(x, trait_matrix %>% select(SPECIES_CD), by='SPECIES_CD') %>%
        mutate(abun_merged = ifelse(is.na(abun_merged), 0, abun_merged)) %>%
        spread(SPECIES_CD, abun_merged, fill =0)),
    data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abun_merged, fill =0)), 
    richness = map( #species richness
      data_wide, 
      function(x) specnumber(x %>% select(-STRAT))),
    simpson = map( #simpson diversity as effective number of species 
      data_wide,
      function(x) 1/(1 - diversity(x %>% select(-STRAT), index = 'simpson'))),
    shannon = map( #shannon diversity as effective number of species 
      data_wide,
      function(x) exp(diversity(x %>% select(-STRAT), index = 'shannon')))) %>% 
  unnest(richness,simpson, shannon)

psu_diversity_no_spp_merged = psu_diversity_no_spp_merged %>%
  unnest(data)
  select(YEAR, PRIMARY_SAMPLE_UNIT, richness,simpson, shannon)

write_csv(psu_diversity_no_spp_merged, "big_csv/diversity_without_spe/psu_diversity_no_spp_merged.csv") 

#bind lat,long, and depth to PSU 


```

```{r psu biodiversity plots and maps}

psu_fk_abun_diversity = read_csv("big_csv/diversity_without_spe/psu_fk_abun_diversity_no_spp.csv",
                                 col_types = cols(
                                   protected_status = col_character())) 

```

# Statistical analysis 

For continuous response variable (diversity - richness, shannon, simpson, functional) and categorical predictor variable (strata)

```{r librarys}

library(MASS)
library(moments)
library(DescTools)
library(rpsychi)
library(AICcmodavg) #AIC
library(ade4) #mantel.rtest
library(ggplot2) #geom_boxplot 
library(nortest) #checks for normality 
library(psych) #pairs.panels 
library(tidyverse)

```

Parametric test require the data to meet assumptions of having a normal distribution and homoscedasticity

- Are diversity values and strata related? 

```{r parametric test}

#1. Read in data 
#strata_diversity = read_csv("diversity_without_spe/strata_fk_abun_diversity_no_spp_merged.csv") 
psu_diversity = read.csv("diversity_without_spe/psu_diversity_no_spp.csv")  

#wide format 
psu_diversity = psu_diversity_no_spp_merged %>%
  unnest(data) %>%
  group_by(YEAR, PRIMARY_SAMPLE_UNIT)

# strata_simp = strata_diversity %>%
#   select(YEAR, STRAT, simpson) %>%
#   group_by(YEAR, STRAT) %>% #103 groups 
#   spread(STRAT, simpson) 
  
#2. Plot data - look at your distribution 
ggplot(data = psu_diversity) +
  geom_boxplot(mapping = aes(x = STRAT, y = simpson)) 

ggplot(data = psu_diversity) +
  geom_boxplot(mapping = aes(x = YEAR, y = simpson)) 

ggplot(data = psu_diversity) + 
  geom_density(aes(x = STRAT))

ggplot(data = psu_diversity) + 
  geom_density(aes(x = simpson))

hist(strata_diversity$simpson, main ="")

ggplot(data = psu_diversity, aes(x = YEAR)) +
  geom_point(aes(y = simpson, color = "simpson"),color = "blue") +
  geom_line(aes(y = simpson, color = "simpson"), color = "blue") +
  facet_grid(STRAT ~ .,) 

summary(psu_diversity$simpson)
#Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#1.000   5.789   8.078   8.221  10.432  22.017 

#3. Check for normality using QQ plot 
qqnorm(strata_diversity$simpson) 
qqline(strata_diversity$simpson, lty=2)  # note that lty is the line type (2=medium dashes; try different values betweem 1-6)

#4. Test whether data is NOT normal using Shapiro-Wilk Normaility Test
shapiro.test(strata_diversity$simpson)
#W = 0.95774, p-value = 0.002318

#5. Check skewness 
skew<-function(x){ 
  m3<-sum((x-mean(x))^3)/length(x) # note that the length expression automatically determines the sample size of your data
  s3<-sqrt(var(x))^3
  m3/s3} 
skew(strata_diversity$simpson) #-0.5078343

#6. Check whether the calculated skew value is significantly different than zero (i.e., not normal) using a t-test. To calulate the t-value, divide the skew by its standard error:
skew(strata_diversity$simpson)/sqrt(6/length(strata_diversity$simpson)) 
#-2.112694
# note that a rough measure of the standard error of the skewness is "sqrt{6/n}" where n is the sample size

#7. check the probability of this t-value by chance alone when the skew is zero using the following:
1-pt(-2.112694, 102) #pt=cummulative density function of t-dist, first number=calculated t-value; second=df
#0.981466

#8. Check for homoscedasticity: can use var.test, fligner.test, or bartlett.test
bartlett.test(strata_diversity$simpson~strata_diversity$YEAR) #between simpson and year 
#Bartlett's K-squared = 15.863, df = 14, p-value = 0.3218

bartlett.test(strata_diversity$simpson~strata_diversity$STRAT) #between simpson and strata
#Bartlett's K-squared = 16.359, df = 6, p-value = 0.01195

#9. Transform data using Boxcox method 
#need non-zero data for boxcox method to add a small amount to every value
strata_diversity$simpson_t = strata_diversity$simpson+0.01
b = boxcox(strata_diversity$simpson_t~strata_diversity$YEAR) 
lamda = b$x #lambda values  
lik = b$y  #y is values of liklihood %>%
bc = cbind(lamda,lik) 
bc[order(-lik),] #use the lamda with the highest likelihood to transform your data #lambda of 1.3939 is best transformation
strata_diversity$simpson_boxcox = strata_diversity$simpson^1.353 #simpson_boxcox is now transformed using the boxcox method 
strata_diversity$simpson_boxcox = strata_diversity$simpson-1
strata_diversity$simpson_boxcox = strata_diversity$simpson/1.353

#10. Check for normality of transformed data 
ggplot() + geom_density(aes(x=strata_diversity$simpson_boxcox))

qqnorm(strata_diversity$simpson_boxcox) 
qqline(strata_diversity$simpson_boxcox, lty=2)
shapiro.test(strata_diversity$simpson_boxcox)
#W = 0.95774, p-value =0.002318
skew(strata_diversity$simpson_boxcox)
#-0.50991
skew(strata_diversity$simpson_boxcox)/sqrt(6/length(strata_diversity$simpson_boxcox)) 
#-2.112694
1-pt(-2.112694, 102)
# 0.981466

# if data is normal after transformation, proceed with parametric test 

#10. Mann-Whitney or t-test parametric test 

#11. ANOVA Parametric test 
  ANOVA = aov(strata_diversity$simpson_boxcox~strata_diversity$YEAR) 
  #plot(ANOVA)
  summary(ANOVA)
#                       Df   Sum Sq    Mean Sq  F value    Pr(>F)
#strata_diversity$YEAR1 1     .5       0.471    0.047      0.829        
#Residuals              101   1020.3  10.102  
#??WHY IS DF 1 AND NOT 14?
drop1(ANOVA,~.,test="F") # type III SS and F Tests 
#                      Df Sum of Sq    RSS    AIC F value Pr(>F)
#<none>                             1020.3 240.19               
#strata_diversity$YEAR  1   0.47104 1020.8 238.24  0.0466 0.8295

ANOVA = aov(strata_diversity$simpson_boxcox~strata_diversity$STRAT) 
  #plot(ANOVA)
  summary(ANOVA)
#                       Df  Sum Sq  Mean Sq   F value   Pr(>F)    
#strata_diversity$STRAT  6  351.2   58.53     8.391     2.54e-07 ***
#Residuals              96  669.6    6.97 
  
ANOVA = aov(simpson~YEAR, data = strata_diversity) 
  #plot(ANOVA)
  summary(ANOVA)
#             Df Sum Sq Mean Sq   F value  Pr(>F)
#YEAR          1    0.9   0.862   0.047    0.829
#Residuals   101 1867.7  18.493
  
ANOVA = aov(simpson~STRAT, data = strata_diversity) 
  #plot(ANOVA)
summary(ANOVA)
#            Df Sum Sq Mean Sq   F value   Pr(>F)    
#STRAT        6  642.8  107.14   8.391     2.54e-07 ***
#Residuals   96 1225.8   12.77

# for post-hoc pairwise tests
#TukeyHSD(ANOVA) # where fit comes from aov()


```

Are there correlations among continuous data (richness, shannon, simpson, functional)?

```{r Correlations}

#Test significance of Spearman rank correlations
  correlation1 = cor.test(strata_diversity$simpson, strata_diversity$shannon, method = "spearman") 
  correlation2 = cor.test(strata_diversity$simpson, strata_diversity$functional, method = "spearman") 
  correlation3 = cor.test(strata_diversity$shannon, strata_diversity$functional, method = "spearman")  
  
#Examine correlations between matrices using Mantel's test
  mantel.rtest(strata_diversity$simpson, strata_diversity$shannon) #R^2
  mantel.rtest(simpson, functional) #R^2
  mantel.rtest(shannon, functional) #R^2

```

## Sample Data Variables
- PRIMARY_SAMPLE_UNIT: A code indicating the primary sample unit in which a sample was collected.
- YEAR: A number indicating the calendar year.
- STATION_NR: A number indicating the secondary sampling unit within a given primary sample unit.
- LAT_DEGREES: Latitude of secondary sampling unit in decimal degrees).
- LON_DEGREES: Longitude of secondary sampling unit in decimal degrees).
- DEPTH: Average depth, in meters, of secondary sampling unit).
- UNDERWATER_VISIBILITY: visibility, in meters, at secondary sampling unit..  
- HABITAT_CD: A code indicating the habitat type. 
- ZONE_NR: A code indicating the distance offshore: 
+ 1 - Inshore
+ 2 - Midchannel
+ 3 - Offshore
+ 4 - Fore-reef
- MPA_NR: A number identifying the marine protected area in which the sample was collected. Zero indicates unprotected status)
- PROT: A boolean value indicating whether a sample was in a protected area or not 
+ 1 - protected area
+ 0 - not protected 
- STRAT: A code indicating the stratum in which a sample was taken. Differs by region.   
+ FMLR
+ FSLR
+ HRRF
+ INPR
+ MCPR 
+ OFPR

```{r sample data}

RVCdata_FK = read_rds("big_csv/RVCdata_FK.rds")

FK_sample_data = RVCdata_FK$sample_data

sample_data = FK_sample_data %>%
  select(YEAR, PRIMARY_SAMPLE_UNIT, LAT_DEGREES, LON_DEGREES, HABITAT_CD, ZONE_NR, MPA_NR, STRAT) %>% #4420893*8
  dplyr::distinct() %>% #8,678*8  
  group_by(YEAR, PRIMARY_SAMPLE_UNIT) #5,241 groups 

psu_diversity = psu_diversity_no_spp_merged %>%
  select(YEAR, PRIMARY_SAMPLE_UNIT, richness, simpson, shannon) %>%
  group_by(YEAR, PRIMARY_SAMPLE_UNIT)

bind_psu_variables = full_join(sample_data, psu_diversity, by="PRIMARY_SAMPLE_UNIT")

write_csv(bind_psu_variables, "big_csv/bind_psu_variables.csv")

##why is this joining funny 
```


```{r shapefiles}
library(foreign)
dbf = read.dbf('FL_files/FLGrid_100m_work8.dbf')
dbf2 = read.dbf('FL_files/Unified_Florida_Hab.dbf')
library(maptools)
shape = readShapePoly("FL_files/Unified_Florida_Hab.shp")
library(shapefiles)
shx = read.shx("FL_files/Unified_Florida_Hab.shx")
library(spatstat)
library(lattice)
library(maps)
library(rgeos)
library(rgdal)
library(ggmap)
library(tmap)
library(marmap) #marine maps 
```

## Evironmental Variables from SERC 
SURV = col_integer(),
BASIN = col_character(),
SEGMENT = col_character(),
ZONE = col_character(),
DATE = col_character(),
TIME = col_time(format = ""),
STATION = col_integer(),
SITE = col_character(),
CHLA-B` = col_character()
LATDEC
LONDEC
DEPTH
NOX-S, NOX-B
NO3_S, NO3_B 
NO2-S, NO2-B
NH4-S, NH4-B 
TN-S, TN-B
DIN-S, DIN-B 
TON-S, TON-B
TP-S, TP-B
SRP-S, SRP-B
APA-S, APA-B
CHLA-S, CHLA-B
TOC-S, TOC-B
SiO2-S, SiO2-B
TURB-S, TURB-B
SAL-S,SAL-B
TEMP-S, TEMP-B
DO-S, DO-B
Kd pH 
TN:TP  
N:P
DIN:TP
Si:DIN
SAT-S, SAT_B
Io
DSIGT
TS, TS_B 
SIGT-S, SIGT-B
SAT-S, SAT_B

```{r SERC environmental variables}

serc_data = read_csv("big_csv/SERC_Water_Quality_FKNMS.csv")


```