---
title: "Florida Keys Reef Fish Biodiversity"
author: "Megan Hepner"
date: "April 28, 2017"
output: html_document
---

## Inputs 
1. species-trait matrix
1. domain abundance dataframe
1. strata abundance dataframe 
1. psu abundance dataframe
1. domain biomass dataframe
1. strata biomass dataframe 
1. psu biomass dataframe

## Outputs 
1.  Functional dendogram
1.  Trait-matrix visualization 
1.  By *domain* for each sampling year and region  
    + Species weighted abundance and plot
    + Species weighted biomass and plot
    + Species biodiversity 
      + species richness and plot
      + Simpson diversity as effective number of species and plot
      + Shannon diversity as effective number of species and plot
      + Functional diversity as effective number of species and plot
1. By *upper keys and lower keys domain* for each sampling year and region 
    + Species weighted abundance and plot
    + Species weighted biomass and plot
    + Species biodiversity 
      + species richness and plot
      + Simpson diversity as effective number of species and plot
      + Shannon diversity as effective number of species and plot
      + Functional diversity as effective number of species and plot
1. By *strata* for each sampling year and region 
    + Species weighted abundance and plot
    + Species weighted biomass and plot
    + Species biodiversity 
      + species richness and plot
      + Simpson diversity as effective number of species and plot
      + Shannon diversity as effective number of species and plot
      + Functional diversity as effective number of species and plot
1. By *primary sampling unit* for each sampling year and region and plot
    + Species weighted abundance and plot
    + Species weighted biomass and plot
    + Species biodiversity 
      + species richness and plot
      + Simpson diversity as effective number of species and plot
      + Shannon diversity as effective number of species and plot
      + Functional diversity as effective number of species and plot
    + statistical tests by strata and year: Shapiro, homocasdicity, 2-way ANOVA
1. Environmental factors from SERC 
1. Heat map of the Florida Keys and Dry Tortugas for the diversity indices 

## Reef Visual Census Background 

Reef fish community data used in this study were collected as part of the multi-agency Reef Visual Census (RVC) [Brandt et al., 2009](https://www.coris.noaa.gov/activities/fish_monitoring_protocol/). Fish communities were visually surveyed underwater annually in the mainland FKNMS (Lower Keys, Middle Keys and Upper Keys) from 1999 to 2012, and biennially from 2012 to 2016. The Dry Tortugas region was surveyed from 1999 to 2000 and biennially from 2004 to 2016.     

The RVC uses a stationary point count method within a randomly selected 7.5 m radius circular plot with a 200m map grid from 1999 to 2012 and 100m map grid from 2014 to 2016 to optimize the observation of conspicuous and diurnally active reef fish, specifically economically and ecologically important reef fish species [Bohnsack and Bannerot, 1986](https://docs.lib.noaa.gov/noaa_documents/NMFS/TR_NMFS/TR_NMFS_41.pdf). 

## Objective 

To analyze and compare changes in species richness, Simpson diversity, Shannon diversity and functional diversity in the Florida Keys National Marine Sanctuary (FKNMS) to reveal changes in temporal and spatial variability from 1999 – 2016. The indices were assessed throughout the Florida Keys, between the upper and lower keys, between habitat types, and for different levels of management (Ecological Reserves (ER), Sanctuary Preservations Areas (SPA), and Special Use/Research Only Areas (SU/RO)). 

## Diversity indicse as effective number of species 

All indices were computed as effective number of species. Effective number of species is the number of equally abundance species needed to produce the observed value of a diversity index (Jost, 2006; Jost et al., 2010; MacArthur, 1965). Jost, 2006 refers to these as true measures of diversity, where prior to the conversion the diversity indices are simply measures of entropy. 

- Species Richness is the number of species in a habitat sampled  

    $Richness = \sum_{i=1}^S p_i^0$  

- Simpson diversity measures the probability that two individuals randomly selected from a sample will belong to different species.   

    $Simpson = 1 - \sum_{i=1}^S p_i^2$  

- Shannon diveristy is a measure of entropy, it is the uncertainty in the species identity of a sample  

    $Shannon = - \sum_{i=1}^S p_i \log_b p_i$  

- Functional diversity (Rao's Q) is a measure of pairwise functional differences between species weighted by their relative abundances   

    $Rao Q = \sum_{i=1}^S-1 \sum_{j=i+1}^S d_i p_i p_j$
    
## Functional distance trait based matrix (from Lefcheck et al.,)

Gowdis computes the Gower (1971) similarity coefficient exactly as described by Podani (1999), then converts it to a dissimilarity coefficient by using D = 1 - S

```{r Functional distance matrix}

#install.packages("clue")
library(clue) #calls cl_ultrametric 
#install.packages("fpc")
library(fpc) #calls pamk
#install.packages("ape")
library(ape)
#install.packages("phytools")
library(phytools)
#install.packages("FD")
library(FD)
## biocLite("BiocUpgrade") ## you may need this
#biocLite("ggtree", type = "source")
devtools::install_github("GuangchuangYu/ggtree") #requires R>= 3.3.2

trait_matrix = read_csv('functional_diversity/species_trait_matrix_348_spp.csv')

# Parsed with column specification:
# cols(
#   Species_code = col_integer(),
#   SPECIES_CD = col_character(),
#   Latin_name = col_character(),
#   Common_name = col_character(),
#   Maxlength = col_double(),
#   Trophic_level = col_double(),
#   Trophic_group = col_character(),
#   Water_column = col_character(),
#   Diel_activity = col_character(),
#   Substrate_type = col_character(),
#   Complexity = col_character(),
#   Gregariousness = col_integer()

trait_matrix = trait_matrix %>%
  as.tibble() %>%
  mutate(
    SPECIES_CD = toupper(as.character(SPECIES_CD))) %>% 
  arrange(SPECIES_CD)%>% 
  select(SPECIES_CD, Maxlength, Trophic_level, Trophic_group, Water_column, Diel_activity, Substrate_type, Complexity, Gregariousness) %>% #348*9
  group_by(
    SPECIES_CD, Maxlength, Trophic_level, Trophic_group, Water_column, Diel_activity, Substrate_type, Complexity, Gregariousness) %>%
  summarize(
    n = n()) %>%
  select(-n) %>%
  ungroup() %>%
  mutate(
    # ordinal traits
    Complexity = ordered(Complexity, levels=c("Low","Medium","High")),
    Gregariousness = ordered(Gregariousness, levels=c("1","2","3"))) %>%
  as.data.frame()

#Set species names as row.names and remove extra columns
rownames(trait_matrix)=trait_matrix$SPECIES_CD 

#Calculate Gower distances. Variables have equal weight. 
traits.dist = gowdis(trait_matrix, ord="podani")

#Use clustering method to produce ultrametric dendrogram because values of Rao's Q can be maximized when fewer than the max number of functional types are present unless distances are ultramtetric 

#to account for sensitivity in clustering use multiple algorithms  (Mouchet et al., 2008) 
tree_methods = c("single","complete","average","mcquitty","ward.D") #average is best clustering method 
trees=lapply(tree_methods,  function(i) hclust(traits.dist, method=i))
par(mfrow=c(3,2))
for(i in 1:length(trees)) {plot(trees[[i]])} #plots the 5 different kinds of trees 

#convert trees to ultrametric
trees.ultra= lapply(trees,function(i) cl_ultrametric(as.hclust(i)))

#Plot each tree
par(mfrow=c(3,2))
for (i in 1:length(trees.ultra)) {plot(trees.ultra[[i]])} #plots the 5 different kinds of trees 

#Build the consensus tree (Mouchet et al 2008 Oikos)  
ensemble.trees=cl_ensemble(list=trees) #list of clusterings 
class(ensemble.trees)
consensus.tree=cl_consensus(ensemble.trees) #synthesizes the information in the elements of a cluster ensemble into a single clustering 
par(mar=c(1,1,1,1))
plot(consensus.tree, horiz=T)

#Calculate dissimilarity values for each tree using 2-norm (Merigot et al 2010 Ecology) to determine which tree best preserves orignial distances 
#spectral norm (2-norm) of the differences of the ultrametrics
all.trees=c(trees.ultra,consensus.tree[1])
names(all.trees)=c(tree_methods,"consensus")
(trees.dissim=lapply(all.trees,function(i) cl_dissimilarity(i,traits.dist,method="spectral"))) 
#Cross-dissimilarities using spectral ultrametric distance:
#Single = 63.17399
#complete = 116.8516
#average = 14.99288
#mcquitty = 36.06599
#ward.D = 3161.581
#consensus =  1065.058 (1067.782) (1091.628)

#Identify best tree and isolate
trees.dissim2=do.call(rbind,trees.dissim)
min.tree=which.min(trees.dissim2) #average 
names(all.trees)[min.tree]
func.dist=all.trees[names(all.trees)==names(all.trees)[min.tree]][[1]]

#Confirm lowest 2-norm value,  spectral norm (2-norm) of the differences if the ultrametrics 
cl_dissimilarity(func.dist,traits.dist,method="spectral") #Cross-dissimilarities using spectral ultrametric distance:  14.99288

#Scale by the max value so that all values are between 0-1
func.dist=func.dist/max(func.dist)

#Plot the best tree - average 
par(mfrow=c(1,1))
par(mar=c(2,1,0,2))
par(pin=c(10,15))
plot(func.dist,horiz=TRUE)
#Save plot: 10" x 15"

#Phylogenetic tree with 348 tips and 347 internal nodes. 
func.dist.phy = as.phylo(as.hclust(func.dist))

#Write newick tree
write.tree(func.dist.phy, "functional_diversity/functional_dendrogram.nwk")
#write.nexus(func.dist.phy, "functional_diversity/functional_dendrogram.nex") 

tree <- read.tree("functional_diversity/functional_dendrogram.nwk")
# List of 4
#  $ edge       : int [1:694, 1:2] 
#  $ Nnode      : int 347
#  $ tip.label  : chr [1:348] 
#  $ edge.length: num
 # - attr(*, "class")= chr "phylo"
 # - attr(*, "order")= chr "cladewise"

# Phylogenetic tree with 348 tips and 347 internal nodes.
# Tip labels:
# 	KYP_SECT, ALE_CILI, ELO_SAUR, HAR_JAGU, CEN_UNDE, TYL_CROC, ...
# Rooted; includes branch lengths.

plot(tree, label.offset=0.2, edge.width = 2, type = "cladogram")  
nodelabels() #add node numbers
tiplabels() #add tip numbers

ggtree(tree, branch.length = T) +
  ggtitle("Functional Dendogram") +
  geom_tiplab()+
  geom_text(aes(x = branch, label=branch.length), vjust = -.5) +
  theme_tree2()

ggtree(tree, layout= "circular") +
  ggtitle("Functional Dendogram") +
  geom_tiplab(size = 1, aes(angle=angle))+
  theme_tree2()

# transformed to a tidy data.frame by fortify method
tree_data <- fortify(tree) #node; parent; branch.length; x; y; label; isTip; branch; angle 
head(tree_data)

#Visualize species' differences in multivariate trait space
#Perform k-means clustering with no a priori specification for k
traits.kclus= pamk(traits.dist, krange=2:10) 

#krange: integer vector. Numbers of clusters which are to be compared by the average silhouette width criterion.
#traits.kclus$nc = 10 
#traits.kclus$crit 0.0000000 0.2374706 0.2362513 0.2355555 0.2681756 0.2581352 0.2759547 0.2852216 0.3006077 0.3062123

# #Perform multidimensional scaling on functional dendrogram
traits_nmds = metaMDS(traits.dist,k=traits.kclus$nc,trymax=20)
#k = Number of dimensions

#*** No convergence -- monoMDS stopping criteria:
   #499: no. of iterations >= maxit
     #1: stress ratio > sratmax

# Dimensions: 10 
# Stress:     0.03545178 
# Stress type 1, weak ties
# No convergent solutions - best solution after 500 tries
# Scaling: centring, PC rotation 
# Species: scores missing


# #Plot in two dimensions - doesn't work 
par(mar=c(4,4,1,1))
ordiplot(traits_nmds,type="n")

#Assign colors to different groups
groups=levels(factor(traits.kclus$pamobject$clustering))
points.symbols=15:16
points.colors=c("firebrick3","cornflowerblue")
for(i in seq_along(groups)) {
   points(traits_nmds$points[traits.kclus$pamobject$clustering==groups[i],],
          pch=points.symbols[i],col=points.colors[i],cex=1.4) }
 ordispider(traits_nmds,factor(traits.kclus$pamobject$clustering),label=F)
 ordihull(traits_nmds,factor(traits.kclus$pamobject$clustering),lty="dotted")
 orditorp(traits_nmds,dis="sites",pcex=0,air=0.5,col="grey10",cex=0.8)

```

## Trait matrix visualization 

[http://jkunst.com/r/pokemon-visualize-em-all/]
[https://d3js.org/]
[http://www.buildingwidgets.com/blog/2015/7/22/week-29-d3treer-v2]

```{r trait visualization}

library(treemap)
devtools::install_github("gluc/data.tree")
devtools::install_github("timelyportfolio/d3treeR")
library(d3treeR)

trait_matrix = read_csv('functional_diversity/species_trait_matrix_348_spp.csv')

trait_matrix_treemap = trait_matrix %>%
  select(Common_name, Maxlength, Trophic_level, Trophic_group, Water_column,Diel_activity,Substrate_type,Complexity, Gregariousness) %>% 
  group_by(Maxlength,
           Trophic_level,
           Trophic_group,
           Water_column,
           Diel_activity,
           Substrate_type,
           Complexity,
           Gregariousness) %>%
  summarise(n = n())%>% 
  ungroup() %>%
  as.data.frame()
  
treemap2 = 
  d3treeR::d3tree2(
    treemap(dtf = trait_matrix_treemap,
            index = c("Trophic_group", "Trophic_level", "Water_column", "Complexity", "Substrate_type", "Diel_activity"),
            vSize = "n", 
            vColor = "Trophic_group",
            type = "index"),
            #title = "Reef Fish Traits",
            #fontsize.title = 14, 
            #algorithm = "squarified"
    rootname = "Species Traits")

```

## Fetch the RVC data from 
[NOAA's Southeast Fisheries Science Center server](https://grunt.sefsc.noaa.gov/rvc_analysis20/) using the [RVC package](https://github.com/jeremiaheb/rvc)
- Fetched weighted abundance and biomass data by domain, strata, and primary sampling units 

### Sample Data Variables
- PRIMARY_SAMPLE_UNIT: A code indicating the primary sample unit in which a sample was collected.
- YEAR: A number indicating the calendar year.
- MONTH: A number indicating the month of the year.
- DAY: A number indicating the day of the month (EST).
- STATION_NR: A number indicating the secondary sampling unit within a given primary sample unit.
- LAT_DEGREES: Latitude of secondary sampling unit in decimal degrees).
- LON_DEGREES: Longitude of secondary sampling unit in decimal degrees).
- DEPTH: Average depth, in meters, of secondary sampling unit).
- UNDERWATER_VISIBILITY: visibility, in meters, at secondary sampling unit.
- MAPGRID_NR: A number indicating the primary sample unit.  
- HABITAT_CD: A code indicating the habitat type. 
- ZONE_NR: A code indicating the distance offshore: 
    + 1 - Inshore
    + 2 - Midchannel
    + 3 - Offshore
    + 4 - Fore-reef
- SUBREGION_NR: A number indicating the subregion.
- MPA_NR: A number identifying the marine protected area in which the sample was collected. Zero indicates unprotected status)
- SPECIES_NR: A number indicating the species for a sample 
- SPECIES_CD: A code indicating the species for a sample. Consists of the first three letters of the generic name and the first four of the specific name   
- LEN: The length, in cm, of a sample.      
- NUM: The number of individuals of a given species and length observed in a secondary sampling unit
- TIME_SEEN: A number indicating when, during sampling, an individual was observed. 1: In the first five minutes, 2: From 5-10 minutes, 3: After 10 minutes. 
- PROT: A boolean value indicating whether a sample was in a protected area or not 
    + 1 - protected area
    + 0 - not protected 
- STRAT: A code indicating the stratum in which a sample was taken. Differs by region.   
    + FMLR
    + FSLR
    + HRRF
    + INPR
    + MCPR 
    + OFPR
- REGION: A code indicating the region in which a sample was taken.
    + FLA KEYS (florida keys)
    + DRY TORT (dry tortugas)

### Stratum Data Variables 
- REGION 
- YEAR 
- PROT 
- STRAT 
- NTOT: The number of possible primary sample units for a given year, region, stratum, and protected status 
- GRID_SIZE: The length (in meters) to a side of a primary sample unit for a given year, region, stratum, and protected status.

### Taxonomic Data Variables 
- SPECIES_CD
- FAMILY
- SCINAME 
- COMNAME
- LC: Minimum length at capture, in centimeters
- LM: Median length at maturity, in centimeters.
- WLEN_A: The linear coefficient of the allometric growth equation in grams per millimeter (g/mm)
- WLEN_B: The exponential coefficient of the allometric growth equation

### Benthic Data Variables 
- REGION: A code indicating the region. 
- YEAR: A number indicating the calendar year.
- PRIMARY_SAMPLE_UNIT: A code indicating the primary sample unit in which a sample was collected.
- STATION_NR: A number indicating the secondary sampling unit within a given primary sample unit.
- DEPTH: Average depth, in meters, of secondary sampling unit.
- MAX_HARD_RELIEF: The maximum height, in meters, of hard relief (e.g. hard corals, rock).
- MAX_SOFT_RELIEF: The maximum height, in meters, of soft relief (e.g. soft corals, sponges).
- AVG_HARD_RELIEF: The average height, in meters, of hard relief.
- HARD_REL_PCT_0: Percentage of hard relief less than 0.2 meters.
- HARD_REL_PCT_1: Percentage of hard relief between 0.2 and 0.5 meters.
- HARD_REL_PCT_2: Percentage of hard relief between 0.5 and 1.0 meters.
- HARD_REL_PCT_3: Percentage of hard relief between 1.0 and 1.5 meters.
- HARD_REL_PCT_4: Percentage of hard relief greater than 1.5 meters.
- PCT_SAND: Percentage of abiotic cover that is sand.
- PCT_HARD_BOTTOM: Percentage of abiotic cover that is hard bottom.
- PCT_RUBBLE: Percentage of abiotic cover that is rubble.
- PCT_CORAL: Percentage of biotic hardbottom that is coral.
- PCT_OCTO: Percentage of biotic hardbottom that is octocoral.
- PCT_SPONGE: Percentage of biotic hardbottom that is sponge.

### Function: getDomainDensity 
- YEAR
- REGION
- SPECIES_CD
- density: Domain-wide mean density for a stratified random survey  
- var: Variance in average density per secondary sampling unit
- n: Number of primary sampling units sampled
- nm: Number of secondary sampling units sampled
- N: Number of total possible primary sample units
- NM: Number of possible secondary sampling units
- length_class: The length class or bin. Only present if length_bins is not NULL. The notation, [lower, upper), is inclusive of the lower bound, but exclusive of the upper bound
- protected_status: The protected status. Only present if merge_protected is FALSE

```{r setup, include= F}
knitr::opts_chunk$set(echo = T)
#rm(list = ls())

#install.packages('devtools')
#devtools::install_github('jeremiaheb/rvc')
library(rvc) #calls RVC data from SEFSC server 
#install.packages("tidyverse")
library(tidyverse)
#install.packages("vegan")
library(vegan) # biodiveristy functions for richness, simpson, and shannon 
#install.packages("dygraphs")
library(dygraphs)
#install.packages("plotly")
library(plotly)
#install.packages("webshot")
library(webshot)
#install.packages("htmlwidgets")
library(htmlwidgets)
#install.packages("FD")

map = purrr::map # override maps::map
select = dplyr::select #override MASS::select 
```

### Florida Keys domain

```{r domain abundance}

domain_fk_abun <- read_csv('big_csv/abundance_domain/domain_fk_abun.csv', 
col_types = cols(
  YEAR = col_integer(),
  REGION = col_character(),
  SPECIES_CD = col_character(),
  abundance = col_double(),
  var = col_double(),
  n = col_integer(),
  nm = col_integer(),
  N = col_double(),
  NM = col_double(),
  protected_status = col_character()
))

domain_fk_abun_no_spp = domain_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species

```

```{r domain biomass}

domain_fk_bio <- read_csv('big_csv/biomass_domain/domain_fk_biomass.csv', 
col_types = cols(
  protected_status = col_character()
))

domain_fk_bio_no_spp = domain_fk_bio %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species
```

```{r domain biodiversity}

trait_matrix = read_csv('functional_diversity/species_trait_matrix_348_spp.csv')

domain_fk_abun <- read_csv('big_csv/abundance_domain/domain_fk_abun.csv', 
col_types = cols(
  YEAR = col_integer(),
  REGION = col_character(),
  SPECIES_CD = col_character(),
  abundance = col_double(),
  var = col_double(),
  n = col_integer(),
  nm = col_integer(),
  N = col_double(),
  NM = col_double(),
  protected_status = col_character()
))

domain_fk_abun_no_spp = domain_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species 

#length(unique(domain_fk_abun_no_spp$SPECIES_CD)) # 348
#length(unique(trait_matrix$SPECIES_CD))          # 348

# check if one species has total abundance of zero (never occurs)
abun.sum2 <- apply(a, 2, sum)
if (any(abun.sum2 == 0) ) stop("At least one species does not occur in any community (zero total abundance across all communities).","\n")

## FK's domain abundance only identified to species level (without spp.) 
fk_domain_abun_diversity_no_spp = domain_fk_abun_no_spp %>% #14,634 * 10
  select(YEAR, protected_status, SPECIES_CD, abundance) %>% #14,634 * 4
  group_by(YEAR, protected_status) %>% ##45 groups (fk 15*3)
  nest(-YEAR) %>%  
  mutate(
    data_wide = map(data, function(x) 
      full_join(x, trait_matrix %>% select(SPECIES_CD), by='SPECIES_CD') %>%
        mutate(abundance = ifelse(is.na(abundance), 0, abundance)) %>%
        spread(SPECIES_CD, abundance, fill =0)), 
    richness = map( #species richness 
      data_wide, 
      function(x) specnumber(x)),
    simpson = map( #simpson diversity as effective number of species 
      data_wide,
      function(x) 1/(1 - diversity(x, index = 'simpson'))),
    shannon = map( #shannon diversity as effective number of species 
      data_wide,
      function(x) exp(diversity(x, index = 'shannon')))) %>% #,
    # func_div = map( 
    #   data_wide, 
    #   function(x) 1/(1 - dbFD(as.matrix(func.dist), x)))) %>% 
  unnest(richness,shannon,simpson)

fk_domain_abun_diversity_no_spp = fk_domain_abun_diversity_no_spp %>%
  select(YEAR, protected_status,richness,shannon,simpson)

write_csv(fk_domain_abun_diversity_no_spp, 'big_csv/diversity_without_spe/domain_fk_abun_diversity_no_spp.csv')

```


```{r dygraphs}

domain_fk_abun_diversity = read_csv('big_csv/abundance_domain/domain_fk_abun_diversity.csv')

## Simpson for protected and unprotected stacked 
fk_simpson = domain_fk_abun_diversity %>% 
  filter(protected_status != 'all') %>%
  select(YEAR, protected_status, dom_simpson) %>%
  mutate(
    protected_status = recode(
      protected_status, '0'='Unprotected', '1'='Protected')) %>%
  spread(protected_status, dom_simpson)

dygraph(fk_simpson, main = "Simpson Reef Fish Diversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 50)) %>%
  dyAxis("x", label = "Year") %>%
  dyOptions(stackedGraph = T) 
  #dyRangeSelector(height = 20) 

## Shannon for protected and unprotected stacked 
fk_shannon = domain_fk_abun_diversity %>% 
  filter(protected_status != 'all') %>%
  select(YEAR, protected_status, dom_shannon) %>%
  mutate(
    protected_status = recode(
      protected_status, '0'='Unprotected', '1'='Protected')) %>%
  spread(protected_status, dom_shannon)

dygraph(fk_shannon, main = "Shannon Reef Fish Diversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 75)) %>%
  dyAxis("x", label = "Year") %>%
  dyOptions(stackedGraph = T)
  #dyRangeSelector(height = 20) 


d = domain_fk_abun_diversity %>% 
  filter(protected_status != 0) %>% 
  filter(protected_status != 1)

#Simpson for "all" filled 
dygraph_simp = d %>%
  select(YEAR, dom_simpson)

dygraph(dygraph_simp, main = "Simpson Diversity of Reef Fish in FKNMS") %>%
  dyOptions(fillGraph = T, fillAlpha = 0.4, drawPoints = T, pointSize = 2) %>%
  dySeries("dom_simpson", label = "Simpson Diversity") %>% 
  dyAxis("y", label = "Effective Number of Species") %>% 
  dyAxis("x", label = "Year")

#Shannon for "all" filled 
dygraph_shannon = d %>%
  select(YEAR, dom_shannon)

dygraph(dygraph_shannon, main = "Shannon Diversity of Reef Fish in FKNMS") %>%
  dyOptions(fillGraph = T, fillAlpha = 0.4, drawPoints = T, pointSize = 2) %>%
  dySeries("dom_shannon", label = "Shannon Diversity") %>% 
  dyAxis("y", label = "Effective Number of Species") %>% 
  dyAxis("x", label = "Year")

#Richness for "all" filled 
dygraph_richness = d %>%
  select(YEAR, dom_richness)

dygraph(dygraph_richness, main = "Species Richness of Reef Fish in FKNMS") %>%
  dyOptions(fillGraph = T, fillAlpha = 0.4, drawPoints = T, pointSize = 2) %>%
  dySeries("dom_richness", label = "Species Richness") %>% 
  dyAxis("y", label = "Number of Species") %>% 
  dyAxis("x", label = "Year")

## Simpson and Shannon stacked for "all"
dygraph_simp_shannon = d %>%
  select(YEAR, dom_simpson, dom_shannon)

dygraph(dygraph_simp_shannon, main = "Florida Keys Reef Fish Biodiversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 70)) %>%
  dyAxis("x", label = "Year") %>%
  dySeries("dom_shannon", label = "Shannon Diversity", color = "blue") %>% 
  dySeries("dom_simpson", label = "Simpson Diversity", color = "red") %>% 
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 1.5) %>% #drawGrid = F
  dyLegend(width = 400)

# m <- plot_ly(x = 1:10)
# saveWidget(as.widget(m), "temp.html")
# webshot("temp.html", file = "test.png",
#         cliprect = "viewport")

#Richness, Shannon, Simpson stacked for "all"
dygraph_rich_simp_shannon = d %>%
  select(YEAR, dom_simpson, dom_shannon, dom_richness)

dygraph(dygraph_rich_simp_shannon, main = "Florida Keys Reef Fish Biodiversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 350)) %>% 
  dyAxis("x", label = "Year") %>%
  dySeries("dom_richness", label = "Richness", color = "purple") %>%
  dySeries("dom_shannon", label = "Shannon", color = "blue") %>% 
  dySeries("dom_simpson", label = "Simpson", color = "red") %>% 
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 1.5) %>% #drawGrid = F
  dyLegend(width = 400)

##without spp 
d_no_spp = fk_domain_abun_diversity_no_spp %>% 
  filter(protected_status != 0) %>% 
  filter(protected_status != 1)

dygraph_rich_simp_shannon_no_spp = d_no_spp %>%
  select(YEAR, simpson, shannon, richness)

dygraph(dygraph_rich_simp_shannon_no_spp, main = "Florida Keys Reef Fish Biodiversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 350)) %>% 
  dyAxis("x", label = "Year") %>%
  dySeries("richness", label = "Richness", color = "purple") %>%
  dySeries("shannon", label = "Shannon", color = "blue") %>% 
  dySeries("simpson", label = "Simpson", color = "red") %>% 
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 1.5) %>% #drawGrid = F
  dyLegend(width = 400)

dygraph_simp_shannon_no_spp = d_no_spp %>%
  select(YEAR, simpson, shannon)

dygraph(dygraph_simp_shannon_no_spp, main = "Florida Keys Reef Fish Biodiversity") %>% 
  dyAxis("y", label = "Effective Number of Species") %>%  #valueRange = c(0, 350)
  dyAxis("x", label = "Year") %>%
  dySeries("shannon", label = "Shannon", color = "blue") %>% 
  dySeries("simpson", label = "Simpson", color = "red") %>% 
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 1.5) %>% #drawGrid = F
  dyLegend(width = 400)

```

```{r domain biodiversity plots}

domain_fk_abun_diversity = read_csv('big_csv/abundance_domain/domain_fk_abun_diversity.csv')
# Parsed with column specification:
# cols(
#   YEAR = col_integer(),
#   protected_status = col_character(),
#   dom_richness = col_integer(),
#   dom_simpson = col_double(),
#   dom_shannon = col_double()
# )
protected_status_names <- c(
  "0" = "Unprotected",
  "1" = "Protected",
  "all" = "All"
  )
#richness, simpson, shannon for domain FKNMS  
FK_abun_RShSim <- ggplot(data = fk_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_richness, color = "dom_richness")) +
  geom_point(aes(y = dom_simpson, color = "dom_simpson")) +
  geom_point(aes(y = dom_shannon, color = "dom_shannon")) +
  geom_line(aes(y = dom_richness, color = "dom_richness")) +
  geom_line(aes(y = dom_simpson, color = "dom_simpson")) +
  geom_line(aes(y = dom_shannon, color = "dom_shannon")) +
  facet_grid(protected_status ~ ., labeller = as_labeller(protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Diversity indices") +
  scale_colour_manual(values = c("darkturquoise", "magenta", "blue"), labels=c("Species Richness", "Shannon Diversity", "Simpson Diversity")) +
  ggtitle("Reef Fish Biodiversity in the FKNMS") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(breaks = c(1999:2016))+
  scale_y_continuous(limits= c(0, 300), breaks = c(0,25,50,75,100,125,150,175,200,225,250,275,300))


#simpson, shannon by protected status for domain FKNMS
FK_abun_ShSim <-ggplot(data = fk_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_simpson, color = "dom_simpson")) +
  geom_point(aes(y = dom_shannon, color = "dom_shannon")) +
  geom_line(aes(y = dom_simpson, color = "dom_simpson")) +
  geom_line(aes(y = dom_shannon, color = "dom_shannon")) +
  facet_grid(protected_status ~ ., labeller = as_labeller(protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Diversity index") +
  scale_colour_manual(values = c("magenta", "blue"), labels=c("Shannon diversity", "Simpson Diversity")) +
  ggtitle("Reef Fish Biodiversity in the Florida Keys") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 250), breaks = c(5,10,15,20,25,30,35,40,45,50))

#species richness by protected status for domain FKNMS
FK_abun_Rich <- ggplot(data = fk_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_richness, color = "dom_richness"),color = "darkturquoise") +
  geom_line(aes(y = dom_richness, color = "dom_richness"), color = "darkturquoise") +
  facet_grid(protected_status ~ ., labeller = as_labeller(protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Level of protection") +
  ggtitle("Species Richness of Reef Fish in the Florida Keys") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 250), breaks = c(0,50, 100, 150, 200, 250, 300))
## wonky 

#shannon by protected status for domain FKNMS
FK_abun_Shan <- ggplot(data = fk_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_shannon, color = "dom_shannon"),color = "magenta") +
  geom_line(aes(y = dom_shannon, color = "dom_shannon"), color = "magenta") +
  facet_grid(protected_status ~ ., labeller = as_labeller(protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Level of protection") +
  #scale_colour_manual(labels=c("Unprotected", "Protected", "All")) +
  ggtitle("Shannon Diveristy of Reef Fish in the Florida Keys") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 50), breaks = c(5,10,15,20,25,30,35,40,45,50))

#simpson by protected status for domain FKNMS
FK_abun_Simp <- ggplot(data = fk_domain_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = dom_simpson, color = "dom_simpson"),color = "blue") +
  geom_line(aes(y = dom_simpson, color = "dom_simpson"), color = "blue") +
  facet_grid(protected_status ~ ., labeller = as_labeller(protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Level of protection") +
  #scale_colour_manual(labels=c("Unprotected", "Protected", "All")) +
  ggtitle("Simpson Diveristy of Reef Fish in the Florida Keys") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 25), breaks = c(5,10,15,20,25))

```

### Florida Keys Strata
1. FSLR - Forereef Shallow Linear Reef (<6m)
2. FMLR - Forereef Midchannel Linear Reef (6-18m)
3. FDLR - Forereef Deep Linear Reef (18-33m)
4. INPR - Inshore patch reef
5. MCPR - Midchannel Patch Reef 
6. OFPR - Offshore Patch Reef
7. HRRF - High Relief Reef (Spur and Groove)

```{r strata abundance}
## FK's abundance by strata 
strata_fk_abun <- read_csv("big_csv/abundance_strata/strata_fk_abun.csv",
col_types = cols(
  YEAR = col_integer(),
  REGION = col_character(),
  STRAT = col_character(),
  SPECIES_CD = col_character(),
  abundance = col_double(),
  var = col_double(),
  n = col_integer(),
  nm = col_integer(),
  N = col_double(),
  NM = col_double(),
  protected_status = col_character()
))

strata_fk_abun_no_spp = strata_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species 
```

```{r strata biomass}

strata_fk_bio <- read_csv('big_csv/biomass_strata/strata_fk_biomass.csv', 
col_types = cols(
  protected_status = col_character()
))

strata_fk_bio_no_spp = strata_fk_bio %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species
```

```{r strata biodiversity}

## FK's abundance by strata 
strata_fk_abun <- read_csv("big_csv/abundance_strata/strata_fk_abun.csv",
col_types = cols(
  YEAR = col_integer(),
  REGION = col_character(),
  STRAT = col_character(),
  SPECIES_CD = col_character(),
  abundance = col_double(),
  var = col_double(),
  n = col_integer(),
  nm = col_integer(),
  N = col_double(),
  NM = col_double(),
  protected_status = col_character()
))

strata_fk_abun_no_spp = strata_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species 

fk_strata_abun_diversity_no_spp = strata_fk_abun_no_spp %>% #133,762 * 12
  filter(protected_status!='all') %>% #66,881 * 12
  select(YEAR, STRAT, protected_status, SPECIES_CD, abundance) %>% #66,881 * 5
  group_by(YEAR, STRAT, protected_status) %>% #294 groups (15 years *7 strata *3prot = 315)??
  nest() %>%  # 294 * 4 
  mutate(
    data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill =0)), 
    richness = map( #species richness
      data_wide, 
      function(x) specnumber(x)),
    simpson = map( #simpson diversity as effective number of species 
      data_wide,
      function(x) 1/(1 - diversity(x, index = 'simpson'))),
    shannon = map( #shannon diversity as effective number of species 
      data_wide,
      function(x) exp(diversity(x, index = 'shannon')))) %>%
  unnest(richness, simpson, shannon)

fk_strata_abun_diversity_no_spp = fk_strata_abun_diversity_no_spp %>%
  select(YEAR, STRAT, protected_status, richness, simpson, shannon) 

write_csv(fk_strata_abun_diversity_no_spp, 'big_csv/diversity_without_spe/strata_fk_abun_diversity_no_spp.csv')

```

```{r strata biodiversity plots, eval = F}

strata_fk_abun_diversity = read_csv("big_csv/abundance_strata/strata_fk_abun_diversity.csv")

# strata_names <- c(
#  "FMLR" = "Forereef Midchannel Linear Reef",
#  "FSLR" = "Forereef Shallow Linear Reef",
#  "HRRF" = "High Relief Reef", #(Spur and Groove)
#  "INPR" = "Inshore patch reef",
#  "MCPR" = "Midchannel Patch Reef",
#  "OFPR" = "Offshore Patch Reef",
#  "FDLR" = "Forereef Deep Linear Reef")

##protected by strata 
fk_strata_protected = strata_fk_abun_diversity %>% 
  filter(protected_status != 1)

#richness protected by strata 
fk_rich_protected = fk_strata_protected %>% 
  select(YEAR, STRAT, richness) %>%
  spread(STRAT, richness)

dygraph(fk_rich_protected, main = "Florida Keys Richness by Strata - NTMR") %>% 
  dyAxis("y", label = "Effective Number of Species") %>%
  dyAxis("x", label = "Year") %>% 
  dyOptions(stackedGraph = T, colors = colorRamps::blue2red(7), fillAlpha = .8, axisLineWidth = 2)

##shannon protected by strata 
fk_shan_protected = fk_strata_protected %>% 
  select(YEAR, STRAT, shannon) %>%
  spread(STRAT, shannon)

dygraph(fk_shan_protected, main = "Florida Keys Shannon Diversity by Strata - NTMR") %>% 
  dyAxis("y", label = "Effective Number of Species") %>%
  dyAxis("x", label = "Year") %>% 
  dyOptions(stackedGraph = T, colors = colorRamps::blue2red(7), fillAlpha = .8, axisLineWidth = 2)

##simpson protected by strata 
fk_simp_protected = fk_strata_protected %>% 
  select(YEAR, STRAT, simpson) %>%
  spread(STRAT, simpson)

dygraph(fk_simp_protected, main = "Florida Keys Simpson Diversity by Strata - NTMR") %>% 
  dyAxis("y", label = "Effective Number of Species") %>%
  dyAxis("x", label = "Year") %>% 
  dyOptions(stackedGraph = T, colors = colorRamps::blue2red(7), fillAlpha = .8, axisLineWidth = 2)#topo.colors(n = 7, alpha = 1) RColorBrewer::brewer.pal(7, "Paired") grDevices::topo.colors(7, alpha = 1) grDevices::rainbow(7) colorspace::rainbow_hcl(7) colorspace::heat_hcl(7)

#unprotected by strata 
fk_strata_unprotected = strata_fk_abun_diversity %>% 
  filter(protected_status != 0)

#richness unprotected by strata 
fk_rich_unprotected = fk_strata_unprotected %>% 
  select(YEAR, STRAT, richness) %>%
  spread(STRAT, richness)

dygraph(fk_rich_unprotected, main = "Florida Keys Richness by Strata - Unprotected") %>% 
  dyAxis("y", label = "Effective Number of Species") %>%
  dyAxis("x", label = "Year") %>% 
  dyOptions(stackedGraph = T, colors = colorRamps::blue2red(7), fillAlpha = .8, axisLineWidth = 2)

#simpson unprotected by strata
fk_simp_unprotected = fk_strata_unprotected %>% 
  select(YEAR, STRAT, simpson) %>%
  spread(STRAT, simpson)

dygraph(fk_simp_unprotected, main = "Florida Keys Simpson Diversity by Strata - Unprotected") %>% 
  dyAxis("y", label = "Effective Number of Species") %>%
  dyAxis("x", label = "Year") %>% 
  dyOptions(stackedGraph = T, colors = colorRamps::blue2red(7), fillAlpha = .8, axisLineWidth = 2)

strata_names <- c(
 "FMLR" = "FMLR",
 "FSLR" = "FSLR",
 "HRRF" = "HRRF", #(Spur and Groove)
 "INPR" = "INPR",
 "MCPR" = "MCPR",
 "OFPR" = "OFPR",
 "FDLR" = "FDLR")

protected_status_names <- c(
  "0" = "Unprotected",
  "1" = "Protected",
  "all" = "All"
  )

#simpson diversity by protected status by strata in FKNMS
fk_strata_ab_simpson <- ggplot(data = strata_fk_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = simpson, color = "simpson"),color = "blue") +
  geom_line(aes(y = simpson, color = "simpson"), color = "blue") +
  facet_grid(STRAT ~ protected_status, 
             labeller = labeller(.rows = strata_names, .cols = protected_status_names)) +
             #labeller(strata_names = label_wrap_gen(3))) + 
  labs(x = "Year", y = "Effective Number of Species") +
  ggtitle("Simpson Diveristy of Reef Fish in the Florida Keys") + 
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12)) +
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 25), breaks = c(5,10,15,20,25))

#shannon diversity by protected status by strata in FKNMS
FK_strata_ab_shannon <- ggplot(data = fk_strata_abun_diversity, aes(x = YEAR)) +
  geom_point(aes(y = shannon, color = "shannon"), color = "magenta") +
  geom_line(aes(y = shannon, color = "shannon"), color = "magenta") +
  facet_grid(STRAT ~ protected_status, 
             labeller = labeller(.rows = strata_names, .cols = protected_status_names)) +
  labs(x = "Year", y = "Effective Number of Species", color = "Level of protection") +
  ggtitle("Shannon Diveristy of Reef Fish in the Florida Keys by Strata") + 
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(1999:2016)) +
  scale_y_continuous(limits= c(0, 50), breaks = c(5,10,15,20,25))

```

### Florida Keys Primary Sampling Unit 

```{r psu abundance}

psu_fk_abun <- read_csv("big_csv/abundance_psu/psu_fk_abun.csv",
col_types = cols(
  YEAR = col_integer(),
  REGION = col_character(),
  STRAT = col_character(),
  PROT = col_integer(),
  PRIMARY_SAMPLE_UNIT = col_character(),
  SPECIES_CD = col_character(),
  m = col_integer(),
  var = col_double(),
  abundance = col_double(),
  protected_status = col_character()))

psu_fk_abun_no_spp = psu_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species 
```

```{r psu biomass}
psu_fk_bio <- read_csv('big_csv/biomass_psu/psu_fk_biomass.csv', 
col_types = cols(
  protected_status = col_character()
))

psu_fk_bio_no_spp = psu_fk_bio %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species
```

```{r psu biodiversity, eval = F}

psu_fk_abun <- read_csv("big_csv/abundance_psu/psu_fk_abun.csv",
col_types = cols(
  YEAR = col_integer(),
  REGION = col_character(),
  STRAT = col_character(),
  PROT = col_integer(),
  PRIMARY_SAMPLE_UNIT = col_character(),
  SPECIES_CD = col_character(),
  m = col_integer(),
  var = col_double(),
  abundance = col_double(),
  protected_status = col_character()))

psu_fk_abun_no_spp = psu_fk_abun %>%
  filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species 

fk_psu_abun_diversity_no_spp = psu_fk_abun_no_spp %>% #3,107,408 x 10
  group_by(YEAR, PRIMARY_SAMPLE_UNIT, STRAT, protected_status) %>% #9,6878 groups 
  nest(-YEAR) %>%
  mutate(
    data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill =0)), 
    richness = map( #species richness
      data_wide, 
      function(x) specnumber(x %>% select(-REGION, - PROT, -m, -var))),
    simpson = map( #simpson diversity as effective number of species 
      data_wide,
      function(x) 1/(1 - diversity(x %>% select(-REGION, - PROT, -m, -var), index = 'simpson'))),
    shannon = map( #shannon diversity as effective number of species 
      data_wide,
      function(x) exp(diversity(x %>% select(-REGION, - PROT, -m, -var), index = 'shannon')))) %>% 
  unnest(richness,simpson, shannon)

write_csv(fk_psu_abun_diversity_no_spp, "big_csv/diversity_without_spe/psu_fk_abun_diversity_no_spp.csv") 

```

```{r psu biodiversity plots and maps}

psu_fk_abun_diversity = read_csv("big_csv/diversity_without_spe/psu_fk_abun_diversity_no_spp.csv",
          col_types = cols(
            protected_status = col_character())) 

```

## Normality

Checking data for normality and homoscedasticity

COLUMN is the column name of the variable of interest that you want to test (e.g., abundance) and DATAFRMAE is the dataframe from which you are working. COLUMN2 is the variable by which you are testing (e.g., year).

```{r normality}

library(MASS)
library(moments)
library(DescTools)
library(rpsychi)

simpson_stat_dom_merged = fk_domain_abun_diversity_no_spp %>%
  select(YEAR, simpson, protected_status) %>%
  filter(protected_status != "0") %>%
  filter(protected_status != "1") 
#ggplot() + geom_density(aes(x=COLUMN), data=DATAFRAME) #first just plot the data to see what your distribution looks like
shapiro.test(simpson_stat_dom_merged$simpson) # test for normality
#data:  simpson_stat_dom_merged$simpson
#W = 0.8214, p-value = 0.007013

#to look at residuals....
r1=lm(simpson_stat_dom_merged$simpson~simpson_stat_dom_merged$YEAR) 
plot(simpson_stat_dom_merged$YEAR, rstandard(r1))
hist(r1$resid)
qqnorm(r1$resid)
qqline(r1$resid)
skewness(r1$resid) #greater than zero is evidence for having skew
#-1.931305
```

## Boxcox method 

Transforming your data using the boxcox method

XXX is where you need to copy in the lamda with the highest likelihood (you can round this number to something more reasonable)

```{r transforming data}

#need non-zero data for boxcox method to add a small amount to every value
simpson_stat_dom_merged$simpson_t <- simpson_stat_dom_merged$simpson+0.01 
b=boxcox(simpson_stat_dom_merged$simpson_t~simpson_stat_dom_merged$YEAR)
b #x is lambda and y is values of likelihood
lamda=b$x #lambda values 
lik=b$y #y is values of liklihood 
bc=cbind(lamda,lik)
bc[order(-lik),] #use the lamda with the highest likelihood to transform your data
#lambda of 2 is best transformation 
simpson_stat_dom_merged$simpson_bc <- simpson_stat_dom_merged$simpson^2 
simpson_stat_dom_merged$simpson_bc <- simpson_stat_dom_merged$simpson_bc-1 
simpson_stat_dom_merged$simpson_bc <- simpson_stat_dom_merged$simpson_bc/2

#simpson_stat_dom_merged$simpson_bc is now transformed using the boxcox method 
```

## Normality 

Checking data for normality after transformation

```{r check for normality}

ggplot() + geom_density(aes(x=simpson_stat_dom_merged$simpson_bc))
shapiro.test(simpson_stat_dom_merged$simpson_bc)
#data:  simpson_stat_dom_merged$simpson_bc
#W = 0.92366, p-value = 0.219
r1=lm(simpson_stat_dom_merged$simpson_bc~simpson_stat_dom_merged$YEAR)
plot(simpson_stat_dom_merged$YEAR, rstandard(r1))
hist(r1$resid)
qqnorm(r1$resid)
qqline(r1$resid)
skewness(r1$resid) #greater than zero is evidence for having skew
#-1.204832

# if data look normal after transformation, proceed with parametric test
# if data do not look normal and homoscedastic after transformation, proceed with non-parametric test

```

## ANOVA

ANOVA: parametric test for differences among groups
Use COLUMN_bc if using transformed data, otherwise use COLUMN

```{r ANOVA}

ANOVA <- aov(simpson_stat_dom_merged$simpson_bc ~ simpson_stat_dom_merged$YEAR) 
plot(ANOVA)
summary(ANOVA) # display Type I ANOVA table
                              #Df Sum Sq Mean Sq F value Pr(>F)
#simpson_stat_dom_merged$YEAR  1    460   459.6   0.277  0.607
#Residuals                    13  21554  1658.0 

#p value 0.6 so not significant 
drop1(ANOVA,~.,test="F") # type III SS and F Tests 
 #                            Df Sum of Sq   RSS    AIC F value Pr(>F)
#<none>                                    21554 113.05               
#simpson_stat_dom_merged$YEAR  1    459.62 22014 111.37  0.2772 0.6074

# for post-hoc pairwise tests
#TukeyHSD(ANOVA) # where fit comes from aov()

```

## Kruskal-Wallis
non-parametric test for differences among groups

```{r Kruskal-Wallis}

kruskal.test(COLUMN~COLUMN2, data=DATAFRAME)

# for post-hoc pairwise tests....
DunnTest(x=DATAFRAME$COLUMN, g=DATAFRAME$COLUMN2, method="fdr") # note that these p-values are one-sided values; for two-sided p-values they need to be doubled

```

```{r shapefiles}
library(foreign)
dbf = read.dbf('FL_files/FLGrid_100m_work8.dbf')
dbf2 = read.dbf('FL_files/Unified_Florida_Hab.dbf')
library(maptools)
shape = readShapePoly("FL_files/Unified_Florida_Hab.shp")
library(shapefiles)
shx = read.shx("FL_files/Unified_Florida_Hab.shx")
library(spatstat)
library(lattice)
library(maps)
library(rgeos)
library(rgdal)
library(ggmap)
library(tmap)
library(marmap) #marine maps 
```

## Evironmental Variables from SERC 

```{r SERC environmental variables}

serc_data = read_csv("big_csv/SERC_Water_Quality_FKNMS.csv")

```