---
title: "Florida Keys Reef Fish Biodiversity"
author: "Megan Hepner"
date: "April 28, 2017"
output: html_document
---

### Inputs 
1. species-trait matrix
1. psu abundance dataframe
1. psu biomass dataframe

### Outputs 
1.  Functional dendogram (Fig: The final ultrametric functional dendrogram for 348 reef fish in the Florida Keys obtained using hierarchial agglometric clustering.)
+ Trait-matrix visualization 

1.  By *strata* for each sampling year (using PSU data)  
+ Species abundance and plot
+ Species biomass and plot
+ species richness and plot
+ Simpson diversity as effective number of species and plot
+ Shannon diversity as effective number of species and plot
+ Functional diversity as effective number of species and plot

1. By *strata and subregion* for each sampling year (using PSU data) 
+ Species abundance and plot
+ Species biomass and plot
+ species richness and plot
+ Simpson diversity as effective number of species and plot
+ Shannon diversity as effective number of species and plot
+ Functional diversity as effective number of species and plot

1. Figure of the FKNMS coral reef tract with strata, subregions, and NTMR 

1. Statistics 
- Interpolate diversity for each year and strata using kriging 
+ Figure. Spatial distribution of each matrices over time. 
- Extract and interpolate the residuals of each index fitted against each other using GAM
+ temporal variable: YEAR
+ spatial variable: Latitude, longitude
+ environmental variable: STRAT, habitat_cd, depth, bottom temperature 
+ management variable: NTMR 
- Calculate partial deviances 

### Reef Visual Census Background 

Reef fish community data used in this study were collected as part of the multi-agency Reef Visual Census (RVC) [Brandt et al., 2009](https://www.coris.noaa.gov/activities/fish_monitoring_protocol/). Fish communities were visually surveyed underwater annually in the mainland FKNMS (Lower Keys, Middle Keys and Upper Keys) from 1999 to 2012, and biennially from 2012 to 2016. The Dry Tortugas region was surveyed from 1999 to 2000 and biennially from 2004 to 2016.     

The RVC uses a stationary point count method within a randomly selected 7.5 m radius circular plot with a 200m map grid from 1999 to 2012 and 100m map grid from 2014 to 2016 to optimize the observation of conspicuous and diurnally active reef fish, specifically economically and ecologically important reef fish species [Bohnsack and Bannerot, 1986](https://docs.lib.noaa.gov/noaa_documents/NMFS/TR_NMFS/TR_NMFS_41.pdf). 

### Objective 

To analyze and compare changes in species richness, Simpson diversity, Shannon diversity and functional diversity in the Florida Keys National Marine Sanctuary (FKNMS) to reveal changes in temporal and spatial variability from 1999 – 2016. The indices were assessed throughout the Florida Keys, between the upper and lower keys, between habitat types, and for different levels of management (Ecological Reserves (ER), Sanctuary Preservations Areas (SPA), and Special Use/Research Only Areas (SU/RO)). 

## Diversity indicse as effective number of species 

All indices were computed as effective number of species. Effective number of species is the number of equally abundance species needed to produce the observed value of a diversity index (Jost, 2006; Jost et al., 2010; MacArthur, 1965). Jost, 2006 refers to these as true measures of diversity, where prior to the conversion the diversity indices are simply measures of entropy. 

- Species Richness is the number of species in a habitat sampled  

$Richness = \sum_{i=1}^S p_i^0$  

- Simpson diversity measures the probability that two individuals randomly selected from a sample will belong to different species.   

$Simpson = 1 - \sum_{i=1}^S p_i^2$  

- Shannon diveristy is a measure of entropy, it is the uncertainty in the species identity of a sample  

$Shannon = - \sum_{i=1}^S p_i \log_b p_i$  

- Functional diversity (Rao's Q) is a measure of pairwise functional differences between species weighted by their relative abundances   

$Rao Q = \sum_{i=1}^S-1 \sum_{j=i+1}^S d_i p_i p_j$

```{r setup, include= F}
knitr::opts_chunk$set(echo = T)
#ln -s ~/Dropbox/big_csv big_csv - write in terminal window to link to 'big_csv' folder
#rm(list = ls())

#install.packages('devtools')
#devtools::install_github('jeremiaheb/rvc')
library(rvc) #calls RVC data from SEFSC server 
#install.packages("tidyverse")
library(tidyverse)
#install.packages("vegan")
library(vegan) # biodiveristy functions for richness, simpson, and shannon 
#install.packages("dygraphs")
library(dygraphs)
#install.packages("plotly")
#library(plotly)
#install.packages("webshot")
library(webshot)
#install.packages("htmlwidgets")
library(htmlwidgets)
#install.packages("FD")
library(FD)
#install.packages("iNEXT")
library(iNEXT) 
RVCdata_FK = read_rds('big_csv/RVCdata_FK.rds')

map = purrr::map # override maps::map
select = dplyr::select #override MASS::select 
group_by =  dplyr::group_by #override plotly::group_by
summarise = dplyr::summarise #override plotly::summarise
```

## Functional distance trait based matrix (from Lefcheck et al., 2014 ChessMap.Rmd)

Gowdis computes the Gower (1971) similarity coefficient exactly as described by Podani (1999), then converts it to a dissimilarity coefficient by using D = 1 - S

```{r Functional distance matrix, eval = F}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("clue")
library(clue) #calls cl_ultrametric 
#install.packages("fpc")
library(fpc) #calls pamk
#install.packages("ape")
library(ape)
#install.packages("phytools")
library(phytools)
#install.packages("FD")
library(FD)
#biocLite("BiocUpgrade") ## you may need this
#biocLite("ggtree", type = "source")
library(tidyverse)

func_dist_rds <- "functional_diversity/func.dist.rds"

if (!file.exists(func_dist_rds)){
  
  trait_matrix = read_csv('functional_diversity/species_trait_matrix_348_spp.csv')
  
  trait_matrix = trait_matrix %>%
    as.tibble() %>%
    mutate(
      SPECIES_CD = toupper(as.character(SPECIES_CD))) %>% 
    arrange(SPECIES_CD)%>% 
    select(SPECIES_CD, Maxlength, Trophic_level, Trophic_group, Water_column, Diel_activity, Substrate_type, Complexity, Gregariousness) %>% #348*9
    group_by(
      SPECIES_CD, Maxlength, Trophic_level, Trophic_group, Water_column, Diel_activity, Substrate_type, Complexity, Gregariousness) %>%
    #summarize(n = n()) %>%
    #select(-n) %>%
    ungroup() %>%
    mutate(
      # ordinal traits
      Complexity = ordered(Complexity, levels=c("Low","Medium","High")),
      Gregariousness = ordered(Gregariousness, levels=c("1","2","3"))) %>%
    as.data.frame()
  
  #Set species names as row.names and remove extra columns
  rownames(trait_matrix)=trait_matrix$SPECIES_CD 
  
  #Calculate Gower distances. Variables have equal weight. 
  traits.dist = gowdis(trait_matrix, ord="podani")
  
  #Use clustering method to produce ultrametric dendrogram because values of Rao's Q can be maximized when fewer than the max number of functional types are present unless distances are ultramtetric 
  
  #to account for sensitivity in clustering use multiple algorithms  (Mouchet et al., 2008) 
  tree_methods = c("single","complete","average","mcquitty","ward.D") #average is best clustering method 
  trees=lapply(tree_methods,  function(i) hclust(traits.dist, method=i))
  par(mfrow=c(3,2))
  for(i in 1:length(trees)) {plot(trees[[i]])} #plots the 5 different kinds of trees 
  
  #convert trees to ultrametric
  trees.ultra= lapply(trees,function(i) cl_ultrametric(as.hclust(i)))
  
  #Plot each tree
  par(mfrow=c(3,2))
  for (i in 1:length(trees.ultra)) {plot(trees.ultra[[i]])} #plots the 5 different kinds of trees 
  
  #Build the consensus tree (Mouchet et al 2008 Oikos)  
  ensemble.trees=cl_ensemble(list=trees) #list of clusterings 
  class(ensemble.trees)
  consensus.tree=cl_consensus(ensemble.trees) #synthesizes the information in the elements of a cluster ensemble into a single clustering 
  par(mar=c(1,1,1,1))
  plot(consensus.tree, horiz=T)
  
  #Calculate dissimilarity values for each tree using 2-norm (Merigot et al 2010 Ecology) to determine which tree best preserves orignial distances 
  #spectral norm (2-norm) of the differences of the ultrametrics
  all.trees=c(trees.ultra,consensus.tree[1])
  names(all.trees)=c(tree_methods,"consensus")
  (trees.dissim=lapply(all.trees,function(i) cl_dissimilarity(i,traits.dist,method="spectral"))) 
  #Cross-dissimilarities using spectral ultrametric distance:
  #Single = 63.17399
  #complete = 116.8516
  #average = 14.99288
  #mcquitty = 36.06599
  #ward.D = 3161.581
  #consensus =  1065.058 (1067.782) (1091.628)
  
  #Identify best tree and isolate
  trees.dissim2=do.call(rbind,trees.dissim)
  min.tree=which.min(trees.dissim2) #average 
  names(all.trees)[min.tree]
  func.dist=all.trees[names(all.trees)==names(all.trees)[min.tree]][[1]]
  
  #Confirm lowest 2-norm value,  spectral norm (2-norm) of the differences if the ultrametrics 
  cl_dissimilarity(func.dist,traits.dist,method="spectral") #Cross-dissimilarities using spectral ultrametric distance:  14.99288
  
  #Scale by the max value so that all values are between 0-1
  func.dist=func.dist/max(func.dist)
  
  saveRDS(func.dist, "functional_diversity/func_dist_rds")
} 
else { #read data 
  func.dist = read_rds("functional_diversity/func_dist_rds")
} 

#Plot the best tree - average 
pdf("functional_dendrogram.pdf", width = 30, height = 70)
par(mfrow=c(1,1))         #nc-by-nr array 
par(mar=c(2,1,0,2))       #number of lines of margins for bottom, left, top, right 
par(pin=c(25,58))         #plot dimensions 
plot(func.dist, horiz=TRUE, main = "Functional Dendrogram", cex = 1, cex.main = 2, cex.lab = 0.5)
dev.off()

#Phylogenetic tree with 348 tips and 347 internal nodes. 
func.dist.phy = as.phylo(as.hclust(func.dist))

#Write newick tree
write.tree(func.dist.phy, "functional_diversity/functional_dendrogram.nwk")
#write.nexus(func.dist.phy, "functional_diversity/functional_dendrogram.nex") 

tree <- read.tree("functional_diversity/functional_dendrogram.nwk")
# List of 4
#  $ edge       : int [1:694, 1:2] 
#  $ Nnode      : int 347
#  $ tip.label  : chr [1:348] 
#  $ edge.length: num
# - attr(*, "class")= chr "phylo"
# - attr(*, "order")= chr "cladewise"

# Phylogenetic tree with 348 tips and 347 internal nodes.
# Tip labels:
# 	KYP_SECT, ALE_CILI, ELO_SAUR, HAR_JAGU, CEN_UNDE, TYL_CROC, ...
# Rooted; includes branch lengths.

plot(tree, label.offset=0.2, edge.width = 2, type = "cladogram")  
nodelabels() #add node numbers
tiplabels() #add tip numbers

ggtree(tree, branch.length = T) +
  ggtitle("Functional Dendogram") +
  geom_tiplab()+
  geom_text(aes(x = branch, label=branch.length), vjust = -.5) +
  theme_tree2()

ggtree(tree, layout= "circular") +
  ggtitle("Functional Dendogram") +
  geom_tiplab(size = 1, aes(angle=angle))+
  theme_tree2()

# transformed to a tidy data.frame by fortify method
tree_data <- fortify(tree) #node; parent; branch.length; x; y; label; isTip; branch; angle 
head(tree_data)

#Visualize species' differences in multivariate trait space
#Perform k-means clustering with no a priori specification for k
traits.kclus= pamk(traits.dist, krange=2:10) 

#krange: integer vector. Numbers of clusters which are to be compared by the average silhouette width criterion.
#traits.kclus$nc = 10 
#traits.kclus$crit 0.0000000 0.2374706 0.2362513 0.2355555 0.2681756 0.2581352 0.2759547 0.2852216 0.3006077 0.3062123

# #Perform multidimensional scaling on functional dendrogram
traits_nmds = metaMDS(traits.dist,k=traits.kclus$nc,trymax=20)
#k = Number of dimensions

#*** No convergence -- monoMDS stopping criteria:
#499: no. of iterations >= maxit
#1: stress ratio > sratmax

# Dimensions: 10 
# Stress:     0.03545178 
# Stress type 1, weak ties
# No convergent solutions - best solution after 500 tries
# Scaling: centring, PC rotation 
# Species: scores missing


# #Plot in two dimensions - doesn't work 
par(mar=c(4,4,1,1))
ordiplot(traits_nmds,type="n")

#Assign colors to different groups
groups=levels(factor(traits.kclus$pamobject$clustering))
points.symbols=15:16
points.colors=c("firebrick3","cornflowerblue")
for(i in seq_along(groups)) {
  points(traits_nmds$points[traits.kclus$pamobject$clustering==groups[i],],
         pch=points.symbols[i],col=points.colors[i],cex=1.4) }
ordispider(traits_nmds,factor(traits.kclus$pamobject$clustering),label=F)
ordihull(traits_nmds,factor(traits.kclus$pamobject$clustering),lty="dotted")
orditorp(traits_nmds,dis="sites",pcex=0,air=0.5,col="grey10",cex=0.8)

```

## Trait matrix visualization 

[http://jkunst.com/r/pokemon-visualize-em-all/]
[https://d3js.org/]
[http://www.buildingwidgets.com/blog/2015/7/22/week-29-d3treer-v2]

```{r trait visualization}

library(tidyverse)
library(treemap)
devtools::install_github("gluc/data.tree")
devtools::install_github("timelyportfolio/d3treeR")
library(d3treeR)
devtools::install_github("GuangchuangYu/ggtree") #requires R>= 3.3.2
library(ggtree)

trait_matrix = read_csv('functional_diversity/species_trait_matrix_348_spp.csv')

trait_matrix_treemap = trait_matrix %>%
  select(Common_name, Maxlength, Trophic_level, Trophic_group, Water_column,Diel_activity,Substrate_type,Complexity, Gregariousness) %>% 
  group_by(Maxlength,
           Trophic_level,
           Trophic_group,
           Water_column,
           Diel_activity,
           Substrate_type,
           Complexity,
           Gregariousness) %>%
  summarise(n = n())%>% 
  ungroup() %>%
  as.data.frame()

treemap2 = 
  d3treeR::d3tree2(
    treemap(
      dtf = trait_matrix_treemap,
      index = c("Trophic_group", "Trophic_level"), #, "Water_column", "Complexity", "Substrate_type", "Diel_activity"),
      vSize = "n", 
      vColor = "Trophic_group",
      type = "index"),
    #title = "Reef Fish Traits",
    #fontsize.title = 14, 
    #algorithm = "squarified"
    rootname = "Species_Traits")

#treemap example 
data(GNI2014)
d3tree2(
  treemap(
    GNI2014
    ,index=c("continent", "iso3")
    ,vSize="population"
    ,vColor="GNI"
    ,type="value"
  )
  , rootname = "World"
)

```

### Fetch the RVC data from 
[NOAA's Southeast Fisheries Science Center server](https://grunt.sefsc.noaa.gov/rvc_analysis20/) using the [RVC package](https://github.com/jeremiaheb/rvc)
- Fetched weighted abundance and biomass data by domain, strata, and primary sampling units 

### Sample Data Variables
- PRIMARY_SAMPLE_UNIT: A code indicating the primary sample unit in which a sample was collected.
- YEAR: A number indicating the calendar year.
- MONTH: A number indicating the month of the year.
- DAY: A number indicating the day of the month (EST).
- STATION_NR: A number indicating the secondary sampling unit within a given primary sample unit.
- LAT_DEGREES: Latitude of secondary sampling unit in decimal degrees).
- LON_DEGREES: Longitude of secondary sampling unit in decimal degrees).
- DEPTH: Average depth, in meters, of secondary sampling unit).
- UNDERWATER_VISIBILITY: visibility, in meters, at secondary sampling unit.
- MAPGRID_NR: A number indicating the primary sample unit.  
- HABITAT_CD: A code indicating the habitat type. 
- ZONE_NR: A code indicating the distance offshore: 
+ 1 - Inshore
+ 2 - Midchannel
+ 3 - Offshore
+ 4 - Fore-reef
- SUBREGION_NR: A number indicating the subregion.
- MPA_NR: A number identifying the marine protected area in which the sample was collected. Zero indicates unprotected status)
- SPECIES_NR: A number indicating the species for a sample 
- SPECIES_CD: A code indicating the species for a sample. Consists of the first three letters of the generic name and the first four of the specific name   
- LEN: The length, in cm, of a sample.      
- NUM: The number of individuals of a given species and length observed in a secondary sampling unit
- TIME_SEEN: A number indicating when, during sampling, an individual was observed. 1: In the first five minutes, 2: From 5-10 minutes, 3: After 10 minutes. 
- PROT: A boolean value indicating whether a sample was in a protected area or not 
+ 1 - protected area
+ 0 - not protected 
- STRAT: A code indicating the stratum in which a sample was taken. Differs by region.   
+ FMLR
+ FSLR
+ HRRF
+ INPR
+ MCPR 
+ OFPR
- REGION: A code indicating the region in which a sample was taken.
+ FLA KEYS (florida keys)
+ DRY TORT (dry tortugas)

### Stratum Data Variables 
- REGION 
- YEAR 
- PROT 
- STRAT 
- NTOT: The number of possible primary sample units for a given year, region, stratum, and protected status 
- GRID_SIZE: The length (in meters) to a side of a primary sample unit for a given year, region, stratum, and protected status.

### Taxonomic Data Variables 
- SPECIES_CD
- FAMILY
- SCINAME 
- COMNAME
- LC: Minimum length at capture, in centimeters
- LM: Median length at maturity, in centimeters.
- WLEN_A: The linear coefficient of the allometric growth equation in grams per millimeter (g/mm)
- WLEN_B: The exponential coefficient of the allometric growth equation

### Benthic Data Variables 
- REGION: A code indicating the region. 
- YEAR: A number indicating the calendar year.
- PRIMARY_SAMPLE_UNIT: A code indicating the primary sample unit in which a sample was collected.
- STATION_NR: A number indicating the secondary sampling unit within a given primary sample unit.
- DEPTH: Average depth, in meters, of secondary sampling unit.
- MAX_HARD_RELIEF: The maximum height, in meters, of hard relief (e.g. hard corals, rock).
- MAX_SOFT_RELIEF: The maximum height, in meters, of soft relief (e.g. soft corals, sponges).
- AVG_HARD_RELIEF: The average height, in meters, of hard relief.
- HARD_REL_PCT_0: Percentage of hard relief less than 0.2 meters.
- HARD_REL_PCT_1: Percentage of hard relief between 0.2 and 0.5 meters.
- HARD_REL_PCT_2: Percentage of hard relief between 0.5 and 1.0 meters.
- HARD_REL_PCT_3: Percentage of hard relief between 1.0 and 1.5 meters.
- HARD_REL_PCT_4: Percentage of hard relief greater than 1.5 meters.
- PCT_SAND: Percentage of abiotic cover that is sand.
- PCT_HARD_BOTTOM: Percentage of abiotic cover that is hard bottom.
- PCT_RUBBLE: Percentage of abiotic cover that is rubble.
- PCT_CORAL: Percentage of biotic hardbottom that is coral.
- PCT_OCTO: Percentage of biotic hardbottom that is octocoral.
- PCT_SPONGE: Percentage of biotic hardbottom that is sponge.

### Function: getDomainDensity 
- YEAR
- REGION
- SPECIES_CD
- density: Domain-wide mean density for a stratified random survey  
- var: Variance in average density per secondary sampling unit
- n: Number of primary sampling units sampled
- nm: Number of secondary sampling units sampled
- N: Number of total possible primary sample units
- NM: Number of possible secondary sampling units
- length_class: The length class or bin. Only present if length_bins is not NULL. The notation, [lower, upper), is inclusive of the lower bound, but exclusive of the upper bound
- protected_status: The protected status. Only present if merge_protected is FALSE

### Florida Keys Strata
1. FSLR - Forereef Shallow Linear Reef (<6m)
2. FMLR - Forereef Midchannel Linear Reef (6-18m)
3. FDLR - Forereef Deep Linear Reef (18-33m)
4. INPR - Inshore patch reef
5. MCPR - Midchannel Patch Reef 
6. OFPR - Offshore Patch Reef
7. HRRF - High Relief Reef (Spur and Groove)

### Florida Keys Primary Sampling Unit 

```{r psu abundance}

if(!file.exists("psu_abun_sum_rds")){
  psu_fk_abun <- read_csv("big_csv/abundance_psu/psu_fk_abun.csv",
                          col_types = cols(protected_status = col_character()))
  #dim(psu_fk_abun) #3,643,306 x 10

  psu_fk_abun_no_spp = psu_fk_abun %>%
    filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) 
  #dim(psu_fk_abun_no_spp) #3,379,964 x 10
  
  psu_abun_sum <- psu_fk_abun_no_spp %>%
    group_by(YEAR, PRIMARY_SAMPLE_UNIT, STRAT) %>%
    summarize(abundance_sum = sum(abundance)) 
  #dim(psu_abun_sum) # 5241 x 4 
  
  saveRDS(psu_abun_sum, "psu_abun_sum_rds")
  
  psu_fk_abun_no_spp_merged = psu_fk_abun_no_spp %>%
  filter(stringr::str_detect(protected_status, 'all')) # 1,689,982 x 10 (this is half the number of rows as the original as it should be)

  write_csv(psu_fk_abun_no_spp_merged, "big_csv/abundance_psu/psu_fk_abundance_no_spp_merged.csv")
}

else{
  psu_abun = read_rds("psu_abun_sum_rds")
}

```

```{r psu biomass}

if(!file.exists("psu_bio_sum_rds")){
  psu_fk_bio <- read_csv('big_csv/biomass_psu/psu_fk_biomass.csv', 
                         col_types = cols( protected_status = col_character())) 
  #dim(psu_fk_bio) 3,429,476*10

  #remove spe.  
  psu_fk_bio_no_spp = psu_fk_bio %>%
    filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #348 species and 
  #dim(psu_fk_bio_no_spp) 3,219,344*10
  
  #summarize 
  psu_bio_sum <- psu_fk_bio_no_spp %>%
    group_by(YEAR, PRIMARY_SAMPLE_UNIT, STRAT, PROT) %>%
    summarize(biomass_sum = sum(biomass)) 
  #dim(psu_bio_sum)  5,241    5
   saveRDS(psu_bio_sum, "psu_bio_sum_rds")

  psu_fk_bio_no_spp_merged = psu_fk_bio_no_spp %>%
    filter(stringr::str_detect(protected_status, 'all')) #1,609,672 x 10 (this is half the number of rows as the original as it should be)
  write_csv(psu_fk_bio_no_spp_merged, "big_csv/biomass_psu/psu_fk_biomass_no_spp_merged.csv")
}

else{ #read data 
  psu_bio = read_rds("psu_bio_sum_rds")
}

```

```{r psu biodiversity}

if(!file.exists("psu_div_rds")){
  
  #read in psu abundance data 
  psu_fk_abun <- read_csv("big_csv/abundance_psu/psu_fk_abun.csv", col_types = cols(protected_status = col_character())) #3,643,306*10
  
  #remove species not identified to species 
  psu_fk_abun_no_spp = psu_fk_abun %>%
    filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$')) #3,379,964*10 
  
  #read in trait matrix 
  trait_matrix = read_csv('functional_diversity/species_trait_matrix_348_spp.csv')
  
  #trait matrix needs to be alphabetically for FD code 
  trait_matrix = trait_matrix %>%
    as.tibble() %>%
    mutate(
      SPECIES_CD = toupper(as.character(SPECIES_CD))) %>% 
    arrange(SPECIES_CD)
  
  #compute diversity indices by PSU  
  psu_diversity_no_spp_merged = psu_fk_abun_no_spp %>% #tibble: 3,379,964 x 10 %>%
    filter(protected_status != "all") %>% #tibble: 1,689,982 x 10
    select(YEAR, PRIMARY_SAMPLE_UNIT,STRAT,PROT,SPECIES_CD,abundance) %>% #tibble: 1,689,982 x 6
    group_by(YEAR, PRIMARY_SAMPLE_UNIT) %>% #4,839 groups
    nest(-YEAR) %>%
    mutate(
      data_wide = map(data, function(x) 
        full_join(x, trait_matrix %>% select(SPECIES_CD), by='SPECIES_CD') %>%
          mutate(abundance = ifelse(is.na(abundance), 0, abundance)) %>%
          spread(SPECIES_CD, abundance, fill =0)),
      data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill =0)), 
      richness = map( #species richness
        data_wide, 
        function(x) specnumber(x %>% select(-STRAT, -PROT))),
      simpson = map( #simpson diversity as effective number of species 
        data_wide,
        function(x) 1/(1 - diversity(x %>% select(-STRAT, -PROT), index = 'simpson'))),
      shannon = map( #shannon diversity as effective number of species 
        data_wide, 
        function(x) exp(diversity(x %>% select(-STRAT, -PROT), index = 'shannon')))) %>% 
    unnest(richness,simpson, shannon)
  
  #get strata and prot information and remove duplicates
  psu_diversity_no_spp_merged_strat = psu_diversity_no_spp_merged %>%
    unnest(data) #tibble: 1,689,982 x 9 
  
  #save dataframe as csv
  write_csv(psu_diversity_no_spp_merged_strat, "big_csv/diversity_without_spe/psu_diversity_no_spp_merged.csv") 
  
  psu_div = psu_diversity_no_spp_merged_strat %>% 
    dplyr::group_by(YEAR, PRIMARY_SAMPLE_UNIT, STRAT, PROT) %>% #5,241*7 #5241 total PSU's over 15 years 
    dplyr::summarize(
      richness = first(richness),
      simpson = first(simpson),
      shannon = first(shannon))

  saveRDS(psu_div, "psu_div_rds")
  
} else{ #read data 
  psu_div = read_rds("psu_div_rds")
}

```

```{r biodiversity 2.0}

if(!file.exists("fd_eveness_rds")){
  
  sample_data = as.tibble(read_csv("big_csv/sample_data/FK/sample_data_all_years.csv"))
  
  psu_fk_abun <- read_csv("big_csv/abundance_psu/psu_fk_abun.csv", col_types = cols(protected_status = col_character())) #3,643,306*10
  
  func.dist = read_rds("functional_diversity/func_dist_rds")
  
  #dim(psu_fk_abun) #3643306      10
  
  psu_fk_abun_no_spp = psu_fk_abun %>%
    filter(!stringr::str_detect(SPECIES_CD, 'SPE\\.$'))  
  dim(psu_fk_abun_no_spp) #3,379,964      10
  
  abund = psu_fk_abun_no_spp %>% 
    filter(protected_status != "all") %>% #1,689,982 x 10
    select(YEAR, PRIMARY_SAMPLE_UNIT,STRAT,PROT, SPECIES_CD, abundance) #1,689,982 * 6
  
  # Cast abundance longways (species as columns)
  abundmat <- spread(abund, SPECIES_CD, abundance, fill = 0)
  #dim(abundmat) #5241  352
  
  # What is up with these species not being in the trait matrix??
  checkthese <- colnames(abundmat)[!colnames(abundmat) %in% rownames(traits)]
  
  # Subset species whose names only appear in traits
  abundmat <- abundmat[colnames(abundmat) %in% rownames(traits)] 
  #dim(abundmat) #5,241    348
  
  # Drop communities with no species (roughly 500 sites)
  abundmat <- abundmat[!rowSums(abundmat) == 0, ] 
  #dim(abundmat) #5,241     348
  
  #Calculate relative values for community matrix
  rel.mat=abundmat/apply(abundmat,1,sum)
  
  #Compute species diversity
  species.dist=matrix(1,ncol(abundmat),ncol(abundmat))-diag(rep(1,ncol(abundmat))) 
  species.div=1/(1-apply(rel.mat,1,function(x) t(x) %*% species.dist %*% x))
  
  #Compute evenness (as in Jost 2010 Diversity)
  evenness=log(species.div)/log(rowSums(abundmat>0))
  
  #Ensure that all evenness calculations are not >1 (bug)
  #func.div=1/(1-apply(rel.mat,1,function(x) t(x) %*% as.matrix(traitsmat) %*% x))
  
  #Ensure that all evenness calculations are not >1 (bug)
  func.div=1/(1-apply(rel.mat,1,function(x) t(x) %*% as.matrix(func.dist) %*% x))
  
  #Bind all to the original dataframe
  list = cbind(sample_data,evenness,species.div,func.div)
  
  fd_evenness = list %>%
    select(PRIMARY_SAMPLE_UNIT, YEAR,STRAT,PROT,species.div,func.div,evenness) %>%
    group_by(YEAR,PRIMARY_SAMPLE_UNIT)
  
  saveRDS(fd_evenness,"fd_eveness_rds")
  
} {else
  fd_evenness = read_rds("fd_eveness_rds")
}

```

```{r bind psu matrices and sample data}

if(!file.exists("psu_matrices_data_rds")){
  psu_abun = read_rds("psu_abun_sum_rds") #5,241*4
  psu_bio = read_rds("psu_bio_sum_rds") #5,241*5
  psu_div = read_rds("psu_div_rds") # #5,241*7
  fd_evenness = read_rds("fd_eveness_rds")
  sample_data = as.tibble(read_csv("big_csv/sample_data/FK/sample_data_all_years.csv")) 
  #PRIMARY_SAMPLE_UNIT, YEAR, MONTH, DAY, STATION_N,R LAT_DEGREES, LON_DEGREES, DEPTH, UNDERWATER_VISIBILITY, MAPGRID_NR, HABITAT_CD, ZONE_NR, SUBREGION_NR, MPA_NR, PROT, STRAT
  
  all_matrices = inner_join(psu_abun, psu_bio) #5,241*6 (+mean abundance and biomass)
  all_matrices = inner_join(all_matrices, psu_div) #5,241*9 (+richness, simpson, shannon)
  all_matrices = inner_join(all_matrices, fd_evenness) #5,241*12(+spec.div, functional.div, evenness)
  all_matrices_data = inner_join(all_matrices, sample_data) #5,241 x 24

  saveRDS(all_matrices_data, "psu_matrices_data_rds")
}
else{
  all_matrices_data = read_rds("psu_matrices_data_rds")
}

```

## Plots

- create multiplot function from (https://stackoverflow.com/questions/24387376/r-weird-error-could-not-find-function-multiplot)

```{r multiplot}

multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
library(tidyverse)
```


```{r Plot by year and strata}

all_matrices_data = read_rds("psu_matrices_data_rds")

if(!file.exists("stats_all_matrices_data_rds")){
  
  stats_all_matrices_data = all_matrices_data %>%
    group_by(YEAR, STRAT) %>% #110 groups 
    summarize(
      abundance_mean = mean(abundance_sum),
      abundance_n = length(abundance_sum),
      abundance_sd = sd(abundance_sum),
      abundance_se = abundance_sd / sqrt(abundance_n),
      abundance_min = min(abundance_sum),
      abundance_max = max(abundance_sum),
      biomass_mean = mean(biomass_sum),
      biomass_n = length(biomass_sum),
      biomass_sd = sd(biomass_sum),
      biomass_se = biomass_sd / sqrt(biomass_n),
      biomass_min = min(biomass_sum),
      biomass_max = max(biomass_sum),
      richness_mean = mean(richness),
      richness_n = length(richness),
      richness_sd = sd(richness),
      richness_se = richness_sd / sqrt(richness_n),
      richness_min = min(richness),
      richness_max = max(richness),
      simpson_mean = mean(simpson),
      simpson_n = length(simpson),
      simpson_sd = sd(simpson),
      simpson_se = simpson_sd / sqrt(simpson_n),
      simpson_min = min(simpson),
      simpson_max = max(simpson),
      shannon_mean = mean(shannon),
      shannon_n = length(shannon),
      shannon_sd = sd(shannon),
      shannon_se = shannon_sd / sqrt(shannon_n),
      shannon_min = min(shannon),
      shannon_max = max(shannon),
      func.div_mean = mean(func.div),
      func.div_n = length(func.div),
      func.div_sd = sd(func.div),
      func.div_se = func.div_sd / sqrt(func.div_n),
      func.div_min = min(func.div),
      func.div_max = max(func.div),
      evenness_mean = mean(evenness),
      evenness_n = length(evenness),
      evenness_sd = sd(evenness),
      evenness_se = evenness_sd / sqrt(evenness_n),
      evenness_min = min(evenness),
      evenness_max = max(evenness))
  
  saveRDS(stats_all_matrices_data, "stats_all_matrices_data_rds")
  
  } else{ 
  stats_all_matrices_data = read_rds("stats_all_matrices_data_rds")
  }

# plot mean and se across strat and year using bar graphs

# BY STRAT and YEAR
#abundance 
ab_ys = ggplot(stats_all_matrices_data,aes(x=YEAR,y=abundance_mean, colour=STRAT, group=STRAT))+ #shape=STRAT
  geom_line(aes(group=STRAT),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymax=abundance_mean+abundance_se,ymin=abundance_mean-abundance_se),width=0.1)+
  labs(title= "A) Mean Abundance", x="Year", y="Abundance of species", colour = "Strata")+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="right", #"none"
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10))
#ggsave(file="abundance_by_year_and_strata.pdf", path="big_csv/plots")

#biomass
bio_ys = ggplot(stats_all_matrices_data,aes(x=YEAR,y=biomass_mean, colour=STRAT, group=STRAT))+ #shape=STRAT
  geom_line(aes(group=STRAT),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin=biomass_mean-biomass_se, ymax=biomass_mean+biomass_se), width=.2) +
  labs(title= "B) Mean Biomass", x="Year", y="Biomass of species", colour = "Strata")+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="right",
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10))
#ggsave(file="biomass_by_year_and_strata.pdf", path="big_csv/plots")

#richness
rich_ys = ggplot(stats_all_matrices_data,aes(x=YEAR,y=richness_mean, colour=STRAT, group=STRAT))+ #shape=STRAT
  geom_line(aes(group=STRAT),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin=richness_mean-richness_se, ymax=richness_mean+richness_se), width=.2)+
  labs(title= "C) Richness", x="Year", y="Effective number of species", colour = "Strata")+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="none",
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10))
#ggsave(file="richness_by_year_and_strata.pdf", path="big_csv/plots")

# evenness
eve_ys = ggplot(stats_all_matrices_data,aes(x=YEAR,y=evenness_mean, colour=STRAT, group=STRAT))+ #shape=STRAT
  geom_line(aes(group=STRAT),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymax=evenness_mean+evenness_se,ymin=evenness_mean-evenness_se),width=0.1)+
  labs(title= "D) Evenness", x="Year", y="Relative evenness",colour = "Strata")+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="none", #right
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10))
#ggsave(file="functional_diversity_by_year_and_strata.pdf", path="big_csv/plots")

#simpson
sim_ys = ggplot(stats_all_matrices_data,aes(x=YEAR,y=simpson_mean, colour=STRAT, group=STRAT))+ #shape=STRAT
  geom_line(aes(group=STRAT),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin=simpson_mean-simpson_se, ymax=simpson_mean+simpson_se), width=.2)+
  labs(title= "E) Simpson", x="Year", y="Effective number of species",colour = "Strata")+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="right",
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10))
#ggsave(file="simpson_diversity_by_year_and_strata.pdf", path="big_csv/plots")

# shannon
shan_ys = ggplot(stats_all_matrices_data,aes(x=YEAR,y=shannon_mean, colour=STRAT, group=STRAT))+ #shape=STRAT
  geom_line(aes(group=STRAT),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymax=shannon_mean+shannon_se,ymin=shannon_mean-shannon_se),width=0.1)+
  labs(title= "F) Shannon", x="Year", y="Effective number of species",colour = "Strata")+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="right",
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10))
ggsave(file="shannon_diversity_by_year_and_strata.pdf", path="big_csv/plots")

# functional
func_ys= ggplot(stats_all_matrices_data,aes(x=YEAR,y=func.div_mean, colour=STRAT, group=STRAT))+ #shape=STRAT
  geom_line(aes(group=STRAT),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymax=func.div_mean+func.div_se,ymin=func.div_mean-func.div_se),width=0.1)+
  labs(title= "G) Functional", x="Year", y="Effective number of species",colour = "Strata")+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="right",
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10))
#ggsave(file="functional_diversity_by_year_and_strata.pdf", path="big_csv/plots")



#########KATIE: I tried using multiplot and grid.arrange but I couldn't get the height and width functions to work correctly. I am thinking we only need one legend of Strat. Preferable enlarged on the right with all 7 plots, but that might be something that has to be done in powerpoint or paint. Maybe also have the bottom read "Year" so that its not below each chart. I am thinking the order should be: 
#           a)Mean abundance, b)Mean Biomass, 
#           c)Richness,        d)Evennnesss, 
#           e)Simpson,         f) Shannon, 
#           g) Functional 


#combine multiple plots
checkit = multiplot(ab_ys, rich_ys, sim_ys, func_ys, bio_ys, eve_ys, shan_ys, cols = 2)
#want to remove STRAT. Just wamt one legend ideally. 

grid.arrange(
  arrangeGrob(ab_ys,bio_ys, ncol=2, nrow=1), 
  arrangeGrob(rich_ys,eve_ys, ncol=2, nrow=1),
  arrangeGrob(sim_ys,shan_ys, ncol=2, nrow=1),
  func_ys,
  #arrangeGrob(func_ys, ncol=2, nrow=1), 
  bottom = "Year") #heights=c(4,4,4), widths=c(16,16,16))

grid.arrange(
  arrangeGrob(ab_ys,bio_ys, eve_ys, ncol=3, nrow=1), 
  arrangeGrob(rich_ys, sim_ys,shan_ys, func_ys, ncol=4, nrow=1), bottom = "Year", heights=c(8,8))

grid.arrange(
  arrangeGrob(rich_ys,sim_ys, ncol=2, nrow=1),
  arrangeGrob(shan_ys,func_ys, ncol=2, nrow=1),bottom = "Year", left = "Effective number of species")
#want to remove STRAT. Just wamt one legend ideally. 


grid.arrange(
  arrangeGrob(ab_ys,bio_ys, ncol=2, nrow=1), 
  arrangeGrob(rich_ys,sim_ys, ncol=2, nrow=1),
  arrangeGrob(shan_ys,func_ys, ncol=2, nrow=1),bottom = "Year") #heights=c(4,4,4), widths=c(16,16,16))
#want to remove STRAT. Just wamt one legend ideally. 

### ***Don't need to combine plots by STRAT and by YEAR just YEAR & STRAT for now***

# by STRAT
#abundance 
ggplot(stats_all_matrices_data, aes(x=STRAT, y=abundance_mean)) +
  geom_bar(position=position_dodge(), stat="identity") +
  geom_errorbar(aes(ymin=abundance_mean-abundance_se, ymax=abundance_mean+abundance_se), width=.2, position=position_dodge(.9)) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~YEAR, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

#biomass
ggplot(stats_all_matrices_data, aes(x=STRAT, y=biomass_mean)) + 
  geom_bar(position=position_dodge(), stat="identity") + 
  geom_errorbar(aes(ymin=biomass_mean-biomass_se, ymax=biomass_mean+biomass_se), width=.2, position=position_dodge(.9)) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~YEAR, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

#richness 
ggplot(stats_all_matrices_data, aes(x=STRAT, y=richness_mean)) +
  geom_bar(position=position_dodge(), stat="identity") +
  geom_errorbar(aes(ymin=richness_mean-richness_se, ymax=richness_mean+richness_se), width=.2, position=position_dodge(.9)) + 
  theme_classic() +
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~YEAR, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

#simpson 
ggplot(stats_all_matrices_data, aes(x=STRAT, y=simpson_mean)) +
  geom_bar(position=position_dodge(), stat="identity") +
  geom_errorbar(aes(ymin=simpson_mean-simpson_se, ymax=simpson_mean+simpson_se), width=.2, position=position_dodge(.9)) + 
  theme_classic() +
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~YEAR, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

#shannon 
ggplot(stats_all_matrices_data, aes(x=STRAT, y=shannon_mean)) +
  geom_bar(position=position_dodge(), stat="identity") +
  geom_errorbar(aes(ymin=shannon_mean-shannon_se, ymax=shannon_mean+shannon_se), width=.2, position=position_dodge(.9)) + 
  theme_classic() +
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~YEAR, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

# functional 
ggplot(stats_all_matrices_data,aes(x=STRAT,y=func.div_mean))+ 
  geom_bar(position=position_dodge(), stat="identity") +
  geom_errorbar(aes(ymin=func.div_mean-func.div_se, ymax=func.div_mean+func.div_se), width=.2,  position=position_dodge(.9))+ 
  theme_classic() +
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~YEAR, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

# evenness
ggplot(stats_all_matrices_data,aes(x=STRAT,y=evenness_mean))+ 
  geom_bar(position=position_dodge(), stat="identity") +
  geom_errorbar(aes(ymin=evenness_mean-evenness_se, ymax=evenness_mean+evenness_se), width=.2,  position=position_dodge(.9))+ 
  theme_classic() +
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~YEAR, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

# by YEAR
#abundance 
ggplot(stats_all_matrices_data, aes(x=YEAR, y=abundance_mean, group=YEAR)) + 
  geom_bar(position=position_dodge(), stat="identity") + 
  geom_errorbar(aes(ymin=abundance_mean-abundance_se, ymax=abundance_mean+abundance_se), width=.2, position=position_dodge(.9)) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~STRAT, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

#biomass 
ggplot(stats_all_matrices_data, aes(x=YEAR, y=biomass_mean, group=YEAR)) + 
  geom_bar(position=position_dodge(), stat="identity") + 
  geom_errorbar(aes(ymin=biomass_mean-biomass_se, ymax=biomass_mean+biomass_se), width=.2, position=position_dodge(.9)) +
  theme_classic() + 
  theme(axis.text.x = element_text(angle=90)) +
  facet_wrap(~STRAT, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

#richness 
ggplot(stats_all_matrices_data, aes(x=YEAR, y=richness_mean, group=YEAR)) +  
  geom_bar(position=position_dodge(), stat="identity") + 
  geom_errorbar(aes(ymin=richness_mean-richness_se, ymax=richness_mean+richness_se), width=.2, position=position_dodge(.9)) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~STRAT, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

#simpson 
ggplot(stats_all_matrices_data, aes(x=YEAR, y=simpson_mean, group=YEAR)) +  
  geom_bar(position=position_dodge(), stat="identity") + 
  geom_errorbar(aes(ymin=simpson_mean-simpson_se, ymax=simpson_mean+simpson_se), width=.2, position=position_dodge(.9)) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~STRAT, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

#shannon
ggplot(stats_all_matrices_data, aes(x=YEAR, y=shannon_mean, group=YEAR)) +  
  geom_bar(position=position_dodge(), stat="identity") + 
  geom_errorbar(aes(ymin=shannon_mean-shannon_se, ymax=shannon_mean+shannon_se), width=.2, position=position_dodge(.9)) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~STRAT, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

#functional
ggplot(stats_all_matrices_data, aes(x=YEAR, y=func.div_mean, group=YEAR)) +  
  geom_bar(position=position_dodge(), stat="identity") + 
  geom_errorbar(aes(ymin=func.div_mean-func.div_se, ymax=func.div_mean+func.div_se), width=.2, position=position_dodge(.9)) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~STRAT, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

#evenness
ggplot(stats_all_matrices_data, aes(x=YEAR, y=evenness_mean, group=YEAR)) +  
  geom_bar(position=position_dodge(), stat="identity") + 
  geom_errorbar(aes(ymin=evenness_mean-evenness_se, ymax=evenness_mean+evenness_se), width=.2, position=position_dodge(.9)) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~STRAT, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

```

Plot each metric by grouped strata and subregion through time 

Group strata by: patch reef, forereef, and high relief 

```{r Plot by year grouped strata and subregion} 

if(!file.exists("matrices_strat_grouped_rds"){
  
  all_matrices_data = read_rds("psu_matrices_data_rds")
  
  matrices_strat_grouped = all_matrices_data %>% 
    mutate(
      STRAT_GROUPED = 
      ifelse(grepl("FMLR|FSLR|FDLR", STRAT), 'LINEAR_REEF', 
             ifelse(grepl("OFPR|MCPR|INPR", STRAT), 'PATCH_REEF',
                    ifelse(grepl('HRRF', STRAT), 'HIGH_RELIEF', "IDK")))) %>%
    group_by(YEAR, STRAT_GROUPED) %>% #add subregion later 
    summarize(
      abundance_mean = mean(abundance_sum),
      abundance_n = length(abundance_sum),
      abundance_sd = sd(abundance_sum),
      abundance_se = abundance_sd / sqrt(abundance_n),
      abundance_min = min(abundance_sum),
      abundance_max = max(abundance_sum),
      biomass_mean = mean(biomass_sum),
      biomass_n = length(biomass_sum),
      biomass_sd = sd(biomass_sum),
      biomass_se = biomass_sd / sqrt(biomass_n),
      biomass_min = min(biomass_sum),
      biomass_max = max(biomass_sum),
      richness_mean = mean(richness),
      richness_n = length(richness),
      richness_sd = sd(richness),
      richness_se = richness_sd / sqrt(richness_n),
      richness_min = min(richness),
      richness_max = max(richness),
      simpson_mean = mean(simpson),
      simpson_n = length(simpson),
      simpson_sd = sd(simpson),
      simpson_se = simpson_sd / sqrt(simpson_n),
      simpson_min = min(simpson),
      simpson_max = max(simpson),
      shannon_mean = mean(shannon),
      shannon_n = length(shannon),
      shannon_sd = sd(shannon),
      shannon_se = shannon_sd / sqrt(shannon_n),
      shannon_min = min(shannon),
      shannon_max = max(shannon),
      func.div_mean = mean(func.div),
      func.div_n = length(func.div),
      func.div_sd = sd(func.div),
      func.div_se = func.div_sd / sqrt(func.div_n),
      func.div_min = min(func.div),
      func.div_max = max(func.div),
      evenness_mean = mean(evenness),
      evenness_n = length(evenness),
      evenness_sd = sd(evenness),
      evenness_se = evenness_sd / sqrt(evenness_n),
      evenness_min = min(evenness),
      evenness_max = max(evenness))
  
  saveRDS(matrices_strat_grouped, "matrices_strat_grouped_rds")
}

else{
  matrices_strat_grouped = read_rds("matrices_strat_grouped_rds")
}


# by STRAT_GROUPED and YEAR
#mean abundance
ggplot(matrices_strat_grouped,aes(x=YEAR,y=abundance_mean, colour=STRAT_GROUPED, group=STRAT_GROUPED))+ #shape=STRAT
  geom_line(aes(group=STRAT_GROUPED),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymax=abundance_mean+abundance_se,ymin=abundance_mean-abundance_se),width=0.1)+
  labs(title= "A) Mean Abundance", x="Year", y="Abundance of species", colour = "Strata")+
  scale_colour_manual(labels=c("High reef", "Linear reef", "Patch reef"), values= c("red", "green", "blue"))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="right",
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10)))
ggsave(file=".pdf", path="big_csv/plots")

#mean biomass
ggplot(matrices_strat_grouped,aes(x=YEAR,y=biomass_mean, colour=STRAT_GROUPED, group=STRAT_GROUPED))+ #shape=STRAT
  geom_line(aes(group=STRAT_GROUPED),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin=biomass_mean-biomass_se, ymax=biomass_mean+biomass_se), width=.2)+
  labs(title= "B) Mean Biomass", x="Year", y="Biomass of species", colour = "Strata")+
  scale_colour_manual(labels=c("High reef", "Linear reef", "Patch reef"), values= c("red", "green", "blue"))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="right",
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10))
ggsave(file="biomass_by_year_strata_subregion.pdf", path="big_csv/plots")
#ggsave(file="", path="big_csv/plots")

#mean richness
ggplot(matrices_strat_grouped,aes(x=YEAR,y=richness_mean, colour=STRAT_GROUPED, group=STRAT_GROUPED))+ #shape=STRAT
  geom_line(aes(group=STRAT_GROUPED),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin=richness_mean-richness_se, ymax=richness_mean+richness_se), width=.2)+
  labs(title= "A) Richness", x="Year", y="Effective number of species", colour = "Strata")+
  scale_colour_manual(labels=c("High reef", "Linear reef", "Patch reef"), values= c("red", "green", "blue"))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="right",
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10)),
ggsave(file="richness_by_year_strata_subregion.pdf", path="big_csv/plots")

#mean simpson
ggplot(matrices_strat_grouped,aes(x=YEAR,y=simpson_mean, colour=STRAT_GROUPED, group=STRAT_GROUPED))+ #shape=STRAT
  geom_line(aes(group=STRAT_GROUPED),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin=simpson_mean-simpson_se, ymax=simpson_mean+simpson_se), width=.2)+
  labs(title= "A) Simpson", x="Year", y="Effective number of species", colour = "Strata")+
  scale_colour_manual(labels=c("High reef", "Linear reef", "Patch reef"), values= c("red", "green", "blue"))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="right",
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10))
 
#mean shannon
ggplot(matrices_strat_grouped,aes(x=YEAR,y=shannon_mean, colour=STRAT_GROUPED, group=STRAT_GROUPED))+ #shape=STRAT
  geom_line(aes(group=STRAT_GROUPED),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin=shannon_mean-shannon_se, ymax=shannon_mean+shannon_se), width=.2)+
  labs(title= "A) Simpson", x="Year", y="Effective number of species", colour = "Strata")+
  scale_colour_manual(labels=c("High reef", "Linear reef", "Patch reef"), values= c("red", "green", "blue"))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="right",
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10))

# by STRAT
#abundance 
ggplot(matrices_strat_grouped, aes(x=STRAT_GROUPED, y=abundance_mean)) +
  geom_bar(position=position_dodge(), stat="identity") +
  geom_errorbar(aes(ymin=abundance_mean-abundance_se, ymax=abundance_mean+abundance_se), width=.2, position=position_dodge(.9)) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~YEAR, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

#biomass
ggplot(matrices_strat_grouped, aes(x=STRAT_GROUPED, y=biomass_mean)) + 
  geom_bar(position=position_dodge(), stat="identity") + 
  geom_errorbar(aes(ymin=biomass_mean-biomass_se, ymax=biomass_mean+biomass_se), width=.2, position=position_dodge(.9)) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~YEAR, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

#mean richness 
ggplot(matrices_strat_grouped, aes(x=STRAT_GROUPED, y=richness_mean)) + 
  geom_bar(position=position_dodge(), stat="identity") + 
  geom_errorbar(aes(ymin=richness_mean-richness_se, ymax=richness_mean+richness_se), width=.2, position=position_dodge(.9)) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~YEAR, scales='free_y')
#ggsave(file="fk_MeanRichness_ByStratumForEachYear.pdf", path="big_csv/plots")

#mean simpson 
ggplot(matrices_strat_grouped, aes(x=STRAT_GROUPED, y=simpson_mean)) +
  geom_bar(position=position_dodge(), stat="identity") +
  geom_errorbar(aes(ymin=simpson_mean-simpson_se, ymax=simpson_mean+simpson_se), width=.2, position=position_dodge(.9)) + 
  theme_classic() +
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~YEAR, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

#mean shannon 
ggplot(matrices_strat_grouped, aes(x=STRAT_GROUPED, y=shannon_mean)) +
  geom_bar(position=position_dodge(), stat="identity") +
  geom_errorbar(aes(ymin=shannon_mean-shannon_se, ymax=shannon_mean+shannon_se), width=.2, position=position_dodge(.9)) + 
  theme_classic() +
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~YEAR, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

# by YEAR
#mean abundance 
ggplot(matrices_strat_grouped, aes(x=YEAR, y=abundance_mean, group=YEAR)) + 
  geom_bar(position=position_dodge(), stat="identity") + 
  geom_errorbar(aes(ymin=abundance_mean-abundance_se, ymax=abundance_mean+abundance_se), width=.2, position=position_dodge(.9)) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~STRAT, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

#mean biomass 
ggplot(matrices_strat_grouped, aes(x=YEAR, y=biomass_mean, group=YEAR)) + 
  geom_bar(position=position_dodge(), stat="identity") + 
  geom_errorbar(aes(ymin=biomass_mean-biomass_se, ymax=biomass_mean+biomass_se), width=.2, position=position_dodge(.9)) +
  theme_classic() + 
  theme(axis.text.x = element_text(angle=90)) +
  facet_wrap(~STRAT, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

#mean richness
ggplot(matrices_strat_grouped, aes(x=YEAR, y=richness_mean, group=YEAR)) +
  geom_bar(position=position_dodge(), stat="identity") +
  geom_errorbar(aes(ymin=richness_mean-richness_se, ymax=richness_mean+richness_se), width=.2, position=position_dodge(.9)) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle=90)) + 
  facet_wrap(~STRAT_GROUPED, scales='free_y')
#ggsave(file=".pdf", path="big_csv/plots")

#mean simpson
ggplot(data = matrices_strat_grouped) + 
  geom_boxplot(mapping = aes(x = YEAR, y = simpson_mean, group = YEAR)) +
  facet_wrap(~STRAT_GROUPED, scales="free_y")
#ggsave(file=".pdf", path="big_csv/plots")

#mean shannon 
ggplot(data = matrices_strat_grouped) + 
  geom_boxplot(mapping = aes(x = YEAR, y = shannon_mean, group = YEAR)) +
  facet_wrap(~STRAT_GROUPED, scales="free_y")
#ggsave(file=".pdf", path="big_csv/plots")

```

```{r plots by NTMR}

if(!file.exists("matrices_strat_grouped_rds"){
  
  all_matrices_data = read_rds("psu_matrices_data_rds")
  
  matrices_ntmr = all_matrices_data %>% 
    group_by(YEAR, PROT) %>% #add subregion later 
    summarize(
      abundance_mean = mean(abundance_sum),
      abundance_n = length(abundance_sum),
      abundance_sd = sd(abundance_sum),
      abundance_se = abundance_sd / sqrt(abundance_n),
      abundance_min = min(abundance_sum),
      abundance_max = max(abundance_sum),
      biomass_mean = mean(biomass_sum),
      biomass_n = length(biomass_sum),
      biomass_sd = sd(biomass_sum),
      biomass_se = biomass_sd / sqrt(biomass_n),
      biomass_min = min(biomass_sum),
      biomass_max = max(biomass_sum),
      richness_mean = mean(richness),
      richness_n = length(richness),
      richness_sd = sd(richness),
      richness_se = richness_sd / sqrt(richness_n),
      richness_min = min(richness),
      richness_max = max(richness),
      simpson_mean = mean(simpson),
      simpson_n = length(simpson),
      simpson_sd = sd(simpson),
      simpson_se = simpson_sd / sqrt(simpson_n),
      simpson_min = min(simpson),
      simpson_max = max(simpson),
      shannon_mean = mean(shannon),
      shannon_n = length(shannon),
      shannon_sd = sd(shannon),
      shannon_se = shannon_sd / sqrt(shannon_n),
      shannon_min = min(shannon),
      shannon_max = max(shannon),
      func.div_mean = mean(func.div),
      func.div_n = length(func.div),
      func.div_sd = sd(func.div),
      func.div_se = func.div_sd / sqrt(func.div_n),
      func.div_min = min(func.div),
      func.div_max = max(func.div),
      evenness_mean = mean(evenness),
      evenness_n = length(evenness),
      evenness_sd = sd(evenness),
      evenness_se = evenness_sd / sqrt(evenness_n),
      evenness_min = min(evenness),
      evenness_max = max(evenness))
  
  saveRDS(matrices_ntmr, "matrices_ntmr_rds")
}

else{
  matrices_ntmr = read_rds("matrices_ntmr_rds")
}

# by YEAR and PROT
#mean abundance
ggplot(matrices_ntmr,aes(x=YEAR,y=abundance_mean, colour=factor(PROT), group=factor(PROT)))+ 
  geom_line(aes(group=PROT),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymax=abundance_mean+abundance_se,ymin=abundance_mean-abundance_se),width=0.1)+
  labs(title= "A) Mean Abundance", x="Year", y="Abundance of species", color = NULL)+
  scale_color_manual(
    values =  c("cyan","blue"),
    breaks = c("1", "0"), 
    labels=c("Protected","Not Protected"))+
  scale_x_continuous(
    limits = c(1999, 2016), 
    breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="right",
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10))
#ggsave(file="abundance_by_year_PROT.pdf", path="big_csv/plots")

#biomass
ggplot(matrices_ntmr,aes(x=YEAR,y=biomass_mean, colour=factor(PROT), group=factor(PROT)))+ #shape=PROT
  geom_line(aes(group=PROT),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin=biomass_mean-biomass_se, ymax=biomass_mean+biomass_se), width=.2) +
  labs(title= "B) Mean Biomass", x="Year", y="Biomass of species", colour = NULL)+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  scale_color_manual(
    values =  c("cyan","blue"),
    breaks = c("1", "0"), 
    labels=c("Protected","Not Protected"))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="right",
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10))
#ggsave(file="biomass_by_year_PROT.pdf", path="big_csv/plots")

#richness
ggplot(matrices_ntmr,aes(x=YEAR,y=richness_mean, colour=factor(PROT), group=factor(PROT)))+ #shape=PROT
  geom_line(aes(group=PROT),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin=richness_mean-richness_se, ymax=richness_mean+richness_se), width=.2)+
  labs(title= "C) Richness", x="Year", y="Effective number of species", color = NULL)+
  scale_color_manual(
    values =  c("cyan","blue"),
    breaks = c("1", "0"), 
    labels=c("Protected","Not Protected"))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="right",
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10))
#ggsave(file="richness_by_year_PROT.pdf", path="big_csv/plots")

#simpson
ggplot(matrices_ntmr,aes(x=YEAR,y=simpson_mean, colour=factor(PROT), group=factor(PROT)))+ #shape=PROT
  geom_line(aes(group=PROT),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin=simpson_mean-simpson_se, ymax=simpson_mean+simpson_se), width=.2)+
  labs(title= "D) Simpson", x="Year", y="Effective number of species",color = NULL)+
  scale_color_manual(
    values =  c("cyan","blue"),
    breaks = c("1", "0"), 
    labels=c("Protected","Not Protected"))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="right",
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10))
#ggsave(file="simpson_diversity_by_year_PROT.pdf", path="big_csv/plots")

# shannon
ggplot(matrices_ntmr,aes(x=YEAR,y=shannon_mean, colour=factor(PROT), group=factor(PROT)))+ #shape=PROT
  geom_line(aes(group=PROT),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymax=shannon_mean+shannon_se,ymin=shannon_mean-shannon_se),width=0.1)+
  labs(title= "E) Shannon", x="Year", y="Effective number of species",color = NULL)+
  scale_color_manual(
    values =  c("cyan","blue"),
    breaks = c("1", "0"), 
    labels=c("Protected","Not Protected"))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="right",
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10))
ggsave(file="shannon_diversity_by_year_PROT.pdf", path="big_csv/plots")


# functional
ggplot(matrices_ntmr,aes(x=YEAR,y=func.div_mean, colour=factor(PROT), group=factor(PROT)))+ #shape=PROT
  geom_line(aes(group=PROT),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymax=func.div_mean+func.div_se,ymin=func.div_mean-func.div_se),width=0.1)+
  labs(title= "F) Functional", x="Year", y="Effective number of species",color = NULL)+
  scale_color_manual(
    values =  c("cyan","blue"),
    breaks = c("1", "0"), 
    labels=c("Protected","Not Protected"))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="right",
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10))
#ggsave(file="functional_diversity_by_year_PROT.pdf", path="big_csv/plots")

# evenness
ggplot(matrices_ntmr,aes(x=YEAR,y=evenness_mean, colour=factor(PROT), group=factor(PROT)))+ #shape=PROT
  geom_line(aes(group=PROT),lwd=1) +
  geom_point(size=3) +
  geom_errorbar(aes(ymax=evenness_mean+evenness_se,ymin=evenness_mean-evenness_se),width=0.1)+
  labs(title= "G) Evenness", x="Year", y="UNITS",color = NULL)+
  scale_color_manual(
    values =  c("cyan","blue"),
    breaks = c("1", "0"), 
    labels=c("Protected","Not Protected"))+
  scale_x_continuous(limits = c(1999, 2016), breaks = c(2000,2002, 2004, 2006, 2008, 2010, 2012, 2014, 2016))+
  theme_bw(base_size=16)+
  theme(
      plot.margin=unit(c(1,1,1,1),"cm"),
      legend.position="right",
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.background=element_blank(),
      strip.text.x=element_text(size=18,hjust=0),
      plot.caption = element_text(size=10))
#ggsave(file="functional_diversity_by_year_PROT.pdf", path="big_csv/plots")

```

## Statistical analysis 

- Are diversity values and strata related? 
- Are doversoty values and subregions related? 
- Are diversity values and protection level related?

```{r stat libraries}

library(MASS) # contains boxcox function
library(moments) # contains skewness function
library(DescTools)
library(rpsychi) #statistical tests using summary data
library(AICcmodavg) #AIC
library(ade4) #mantel.rtest
library(ggplot2) #geom_boxplot and other plots
library(nortest) #checks for normality 
library(psych) #pairs.panels 
library(tidyverse)
library(magrittr) # needed to call this library separately from one in tidyverse in order to get functionality of github version not found in CRAN version
library(parallel)
#Create socket cluster of multiple cores 
cl=makeCluster(detectCores())
#Load automap on all cores (to access function `autoKrige`)
clusterEvalQ(cl,c(library(automap),library(mgcv),library(plotrix)) )

```

```{r Map matrices}
#Generate figure legend based on FKNMS map 
#Read in shapefile and fortify
if (!file.exists('fk.shp.fortify_rds')){
fk.shp = readOGR( dsn = "FKNMS_sample_grid", layer = "FlaKeys_Grid") 
fk.shp.fortify = fortify(fk.shp) 
saveRDS(fk.shp.fortify, 'fk.shp.fortify_rds')
} 
else{ 
  fk.shp.fortify = readRDS('fk.shp.fortify_rds')
}

all_matrices_data = read_rds("psu_matrices_data_rds")

#Create data frame for legend points and titles
data=data.frame(labels=c("Lower","Middle","Upper"),
                shape=c(15:18,6),color=c("grey15","grey30","grey45"),
                long=c(-76.15,-76.135,-76.21,-76.41,-76.34),
                lat=c(37.2,37.7,38.15,38.67,39.16),
                lat2=c(37.08,37.54,38.03,38.51,39.04))
data$labels=factor(data$labels,levels=c("Lower","Middle","Upper"))

#Generate map legend for inset
inset.map=
  ggplot()+
  #Plot outline of Florida Keys 
  geom_polygon(data=subset(fk.shp.fortify,group==0.1),aes(x=long,y=lat,group=group),col="grey50",fill="white",lwd=0.4)+
  geom_segment(aes(x=-75.98281,y=37.39372,xend=-76.24769,yend=37.39372),col="grey50",lwd=1)+
  geom_segment(aes(x=-76.24572,y=37.87455,xend=-76.01332,yend=37.87455),col="grey50",lwd=1)+
  geom_segment(aes(x=-76.41753,y=38.35538,xend=-76.26395,yend=38.35538),col="grey50",lwd=1)+
  geom_segment(aes(x=-76.49554,y=38.83621,xend=-76.35090,yend=38.83621),col="grey50",lwd=1)+
  #Add legend points
  geom_point(data=data,aes(x=long,y=lat,shape=labels,color=labels),size=8)+
  scale_color_manual(values=c("grey45","grey30","grey15"))+
  scale_shape_manual(values=c(6,18:15))+
  #Add legend labels
  geom_text(data=data,aes(x=long,y=lat2,label=labels),size=4.6,fontface="bold")+
  theme(
    plot.margin=unit(c(0,0,0,0),"cm"),
    panel.background=element_blank(),
    legend.position="none",panel.border=element_blank(),
    panel.grid.major=element_blank(),panel.grid.minor=element_blank(),
    axis.title.x=element_blank(),axis.title.y=element_blank(),
    axis.text.x=element_blank(),axis.text.y=element_blank(),
    axis.line=element_blank(),axis.ticks=element_blank())  

#Plot diversity by region, strata and year for each diversity index
lapply(alphadiv.list,function(i) {
  #Melt dataframe so diversity indices are in a single column
  x=melt(i,id.vars=c("Month","Year","Reg"),measure.vars=c("abundance", "biomass", "richness","simpson","shannon", "raoQ"))
  #Summarize means and SEs by month, region, and variable
  x=ddply(x,c("Month","Reg","variable"),summarize,value.mean=mean(value),value.se=std.error(value))
  #Rename levels for figure plotting
  levels(x$variable)=c("A) Abundance","B) Biomass","C) Richness ","D) Shannon","E) Gini-Simpson","F) Functional")
  #Rename regions for figure plotting
  x$Reg=factor(x$Reg); levels(x$Reg)=c("Upper","Middle","Lower")
  #Plot line graph
  p=ggplot(x,aes(x=as.factor(Month),y=value.mean,col=as.factor(Reg),shape=as.factor(Reg)))+
    #annotation_custom(grob=inset.map,xmin=19,xmax=22.5,ymin=0,ymax=1.5)+
    geom_line(aes(group=as.factor(Reg)),lwd=1)+geom_point(size=5)+
    scale_color_manual(values=c("grey15","grey30","grey45","grey60","black"),name="Region")+
    scale_shape_manual(values=c(15:18,6),name="Region")+
    geom_errorbar(aes(ymax=value.mean+value.se,ymin=value.mean-value.se),width=0.1)+facet_wrap(~variable,scales="free")+
    labs(x="",y="Diversity")+theme_bw(base_size=16)+
    theme(
      plot.margin=unit(c(0,1,0,0),"cm"),legend.position="none",
      panel.grid.major=element_blank(),panel.grid.minor=element_blank(),panel.margin=unit(0.8,"lines"),
      strip.background=element_blank(),strip.text.x=element_text(size=18,hjust=0))
  #Add legend map
  grid.arrange(p,inset.map,nrow=1,widths=c(1,0.2))
} )

#Save PDF (13" x 8")
```

Examine spatial and temporal trends in the different dimensions of reef fish diversity. 

- Interpolate diversity for each strata/region/year groupings using kriging 
```{r Kriging}

library(parallel) #clusterExport 
library(rgdal) #readOGR 
library(sp) #Calls: coordinates, spsample
all_matrices_data = read_rds("psu_matrices_data_rds")

#Create new list of data frames to convert to SpatialPointsDataFrame
alphasp.list=all_matrices_data
names(alphasp.list)=names(all_matrices_data)
#Convert class("dataframe") into class("SpatialPointsDataFrame")
for(i in seq_along(alphasp.list)) { sp:::coordinates(alphasp.list[[i]])=~LON_DEGREES+LAT_DEGREES } 
lapply(alphasp.list,class)
#Project lat long coordinates onto an ellipse
for(i in seq_along(alphasp.list)) { 
  proj4string(alphasp.list[[i]])="+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs" }
#Transform the projection into cartesian coordinates
alphasp.list=lapply(seq_along(alphasp.list),function(i) {
  spTransform(alphasp.list[[i]],CRS("+proj=utm +zone=18 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0")) } )

#Create a spatial grid onto which interpolated values can be plotted

#read in shapefile 
fk.shp = readOGR( dsn = "FKNMS_sample_grid", layer = "FlaKeys_Grid") 

#Generate interpolation grid
fknms.shp=spTransform(fk.shp,CRS("+proj=utm +zone=18 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0 ")) 
grd=spsample(fknms.shp,type="regular",n=2500,nsig=2) #n is sample size
par(mfrow=c(1,1))
plot(grd)

#Send interpolation grid to each core
parallel::clusterExport(cl, varlist=c("grd"))

all_matrices_data = as.matrix(all_matrices_data)

#Perform kriging interpolation for each matrix
#responses are log transformed to better meet assumptions of normality of errors and constant variance.
krigeall.list=parLapply(cl,alphasp.list,function(i) {
lapply(names(i@data)[66:71],function(j) {
as.data.frame(autoKrige(formula(paste("log10(",j,")~1")),i,grd)$krige_output) } ) } )

#Apply names for titling during plotting
#names(krigeall.list)=c("abund","biomass","pres.abs")
for(i in seq_along(krigeall.list)) {
  names(krigeall.list[[i]])=c("abundance_sum", "biomass_sum", "richness", "shannon","simpspn") 
  }  

#Plot all components across all years 
lapply(seq_along(krigeall.list),function(i) {
  do.call(grid.arrange,lapply(seq_along(krigeall.list[[i]]),function(j) {
    ggplot()+
      geom_raster(data=krigeall.list[[i]][[j]],aes(x=x1,y=x2,fill=var1.pred))+
      scale_fill_gradientn(colours=rev(rainbow(3)))+
      coord_equal()+labs(x="",y="",title=paste(names(krigeall.list[[i]])[j]))+ theme_bw()+theme(plot.margin=unit(c(0,0,0,0),"cm"),
                                                                                                axis.text.x=element_blank(),axis.text.y=element_blank(),
                                                                                                axis.ticks=element_blank(),legend.title=element_blank()) 
    } ) ) 
  } )

#Now repeat but instead interpolate by grouped years
krigemonth.list=parLapply(cl,alphasp.list,function(i) {
  lapply(names(i@data)[66:71],function(j) {
    lapply(unique((i@data)$STRAT),function(k) {
      as.data.frame(autoKrige(formula(paste("log(",j,")~1")),subset(i,i@data$STRAT==k),grd)$krige_output)  
      } ) 
    } ) 
  } )

#Apply names for titling during plotting
for(i in seq_along(krigemonth.list)) {
  names(krigemonth.list[[i]])=c("A) Mean Abundance","B) Mean Biomass","C) Richness","D) Shannon","E) Gini-Simpson","F) Functional") 
  }
for(i in seq_along(krigemonth.list)) {
  for(j in seq_along(krigemonth.list[[i]])) {
    names(krigemonth.list[[i]][[j]])=c("1999-2004","2005-2010","2011-2016") #change to year groupings 
  } 
}

#Collapse lists into data frame
krigemonth.list=lapply(krigemonth.list,function(i) lapply(i,function(j) do.call(rbind,j)) )

#Split rownames into month value and bind to dataframe
krigemonth.list=lapply(krigemonth.list,function(i) {
  lapply(i,function(j) {
    cbind(data.frame(str_split_fixed(rownames(j),"\\.",2)[,1]),j) } ) } )
for(i in seq_along(krigemonth.list)) {
  for(j in seq_along(krigemonth.list[[i]])) { names(krigemonth.list[[i]][[j]])[1]=c("month") } }
#Order levels of month for facetting
for(i in seq_along(krigemonth.list)) {
  for(j in seq_along(krigemonth.list[[i]])) { 
    krigemonth.list[[i]][[j]]$month=factor(krigemonth.list[[i]][[j]]$month,levels=c("Mar","May","July","Sept","Nov")) } }

#Plot and facet by month (12" x 18")
lapply(seq_along(krigemonth.list),function(i) {
  do.call(grid.arrange,c(lapply(seq_along(krigemonth.list[[i]]),function(j) {
    if(j!=2) {
      ggplot()+
        geom_raster(data=krigemonth.list[[i]][[j]],aes(x=x1,y=x2,fill=exp(var1.pred)))+
        scale_fill_gradientn(colours=rev(rainbow(3)))+
        facet_wrap(~month,nrow=1)+
        coord_equal()+labs(x="",y="",title=names(krigemonth.list[[i]])[j])+
        theme_bw()+theme(plot.margin=unit(c(0,0,0,0),"cm"),
                   plot.title=element_text(size=18,hjust=0,vjust=1),
                   strip.text=element_text(size=12),
                   panel.grid.major=element_blank(),panel.grid.minor=element_blank(),
                   axis.text.x=element_blank(),axis.text.y=element_blank(),
                   axis.ticks=element_blank(),legend.title=element_blank()) 
    } else {
      ggplot()+
        geom_raster(data=krigemonth.list[[i]][[j]],aes(x=x1,y=x2,fill=sin(var1.pred)^2))+
        scale_fill_gradientn(colours=rev(rainbow(3)))+
        facet_wrap(~month,nrow=1)+
        coord_equal()+labs(x="",y="",title=names(krigemonth.list[[i]])[j])+
        theme_bw()+theme(plot.margin=unit(c(0,0,0,0),"cm"),
                     plot.title=element_text(size=18,hjust=0,vjust=1),
                     strip.text=element_text(size=12),
                     panel.grid.major=element_blank(),panel.grid.minor=element_blank(),
                     axis.text.x=element_blank(),axis.text.y=element_blank(),
                     axis.ticks=element_blank(),legend.title=element_blank()) }
  } ), list(ncol=2)) )
} )

```

Build GAM model with temporal, spatial, environmental and management variables 

$y_i = \alpha + f_1(X_1i) + f_2(X_2i) + g_1(X_3i) + g_2(X_4i) + \epsilon_i$
$y_i = \alpha + f_1 (YEAR)+ f_2(PROT) + g_1(LONG, LAT)* YEAR + g_2(STRAT) + \epsilon_i$

$y_i$ is the calculated diversity for station i (response variable)
$\alpha$ is the intercept that scales the model prediction to the appropriate level of the response
$f_1 (X_1i)$ is a smoother
$g_1 (X_1i)$ is nonparametric smoother for covariates strata
$\epsilon_i$ the residual error

Temporal variables:
- YEAR: Year (categorical factor)

Spatial variables:
- latitude and longitude (continuous covariate)

Environmental variables: 
- strata (categorical factor)
- Rugosity (categorical factor)
- bottom depth (continuous covariate)
- vertical relief (continuous covariate)
- % coral cover (continuous covariate)
- % hard bottom (continuous covariate)

Management variable:
- PROT: No take marine reserve  (categorical factor)

Conitinuous covariates: modeled using non-parametric smoothing 
- thin plate regression splines with shrinkage terms (Ciannelli et al., 2008)

Categorical factors: models parametrically to determine their mean effect sizes (Zurr et al., 2009)

- fit GAM to each index separately. 

```{r practice GAM}

#Build full model and null model based on (transformed) response variable
#s() smooth term in gam model 
#bs() a two letter character string indicating the smoothing basis to use (ts plate regression splines)

# 1. Read in data and remove years (optional)
all_matrices_data = read_rds("psu_matrices_data_rds")
as.data.frame(all_matrices_data)
#names(all_matrices_data)

#Send all_matrices_data to each core
clusterExport(cl,varlist=c("all_matrices_data")) #varlist - character vector of names of objects to export


partial.deviance.list=parLapply(cl,seq_along(all_matrices_data),function(i) {
  do.call(rbind,lapply(list("abundance_sum","biomass_sum","richness","simpson","shannon"),function(j) {
    
    #Subset only columns that are necessary to run the models and omit any NAs so all models are run on the same data
    data=na.omit(all_matrices_data[[i]][,c("YEAR","LONG_DEGREES","LAT_DEGREES","DEPTH","STRAT","PROT","abundance_sum","biomass_sum","richness","simpson","shannon")])
    
    #Build full model and null model based on (transformed) response variable
    if(j=="richness") {
      fullmod = gam(formula(paste(j,"~YEAR+STRAT+PROT+s(LAT_DEGREES, LONG_DEGREES, by=YEAR, bs='ts')+s(DEPTH, by='ts')")), family=poisson, data=data)
      nullmod = gam(formula(paste(j,"~1")), family=poisson, data=data)
    } else {
      fullmod = gam(formula(paste("log(",j,")~YEAR+STRAT+PROT+s(LAT_DEGREES, LONG_DEGREES, by=YEAR, bs='ts')+s(DEPTH, by='ts')")), data=data)
      nullmod = gam(formula(paste("log(",j,")~1")), data=data)
    }
    
    #Set up empty vectors for deviances
    time.dev=c(); space.dev=c(); env.dev=c(); mngt.dev=c()
    
    #Sequence 1: time->space->environment
    time.dev[1]=
      (deviance(update(fullmod,.~.-YEAR,sp=fullmod$sp))-
         deviance(fullmod))/deviance(nullmod)
    space.dev[1]=
      (deviance(update(fullmod,.~.-YEAR-s(LONG_DEGREES,LAT_DEGREES,by=YEAR,bs='ts'),sp=fullmod$sp[6:9]))-
         deviance(update(fullmod,.~.-YEAR,sp=fullmod$sp)))/deviance(nullmod)
    env.dev[1]=
      (deviance(nullmod)-
         deviance(update(fullmod,.~.-YEAR-s(LONG_DEGREES,LAT_DEGREES,by=YEAR,bs='ts'),sp=fullmod$sp[6:9])))/deviance(nullmod)
    
    #Sequence 2: time->environment->space
    time.dev[2]=
      (deviance(update(fullmod,.~.-YEAR,sp=fullmod$sp))-
         deviance(fullmod))/deviance(nullmod)
    env.dev[2]=
      (deviance(update(fullmod,.~.-YEAR-STRAT-s(DEPTH,bs='ts'),sp=fullmod$sp[1:5]))-
         deviance(update(fullmod,.~.-YEAR,sp=fullmod$sp)))/deviance(nullmod)
    space.dev[2]=
      (deviance(nullmod)-
         deviance(update(fullmod,.~.-YEAR-STRAT-s(DEPTH,bs='ts'),sp=fullmod$sp[1:5])))/deviance(nullmod)
    
    #Sequence 3: space->time->environment
    space.dev[3]=
      (deviance(update(fullmod,.~.-s(LONG_DEGREES,LAT_DEGREES,by=YEAR,bs='ts'),sp=fullmod$sp[6:9]))-
         deviance(fullmod))/deviance(nullmod)
    time.dev[3]=
      (deviance(update(fullmod,.~.-s(LONG_DEGREES,LAT_DEGREES,by=YEAR,bs='ts')-YEAR,sp=fullmod$sp[6:9]))-
         deviance(update(fullmod,.~.-s(LONG_DEGREES,LAT_DEGREES,by=YEAR,bs='ts'),sp=fullmod$sp[6:9])))/deviance(nullmod)
    env.dev[3]=
      (deviance(nullmod)-
         deviance(update(fullmod,.~.-s(LONG_DEGREES,LAT_DEGREES,by=YEAR,bs='ts')-YEAR,sp=fullmod$sp[6:9])))/deviance(nullmod)
    
    #Sequence 4: space->environment->time
    space.dev[4]=
      (deviance(update(fullmod,.~.-s(LONG_DEGREES,LAT_DEGREES,by=YEAR,bs='ts'),sp=fullmod$sp[6:9]))-
         deviance(fullmod))/deviance(nullmod)
    env.dev[4]=
      (deviance(update(fullmod,.~.-s(LONG_DEGREES,LAT_DEGREES,by=YEAR,bs='ts')-STRAT-s(DEPTH,bs='ts')))-
         deviance(update(fullmod,.~.-s(LONG_DEGREES,LAT_DEGREES,by=YEAR,bs='ts'),sp=fullmod$sp[6:9])))/deviance(nullmod)
    time.dev[4]=
      (deviance(nullmod)-
         deviance(update(fullmod,.~.-s(LONG_DEGREES,LAT_DEGREES,by=YEAR,bs='ts')-STRAT-s(DEPTH,bs='ts'))))/deviance(nullmod)
    
    #Sequence 5: environment->time->space
    env.dev[5]=
      (deviance(update(fullmod,.~.-STRAT-s(DEPTH,bs='ts'),sp=fullmod$sp[1:5]))-
         deviance(fullmod))/deviance(nullmod)
    time.dev[5]=
      (deviance(update(fullmod,.~.-STRAT-s(DEPTH,bs='ts')-YEAR,sp=fullmod$sp[1:5]))-
         deviance(update(fullmod,.~.-STRAT-s(DEPTH,bs='ts'),sp=fullmod$sp[1:5])))/deviance(nullmod)
    space.dev[5]=
      (deviance(nullmod)-
         deviance(update(fullmod,.~.-STRAT-s(DEPTH,bs='ts')-YEAR,sp=fullmod$sp[1:5])))/deviance(nullmod)
    
    #Sequence 6: environment->space->time
    env.dev[6]=
      (deviance(update(fullmod,.~.-STRAT-s(DEPTH,bs='ts'),sp=fullmod$sp[1:5]))-
         deviance(fullmod))/deviance(nullmod)
    space.dev[6]=
      (deviance(update(fullmod,.~.-STRAT-s(DEPTH,bs='ts')-s(LONG_DEGREES,LAT_DEGREES,by=YEAR,bs='ts')))-
         deviance(update(fullmod,.~.-STRAT-s(DEPTH,bs='ts'),sp=fullmod$sp[1:5])))/deviance(nullmod)
    time.dev[6]=
      (deviance(nullmod)-
         deviance(update(fullmod,.~.-STRAT-s(DEPTH,bs='ts')-s(LONG_DEGREES,LAT_DEGREES,by=YEAR,bs='ts'))))/deviance(nullmod)

    #Bind results into data.frame
    prop.deviance=data.frame(
      Diversity=paste(j),
      Predictor=c("All","Space","Time","Environment","Management"),
      Partial.Deviance=c(summary(fullmod)$dev.expl,mean(space.dev),mean(time.dev),mean(env.dev),mean(mngt.dev)),
      Partial.Deviance.SE=c(0,std.error(space.dev),std.error(time.dev),std.error(env.dev),std.error(mngt.dev)))

  } 
  ) ) 
} 
)
    
#Close the cluster
stopCluster(cl)    
  
#   summary(shannon_fullmodel) #explained deviance (R2), variance of residuals, etc
#   anova(shannon_fullmodel)#p-value and F-test
#   gam.check(shannon_fullmodel) #produces graphical diagnosis for the model. Use to 1) assess normality with QQ plot and the histogram 2) homogeneity (residuals vs fitted values) 3) model fit (fitted values versus observed values). Ideally the response against fitted values should show a straight line. 
# 
#     rich_model = gam(formula(richness~YEAR+rugosity+s(longitude,latitude,by=YEAR,bs='ts')+s(depth,bs='ts')+s(vert_relief,bs='ts')+s(per_coral_cover,bs='ts')+s(per_hardbottom,bs='ts')), family=poisson),
#     rich_nullmodel = gam(formula(richness,"~1"), family=poisson) 
#     
#     shan_model = gam(formula(shannon~YEAR+rugosity+s(longitude,latitude,by=YEAR,bs='ts')+s(depth,bs='ts')+s(vert_relief,bs='ts')+s(per_coral_cover,bs='ts')+s(per_hardbottom,bs='ts'))) 
#     shannon_nullmodel = gam(formula("log(",shannon,")~1"))) 
# 
# mgcv::summary(gam$rich_model) #explained deviance (R2), variance of residuals, etc. 
# mgcv::anova(gam$rich_model)#p-value and F-test
# mgcv::plot(gam$rich_model) #produces fitted smoother
# gam.check(gam$rich_model) #produces graphical diagnosis for the model. Use to 1) assess normality with QQ plot and the histogram 2) homogeneity (residuals vs fitted values) 3) model fit (fitted values versus observed values). Ideally the response against fitted values should show a straight line. 
# 
# #Check how well the model fits using graphical evidence or if it isn't clear with AIC, which measures goodness of fit and model complexity. The lower the AIC, the better the model fits the data. Can also look at GCV value in summary output. 
# AIC(gam$rich_model)

```

- To assess the explanatory power of each variable calculate the partial deviances by sequentially removing suites of predictors from the full gam model corresponding to indicators of space, time, environment, or management. 

- Repeat this for all possible permutations 

- Average the deviances for all models in which the predictor appeared and calculate the standard error 

- the partial deviances are the proportion of total explained deviance of the model uniquely explained by space, time, environment or management, accounting for the other predictors in the model. 

```{r partial deviances}

#Name objects in list
names(partial.deviance.list)=c("abund","biomass","pres.abs")
#Rename diversity levels for better plotting
for(i in seq_along(partial.deviance.list)) {
  partial.deviance.list[[i]]=transform(partial.deviance.list[[i]],Diversity=recode(Diversity,
                                                                                   "'richness'='Richness';'evenness'='Evenness';'species.div'='Gini-Simpson';
     'func.div'='Functional';'phylo.div'='Phylogenetic';'taxo.div'='Taxonomic'")) }
#Reorder diversity levels for better plotting
for(i in seq_along(partial.deviance.list)) {
  partial.deviance.list[[i]][,"Diversity"]=factor(partial.deviance.list[[i]][,"Diversity"],
                                                  levels=c("Richness","Evenness","Gini-Simpson","Functional","Phylogenetic","Taxonomic")) }

#Create barplot (8" x 6")
lapply(seq_along(partial.deviance.list),function(i) {
  data=subset(partial.deviance.list[[i]],Predictor!="All")
  ggplot(data,aes(x=Diversity,y=Partial.Deviance*100,fill=Predictor))+
    geom_bar(position="dodge",color="black")+
    geom_errorbar(aes(max=(Partial.Deviance+Partial.Deviance.SE)*100,
                      min=(Partial.Deviance-Partial.Deviance.SE)*100),
                  width=0.3,size=0.6,position=position_dodge(width=0.9))+
    labs(x="",y="Total % Deviance Explained\n")+
    scale_fill_manual(values=c("grey50","grey80","white"))+
    theme_bw(base_size=18)+theme(
      panel.grid.major=element_blank(),panel.grid.minor=element_blank(),
      legend.position="bottom",legend.title=element_blank()) } )

```

