title: "MBON SDG14 FKNMS Ecological Classifications"
author: "Megan Hepner"
date: "May 5, 2017"
output: html_document
---

## Sustainable Development Goal 

SDG14: Conserve and sustainably use the oceans, seas and marine resources
    + SDG14.2:VBy 2020, sustainably manage, and protect marine and coastal ecosystems to avoid significant adverse impacts, including by strengthening their resilience and take action for their restoration, to achieve healthy and productive oceans

## National Marine Sanctuary Condition Report Questions 

### Water Quality 
1. What is the eutrophic condition of the sanctuary waters and how is it changing?
2. Do sanctuary waters pose risks to human health and how are they changing?
3. Have recent changes in climate alter water conditions and how are they changing?
4. Are other stressors, individually or in combination, affecting water quality, and how are they changing?

### Habitat Resources 
5. What is the integrity of major habitat types and how are they changing?
6. What are contaminant concentrations in sanctuary habitats and how are they changing?

### Living Resources 
7. What is the status and trend of keystone and foundation species and how is it changing?
8. What is the status and trend of other focal species and how is it changing?
9. What is the status of non-indigenous species and how is it changing?
10. What is the status of biodiversity and how is it changing?

### Maritime Archeological Resources  
11. What is the archaeological integrity of known maritime archaeological resources and how is it changing? 
12. Do known maritime archaeological resources pose an environmental hazard and how is this threat changing?

### Human Demensions 
13. What are the levels of human activities that may adversely influence water quality and how are they changing? 
14. What are the levels of human activities that may adversely influence habitats and how are they changing?
15. What are the levels of human activities that may adversely influence living resource quality and how are they changing? 
16. What are the levels of human activities that may adversely influence maritime archaeological resource quality and how are they changing?
17. What are the states of influential human drivers and how are they changing?

## FKNMS Infographic data 

| Classification level           | Icon                   | Metric                                        | Dataset | 
|:-------------------------------|:-----------------------|:----------------------------------------------|:--------|
| Trophic groups                 | Forage fish            | biomass, density                              | RVC     | 
| Trophic groups                 | Grouper                | biomass, density, richness simpson, shannon   | RVC     | 
| Trophic groups                 | Grunt                  | biomass, density, richness simpson, shannon   | RVC     | 
| Trophic groups                 | Higher level reef fish | biomass, density, richness simpson, shannon   | RVC     |
| Trophic groups                 | Hogfish                | biomass, density                              | RVC     |
| Trophic groups                 | Lionfish               | biomass, density                              | RVC     |
| Trophic groups                 | Opportunists           | biomass, density, richness simpson, shannon   | RVC     |
| Trophic groups                 | Parrotfish             | biomass, density, richness simpson, shannon   | RVC     |
| Trophic groups                 | Rays                   | biomass, density                              | RVC     |
| Trophic groups                 | Seabass and Hamlets    | biomass, density, richness simpson, shannon   | RVC     |
| Trophic groups                 | Sharks                 | biomass, density                              | RVC     |
| Trophic groups                 | Small reeffish         | biomass, density, richness simpson, shannon   | RVC     |
| Trophic groups                 | Snapper                | biomass, density, richness simpson, shannon   | RVC     |
| Trophic level                  | Trophic pyramid        | biomass                                       | RVC     |
| Focal species                  | Exploited fish icon    | biomass, density                              | RVC     |
| Biodiversity                   | Reef fish biodiversity | richness, simpson, shannon                    | RVC     |
| Foundation species             | Stony coral            | percent cover & diseased                      | CREMP   |
| Foundation species             | Sea fan                | percent cover & diseased                      | CREMP   |
| Foundation species             | Macroalgae             | percent cover & diseased                      | CREMP   |       
| Foundation species             | Sponges                | percent cover & diseased                      | CREMP   |       
| Keystone species               | Long-spined sea urchin | abundance                                     | TBD     |
| Focal species                  | Caribbean spiny lobster| abundance                                     | TBD     |
| Focal species                  | Queen Conch            | abundance                                     | TBD     |
| Focal species                  | Sea turtles            | abundance                                     | TBD     |
    
```{r librarys}
#rm(list = ls())

library(knitr)
library(tidyverse)
library(rvc)
library(vegan)
library(dygraphs)
library(grDevices)
library(RColorBrewer)

```

# Reef Visual Census (RVC)

In order to get the big RDS files, Megan shared her Dropbox folder big_csv, so on Mac in Terminal shell (or directly in Git pane, gear icon, Shell...), run this command to symbolically link the `big_csv` folder: 

```
ln -s ~/Dropbox/big_csv big_csv
```

```{r call RVC data}
#rds files to read 
RVCdata_FK_rds = 'big_csv/RVCdata_FK.rds'

#rvc94_16_csv = 'big_csv/rvc94_16.csv'

if (!all(file.exists(RVCdata_FK_rds))){
  RVCdata_FK <- getRvcData(1999:2016, "FLA KEYS", server = 'https://www.sefsc.noaa.gov/rvc_analysis20/')
  write_rds(RVCdata_FK, 'big_csv/RVCdata_FK.rds')

} else{ #read data
RVCdata_FK = read_rds('big_csv/RVCdata_FK.rds')}

# add 1994 - 1998 RVC data and reweight by strata
# if (!file.exists(rvc94_16_csv)){
#   rvc98 <-read.csv("90srvc/fk1998_dat2.csv")
#   rvc97 <-read.csv("90srvc/fk1997_dat2.csv")
#   rvc96 <-read.csv("90srvc/fk1996_dat2.csv")
#   rvc95 <-read.csv("90srvc/fk1995_dat2.csv")
#   rvc94 <-read.csv("90srvc/fk1994_dat2.csv")
#   rvc94_16 <- rbind(RVCdata_FK$sample_data, rvc98, rvc97, rvc96, rvc95, rvc94)
#   write_csv(rvc94_16, "big_csv/rvc94_16.csv")
# } else {
#   rvc94_16 = read.csv("big_csv/rvc94_16.csv")
# }
# 
# strat94 <- read.csv("90srvc/FKeys_NTOTS_Reweight_1994.csv")
# strat <- strat94[which(strat94$YEAR<1999),]
# RVCdata_FK$sample_data <- rvc94_16
# RVCdata_FK$stratum_data <- rbind(strat,RVCdata_FK$stratum_data)
```

## Trophic Groups
1. Forage fish 
2. Grouper
3. Grunts
4. Higher Level Reef Fish
5. Hogfish
6. Lionfish
7. Opportunists 
8. Parrotfish
9. Rays
10. Sea bass and Hamlets
11. Sharks
12. Small Reef Fish
13. Snappers

#### Forage fish

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Hypoatherina harringtonensis* | reef silverside	    | HYP HARR  |
|*Atherinomorus stipes*	        | hardhead silverside	| ATH STIP  |
|*Jenkinsia sp.*	              | herring species	    | JEN SPE.  |
|*Harengula jaguana*	          | scaled sardine	    | HAR JAGU  |
|*Anchoa lyolepis*	            | dusky anchovy	      | ANC LYOL  |
|*Inermia vittata*	            | boga	              | INE VITT  |
|*Hemiramphus brasiliensis*	    | ballyhoo	          | HEM BRAS  |
|*Sardinella aurita*	          | spanish sardine	    | SAR AURI  |
|*Harengula humeralis*	        | redear sardine	    | HAR HUME  |
|*Chriodorus atherinoides*	    | hardhead halfbeak	  | CHR ATHE  |


```{r Forage fish}

forage_fk_abun_csv     = "trophic_groups/abundance/forage_fk_abun.csv"        #abundance
forage_fk_den_csv      = "trophic_groups/density/forage_fk_den.csv"           #density
forage_fk_bio_csv      = "trophic_groups/biomass/forage_fk_bio_merged.csv"    #biomass 
forage_fk_diversity_csv= "trophic_groups/diversity/forage_fk_diversity.csv"   #diversity 

if (!all(
  file.exists(
    forage_fk_abun_csv, forage_fk_den_csv, forage_fk_bio_csv,forage_fk_diversity_csv))){
  
  forage_fish_spp_list= c("HYP HARR", "ATH STIP", "JEN SPE.", "HAR JAGU", "ANC LYOL", "INE VITT", "HEM BRAS", "SAR AURI", "HAR HUME", "CHR ATHE")

  #abundance
  forage_fk_abun = getDomainAbundance(RVCdata_FK, forage_fish_spp_list)
  write_csv(forage_fk_abun, "trophic_groups/abundance/forage_fk_abun.csv")
 
  #density
  forage_fk_den = getDomainDensity(RVCdata_FK, forage_fish_spp_list)
  write_csv(forage_fk_den, "trophic_groups/density/forage_fk_den.csv")
  
  #biomass
  forage_fk_bio = getDomainBiomass(RVCdata_FK, forage_fish_spp_list, RVCdata_FK$taxonomic_data)
  write_csv(forage_fk_bio, "trophic_groups/biomass/forage_fk_bio_merged.csv")
 
  #fk diversity metrics 
  forage_fk_abundance = forage_fk_abun %>% 
    select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
    group_by(YEAR, protected_status) %>% #48 groups 
    nest(-YEAR) %>%  
    mutate(
      data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
      #species richness 
      richness = map(
        data_wide, 
        function(x) specnumber(x)),
      #simpson diversity as effective number of species 
      simpson = map( 
        data_wide,
        function(x) 1/(1 - diversity(x, index = 'simpson'))),
      #shannon diversity as effective number of species 
      shannon = map(
        data_wide,
        function(x) exp(diversity(x, index = 'shannon')))) %>%
    unnest(richness, simpson, shannon)
  
  forage_fk_diversity = forage_fk_abundance %>%
    select(YEAR, protected_status, richness, shannon, simpson) 
  write_csv(forage_fk_diversity, "big_csv/trophic_groups/diversity/forage_fk_diversity.csv")
} 

```

#### Grouper

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Cephalopholis cruentata*      |	graysby	            | CEP CRUE  |
|*Cephalopholis fulva*          |	coney	              | CEP FULV  |
|*Dermatolepis inermis*         |	marbled grouper	    | DER INER  |
|*Epinephelus flavolimbatus*    |	yellowedge grouper	| EPI FLAV  |
|*Epinephelus itajara*          |	goliath grouper	    | EPI ITAJ  |
|*Epinephelus morio*            |	red grouper	        | EPI MORI  |
|*Epinephelus niveatus*	        | snowy grouper	      | EPI NIVE  |
|*Epinephelus striatus*         |	nassau grouper	    | EPI STRI  |
|*grouper-sea bass species*     |	grouper-sea bass sp |	SRR SPE.  |
|*Mycteroperca acutirostris*    |	western comb grouper| MYC ACUT  |
|*Mycteroperca bonaci*          |	black grouper	      | MYC BONA  |
|*Mycteroperca interstitialis*  |	yellowmouth grouper	| MYC INTE  |
|*Mycteroperca microlepis*      |	gag	                | MYC MICR  |
|*Mycteroperca phenax*          |	scamp	              | MYC PHEN  |
|*Mycteroperca tigris*          |	tiger grouper	      | MYC TIGR  |
|*Mycteroperca venenosa*        |	yellowfin grouper	  | MYC VENE  |


```{r Grouper}

grouper_fk_abun_csv       = "trophic_groups/abundance/grouper_fk_abun.csv"
grouper_fk_den_csv        = "trophic_groups/density/grouper_fk_den.csv"
grouper_fk_bio_csv        = "trophic_groups/biomass/grouper_fk_bio_merged.csv"
grouper_fk_diversity_csv  = "trophic_groups/diversity/grouper_fk_diversity.csv"

if(!all(file.exists(grouper_fk_abun_csv,grouper_fk_den_csv,grouper_fk_bio_csv, grouper_fk_diversity_csv))){
  
  grouper_spp_list= c("CEP CRUE", "MYC BONA", "EPI MORI", "EPI STRI", "EPI ITAJ", "CEP FULV", "MYC MICR", "MYC PHEN", "MYC VENE", "MYC INTE", "MYC TIGR", "EPI FLAV", "DER INER", "SRR SPE.", "Epi nive", "Myc acut")
  
  #abundance
  grouper_fk_abun = getDomainAbundance(RVCdata_FK, grouper_spp_list)
  write_csv(grouper_fk_abun, "trophic_groups/abundance/grouper_fk_abun.csv")
  
  #density
  grouper_fk_den = getDomainDensity(RVCdata_FK, grouper_spp_list)
  write_csv(grouper_fk_den, "trophic_groups/density/grouper_fk_den.csv")
  
  #biomass
  grouper_fk_bio = getDomainBiomass(RVCdata_FK, grouper_spp_list, RVCdata_FK$taxonomic_data)
  write_csv(grouper_fk_bio, "trophic_groups/biomass/grouper_fk_bio_merged.csv")
  
  #diversity metrics 
  grouper_fk_abundance = grouper_fk_abun %>% 
    select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
    group_by(YEAR, protected_status) %>% 
    nest(-YEAR) %>%  
    mutate(
      data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill =0)), 
      #species richness 
      richness = map(
        data_wide, 
        function(x) specnumber(x)),
      #simpson diversity as effective number of species 
      simpson = map( 
        data_wide,
        function(x) 1/(1 - diversity(x, index = 'simpson'))),
      #shannon diversity as effective number of species 
      shannon = map(
        data_wide,
        function(x) exp(diversity(x, index = 'shannon')))) %>%
    unnest(richness, simpson, shannon)
  
  grouper_fk_diversity = grouper_fk_abundance %>%
    select(YEAR, protected_status, richness, shannon, simpson) 
  write_csv(grouper_fk_diversity, "big_csv/trophic_groups/diversity/grouper_fk_diversity.csv")
}

```

#### Grunts 

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Anisotremus surinamensis*	    | black margate	      | ANI SURI  |
|*Anisotremus virginicus*	      | porkfish	          | ANI VIRG  |
|*Haemulon album*	              | margate	            | HAE ALBU  |
|*Haemulon aurolineatum*	      | tomtate	            | HAE AURO  | 
|*Haemulon carbonarium*	        | caesar grunt	      | HAE CARB  |
|*Haemulon chrysargyreum*       |	smallmouth grunt	  | HAE CHRY  |
|*Haemulon flavolineatum*       |	french grunt	      | HAE FLAV  |
|*Haemulon macrostomum*	        | spanish grunt	      | HAE MACR  |
|*Haemulon melanurum*	          | cottonwick	        | HAE MELA  |
|*Haemulon parra*               |	sailors choice	    | HAE PARR  |
|*Haemulon plumierii*           |	white grunt	        | HAE PLUM  | 
|*Haemulon sciurus*             |	bluestriped grunt	  | HAE SCIU  |
|*Haemulon sp.*                 |	grunt species	      | HAE SPE.  |
|*Haemulon striatum*            |	striped grunt	      | HAE STRI  | 

```{r Grunts}

grunt_fk_abun_csv     = "trophic_groups/abundance/grunt_fk_abun.csv"
grunt_fk_den_csv      = "trophic_groups/density/grunt_fk_den.csv"
grunt_fk_bio_csv      = "trophic_groups/biomass/grunt_fk_bio_merged.csv"
grunt_fk_diversity_csv= "trophic_groups/diversity/grunt_fk_diversity.csv"

if(!all(file.exists(grunt_fk_abun_csv,grunt_fk_den_csv,grunt_fk_bio_csv, grunt_fk_diversity_csv))){
  
  grunt_spp_list = c("HAE SCIU", "HAE AURO", "HAE PLUM", "HAE SPE.", "HAE FLAV", "HAE CHRY", "ANI VIRG","HAE MELA", "HAE CARB", "HAE PARR", "HAE STRI", "HAE MACR", "ANI SURI", "HAE ALBU")
  
  #abundance
  grunt_fk_abun = getDomainAbundance(RVCdata_FK, grunt_spp_list)
  write_csv(grunt_fk_abun, "trophic_groups/abundance/grunt_fk_abun.csv")
  
  #density
  grunt_fk_den = getDomainDensity(RVCdata_FK, grunt_spp_list)
  write_csv(grunt_fk_den, "trophic_groups/density/grunt_fk_den.csv")
  
  #biomass
  grunt_fk_bio = getDomainBiomass(RVCdata_FK, grunt_spp_list, RVCdata_FK$taxonomic_data)
  write_csv(grunt_fk_bio, "trophic_groups/biomass/grunt_fk_bio_merged.csv")
  
  #diversity metrics 
  grunt_fk_abundance = grunt_fk_abun %>% 
    select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
    group_by(YEAR, protected_status) %>% #48 groups 
    nest(-YEAR) %>%  
    mutate(
      data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
      #species richness 
      richness = map(
        data_wide, 
        function(x) specnumber(x)),
      #simpson diversity as effective number of species 
      simpson = map( 
        data_wide,
        function(x) 1/(1 - diversity(x, index = 'simpson'))),
      #shannon diversity as effective number of species 
      shannon = map(
        data_wide,
        function(x) exp(diversity(x, index = 'shannon')))) %>%
    unnest(richness, simpson, shannon)
  
  grunt_fk_diversity = grunt_fk_abundance %>%
    select(YEAR, protected_status, richness, simpson, shannon) 
  write_csv(grunt_fk_diversity, "trophic_groups/diversity/grunt_fk_diversity.csv")
}

```

#### Higher Level Reef Fish

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|

```{r Higher Level Reef Fish}

higher_reef_fk_abun_csv     = "trophic_groups/abundance/higher_reef_fk_abun.csv"
higher_reef_fk_den_csv      = "trophic_groups/density/higher_reef_fk_den.csv"
higher_reef_fk_bio_csv      = "trophic_groups/biomass/higher_reef_fk_bio_merged.csv"
higher_reef_fk_diversity_csv= "trophic_groups/diversity/higher_reef_fk_diversity.csv"

if(!all(file.exists(higher_reef_fk_abun_csv,higher_reef_fk_den_csv,higher_reef_fk_bio_csv,higher_reef_fk_diversity_csv))){
  
  higher_reef_spp_list = c("THA BIFA","STE PART","GYM MORI","PRI AREN","ABU SAXA","HAL BIVI","HAL GARN","HAL MACU","CHR MULT","CLE PARR","STE PLAN","CHR CYAN","MUL MART","MIC CHRY","STE ADUS","CHA CAPI","STE VARI","PEM SCHO","POM ARCU","CAN ROST","CHA OCEL","PSE MACU","HOL TRIC","CHR SCOT","HAL RADI","CAL CALA","CHA SEDE","XYR SPLE","STE LEUC","HOL CILI","POM PARU","AUL MACU","CHA STRI","OPI AURI","HAL POEY","CHR INSO","HOL ADSC","ODO DENT","STE DIEN","HOL BERM","CHA FABE","XYR MART","GER CINE","HOL RUFU","LAC TRIQ","PAR ACUM","ALU SCRI","BAL CAPR","MAL PLUM","CAN SUFF","CAN PULL","HET CRUE","OGC SPE.","XYR SPE.","ACA QUAD","CAL BAJO","MON TUCK","SPH SPEN","ARC RHOM","DIO HOLO","LAC BICA","XYR NOVA","BAL VETU","DIO HYST","HAL CYAN","GYM FUNE","DIP HOLB","APO PSEU","HOL TOWN","EQU PUNC","LAC TRIG","HAL PICT","MYR JACO","SAR VEXI","SYN INTE","HAL CAUD","AMB PINO","GYM MILI","ALU SCHO","ACA POLY","CAL PROR","CHI SCHO","HET LONG","SCO PLUM","CHR ENCH","CAL SPE.","APO MACU","CAN MACR","PAR UMBR","LAG RHOM","EQU LANC","CAL PENN","STE HISP","APO BINO","NEO MARI","ARC PROB","CHI ANTE","APO TOWN","DOR MEGA","REM REMO","OPI WHIT","CEN ARGI","SYN FOET","AST SPE.","CAR HIPP","SAR CORU","OPI MACR","GYM VICI","DIP ARGE","PAG PAGR","PAR BAIR","BOT LUNA","EUC ARGE","PRO ACUL","LAB GOBI","FIS TABA","ALU SPE.","GRA LORE","MON CILI","ECH NEUC","ALU MONO","BOT OCEL","BAL SPE.","PAR ALBI","STE SPE.","APO QUAD","ENC NIGR","GYM SAXI","OPI SPE.","ELO SAUR","MAN BIRO","APO AURO","SPH TEST","CAL NODO","DIO SPE.","SCO CARI","LAB FILA","AHL EGMO","ANT OCEL","AST GUTT","LAB SPE.","LAB SPE.","MUR RETI","AST STEL","EUC GULA","HOL SPE.","ORT CHRY","PRI OPHR","LAB BUCC","ACA SOLA","CAL LEUC","ENC CARY","GYM NIGR","MYR BREV","SAR BULL","TRA GOOD","TRA LATH","UMB CORO","EUC JONE","EUCLEFR","STY LATE","SYA MICR","SYN SCOV","APO PHEN","HIP REID","PHA XENU","AST PUNC","CHI ATIN","CHL CHRY","HIP EREC","OGC NASU","SER PHOE","SPH NEPH","MYR OCEL","PRI RUBI","RHI LENT","ALB VULP","DAC VOLI","RYP BIST","BRO BARB","HYP GUMM","OPS TAU_","EMM ATLA","UPE PARV","SYN SYNO")
  
  #abundance
  higher_reef_fk_abun = getDomainAbundance(RVCdata_FK, higher_reef_spp_list)
  write_csv(higher_reef_fk_abun, "trophic_groups/abundance/higher_reef_fk_abun_merged.csv")
  
  #density
  higher_reef_fk_den = getDomainDensity(RVCdata_FK, higher_reef_spp_list)
  write_csv(higher_reef_fk_den, "trophic_groups/density/higher_reef_fk_den_merged.csv")
  
  #biomass
  higher_reef_fk_bio = getDomainBiomass(RVCdata_FK, higher_reef_spp_list, RVCdata_FK$taxonomic_data)
  write_csv(higher_reef_fk_bio, "trophic_groups/biomass/higher_reef_fk_bio_merged.csv")
  
  #fk diversity
  higher_reef_fk_abundance = higher_reef_fk_abun %>% 
    select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
    group_by(YEAR, protected_status) %>% #48 groups 
    nest(-YEAR) %>%  
    mutate(
      data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
      #species richness 
      richness = map(
        data_wide, 
        function(x) specnumber(x)),
      #simpson diversity as effective number of species 
      simpson = map( 
        data_wide,
        function(x) 1/(1 - diversity(x, index = 'simpson'))),
      #shannon diversity as effective number of species 
      shannon = map(
        data_wide,
        function(x) exp(diversity(x, index = 'shannon')))) %>%
    unnest(richness, simpson, shannon)
  
  higher_reef_diversity = higher_reef_fk_abundance %>%
    select(YEAR, protected_status, richness, simpson, shannon) 
  write_csv(higher_reef_diversity, "trophic_groups/diversity/higher_reef_fk_diversity.csv")
}

```

#### Hogfish

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Bodianus pulchellus*	        | spotfin hogfish	    | BOD PULC  |
|*Bodianus rufus*	              | Spanish hogfish	    | BOD RUFU  |
|*Lachnolaimus maximus*	        | hogfish	            | LAC MAXI  |

```{r Hogfish}

hogfish_fk_abun_csv     = "trophic_groups/abundance/hogfish_fk_abun.csv"
hogfish_fk_den_csv      = "trophic_groups/density/hogfish_fk_den.csv"
hogfish_fk_bio_csv      = "trophic_groups/biomass/hogfish_fk_bio_merged.csv"

if(!all(file.exists(hogfish_fk_abun_csv,hogfish_fk_den_csv))){
  
  hogfish_spp_list = c("LAC MAXI", "BOD RUFU", "BOD PULC")
  
  #abundance
  hogfish_fk_abun = getDomainAbundance(RVCdata_FK, hogfish_spp_list, merge_protected = F)
  write_csv(hogfish_fk_abun, "trophic_groups/abundance/hogfish_fk_abun.csv")
  
  #density
  hogfish_fk_den = getDomainDensity(RVCdata_FK, hogfish_spp_list, merge_protected = F)
  write_csv(hogfish_fk_den, "trophic_groups/density/hogfish_fk_den.csv")
  
  #biomass
  hogfish_fk_bio = getDomainBiomass(RVCdata_FK, hogfish_spp_list, RVCdata_FK$taxonomic_data)
  write_csv(hogfish_fk_bio, "trophic_groups/biomass/hogfish_fk_bio_merged.csv")
  
  #FK diversity
  hogfish_fk_abundance = hogfish_fk_abun %>% 
    select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
    group_by(YEAR, protected_status) %>% #48 groups 
    nest(-YEAR) %>%  
    mutate(
      data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
      #species richness 
      richness = map(
        data_wide, 
        function(x) specnumber(x)),
      #simpson diversity as effective number of species 
      simpson = map( 
        data_wide,
        function(x) 1/(1 - diversity(x, index = 'simpson'))),
      #shannon diversity as effective number of species 
      shannon = map(
        data_wide,
        function(x) exp(diversity(x, index = 'shannon')))) %>%
    unnest(richness, simpson, shannon)
  
  hogfish_fk_diversity = hogfish_fk_abundance %>%
    select(YEAR, protected_status, richness, simpson, shannon) 
  write_csv(hogfish_fk_diversity, "trophic_groups/diversity/hogfish_fk_diversity.csv")
}
```

#### Lionfish

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Pterois volitans*             | red lionfish        | PTE VOLI  |

```{r Lionfish}

lionfish_fk_abun_csv     = "trophic_groups/abundance/lionfish_fk_abun.csv"
lionfish_fk_den_csv      = "trophic_groups/density/lionfish_fk_den.csv"
lionfish_fk_bio_csv      = "trophic_groups/biomass/lionfish_fk_bio_merged.csv"

if(!all(file.exists(lionfish_fk_abun_csv,lionfish_fk_den_csv,lionfish_fk_bio_csv))){
  #"PTE VOLI" - lionfish
  
  #abundance
  lionfish_fk_abun = getDomainAbundance(RVCdata_FK, "PTE VOLI", merge_protected = F)
  write_csv(lionfish_fk_abun, "trophic_groups/abundance/lionfish_fk_abun.csv")
  
  #density
  lionfish_fk_den = getDomainDensity(RVCdata_FK, "PTE VOLI", merge_protected = F)
  write_csv(lionfish_fk_den, "trophic_groups/density/lionfish_fk_den.csv")
  
  #biomass
  lionfish_fk_bio = getDomainBiomass(RVCdata_FK, "PTE VOLI", RVCdata_FK$taxonomic_data)
  write_csv(lionfish_fk_bio, "trophic_groups/biomass/lionfish_fk_bio_merged.csv")
}

```

#### Opportunists 

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Alectis ciliaris*	            | african pompano	    | ALE CILI  |
|*Caranx crysos*	              | blue runner	        | CAR CRYS  |
|*Caranx bartholomaei*          |	yellow jack	        | CAR BART  |
|*Caranx latus*                 |	horse-eye jack	    | CAR LATU  |
|*Caranx lugubris*              |	black jack	        | CAR LUGU  |
|*Caranx ruber*	                | bar jack	          | CAR RUBE  |
|*Caranx sp.*	                  | jack species	      | CAR SPE.  |
|*Decapterus punctatus*	        | round scad	        | DEC PUNC  |
|*Decapterus macarellus*	      | mackerel scad	      | DEC MACA  |
|*Elagatis bipinnulata*	        | rainbow runner	    | ELA BIPI  |
|*Euthynnus alletteratus*	      | little tunny	      | EUT ALLE  |
|*Oligoplites saurus*           |	leatherjack	        | OLI SAUR  |
|*Sphyraena barracuda*	        | great barracuda	    | SPH BARR  |
|*Rachycentron canadum*	        | cobia	              | RAC CANA  |
|*Scomberomorus regalis*	      | cero	              | SCO REGA  |
|*Sphyraena picudilla*	        | southern sennet	    | SPH PICU  |
|*Selene vomer*	                | lookdown	          | SEL VOME  |
|*Seriola rivoliana*	          | almaco jack       	| SER RIVO  |
|*Scomberomorus maculatus*	    | spanish mackerel  	| SCO MACU  |
|*Seriola dumerili*	            | greater amberjack 	| SER DUME  |
|*Scomberomorus cavalla*	      | king mackerel	      | SCO CAVA  |
|*Sphyraena guachancho*	        | guaguanche	        | SPH GUAC  |
|*Seriola sp.*	                | jack species        |	SER SPE.  |
|*Seriola zonata*	              | banded rudderfish	  | SER ZONA  |
|*Trachinotus falcatus*	        | permit	            | TRA FALC  |

```{r Opportunists}

opportunists_fk_abun_csv     = "trophic_groups/abundance/opportunists_fk_abun.csv"
opportunists_fk_den_csv      = "trophic_groups/density/opportunists_fk_den.csv"
opportunists_fk_bio_csv      = "trophic_groups/biomass/opportunists_fk_den_merged.csv"
opportunists_fk_diversity_csv= "trophic_groups/diversity/opportunists_fk_diversity.csv"

if(!all(file.exists(opportunists_fk_abun_csv,opportunists_fk_den_csv,opportunists_fk_bio_csv ,opportunists_fk_diversity_csv))){
  
  opportunists_spp_list = c("CAR RUBE", "SPH BARR", "ALE CILI","DEC PUNC", "DEC MACA", "CAR CRYS", "CAR BART", "SCO REGA","SPH PICU", "CAR LATU", "SEL VOME", "ELA BIPI", "CAR SPE.", "TRA FALC", "EUT ALLE", "SER RIVO", "SCO MACU", "SER DUME", "SCO CAVA", "SPH GUAC", "CAR LUGU", "OLI SAUR", "SER SPE.", "Rac cana", "Ser zona")
  
  #abundance
  opportunists_fk_abun = getDomainAbundance(RVCdata_FK, opportunists_spp_list, merge_protected = F)
  write_csv(opportunists_fk_abun, "trophic_groups/abundance/opportunists_fk_abun.csv")
  
  #density
  opportunists_fk_den = getDomainDensity(RVCdata_FK, opportunists_spp_list, merge_protected = F)
  write_csv(opportunists_fk_den, "trophic_groups/density/opportunists_fk_den.csv")
  
  #biomass
  opportunists_fk_bio = getDomainBiomass(RVCdata_FK, opportunists_spp_list, RVCdata_FK$taxonomic_data)
  write_csv(opportunists_fk_bio, "trophic_groups/biomass/opportunists_fk_bio_merged.csv")
  
  #diversity
  opportunists_fk_abundance = opportunists_fk_abun %>% 
    select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
    group_by(YEAR, protected_status) %>% #48 groups 
    nest(-YEAR) %>%  
    mutate(
      data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
      #species richness 
      richness = map(
        data_wide, 
        function(x) specnumber(x)),
      #simpson diversity as effective number of species 
      simpson = map( 
        data_wide,
        function(x) 1/(1 - diversity(x, index = 'simpson'))),
      #shannon diversity as effective number of species 
      shannon = map(
        data_wide,
        function(x) exp(diversity(x, index = 'shannon')))) %>%
    unnest(richness, simpson, shannon)
  
  opportunists_fk_diversity = opportunists_fk_abundance %>%
    select(YEAR, protected_status, richness, simpson, shannon) 
  write_csv(opportunists_fk_diversity, "trophic_groups/diversity/opportunists_fk_diversity.csv")
}

```

#### Parrotfish

| Scientific name               | Common name           | RVC code  |  
|:------------------------------|:----------------------|:----------|
|*Cryptotomus roseus*	          | bluelip parrotfish	  | CRY ROSE  |
|*Nicholsina usta*	            | emerald parrotfish	  | NIC USTA  |
|*Scarus coelestinus*	          | midnight parrotfish	  | SCA COEL  |
|*Scarus coeruleus*	            | blue parrotfish	      | SCA COER  |
|*Scarus guacamaia*	            | rainbow parrotfish	  | SCA GUAC  |
|*Scarus iseri*	                | striped parrotfish	  | SCA ISER  |
|*Scarus sp.*	                  | parrotfish species	  | SCA SPE.  |
|*Scarus taeniopterus*          |	princess parrotfish	  | SCA TAEN  |
|*Scarus vetula*	              | queen parrotfish	    | SCA VETU  |
|*Sparisoma atomarium*	        | greenblotch parrotfish|	SPA ATOM  |
|*Sparisoma aurofrenatum*	      | redband parrotfish	  | SPA AURO  |
|*Sparisoma chrysopterum*	      | redtail parrotfish	  | SPA CHRY  |
|*Sparisoma radians*	          | bucktooth parrotfish	| SPA RADI  |
|*Sparisoma rubripinne*	        | yellowtail parrotfish	| SPA RUBR  |
|*Sparisoma sp.*	              | parrotfish species	  | SPA SPE.  |
|*Sparisoma viride*	            | stoplight parrotfish	| SPA VIRI  |

```{r Parrotfish}

parrotfish_fk_abun_csv      = "trophic_groups/abundance/parrotfish_fk_abun.csv"
parrotfish_fk_den_csv       = "trophic_groups/density/parrotfish_fk_den.csv"
parrotfish_fk_bio_csv       = "trophic_groups/biomass/parrotfish_fk_bio_merged.csv"
parrotfish_fk_diversity_csv = "trophic_groups/diversity/parrotfish_fk_diversity.csv"

if(!all(file.exists(parrotfish_fk_abun_csv,parrotfish_fk_den_csv,parrotfish_fk_bio_csv, parrotfish_fk_diversity_csv))){
  
  parrotfish_spp_list = c('SCA ISER', "SCA COEL", "SPA AURO", "SPA VIRI", "SPA ATOM", "SCA TAEN", 'SPA RUBR', "SPA CHRY", "SCA VETU", "SCA COER", "SCA GUAC", "CRY ROSE", "SPA RADI", "SCA SPE.", "NIC USTA", "SPA SPE.")
  
  #abundance
  parrotfish_fk_abun = getDomainAbundance(RVCdata_FK, parrotfish_spp_list, merge_protected = F)
  write_csv(parrotfish_fk_abun, "trophic_groups/abundance/parrotfish_fk_abun.csv")
  
  #density
  parrotfish_fk_den = getDomainDensity(RVCdata_FK, parrotfish_spp_list, merge_protected = F)
  write_csv(parrotfish_fk_den, "trophic_groups/density/parrotfish_fk_den.csv")
  
  #biomass
  parrotfish_fk_bio = getDomainBiomass(RVCdata_FK, parrotfish_spp_list, RVCdata_FK$taxonomic_data)
  write_csv(parrotfish_fk_bio, "trophic_groups/biomass/parrotfish_fk_bio_merged.csv")
  
  #diversity
  parrotfish_fk_abundance = parrotfish_fk_abun %>% 
    select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
    group_by(YEAR, protected_status) %>% #48 groups 
    nest(-YEAR) %>%  
    mutate(
      data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
      #species richness 
      richness = map(
        data_wide, 
        function(x) specnumber(x)),
      #simpson diversity as effective number of species 
      simpson = map( 
        data_wide,
        function(x) 1/(1 - diversity(x, index = 'simpson'))),
      #shannon diversity as effective number of species 
      shannon = map(
        data_wide,
        function(x) exp(diversity(x, index = 'shannon')))) %>%
    unnest(richness, simpson, shannon)
  
  parrotfish_fk_diversity = parrotfish_fk_abundance %>%
    select(YEAR, protected_status, richness, simpson, shannon) 
  write_csv(parrotfish_fk_diversity, "trophic_groups/diversity/parrotfish_fk_diversity.csv")
}

```

#### Rays

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Aetobatus narinari*	          | spotted eagle ray	  | AET NARI  |
|*Dasyatis americana*	          | southern stingray	  | DAS AMER  |
|*Echeneis naucrates*	          | sharksucker	        | ECH NAUC  |
|*Narcine bancroftii*	          | lesser electric ray	| NAR BANC  |
|*Rhinoptera bonasus*	          | cownosed ray	      | RHI BONA  |
|*Urobatis jamaicensis*	        | yellow stingray	    | URO JAMA  |

```{r Rays}

rays_fk_abun_csv     = "trophic_groups/abundance/rays_fk_abun.csv"
rays_fk_den_csv      = "trophic_groups/density/rays_fk_den.csv"
rays_fk_bio_csv      = "trophic_groups/biomass/rays_fk_bio_merged.csv"

if(!all(file.exists(rays_fk_abun_csv,rays_fk_den_csv, rays_fk_bio_csv))){
  
  rays_spp_list = c("URO JAMA", "ECH NAUC", "DAS AMER", "AET NARI", "NAR BANC", "Rhi bona")
  
  #abundance
  rays_fk_abun = getDomainAbundance(RVCdata_FK, rays_spp_list, merge_protected = F)
  write_csv(rays_fk_abun, "trophic_groups/abundance/rays_fk_abun.csv")
  
  #density
  rays_fk_den = getDomainDensity(RVCdata_FK, rays_spp_list, merge_protected = F)
  write_csv(rays_fk_den, "trophic_groups/density/rays_fk_den.csv")
  
  #biomass
  rays_fk_bio = getDomainBiomass(RVCdata_FK, rays_spp_list, RVCdata_FK$taxonomic_data)
  write_csv(rays_fk_bio, "trophic_groups/biomass/rays_fk_bio_merged.csv")
}
```

#### Sea bass and Hamlets

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Alphestes afer*	              | mutton hamlet	      | ALP AFER  |
|*Centropristis ocyurus*	      | bank sea bass	      | CEN OCYU  |
|*Centropristis striata*        |	black sea bass	    | CEN STRI  |
|*Diplectrum formosum*          |	sand perch	        | DIP FORM  |
|*Epinephelus adscensionis*     |	rock hind	          | EPI ADSC  |
|*Epinephelus drummondhayi*     |	speckled hind	      | EPI DRUM  |
|*Epinephelus guttatus*         |	red hind	          | EPI GUTT  |
|*Hypoplectrus chlorurus*	      | yellowtail hamlet	  | HYP CHLO  |
|*Hypoplectrus gemma*	          | blue hamlet	        | HYP GEMM  |
|*Hypoplectrus guttavarius*	    | shy hamlet	        | HYP GUTT  |
|*Hypoplectrus hybrid*          |	hybrid hamlet	      | HYP HYBR  |
|*Hypoplectrus indigo*          |	indigo hamlet	      | HYP INDI  |
|*Hypoplectrus nigricans*       |	black hamlet	      | HYP NIGR  |
|*Hypoplectrus puella*          |	barred hamlet	      | HYP PUEL  |
|*Hypoplectrus sp.*             |	hamlet species	    | HYP SPE.  |
|*Hypoplectrus tann*            |	tan hamlet	        | HYP TANN  |
|*Hypoplectrus unicolor*        |	butter hamlet	      | HYP UNIC  |
|*Liopropoma eukrines*          |	wrasse basslet	    | LIO EUKR  |
|*Liopropoma mowbrayi*          |	cave basslet	      | LIO MOWB  |
|*Liopropoma rubre*	            | peppermint basslet	| LIO RUBE  |
|*Paranthias furcifer*	        | atlantic creolefish	| PAR FURC  |
|*Pristipomoides aquilonaris*	  | wenchman	          | PRI AQUI  |
|*Rypticus maculatus*	          | whitespotted soapfish	| RYP MACU  |
|*Rypticus saponaceus*	        | greater soapfish	  | RYP SAPO  |
|*Schultzea beta*	              | school bass	        | SCH BETA  |
|*Serranus annularis*	          | orangeback bass	    | SER ANNU  |
|*Serranus baldwini*	          | lantern bass	      | SER BALD  |
|*Serranus subligarius*	        | belted sandfish	    | SER SUBL  |
|*Serranus tabacarius*	        | tobaccofish	        | SER TABA  |
|*Serranus tigrinus*	          | harlequin bass	    | SER TIGR  |
|*Serranus tortugarum*	        | chalk bass	        | SER TORT  |

```{r Sea bass and Hamlets}

seabass_fk_abun_csv     = "trophic_groups/abundance/seabass_fk_abun.csv"
seabass_fk_den_csv      = "trophic_groups/density/seabass_fk_den.csv"
seabass_fk_bio_csv      = "trophic_groups/biomass/seabass_fk_bio_merged.csv"
seabass_fk_diversity_csv= "trophic_groups/diversity/seabass_fk_diversity.csv"

if(!all(file.exists(seabass_fk_abun_csv,seabass_fk_den_csv,seabass_fk_bio_csv, seabass_fk_diversity_csv))){
  
  sea_bass_hamlet_spp_list = c("SER TIGR","HYP GEMM","HYP UNIC","SER TABA","SER BALD","HYP PUEL","EPI ADSC","SER TORT","EPI GUTT","HYP NIGR","DIP FORM","HYP SPE.","SCH BETA","HYP TANN", "RYP SAPO","PAR FURC","HYP INDI","HYP GUTT","ALP AFER","LIO EUKR","SER ANNU","LIO RUBE","HYP HYBR","HYP CHLO","PRI AQUI","RYP MACU","SER SUBL","CEN OCYU","CEN STRI","Epi drum","Lio mowb")
  
  #abundance
  seabass_fk_abun = getDomainAbundance(RVCdata_FK, sea_bass_hamlet_spp_list, merge_protected = F)
  write_csv(seabass_fk_abun, "trophic_groups/abundance/seabass_fk_abun.csv")
  
  #density
  seabass_fk_den = getDomainDensity(RVCdata_FK, sea_bass_hamlet_spp_list, merge_protected = F)
  write_csv(seabass_fk_den, "trophic_groups/density/seabass_fk_den.csv")
  
  #biomass
  seabass_fk_bio = getDomainBiomass(RVCdata_FK, sea_bass_hamlet_spp_list,RVCdata_FK$taxonomic_data)
  write_csv(seabass_fk_bio, "trophic_groups/biomass/seabass_fk_bio_merged.csv")
  
  #diversity
  seabass_fk_abundance = seabass_fk_abun %>% 
    select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
    group_by(YEAR, protected_status) %>% #48 groups 
    nest(-YEAR) %>%  
    mutate(
      data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
      #species richness 
      richness = map(
        data_wide, 
        function(x) specnumber(x)),
      #simpson diversity as effective number of species 
      simpson = map( 
        data_wide,
        function(x) 1/(1 - diversity(x, index = 'simpson'))),
      #shannon diversity as effective number of species 
      shannon = map(
        data_wide,
        function(x) exp(diversity(x, index = 'shannon')))) %>%
    unnest(richness, simpson, shannon)
  
  seabass_fk_diversity = seabass_fk_abundance %>%
    select(YEAR, protected_status, richness, simpson, shannon) 
  write_csv(seabass_fk_diversity, "big_csv/trophic_groups/diversity/seabass_fk_diversity.csv")
}

```

#### Sharks

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Carcharhinus falciformis*	    | silky shark	        | CAR FALC  |
|*Carcharhinus leucas*	        | bull shark	        | CAR LEUC  |
|*Carcharhinus limbatus*	      | blacktip shark	    | CAR LIMB  |
|*Carcharhinus obscurus*	      | dusky shark	        | CAR OBSC  |
|*Carcharhinus plumbeus*	      | sandbar shark	      | CAR PLUM  |
|*Carcharhinus perezii*	        | reef shark	        | CAR PERE  |
|*Galeocerdo cuvieri*	          | tiger shark	        | GAL CUVI  | 
|*Ginglymostoma cirratum*	      | nurse shark	        | GIN CIRR  |
|*Negaprion brevirostris*	      | lemon shark	        | Neg BREV  |
|*Sphyrna mokarran*	            | great hammerhead	  | SPH MOKA  |
|*Sphyrna tiburo*	              | bonnethead	        | SPH TIBU  |
|*Sphyrna lewini*	              | scalloped hammerhead| SPH LEWI  |

```{r Sharks}

sharks_fk_abun_csv      = "trophic_groups/abundance/sharks_fk_abun.csv"
sharks_fk_den_csv       = "trophic_groups/density/sharks_fk_den.csv"
sharks_fk_bio_csv       = "trophic_groups/biomass/sharks_fk_bio_merged.csv"

if(!all(file.exists(sharks_fk_abun_csv,sharks_fk_den_csv,sharks_fk_bio_csv))){
  
  sharks_spp_list = c("GIN CIRR", "CAR PERE", "CAR LEUC", "CAR LIMB", "SPH MOKA", "SPH TIBU", "CAR FALC", "SPH LEWI", "CAR OBSC", "Neg brev", "CAR PLUM", 'Gal cuvi')
  
  #abundance
  sharks_fk_abun = getDomainAbundance(RVCdata_FK, sharks_spp_list, merge_protected = F)
  write_csv(sharks_fk_abun, "trophic_groups/abundance/sharks_fk_abun.csv")
  
  #density
  sharks_fk_den = getDomainDensity(RVCdata_FK, sharks_spp_list, merge_protected = F)
  write_csv(sharks_fk_den, "trophic_groups/density/sharks_fk_den.csv")
  
  #biomass
  sharks_fk_bio = getDomainBiomass(RVCdata_FK, sharks_spp_list, RVCdata_FK$taxonomic_data)
  write_csv(sharks_fk_bio, "trophic_groups/biomass/sharks_fk_bio_merged.csv")
}

```

#### Small Reef Fish

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|

```{r Small Reef Fish}

small_reef_fk_abun_csv      = "trophic_groups/abundance/small_reef_fk_abun.csv"
small_reef_fk_den_csv       = "trophic_groups/density/small_reef_fk_den.csv"
small_reef_fk_bio_csv       = "trophic_groups/biomass/small_reef_fk_bio_merged.csv"
small_reef_fk_diversity_csv = "trophic_groups/diversity/small_reef_fk_diversity.csv"

if(!all(file.exists(small_reef_fk_abun_csv,small_reef_fk_den_csv,small_reef_fk_bio_csv, small_reef_fk_diversity_csv))){ 
  
  small_reef_fish_spp_list = c("COR PERS","ACA COER","KYP SECT","ACA BAHI","ACA CHIR","COR GLAU","ELA OCEA","GNA THOM","PTE CALL","ACA SPE.","MAL TRIA","COR DICR","PAR MARM","MAL MACR","OPH MACC","CTE SAEP","MIC CARR","PTE HELE","HEM SIMU","SCA CRIS","COR EIDO","GOB SPE.","TYL CROC","MIC MICR","ACA ASPE","COR SPE.","BLE SPE.","STR NOTA","MEL NIGE","HYP BERM","EMB PAND","ELA EVEL","ACA CHAP","EMB BAHA","COR LIPE","NES LONG","ENN BOEH","ACA MARI","LAB NUCH","MAL VERS","MAL GILL","ELA MACR","PRI HIPO","GOB DILE","ELA SAUC","ELA XANT","STR TIMU","MAL AURO","PAR MARN","CHA LIMB","OXY STIG","ELA HORS","PAR NIGR","COR PUNC","ACA SPIN","BOL BOQU","ElA RAND","LAB KALI","LAB NIGR","ENN ALTI")
  
  #abundance
  small_reef_fk_abun = getDomainAbundance(RVCdata_FK, small_reef_fish_spp_list, merge_protected = F)
  write_csv(small_reef_fk_abun, "trophic_groups/abundance/small_reef_fk_abun.csv")
  
  #density
  small_reef_fk_den = getDomainDensity(RVCdata_FK, small_reef_fish_spp_list, merge_protected = F)
  write_csv(small_reef_fk_den, "trophic_groups/density/small_reef_fk_den.csv")
  
  #biomass
  small_reef_fk_bio = getDomainBiomass(RVCdata_FK, small_reef_fish_spp_list, RVCdata_FK$taxonomic_data)
  write_csv(small_reef_fk_bio, "trophic_groups/biomass/small_reef_fk_bio_merged.csv")
  
  #fk diversity
  small_reef_fk_abundance = small_reef_fk_abun %>% 
    select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
    group_by(YEAR, protected_status) %>% #48 groups 
    nest(-YEAR) %>%  
    mutate(
      data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
      #species richness 
      richness = map(
        data_wide, 
        function(x) specnumber(x)),
      #simpson diversity as effective number of species 
      simpson = map( 
        data_wide,
        function(x) 1/(1 - diversity(x, index = 'simpson'))),
      #shannon diversity as effective number of species 
      shannon = map(
        data_wide,
        function(x) exp(diversity(x, index = 'shannon')))) %>%
    unnest(richness, simpson, shannon)
  
  small_reef_fk_diversity = small_reef_fk_abundance %>%
    select(YEAR, protected_status, richness, simpson, shannon) 
  write_csv(small_reef_fk_diversity, "big_csv/trophic_groups/diversity/small_reef_fk_diversity.csv")
}

```

#### Snappers

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Lutjanus analis*              |	mutton snapper	    | LUT ANAL  |
|*Lutjanus apodus*	            | schoolmaster	      | LUT APOD  |
|*Lutjanus buccanella*	        | blackfin snapper	  | LUT BUCC  |
|*Lutjanus campechanus*	        | red snapper	        | Lut CAMP  |
|*Lutjanus cyanopterus*	        | cubera snapper	    | LUT CYAN  |
|*Lutjanus griseus*	            | gray snapper	      | LUT GRIS  |
|*Lutjanus jocu*	              | dog snapper	        | LUT JOCU  |
|*Lutjanus mahogoni*	          | mahogany snapper	  | LUT MAHO  |
|*Lutjanus synagris*	          | lane snapper	      | LUT SYNA  |
|*Ocyurus chrysurus*	          | yellowtail snapper	| OCY CHRY  |
|*Rhomboplites aurorubens*	    | vermilion snapper	  | RHO AURO  |
|*snapper species*	            | snapper species	    | LUT SPE.  |

```{r Snappers}

snapper_fk_abun_csv      = "trophic_groups/abundance/snapper_fk_abun.csv"
snapper_fk_den_csv       = "trophic_groups/density/snapper_fk_den.csv"
snapper_fk_bio_csv       = "trophic_groups/biomass/snapper_fk_bio_merged.csv"
snapper_fk_diversity_csv = "trophic_groups/diversity/snapper_fk_diversity.csv"

if(!all(file.exists(snapper_fk_abun_csv,snapper_fk_den_csv,snapper_fk_bio_csv,snapper_fk_diversity_csv))){
  
  snapper_spp_list = c("OCY CHRY", "LUT GRIS", "LUT ANAL", "LUT APOD", "LUT SYNA", "LUT MAHO", "LUT JOCU", "LUT BUCC", 'LUT SPE.', "LUT CYAN", "RHO AURO", "Lut camp")
  
  #abundance
  snapper_fk_abun = getDomainAbundance(RVCdata_FK, snapper_spp_list, merge_protected = F)
  write_csv(snapper_fk_abun, "trophic_groups/abundance/snapper_fk_abun.csv")
  
  #density
  snapper_fk_den = getDomainDensity(RVCdata_FK, snapper_spp_list, merge_protected = F)
  write_csv(snapper_fk_den, "trophic_groups/density/snapper_fk_den.csv")
  
  #biomass
  snapper_fk_bio = getDomainBiomass(RVCdata_FK, snapper_spp_list, RVCdata_FK$taxonomic_data)
  write_csv(snapper_fk_bio, "trophic_groups/biomass/snapper_fk_bio_merged.csv")
  
  #diversity
  snapper_fk_abundance = snapper_fk_abun %>% 
    select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
    group_by(YEAR, protected_status) %>% #48 groups 
    nest(-YEAR) %>%  
    mutate(
      data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
      #species richness 
      richness = map(
        data_wide, 
        function(x) specnumber(x)),
      #simpson diversity as effective number of species 
      simpson = map( 
        data_wide,
        function(x) 1/(1 - diversity(x, index = 'simpson'))),
      #shannon diversity as effective number of species 
      shannon = map(
        data_wide,
        function(x) exp(diversity(x, index = 'shannon')))) %>%
    unnest(richness, simpson, shannon)
  
  snapper_fk_diversity = snapper_fk_abundance %>%
    select(YEAR, protected_status, richness, simpson, shannon) 
  write_csv(snapper_fk_diversity, "trophic_groups/diversity/snapper_fk_diversity.csv")
}
```

## Dygraphs 

### Trophic Groups diversity dygraphs 
```{r Density dygraphs, eval = F}

# read in all density csv files for trophic groups
path = "trophic_groups/density/"
files = list.files(path=path, pattern = "*.csv")
for(file in files){
  perpos <- which(strsplit(file, "")[[1]]==".")
  assign(
    gsub(" ","",substr(file, 1, perpos-1)), 
    read.csv(paste(path, file, sep= "")))  
}

#summarize density and spread so it is in data wide format 
for (i in 1:length(files)){
  select(YEAR, protected_status, SPECIES_CD, density) %>% 
    group_by(YEAR, protected_status) %>%
    summarise(
      ntot = sum(density)) %>%
    filter(protected_status != "0") %>%
    filter(protected_status != "1") %>% 
    mutate(
      protected_status = recode(
        protected_status, '0'='unprotected', '1'='protected')) %>%
    spread(protected_status, ntot)
} 

dygraph(higher_reef_fk_density, main = 'Higher Level Reef Fish Density') %>% 
    dyAxis("y", label = "Density") %>%
    dyAxis("x", label = "Year") %>%
    dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 2)

dygraph(higher_reef_fk_density, main = 'Higher Level Reef Fish Density') %>% 
    dyAxis("y", label = "Density") %>%
    dyAxis("x", label = "Year") %>%
    dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 2)

dygraph(parrotfish_fk_density, main = 'Parrotfish Density') %>% 
    dyAxis("y", label = "Density") %>%
    dyAxis("x", label = "Year") %>%
    dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 2)

dygraph(sharks_fk_density, main = 'Sharks Density') %>% 
    dyAxis("y", label = "Density") %>%
    dyAxis("x", label = "Year") %>%
    dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 2)

dygraph(snapper_fk_density, main = 'Snapper Density') %>% 
    dyAxis("y", label = "Density") %>%
    dyAxis("x", label = "Year") %>%
    dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 2)


```

### Trophic Groups biomass dygraphs 
```{r Biomass dygraphs}
# read in all biomass csv files for trophic groups
path = "trophic_groups/biomass/"
files = list.files(path=path, pattern = "*.csv")
for(file in files){
  perpos <- which(strsplit(file, "")[[1]]==".")
  assign(
    gsub(" ","",substr(file, 1, perpos-1)), 
    read.csv(paste(path, file, sep= "")))  
}

```

### Trophic Groups diversity dygraphs 
```{r Diversity dygraphs}
# read in all diversity csv files for trophic groups
path = "trophic_groups/diversity/"
files = list.files(path=path, pattern = "*.csv")
for(file in files){
  perpos <- which(strsplit(file, "")[[1]]==".")
  assign(
    gsub(" ","",substr(file, 1, perpos-1)), 
    read.csv(paste(path, file, sep= "")))  
}

```

## Grouper Test w/ BB

Metrics:
- abundance
- density 
- richness
- Simpson
- Shannon

- Q: Have multiple groups, so how to look at single metric across groups?
  A: TrelliscopeJS to the rescue!
  
- Protected areas: in/out

For `big_csv/trophic_groups/abundance/*_*_abun.csv`: 
- `YEAR`
- `REGION`: FLA KEYS, or Dry Tortugas
- `SPECIES_CD`: species code
- `abundance`: units?
- `var`: Variance in average density per secondary sampling unit
- `n`: Number of primary sampling units sampled
- `nm`: Number of secondary sampling units sampled
- `N`: Number of possible primary sample units
- `NM`: Number of possible secondary sampling units
- `protected_status`: The protected status. Only present if merge_protected is FALSE
- `density`: Average density per secondary sampling unit

So far, per species. Then when summing to # of individuals across species, lose biomass idea.

TODO:
- get weighted biomass as density!
- sum abundance across species

Plotting with [htmlwidgets: leaflet](http://www.htmlwidgets.org/showcase_leaflet.html):
- [dygraphs for R](http://rstudio.github.io/dygraphs/)

```{r test plot, eval = F}

#csv = "big_csv/trophic_groups/abundance/grouper_fk_abun.csv"
grouper = read_csv("trophic_groups/diversity/grouper_fk_diversity.csv")
grouper_abun = read_csv("trophic_groups/abundance/grouper_fk_abun.csv")

grouper_abundance = grouper_abun %>%
  select(YEAR,protected_status, SPECIES_CD,abundance) %>% 
  group_by(YEAR, protected_status) %>%
  summarise(
  ntot = sum(abundance)) %>%
  filter(protected_status != "all") %>%
  mutate(
    protected_status = recode(
      protected_status, '0'='unprotected', '1'='protected')) %>%
  spread(protected_status, ntot)

dygraph(grouper_abundance, main = 'Grouper Abundance') %>% 
    dyAxis("y", label = "Abundance") %>%
    dyAxis("x", label = "Year") %>%
    dyOptions(stackedGraph = T)
  

source('~/github/cr-metrics/functions.R')

plot_timeseries(
  title   = 'Grouper Richness',
  y_label = 'Richness',
  x_label = 'Year',
  csv_tv  = csv,
  col_t   = 'YEAR',
  filter  = 'protected_status == "all"',
  col_y   = 'grouper_richness',
  skip    = 0)

grouper = grouper %>%
  filter(protected_status != 'all') %>%
  select(YEAR, protected_status, shannon) %>%
  mutate(
    protected_status = recode(
      protected_status, '0'='unprotected', '1'='protected')) %>%
  spread(protected_status, shannon)

dygraph(grouper, main = 'Grouper Shannon Diversity') %>% 
    dyAxis("y", label = "Effective Number of Species") %>%
    dyAxis("x", label = "Year") %>%
     dyOptions(stackedGraph = T)%>% 
    dyRangeSelector(height = 20) 
  
```

## Trophic Level 
```{r Trophic levels, eval = F}

trophic_level_bio_csv = "trophic_level_bio.csv"
  
if(!file.exists(trophic_level_bio_csv)){ 
  
#potential visualize as a stacked chart 
trophic_levels = read_csv('trophic_level/RVC_trophic_level.csv')
# Parsed with column specification:
# cols(
#   Latin_Name = col_character(),
#   Common_Name = col_character(),
#   RVC_CODE = col_character(),
#   Trophic_Level = col_integer()
# )

trophic_level = trophic_levels %>%
  select(RVC_CODE, Trophic_Level) %>%  #391 species 
  mutate(
    TL = # 3 trophic levels 
      ifelse(Trophic_Level == 2, "TL2",
                    ifelse(Trophic_Level == 3, "TL3",
                           ifelse(Trophic_Level == 4, "TL4", NA))),
    SPECIES_CD = RVC_CODE) %>% 
  arrange(SPECIES_CD) %>% 
  select(SPECIES_CD, TL) 

#Get Biomass for the 391 species 
trophic_level_bio = getDomainBiomass(RVCdata_FK, trophic_level$RVC_CODE, RVCdata_FK$taxonomic_data)
write_csv(trophic_level_bio, "trophic_level/trophic_level_bio.csv")
#Warning messages: 
#1: In .checkSpeciesMatch(x, species, growth_parameters) : growth_parameters for ACA SOLA, ALB VULP, AST GUTT, BOL BOQU, BRO BARB, CAL LEUC, CAR OBSC, CAR PLUM, CEN OCYU, CEN STRI, CHI ATIN, CHL CHRY, EMM ATLA, ENC CARY, ENN ALTI, EPI DRUM, EUC JONE, EUC LEFR, GAL CUVI, HIP EREC, HIP REID, LIO MOWB, LUT CAMP, NAR BANC, NEG BREV, OGC NASU, OPI SPE., OPS TAU_, PRI OPHR, PRI RUBI, RAC CANA, RHI BONA, RYP BIST, RYP MACU, SAR BULL, SER ZONA, SPH NEPH, SRR SPE., STE SPE., STY LATE, SYN SCOV, SYN SYNO, TRA GOOD, TRA LATH, UMB CORO, UPE PARV unavailable
# 2: In .wrapperProto(x, species, length_bins, merge_protected, getDomainBiomass,: length mismatch between species provided and species found.
# 391 species provided 390 species found.

trophic_level_bio = read_csv("trophic_level/trophic_level_bio.csv")
# Parsed with column specification:
# cols(
#   YEAR = col_integer(),
#   REGION = col_character(),
#   SPECIES_CD = col_character(),
#   biomass = col_double(),
#   var = col_double(),
#   n = col_integer(),
#   nm = col_integer(),
#   N = col_integer(),
#   NM = col_double()
# )

trophic_level_bio = trophic_level_bio %>%
  select(YEAR, biomass, SPECIES_CD) %>% 
  left_join(trophic_level, trophic_level_bio, by = 'SPECIES_CD')

write_csv(trophic_level_bio, "trophic_level/tlbio.csv")
}

trophic_level_bio = read_csv("trophic_level/tlbio.csv")
# Parsed with column specification:
# cols(
#   YEAR = col_integer(),
#   biomass = col_double(),
#   SPECIES_CD = col_character(),
#   TL = col_character()

trophic_level_biomass = trophic_level_bio %>% 
  group_by(YEAR, TL) %>%
  summarise(
  ntot = sum(biomass)) %>% 
  spread(TL, ntot)

```

## Exploited reef fish 

| Scientific name               | Common name         | RVC code  | Comment     |
|:------------------------------|:--------------------|:----------|-------------|
|*Cephalopholis cruentata*      | Graysby             | CEP CRUE  | exploited   |
|*Epinephelus itajara*          | Goliath grouper     | EPI ITAJ  | moritorium  |
|*Epinephelus striatus*         | Nassau grouper      | EPI STRI  | moritorium  |
|*Epinephelus Morio*            | Red grouper         | EPI MORI  | exploited   |
|*Lachnolaimus maximus*         | Hogfish             | LAC MAXI  | exploited   |
|*Lutjanus analis*              | Mutton snapper      | LUT ANAL  | exploited   |
|*Lutjanus griseus*             | Grey snapper        | LUT GRIS  | exploited   |
|*Mycteroperca bonaci*          | Black grouper       | MYC BONA  | exploited   |
|*Ocyurus chrysurus*            | Yellowtail snapper  | OCY CHRY  | exploited   |

```{r Exploited reef fish}

exploited_fk_abun_csv = "exploited_species/exploited_fk_abun.csv"
exploited_fk_den_csv  = "exploited_species/exploited_fk_den.csv"
exploited_fk_bio_csv  = "exploited_species/exploited_fk_bio.csv"

if(!all(file.exists(exploited_fk_abun_csv, exploited_fk_den_csv, exploited_fk_bio_csv))){
  
  exploited_spp_list= c(
    "CEP CRUE", #Graysby
    "EPI ITAJ", #Goliath Grouper 
    "EPI MORI", #Red Grouper
    "EPI STRI", #Nassau Grouper 
    "LAC MAXI", #Hogfish
    "LUT ANAL", #Mutton Snapper
    "LUT GRIS", #Grey Snapper    
    "MYC BONA", #Black Grouper
    "OCY CHRY" #Yellowtail Snapper 
  )
  #abundance
  exploited_fk_abun <- getDomainAbundance(RVCdata_FK, species = exploited_spp_list, merge_protected = F)
  write_csv(exploited_fk_abun,"exploited_species/exploited_fk_abun.csv")
  
  #density 
  exploited_fk_den <- getDomainDensity(RVCdata_FK, species = exploited_spp_list, merge_protected = F)
  write_csv(exploited_fk_den,"exploited_species/exploited_fk_den.csv")
  
  #biomass 
  exploited_fk_bio <- getDomainBiomass(RVCdata_FK, species = exploited_spp_list, growth_parameters = RVCdata_FK$taxonomic_data, merge_protected = F)
  write_csv(exploited_fk_bio,"exploited_species/exploited_fk_bio.csv")
}

exploited_den = read_csv("exploited_species/exploited_fk_den.csv")

exploited_density = exploited_den %>%
  mutate(
    protected_status = recode(
      protected_status, '0'='unprotected', '1'='protected'), 
    common_name = recode(
      SPECIES_CD,
      'CEP CRUE' = 'Graysby',
      'EPI ITAJ' = 'Goliath_Grouper',
      'EPI MORI' = 'Red_Grouper',
      'EPI STRI' = 'Nassau_Grouper',
      'LAC MAXI' = 'Hogfish',
      'LUT ANAL' = 'Mutton_Snapper',
      'LUT GRIS' = 'Grey_Snapper',
      'MYC BONA' = 'Black_Grouper',
      'OCY CHRY' = 'Yellowtail_Snapper')) %>% 
  select(YEAR, common_name, protected_status, density) %>%
  group_by(YEAR, common_name, protected_status) %>%
  filter(protected_status != "all") %>%
  summarise(
    ntot = sum(density)) %>%
  spread(common_name, ntot)

#Graysby density    
Black_Grouper = exploited_density %>%
  select(YEAR, protected_status, Black_Grouper) %>%
  spread(protected_status, Black_Grouper)

dygraph(Black_Grouper, main = 'Black Grouper Density') %>% 
  dyAxis("y", label = "Density") %>%
  dyAxis("x", label = "Year") %>%
  dyOptions(stackedGraph = T)

ggplot(Black_Grouper, aes(x = YEAR)) +
  geom_point(aes(y = protected, color = "protected")) +
  geom_point(aes(y = unprotected, color = "unprotected")) +
  geom_line(aes(y = protected, color = "protected")) +
  geom_line(aes(y = unprotected, color = "unprotected"))+
  ggtitle("Black Grouper Density") +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold', size = 12))+
  labs(x = "Year", y = "Density", color = "Level of Protection") +
  scale_x_continuous(breaks = c(1999:2016))

#Graysby density    
Graysby = exploited_density %>%
  select(YEAR, protected_status, Graysby) %>%
  spread(protected_status, Graysby)

dygraph(Graysby, main = 'Graysby Density') %>% 
  dyAxis("y", label = "Density") %>%
  dyAxis("x", label = "Year") %>%
  dyOptions(stackedGraph = T)

#Grey_Snapper density    
Grey_Snapper = exploited_density %>%
  select(YEAR, protected_status, Grey_Snapper) %>%
  spread(protected_status, Grey_Snapper)

dygraph(Grey_Snapper, main = 'Grey Snapper Density') %>% 
  dyAxis("y", label = "Density") %>%
  dyAxis("x", label = "Year") %>%
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 2)

#Goliath grouper density 
Goliath_Grouper = exploited_density %>%
  select(YEAR, protected_status, Goliath_Grouper) %>%
  spread(protected_status, Goliath_Grouper)

dygraph(Goliath_Grouper, main = 'Goliath Grouper Density') %>% 
  dyAxis("y", label = "Density") %>%
  dyAxis("x", label = "Year") %>%
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 2)

#Hogfish density    
Hogfish = exploited_density %>%
  select(YEAR, protected_status, Hogfish) %>%
  spread(protected_status, Hogfish)

dygraph(Hogfish, main = 'Hogfish Density') %>% 
  dyAxis("y", label = "Density") %>%
  dyAxis("x", label = "Year") %>%
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 2)

##Nassau Grouper
Nassau_Grouper = exploited_density %>%
  select(YEAR, protected_status, Nassau_Grouper) %>%
  spread(protected_status, Nassau_Grouper)

dygraph(Nassau_Grouper, main = 'Nassau Grouper Density') %>% 
  dyAxis("y", label = "Density") %>%
  dyAxis("x", label = "Year") %>%
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = )

##Yellowtail Snapper 
yellowtail_snapper = exploited_density %>%
  select(YEAR, protected_status, Yellowtail_Snapper) %>%
  spread(protected_status, Yellowtail_Snapper)

dygraph(yellowtail_snapper, main = 'Yellowtail Snapper Density') %>% 
  dyAxis("y", label = "Density") %>%
  dyAxis("x", label = "Year") %>%
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 2)

#exploited reef fish density in protected areas 
exploited_density_protected = exploited_density %>%
  filter(protected_status == 'protected') %>% 
  select(YEAR, Black_Grouper, Goliath_Grouper, Graysby, Grey_Snapper, Hogfish, Mutton_Snapper, Nassau_Grouper,Red_Grouper, Yellowtail_Snapper)

dygraph(exploited_density_protected, main = "Exploited Reef Fish Density in Protected Areas") %>% 
  dyAxis("y", label = "Density") %>%
  dyAxis("x", label = "Year") %>%
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 2, colors = rainbow(9))

exploited_density_unprotected = exploited_density %>%
  filter(protected_status == 'unprotected') %>% 
  select(YEAR, Black_Grouper, Goliath_Grouper, Graysby, Grey_Snapper, Hogfish, Mutton_Snapper, Nassau_Grouper,Red_Grouper, Yellowtail_Snapper)

dygraph(exploited_density_unprotected, main = "Exploited Reef Fish Density in Protected Areas") %>% 
  dyAxis("y", label = "Density") %>%
  dyAxis("x", label = "Year") %>%
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 2, colors = grDevices::rainbow(9, start = 0, end = 1, alpha = 0.6))


##Exploited Biomass 
exploited_bio = read_csv("exploited_species/exploited_fk_bio.csv")
# Parsed with column specification:
# cols(
#   YEAR = col_integer(),
#   REGION = col_character(),
#   SPECIES_CD = col_character(),
#   biomass = col_double(),
#   var = col_double(),
#   n = col_integer(),
#   nm = col_integer(),
#   N = col_integer(),
#   NM = col_double(),
#   protected_status = col_character()
# )

exploited_biomass = exploited_bio %>%
  mutate(
    protected_status = recode(
      protected_status, '0'='unprotected', '1'='protected'), 
    common_name = recode(
      SPECIES_CD,
      'CEP CRUE' = 'Graysby',
      'EPI ITAJ' = 'Goliath_Grouper',
      'EPI MORI' = 'Red_Grouper',
      'EPI STRI' = 'Nassau_Grouper',
      'LAC MAXI' = 'Hogfish',
      'LUT ANAL' = 'Mutton_Snapper',
      'LUT GRIS' = 'Grey_Snapper',
      'MYC BONA' = 'Black_Grouper',
      'OCY CHRY' = 'Yellowtail_Snapper')) %>% 
  select(YEAR, common_name, protected_status, biomass) %>%
  group_by(YEAR, common_name, protected_status) %>%
  filter(protected_status != "all") %>%
  summarise(
    ntot = sum(biomass)) %>%
  spread(common_name, ntot)


exploited_biomass_all = exploited_bio %>% #432*10
  filter(protected_status == 'all') %>%   #144*10 
  select(YEAR, SPECIES_CD, biomass) %>%   #144*3
  mutate(
    common_name = recode(
      SPECIES_CD,
      'CEP CRUE' = 'Graysby',
      'EPI ITAJ' = 'Goliath Grouper',
      'EPI MORI' = 'Red Grouper',
      'EPI STRI' = 'Nassau Grouper',
      'LAC MAXI' = 'Hogfish',
      'LUT ANAL' = 'Mutton Snapper',
      'LUT GRIS' = 'Grey Snapper',
      'MYC BONA' = 'Black Grouper',
      'OCY CHRY' = 'Yellowtail Snapper')) %>% 
  select(YEAR, common_name, biomass) %>%
  spread(common_name, biomass)

dygraph(exploited_biomass_all, main = "Exploited Reef Fish Biomass in Protected Areas") %>% 
  dyAxis("y", label = "Density") %>%
  dyAxis("x", label = "Year") %>%
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 2) #colors = grDevices::rainbow(9, alpha = 0.6)

```

## Reef Fish Biodiversity 

```{r Biodiversity}

domain_fk_abun_diversity = read_csv('big_csv/abundance_domain/domain_fk_abun_diversity.csv')
#strata_fk_abun_diversity = read_csv("big_csv/abundance_strata/strata_fk_abun_diversity.csv")
#psu_fk_abun_diversity = read_csv("big_csv/abundance_psu/psu_fk_abun_diversity.csv",  col_types = cols(protected_status = col_character())) 

domain_fk_abun_diversity = read_csv('big_csv/abundance_domain/domain_fk_abun_diversity.csv')

## Simpson for protected and unprotected stacked 
fk_simpson = domain_fk_abun_diversity %>% 
  filter(protected_status != 'all') %>%
  select(YEAR, protected_status, dom_simpson) %>%
  mutate(
    protected_status = recode(
      protected_status, '0'='Unprotected', '1'='Protected')) %>%
  spread(protected_status, dom_simpson)

dygraph(fk_simpson, main = "Simpson Reef Fish Diversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 50)) %>%
  dyAxis("x", label = "Year") %>%
  dySeries("Protected", strokeWidth = 2, strokePattern = "dashed") %>% 
  dySeries("Unprotected", strokeWidth = 2) %>%
  dyOptions(stackedGraph = T, fillAlpha = 0.8, axisLineWidth = 2)

  #dyRangeSelector(height = 20) 

## Shannon for protected and unprotected stacked 
fk_shannon = domain_fk_abun_diversity %>% 
  filter(protected_status != 'all') %>%
  select(YEAR, protected_status, dom_shannon) %>%
  mutate(
    protected_status = recode(
      protected_status, '0'='Unprotected', '1'='Protected')) %>%
  spread(protected_status, dom_shannon)

dygraph(fk_shannon, main = "Shannon Reef Fish Diversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 75)) %>%
  dyAxis("x", label = "Year") %>%
  dySeries("Protected", strokeWidth = 2, strokePattern = "dashed") %>% 
  dySeries("Unprotected", strokeWidth = 2) %>%
  dyOptions(stackedGraph = T, fillAlpha = 0.8, axisLineWidth = 2)
  #dyRangeSelector(height = 20) 

d = domain_fk_abun_diversity %>% 
  filter(protected_status != 0) %>% 
  filter(protected_status != 1)

#Simpson for "all" filled 
dygraph_simp = d %>%
  select(YEAR, dom_simpson)

dygraph(dygraph_simp, main = "Simpson Diversity of Reef Fish in FKNMS") %>%
  dyOptions(fillGraph = T, fillAlpha = 0.4, drawPoints = T, pointSize = 2) %>%
  dySeries("dom_simpson", label = "Simpson Diversity", color = "red") %>% 
  dyAxis("y", label = "Effective Number of Species") %>% 
  dyAxis("x", label = "Year") %>%
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 2)

#Shannon for "all" filled 
dygraph_shannon = d %>%
  select(YEAR, dom_shannon)

dygraph(dygraph_shannon, main = "Shannon Diversity of Reef Fish in FKNMS") %>%
  dyOptions(fillGraph = T, fillAlpha = 0.4, drawPoints = T, pointSize = 2) %>%
  dySeries("dom_shannon", label = "Shannon Diversity", color = "blue") %>% 
  dyAxis("y", label = "Effective Number of Species") %>% 
  dyAxis("x", label = "Year") %>% 
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 2)

#Richness for "all" filled 
dygraph_richness = d %>%
  select(YEAR, dom_richness)

dygraph(dygraph_richness, main = "Species Richness of Reef Fish in FKNMS") %>%
  dyOptions(fillGraph = T, fillAlpha = 0.4, drawPoints = T, pointSize = 2) %>%
  dySeries("dom_richness", label = "Species Richness", color = "purple") %>% 
  dyAxis("y", label = "Number of Species") %>% 
  dyAxis("x", label = "Year") %>% 
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 2)

## Simpson and Shannon stacked for "all"
dygraph_simp_shannon = d %>%
  select(YEAR, dom_simpson, dom_shannon)

dygraph(dygraph_simp_shannon, main = "Florida Keys Reef Fish Biodiversity") %>% 
  dyAxis("y", label = "Effective Number of Species", valueRange = c(0, 70)) %>%
  dyAxis("x", label = "Year") %>%
  dySeries("dom_shannon", label = "Shannon Diversity", color = "blue") %>% 
  dySeries("dom_simpson", label = "Simpson Diversity", color = "red") %>% 
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 1.5) %>% #drawGrid = F
  dyLegend(width = 400)

#Richness, Shannon, Simpson stacked for "all"
dygraph_rich_simp_shannon = d %>%
  select(YEAR, dom_simpson, dom_shannon, dom_richness)

dygraph(dygraph_rich_simp_shannon, main = "Florida Keys Reef Fish Biodiversity") %>% 
  dyAxis("y", label = "Effective Number of Species") %>% 
  dyAxis("x", label = "Year") %>%
  dySeries("dom_richness", label = "Richness", color = "purple") %>%
  dySeries("dom_shannon", label = "Shannon",  color = "blue") %>% 
  dySeries("dom_simpson", label = "Simpson", color = "red") %>% 
  dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 2) %>% #drawGrid = F
  dyLegend(width = 400)

```

## Other Groups
```{r Other Groups}
#### Corals, Lobster, Queen Conch

#cremp_abund       = read.csv("C:/Users/luke.mceachron/Documents/Git/RVC/OtherGroups/CREMP_Pcover_group.csv")
#lobster = read.csv("C:/Users/luke.mceachron/Documents/Git/RVC/OtherGroups/Lobster_landings.csv")
#conch = read.csv("C:/Users/luke.mceachron/Documents/Git/RVC/OtherGroups/QueenConchData.csv")
#crichness<-read.csv("C:/Users/luke.mceachron/Documents/Git/RVC/OtherGroups/CREMP_Domain_Coral_Richness.csv")

cremp_abund = read_csv("SDG14/OtherGroups/CREMP_Pcover_group.csv")
#Parsed with column specification:
#cols(
  #Year = col_integer(),
  #Macroalgae = col_double(),
  #Cyano = col_double(),
  #Octocoral = col_double(),
  #Others = col_double(),
  #Porifera = col_double(),
  #Seagrass = col_double(),
  #`Stony Coral` = col_double(),
  #Zoanthidea = col_double(),
  #Urchins = col_double())

head(sc_pcov_spp)
names(crichness)

#Estimate Coral shannon/simpson diversity using Average site level percent cover. Richness is estimated from the full domain of CREMP samples, rather than average site level richness.

library(vegan)

sc_pcov_spp       = read.csv("C:/Users/luke.mceachron/Documents/Git/RVC/OtherGroups/CREMP_Pcover_spp_tpose.csv")
sc_pcov_spp=sc_pcov_spp[-c(1)]

for (i in 1:nrow(sc_pcov_spp)) {
  Richness<-(crichness$Domain.Richness)
  Shannon<-(diversity(sc_pcov_spp,index="shannon"))
  Shannon_Effective_Number<-(exp(diversity(sc_pcov_spp,index="shannon")))
  Simpson<-(diversity(sc_pcov_spp,index="simpson"))
  Simpson_Effective_Number<-(1/(1-(diversity(sc_pcov_spp,index="simpson"))))
  coral1<-cbind(Richness,Shannon,Shannon_Effective_Number,Simpson,Simpson_Effective_Number)
  }

years<-rbind(1996,1997,1998,1999,2000,2001,2002,2003,2004,2005,2006,2007,2008,2009,2010,2011,2012,2013,2014,2015)
colnames(years)[1]<-"Year"
coral_diversity<-cbind(coral1,years)
coral_diversity

## Everything else is basically just relative abundance data via mean site level % cover (CREMP data), landings (lobster), and Abundance per aggregation (conch)

algaecover<-cremp_abund[c(1:2)]
octocover<-cremp_abund[c(1,4)]
stonycover<-cremp_abund[c(1,8)]
lobster
conch

dygraph(algaecover, main = 'Macroalgae Mean % Cover') %>% 
    dyAxis("y", label = "Macroalgae") %>%
    dyAxis("x", label = "Year") %>%
    dyOptions(stackedGraph = T)

dygraph(octocover, main = 'Octocoral Mean % Cover') %>% 
    dyAxis("y", label = "Octocoral") %>%
    dyAxis("x", label = "Year") %>%
    dyOptions(stackedGraph = T)

dygraph(stonycover, main = 'Stony Coral Mean % Cover') %>% 
    dyAxis("y", label = "Stony Coral") %>%
    dyAxis("x", label = "Year") %>%
    dyOptions(stackedGraph = T)


lobster1=lobster[-c(4)]
dygraph(lobster1, main = 'Lobster Commercial and Recreational Landings') %>% 
    dyAxis("y", label = "Landings (Lbs.)") %>%
    dyAxis("x", label = "Year") %>%
    dySeries("Commercial_lbs", label = "Commercial Landings", color = "blue") %>% 
    dySeries("Recreational_lbs", label = "Recreational Landings", color = "red") %>% 
    dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 1.5) %>% #drawGrid = F
    dyLegend(width = 400)


conch1=conch[-c(3:4)]
dygraph(conch1, main = 'Queen Conch, Relative Abundance in Sampled Breeding Aggregations') %>% 
    dyAxis("y", label = "Abundance") %>%
    dyAxis("x", label = "Year") %>%
    dyOptions(stackedGraph = T)

#For some reason, the Y axis values are not representing the data. E.g., In 1996, the Species Richness line starts at 60 on the y axis, but the data (when you roll over) indicate Richness equals 43.

coraldiv<-as.data.frame(coral_diversity)
coraldiv1=coraldiv[-c(2,4)]

dygraph(coraldiv1, main = 'Coral Diversity') %>% 
       dyAxis("x", label = "Year") %>%
     dyAxis("y", label = "Diversity") %>%
    dySeries("Richness", label = "Species Richness", color = "blue") %>% 
    dySeries("Shannon_Effective_Number", label = "Shannon Effective Number", color = "red") %>% 
    dySeries("Simpson_Effective_Number", label = "Simpson Effective Number", color = "green") %>% 
    dyOptions(stackedGraph = T, fillAlpha = 0.6, axisLineWidth = 1.5) %>% #drawGrid = F
    dyLegend(width = 400)

```
